### 2024-03-18 paper/main.tex
import paper files

\documentclass[11pt]{article}
\usepackage{hyperref}
\usepackage{amsmath, amsfonts, amssymb, mathrsfs}
\usepackage{dcolumn}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{filemod}
\usepackage{natbib}
\usepackage[ruled, vlined]{algorithm2e}
\usepackage{floatrow}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{graphicx}

\hypersetup{
    colorlinks=true,
    linkcolor=blue, 
    urlcolor=black,
    citecolor=blue, 
    }

\oddsidemargin=0.25in
\evensidemargin=0.25in
\textwidth=7in
\textheight=8.75in
\topmargin=-.5in
\addtolength{\oddsidemargin}{-.5in}
\addtolength{\evensidemargin}{-.5in}
\footskip=0.5in
%\doublespacing

\title{Article Title}
\author{Stephen A. Walsh\thanks{Corresponding author: Division of Natural Sciences, 
        Math and Technology, Elms College, {\tt walshst@elms.edu}} \and 
        Annie S. Booth\thanks{Department of Statistics, North Carolina State University} \and
        David Higdon\thanks{Department of Statistics, Virginia Tech} \and
        Marco A.R. Ferreira\footnotemark[3]}
\date{\today}

\begin{document}

\maketitle
\bigskip

\begin{abstract} 
This document serves as a tentative outline of our anticipated DGP-cosmology paper.
\end{abstract}

%\vfill

\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{itemize}
    \item Introduce the cosmology application - why is it important?  What type of data do we have?  What do we want to do with it?
    \item Review what has been done (Mira Titan, cosmic EMU, process convolutions?)
    \item Overview our contribution (two-fold, see below) and describe the layout of the manuscript
    \item Annie says: I would prefer not to have any section labelled "review", but rather to provide relevant review and citations as we go along.  Our contribution is in the folding together of many methods, not in proposing a true novel method, so I think this will flow more nicely.
    \item Provide the link to the github repository where all our codes are provided.
\end{itemize}

\section{The data}

Example reference: \citep{damianou2013deep}

Here, we provide further details on the form of the data (in the introduction, we only briefly introduce it).  Include a figure of the raw data for a particular cosmology.  We describe the different modeling components and the regions of the space where they are trusted.  We also describe how each cosmology is a simulation based on 8 inputs.

We will handle the data in two stages.  First, we will focus on an individual cosmology and obtain the best predicted curve.  Second, we will use these curves to predict the curve for a held-out cosmology.  These two stages form the main two sections of this paper.

\section{Bayesian Hierarchical Modeling for Particular Cosmologies}

The main purpose of this section is to describe our first contribution: the use of a Bayesian hierarchical model to estimate the true curve (and provide UQ) for a particular cosmology.  We focus solely on one cosmology, independent of the others.

\subsection{Gaussian Process Model}

Here, we introduce the Bayesian GP model.  We mention the tapering and the formulation of the block covariance matrix.  As well as the posterior conditioning that provides the closed-form solution for the posterior mean and variance.  We can cite Steve's dissertation.

\subsection{Deep Gaussian Process Model}

Here, we expand on the previous section by incorporating a latent layer warping.  We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.

\subsection{Simulation Study}

Here, we can provide results from a sim study comparing the GP and DGP, in which case the true curve is known.

\subsection{Actual results (but more aptly named)}

Here we can provide figures of the predicted mean and variances for a subset of the actual cosmologies.

\section{Prediction for new cosmologies}

This section details our second contribution - the use of the Bayesian DGP model of the previous section to predict the curve for held-out cosmologies.

\subsection{The model}

Here we describe the PCA-GP model.  We discuss how we use the predicted curves that we got previously as training data.

\subsection{Actual results (but more aptly named)}

Here we show results on held-out cosmologies.  We compare to cosmicEMU and Jared's convolution approach.

\section{Discussion}

Brief recap, directions for future research.

\bibliographystyle{jasa}
\bibliography{references}

\end{document}

### 2024-04-30 paper/main.tex
add more details to outline, note potential images

@@
     \item Introduce the cosmology application - why is it important?  What type of data do we have?  What do we want to do with it?
-    \item Review what has been done (Mira Titan, cosmic EMU, process convolutions?)
+    \item Review what has been done (Mira Titan, cosmic EMU, process convolutions)
     \item Overview our contribution (two-fold, see below) and describe the layout of the manuscript
-    \item Annie says: I would prefer not to have any section labelled "review", but rather to provide relevant review and citations as we go along.
-Our contribution is in the folding together of many methods, not in proposing a true novel method, so I think this will flow more nicely.
+    \item Our contribution is in the synthesis of hierarchical modeling, DGPs, PCs.
     \item Provide the link to the github repository where all our codes are provided.
@@
 We describe the different modeling components and the regions of the space where they are trusted.
-We also describe how each cosmology is a simulation based on 8 inputs.
+We also describe how each cosmology is a simulation based on 8 inputs, and the output $\mathcal{P}(k)$ is functional.
 
@@
 These two stages form the main two sections of this paper.
+
+Images:
+\begin{itemize}
+    \item Example cosmology showing 16 low res, 1 high res, 1 pert, wt avg, and where each output is deemed unbiased
+\end{itemize}
 
@@
 
+Estimate one cosmology with 50,000 runs, and use this as a starting point for each of the others
 Here we can provide figures of the predicted mean and variances for a subset of the actual cosmologies.
+
+Images:
+\begin{itemize}
+    \item Illustration of DGP fit (post mean and UQ, could show cosmicEMU as well?)
+    \item estimated warping (or, how all 111 warpings compare)
+\end{itemize}
 
@@
 
+\begin{itemize}
+    \item show breakdown of estimated mean trend, basis functions/PCs, and weights of a particular PC/BF.
+\end{itemize}
+
 \subsection{Actual results (but more aptly named)}
 
-Here we show results on held-out cosmologies.  We compare to cosmicEMU and Jared's convolution approach.
+Here we show results on held-out cosmologies.
+We compare to cosmicEMU and Jared's convolution approach.
+We can also discuss estimation of main effects, interactions amongst the 8 parameters, and variance decomposition.
+
+\begin{itemize}
+    \item compare the 4 with boxplot: our DGP training, DGP testing, vs proc conv training (aka Jared's files), proc conv testing (aka Cosmic EMU)
+    \item compare the 4's RMSEs as a function of $k$
+    \item Main effects
+\end{itemize}
 
@@
 
+\section{Appendix}
+
+The different appendices from dissertation as necessary. Can also put details of dataset in supplementary material.
+
 \bibliographystyle{jasa}

### 2024-05-15 paper/main.tex
init commit intro for paper

@@
 
-\title{Article Title}
+\title{Deep Gaussian Processes and Hierarchical Modeling with Functional Data: A Case Study in Cosmological Power Spectra}
 \author{Stephen A. Walsh\thanks{Corresponding author: Division of Natural Sciences, 
@@
 \begin{abstract} 
-This document serves as a tentative outline of our anticipated DGP-cosmology paper.
+Understanding the structure of our universe and how matter is spread out and expanding is an area of active research.
+As cosmological surveys grow in complexity, developing surrogate models to efficiently predict matter power spectra is an important research area in cosmology.
+In this work, we synthesize methods from deep Gaussian processes, Bayesian hierarchical modeling, and basis functions in a novel way to estimate and predict matter power spectra (which are functional in nature) for different cosmologies.
+Our method has favorable results compared to the benchmark cosmological emulator (Cosmic Emu), and our code is available in a repository.
 \end{abstract}
@@
 
-\begin{itemize}
-    \item Introduce the cosmology application - why is it important?  What type of data do we have?  What do we want to do with it?
-    \item Review what has been done (Mira Titan, cosmic EMU, process convolutions)
-    \item Overview our contribution (two-fold, see below) and describe the layout of the manuscript
-    \item Our contribution is in the synthesis of hierarchical modeling, DGPs, PCs.
-    \item Provide the link to the github repository where all our codes are provided.
-\end{itemize}
+To further our understanding of the structure and movement of the universe, researchers utilize numerous cosmological surveys such as the Sloan Digital Sky Survey \citep{york2000sloan} and the upcoming Nancy Grace Roman Space Telescope \citep{Dore2019WFIRST} which are continuously growing in complexity.
+With these tools, cosmologists aim to continue learning more about cosmic acceleration \citep{caldwell2009physics}.
+To this end, one important ingredient to aid in this understanding of our universe is the matter power spectrum.
 
-\section{The data}
+The matter power spectrum is a fundamental concept in cosmology and describes the distribution of matter as a function of spatial scale.
+It is often represented as a function of wavenumber $k$ (units Mpc$^{-1}$), which is inversely related to spatial scale.
+On large scales (i.e., small $k$ values), cosmic expansion dynamics and the matter power spectrum behave according to linear perturbation theory \citep{pietroni2008flowing, lesgourgues2009non}.
+On smaller scales, non-linear dynamics require the use of computationally intensive simulations; the Coyote Universe \citep{lawrence2010coyote} and more recently the Mira-Titan Universe \citep{moran2023mira} are two simulation suites dedicated to this effort.
 
-Example reference: \citep{damianou2013deep}
+Leveraging these simulations allows for emulation and prediction of the matter power spectrum under varying specifications for eight different cosmological parameters (i.e., different cosmologies).
+Estimating and understanding the influence of these parameters on cosmic expansion are areas of active research.
+In addition to presenting the Mira-Titan simulation suite, \cite{moran2023mira} builds off of previous spectrum emulation \citep{lawrence2017mira} to provide the final emulator (Cosmic Emu) based on the full suite of simulations.
+
+The Cosmic Emu emulator is constructed by (process convolutions, ...). Segue to our method.
+
+Due to the nonstationary behavior expected to be seen in power spectra (e.g., from baryonic acoustic oscillations), we expand on the previous emulation methods and propose a novel synthesis of methods to quantify uncertainty and predict matter power spectra for different cosmologies using deep Gaussian processes (DGPs).
+In order to handle multiple realizations of functional output (from the Mira-Titan simulation suite), we use hierarchical and basis functions modeling within our DGP framework to accurately estimate and predict power spectra for different cosmologies.
+When comparing to Cosmic Emu, we see favorable results for our method.
+The code to reproduce these results are available at \url{https://github.com/stevewalsh124/dgp.hm}.
+
+\section{Data}
 
@@
 
-Here, we expand on the previous section by incorporating a latent layer warping.
+Here, we expand on the previous section by incorporating a latent layer warping \citep{damianou2013deep}.
 We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.
@@
 
-Brief recap, directions for future research.
+Brief recap, directions for future research: BSS-ANOVA in lieu of DGPs, hydrodynamical sims or TCs as other applications of this methodology.
+A more detailed sensitivity analysis for the parameters, modeling $W$ across different cosmologies.
 

### 2024-05-16 paper/main.tex
add Data section draft

@@
 
-The Cosmic Emu emulator is constructed by (process convolutions, ...). Segue to our method.
+\textcolor{blue}{The Cosmic Emu emulator is constructed by (process convolutions, ...). Segue to our method after and compare/contrast with Cosmic Emu.}
 
@@
 The code to reproduce these results are available at \url{https://github.com/stevewalsh124/dgp.hm}.
+\textcolor{blue}{Please note: this repository is currently private until we hear back regarding permissions for sharing the Mira-Titan data used to train and predict this model.}
+
+The remainder of the paper is organized as follows: Section \ref{sec:data} provides more details regarding the Mira-Titan simulation suite.
+Section \ref{sec:hm_fit} describes our model and the training procedure for both toy data as well as the Mira-Titan dataset.
+With this trained model, Section \ref{sec:pred} explains our modeling approach for predictions at held-out cosmologies and compares with other methods.
+Section \ref{sec:disc} concludes with a discussion of our contributions and avenues for future research.
 
 \section{Data}
+\label{sec:data}
 
-Here, we provide further details on the form of the data (in the introduction, we only briefly introduce it).
-Include a figure of the raw data for a particular cosmology.
-We describe the different modeling components and the regions of the space where they are trusted.
-We also describe how each cosmology is a simulation based on 8 inputs, and the output $\mathcal{P}(k)$ is functional.
+The Mira-Titan simulation suite is a dataset consisting of simulated power spectra for 117 different cosmologies.
+Each of these cosmologies is specified by 8 cosmological parameter values, which we will denote as $\theta \in \mathbb{R}^8$.
+Different simulation output exists for different values of a ninth value, the redshift; in this work, we fix this value at 0.
+For each cosmology, we have a set of 18 simulated spectra: an inexpensive spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from lower-resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from a high resolution simulation ($y_h$).
+
+These spectra are represented as a function of wavenumber $k$, and hence is denoted $P(k)$.
+Following \cite{moran2023mira}, we model on the emulation space $\mathcal{P}(k)=\log_{10}\left(\frac{k^{1.5}P(k)}{2\pi^2}\right)$.
+The output is available for $0.001 \leq k \leq 5$ for $n=351$ different values of $k$, but each data type has differing values where it is deemed approximately unbiased.
+For $k<0.04$, only $y_p$ provides a reliable estimate of the spectra.
+$y_{\ell_r}$ are valid for $0.04 \leq k \leq 0.25$, and $y_h$ is valid for $0.04 \leq k \leq 5$.
+Figure \ref{fig:plot_data} shows an example of the output for a particular cosmology.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=4in]{plot_data.png}
+    \caption{The perturbation theory, low resolution runs, and high resolution run for the first training cosmology.
+The weighted average is also shown as a dashed line.
+Dotted lines at the bottom indicate where each data type is deemed approximately unbiased.}
+    \label{fig:plot_data}
+\end{figure}
+
+From the multiple low- and high-resolution spectra across all cosmologies, \cite{moran2023mira} obtain estimates for the precision of these spectra across the $n$ wavenumber values ($p_1,\dots,p_n$) using a log-log regression model, and a multiplier $c$ for the increase in precision from the low- to high- resolution output.
+Given that the perturbation theory spectra is non-stochastic, we treat this as having a precision of $10^6$.
+We leverage these precision estimates within our model and calculate a weighted average for each cosmology: $\bar y = AddThisIn$, \textcolor{blue}{where we have four diagonal covariance matrices.}
 
 We will handle the data in two stages.
-First, we will focus on an individual cosmology and obtain the best predicted curve.
-Second, we will use these curves to predict the curve for a held-out cosmology.
-These two stages form the main two sections of this paper.
-
-Images:
-\begin{itemize}
-    \item Example cosmology showing 16 low res, 1 high res, 1 pert, wt avg, and where each output is deemed unbiased
-\end{itemize}
+First, we will focus on 111 training cosmologies and obtain the best esimated spectrum for each.
+Second, we will use these estimates to predict the spectra for six held-out cosmologies.
 
 \section{Bayesian Hierarchical Modeling for Particular Cosmologies}
+\label{sec:hm_fit}
 
@@
 \subsection{Simulation Study}
+\label{subsec:sim}
 
@@
 
-\subsection{Actual results (but more aptly named)}
+\subsection{Estimating Spectra with Mira-Titan Data}
+\label{subsec:mira_fit}
 
@@
 
-\section{Prediction for new cosmologies}
+\section{Prediction for Unobserved Cosmologies}
+\label{sec:pred}
 
@@
 
-\subsection{The model}
+\subsection{PCA-GP Model}
+\label{subsec:pca}
 
@@
 
-\subsection{Actual results (but more aptly named)}
+\subsection{Predicting Spectra from Mira-Titan Data}
+\label{subsec:mira_pred}
 
@@
 \section{Discussion}
+\label{sec:disc}
 
 Brief recap, directions for future research: BSS-ANOVA in lieu of DGPs, hydrodynamical sims or TCs as other applications of this methodology.
-A more detailed sensitivity analysis for the parameters, modeling $W$ across different cosmologies.
+A more detailed sensitivity analysis for the parameters, modeling $W$ across different cosmologies, consider modeling for different redshift values other than $z=0$.
 
 \section{Appendix}
+\label{sec:apdx}
 

### 2024-05-17 plotting/plot_data_fit.R
tidy plots of data and fit

@@ -52,21 +52,22 @@ y_avg <- (1 / Lam_z) * (Lam_pt * y_pt + Lam_lo * y_lra + Lam_hi * y_hi)
 # Plot data types #
 ###################
 
-matplot(x, y_lo, col = "gray", type = "l", lty = 1, ylim = range(y_pt, y_lo, y_hi),
-        xlab = "x", ylab = "y")
-lines(x, y_pt, col = "green")
-lines(x, y_lra, col = "blue")
-lines(x, y_hi, col = "red")
-lines(x[index_list$pert.ix], rep(-1, length(index_list$pert.ix)), 
-      col = "green", lty = 3)
-lines(x[index_list$lowres.ix], rep(-1.1, length(index_list$lowres.ix)), 
-      col = "blue", lty = 3)
-lines(x[index_list$highres.ix], rep(-1.2, length(index_list$highres.ix)), 
-      col = "red", lty = 3)
-lines(x, y_avg, col = "orange", lty = 2, lwd = 4)
-legend(x = "right", legend = c("pert","low res","low res avg","hi res", "wt avg"),
-       col = c("green","gray","blue","red","orange"), lty = c(1,1,1,1,2), 
-       lwd = c(1,1,1,1,4), cex=0.75)
+matplot(log10(k), (y_lo + loess_fit$fitted)*sd_y+mean_y, col = "gray", 
+        type = "l", lty = 1, ylim = range(y_avg*sd_y+mean_y),
+        xlab=expression(log[10](k)), 
+        ylab=expression(paste("\U1D4AB","(k)")))
+lines(log10(k), y_pt, col = "green")
+lines(log10(k), y_hi, col = "red")
+lines(log10(k)[index_list$pert.ix], rep(-1, length(index_list$pert.ix)), 
+      col = "green", lty = 3, lwd=3)
+lines(log10(k)[index_list$lowres.ix], rep(-1.1, length(index_list$lowres.ix)), 
+      col = "gray", lty = 3, lwd=3)
+lines(log10(k)[index_list$highres.ix], rep(-1.2, length(index_list$highres.ix)), 
+      col = "red", lty = 3, lwd=3)
+lines(log10(k), y_avg*sd_y+mean_y, col = "orange", lty = 2, lwd = 4)
+legend(x = "right", legend = c("pert","low res","hi res", "wt avg","ind"),
[…]

### 2024-05-21 paper/main.tex
update photos in draft
add in weighted avg, fix figure label

@@
 \usepackage{hyperref}
-\usepackage{amsmath, amsfonts, amssymb, mathrsfs}
+\usepackage{amsmath, amsfonts, amssymb, mathrsfs, dsfont}
 \usepackage{dcolumn}
@@
     \centering
-    \includegraphics[width=4in]{plot_data.png}
+    \includegraphics[width=6in]{plot_data.png}
     \caption{The perturbation theory, low resolution runs, and high resolution run for the first training cosmology.
@@
 
-From the multiple low- and high-resolution spectra across all cosmologies, \cite{moran2023mira} obtain estimates for the precision of these spectra across the $n$ wavenumber values ($p_1,\dots,p_n$) using a log-log regression model, and a multiplier $c$ for the increase in precision from the low- to high- resolution output.
-Given that the perturbation theory spectra is non-stochastic, we treat this as having a precision of $10^6$.
-We leverage these precision estimates within our model and calculate a weighted average for each cosmology: $\bar y = AddThisIn$, \textcolor{blue}{where we have four diagonal covariance matrices.}
+From the multiple low- and high-resolution spectra across all cosmologies, \cite{moran2023mira} obtained estimates for the precision of these spectra across the $n$ wavenumber values ($p_1,\dots,p_n$) using a log-log regression model, and a multiplier $c$ for the increase in precision from the low- to high- resolution output.
+Given that the perturbation theory spectra is non-stochastic, we treat this as having a precision of $10^5$.
+We leverage these precision estimates within our model at each data type's unbiased wavenumber values to define $\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, $\Lambda_{\ell} = \text{diag}\left((p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, $\Lambda_h = \text{diag}\left(c(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$.
+Here, each precision matrix is diagonal and the diagonal elements consists of the precision values or zero, depending on whether the $k$ value is approximately unbiased for that data type.
+With this, we calculate a weighted average for each cosmology: $\bar y = \Lambda_S^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, where $\Lambda_S = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for the weighted average for a particular cosmology.
 
@@
 We focus solely on one cosmology, independent of the others.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=6in]{plot_fit.png}
+    \caption{95\% credible intervals for the power spectrum of model 1 (posterior mean subtracted).
+Plots of the perturbation theory, low-resolution and high-resolution runs are also shown, alongside the weighted average.}
+    \label{fig:plot_fit}
+\end{figure}
 

### 2024-05-23 paper/main.tex
first part of section 3

@@
 The code to reproduce these results are available at \url{https://github.com/stevewalsh124/dgp.hm}.
-\textcolor{blue}{Please note: this repository is currently private until we hear back regarding permissions for sharing the Mira-Titan data used to train and predict this model.}
+\textcolor{blue}{Please note: this repository is currently private until we hear back regarding permissions for sharing the Mira-Titan data alongside our code.}
 
@@
 Given that the perturbation theory spectra is non-stochastic, we treat this as having a precision of $10^5$.
-We leverage these precision estimates within our model at each data type's unbiased wavenumber values to define $\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, $\Lambda_{\ell} = \text{diag}\left((p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, $\Lambda_h = \text{diag}\left(c(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$.
+We leverage these precision estimates within our model at each data type's unbiased wavenumber values to define $\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, $\Lambda_{\ell} = \text{diag}\left(16(p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, $\Lambda_h = \text{diag}\left(c(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$.
 Here, each precision matrix is diagonal and the diagonal elements consists of the precision values or zero, depending on whether the $k$ value is approximately unbiased for that data type.
-With this, we calculate a weighted average for each cosmology: $\bar y = \Lambda_S^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, where $\Lambda_S = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for the weighted average for a particular cosmology.
+With this, we calculate a weighted average for each cosmology: $\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for the weighted average for a particular cosmology.
 
@@
 
-The main purpose of this section is to describe our first contribution: the use of a Bayesian hierarchical model to estimate the true curve (and provide UQ) for a particular cosmology.
+Here we describe our first contribution: the use of a Bayesian hierarchical model to estimate the underlying spectra (and quantify uncertainty) for a particular cosmology.
 We focus solely on one cosmology, independent of the others.
+
+\subsection{Gaussian Process Model}
+
+Here, we introduce the Bayesian Gaussian Process (GP) model leveraging the aggregated data $D_n \equiv \{X_n, \bar Y, \Lambda\}$ for a particular cosmology.
+We assume the matter power spectrum $S$ for a given cosmology follows a GP with mean $\mu$ and covariance matrix $\Sigma_S$.
+
+From plots of the multiple low-resolution spectra, we find they vary smoothly about some mean $S$, suggesting that a dense covariance matrix $\Sigma_\varepsilon$ may outperform a diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$.
+In previous work, we have tried multiple candidates for $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and found that a Mat\'ern covariance $\Sigma_\ell$ trained on the low-resolution runs (pre-scaled by the precision values and subtracting a LOESS-smoothed average) performs the best.
+We specify this structure within $\Sigma_\varepsilon$ and refer readers to \cite{walsh2023bayesian} for a more detailed exposition.
+This yields $\Sigma_\varepsilon=\left(\Lambda_p^{-1} + \Sigma_\ell + \Lambda_h^{-1}\right)^{-1}$, where $\Lambda_p, \Sigma_\ell^{-1}, \text{ and } \Lambda_h$ are all $n\times n$ and are 0 wherever the corresponding data type is biased.
+
+% $$
+% \mathbf{\Sigma_\varepsilon} =
+% \begin{bmatrix}
+% \Lambda_p^{-1} & \mathbf{0} & \mathbf{0} \\
+% \mathbf{0} & \mathbf{\Sigma}_{\ell} & \mathbf{0} \\
+% \mathbf{0} & \mathbf{0} & \Lambda_h^{-1}
+% \end{bmatrix}
+% $$
+
+Therefore, we model $\bar Y|S$ as a GP with mean $S$ and block-diagonal covariance matrix $\Sigma_\varepsilon$:
+
+$$
+\begin{aligned}
+\bar Y|S &\sim \mathcal{N}_n(S,\Sigma_\varepsilon) \\
+S &\sim GP\left(\mu, \Sigma_S(x,x')\right)
+\end{aligned}
+$$
+
+Upon integrating out $S$, we obtain $\bar Y \sim GP(\mu, \Sigma_S+\Sigma_\varepsilon)$.
+Additionally, this model yields the posterior distribution $S|\bar Y \sim N(m, C)$, where $C=\left(\Sigma_S^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1}$ and $m=C\left(\Sigma_\varepsilon^{-1}\bar Y+\Sigma_S^{-1}\mu\right)$.
+
+This distribution provides the posterior mean and corresponding uncertainty for $S$ within a given cosmology.
+Without this hierarchy, uncertainty would not be properly addressed; the weighted average would simply be nearly interpolated across the relatively dense set of $k$ values.
+
+\subsection{Deep Gaussian Process Model}
+
+Considering the Mira-Titan data, some nonstationarity is expected (e.g., due to baryonic acoustic oscillations when $-2 \leq \log_{10}(k) \leq -1$).
+To account for this, we incorporate a latent layer warping $W$ within the hierarchical model to incorporate nonstationary dynamics of the matter power spectra.
+When $W$ is a GP, this creates a deep Gaussian process (DGP) model \citep{damianou2013deep}.
+Additional latent layers could be considered, although one is sufficient for this dataset \citep{dunlop2018deep}.
+
+$$
+\begin{aligned}
+\bar Y|S,W &\sim \mathcal{N}_n(S,\Sigma_\varepsilon) \\
+S|W &\sim GP\left(\mu, \Sigma_S(w,w')\right) \\
+W &\sim GP\left(0, \Sigma_W(x,x')\right)
+\end{aligned}
+$$
+
+We have found that whether the prior mean of $W$ is set to 0 or $X$ \citep[which would indicate stationarity apriori, ][]{schmidt2003bayesian} does not have a substantial impact on the model fit.
+We construct this model in the compositional form, and also use MCMC to obtain full uncertainty for each of the model parameters \citep{sauer2023active}.
+
+We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.
+
+\subsection{Simulation Study}
+\label{subsec:sim}
+
+Here, we provide results from a simulation study comparing the GP and DGP models, in which case the true curve is known.
+\textcolor{blue}{I can include details for a sim study based on simulated data I construct, or I can use the dataset Dave mentioned getting to use here.}
+
+\subsection{Estimating Spectra with Mira-Titan Data}
+\label{subsec:mira_fit}
+
+Estimate one cosmology with 50,000 runs, and use this as a starting point for each of the others
+
+Here we can provide figures of the predicted mean and variances for a subset of the actual cosmologies.
+Figure \ref{fig:plot_fit} shows an example of the uncertainty for our posterior mean for the first (of 111) training cosmologies; the posterior mean is subtracted in order to visualize this uncertainty.
 
@@
 \end{figure}
-
-\subsection{Gaussian Process Model}
-
-Here, we introduce the Bayesian GP model.
-We mention the tapering and the formulation of the block covariance matrix.
-As well as the posterior conditioning that provides the closed-form solution for the posterior mean and variance.
-We can cite Steve's dissertation.
-
-\subsection{Deep Gaussian Process Model}
-
-Here, we expand on the previous section by incorporating a latent layer warping \citep{damianou2013deep}.
-We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.
-
-\subsection{Simulation Study}
-\label{subsec:sim}
-
-Here, we can provide results from a sim study comparing the GP and DGP, in which case the true curve is known.
-
-\subsection{Estimating Spectra with Mira-Titan Data}
-\label{subsec:mira_fit}
-
-Estimate one cosmology with 50,000 runs, and use this as a starting point for each of the others
-Here we can provide figures of the predicted mean and variances for a subset of the actual cosmologies.
 

### 2024-05-24 paper/main.tex
add initial pred figures

@@
 \footskip=0.5in
-%\doublespacing
+\doublespacing
 
@@
 
-\textcolor{blue}{The Cosmic Emu emulator is constructed by (process convolutions, ...). Segue to our method after and compare/contrast with Cosmic Emu.}
-
-Due to the nonstationary behavior expected to be seen in power spectra (e.g., from baryonic acoustic oscillations), we expand on the previous emulation methods and propose a novel synthesis of methods to quantify uncertainty and predict matter power spectra for different cosmologies using deep Gaussian processes (DGPs).
+In this work, we expand on the previous emulation methods and propose a novel synthesis of methods to quantify uncertainty and predict matter power spectra for different cosmologies using deep Gaussian processes (DGPs).
 In order to handle multiple realizations of functional output (from the Mira-Titan simulation suite), we use hierarchical and basis functions modeling within our DGP framework to accurately estimate and predict power spectra for different cosmologies.
@@
 
-Here we show results on held-out cosmologies.
-We compare to cosmicEMU and Jared's convolution approach.
-We can also discuss estimation of main effects, interactions amongst the 8 parameters, and variance decomposition.
+Here we show results on held-out cosmologies.  The Cosmic Emu emulator is constructed by (process convolutions, ...). Segue to our method after and compare/contrast with Cosmic Emu.
 
-\begin{itemize}
-    \item compare the 4 with boxplot: our DGP training, DGP testing, vs proc conv training (aka Jared's files), proc conv testing (aka Cosmic EMU)
-    \item compare the 4's RMSEs as a function of $k$
-    \item Main effects
-\end{itemize}
+In addition to predicting for held-out cosmologies, we can also predict over a large uniform or full factorial design to obtain estimated main effects for each parameter in $\theta$, interactions amongst the 8 parameters, and variance decomposition.
+Should this be included?
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=6in]{rmse_by_k.png}
+    \caption{Average RMSE of each method across all $k$ values.}
+    \label{fig:plot_rmse_k}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=6in]{rmse_box.png}
+    \caption{Boxplot of all RMSEs of each method across all six test models.
+From top to bottom: box plots for our training method, process convolution training method, Cosmic EMU predictions, and our DGP-PC predictions.}
+    \label{fig:plot_rmse_box}
+\end{figure}
 

### 2024-05-24 plotting/plot_pca.R

@@ -117,7 +117,7 @@ for(i in 1:6){
 # compare RMSEs for each k value, by method
 plot(log10(k),colMeans(sqrt(mses)), type="l", col="magenta", lwd=2,
      ylim = range(0,colMeans(sqrt(mses)),colMeans(sqrt(mses_emu))),
-     main = "average RMSE by k value")
+     ylab = "Average RMSE")
 lines(log10(k),colMeans(sqrt(mses_emu)), col="orange", lwd=2)
 lines(log10(k),colMeans(sqrt(mses_conv_t)), col="pink")
 lines(log10(k),colMeans(sqrt(mses_dgp_t)), col="blue")
@@ -126,8 +126,7 @@ legend("topright",legend = c("cosmicEMU","DGP-GPPC","conv_t","dgp_t"),
 
 # boxplot of RMSEs by method
 boxplot(c(sqrt(mses)),c(sqrt(mses_emu)),c(sqrt(mses_conv_t)),c(sqrt(mses_dgp_t)), 
-        names=c("DGP-GPPC","EMU","conv_t","dgp_t"), ylab="RMSE",
-        main = "all RMSEs across k values and cosmologies", 
+        names=c("DGP-GPPC","EMU","conv_t","dgp_t"), xlab="RMSE", horizontal = T,
         col = c("magenta","orange","pink","blue"))
 
 # compare average MSE for both prediction methods

### 2024-06-03 paper/main.tex
init commit run_sim.R; minor changes to draft

@@
 From plots of the multiple low-resolution spectra, we find they vary smoothly about some mean $S$, suggesting that a dense covariance matrix $\Sigma_\varepsilon$ may outperform a diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$.
-In previous work, we have tried multiple candidates for $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and found that a Mat\'ern covariance $\Sigma_\ell$ trained on the low-resolution runs (pre-scaled by the precision values and subtracting a LOESS-smoothed average) performs the best.
+In previous work, we have modeled multiple candidates for $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and found that a Mat\'ern covariance $\Sigma_\ell$ trained on the low-resolution runs (pre-scaled by the precision values and subtracting a LOESS-smoothed average) performs the best.
 We specify this structure within $\Sigma_\varepsilon$ and refer readers to \cite{walsh2023bayesian} for a more detailed exposition.
 This yields $\Sigma_\varepsilon=\left(\Lambda_p^{-1} + \Sigma_\ell + \Lambda_h^{-1}\right)^{-1}$, where $\Lambda_p, \Sigma_\ell^{-1}, \text{ and } \Lambda_h$ are all $n\times n$ and are 0 wherever the corresponding data type is biased.
+
+We transform $X$ values to fall within $[0,1]$ and $Y$ so that the mean and variance are 0 and 1, respectively.
+This allows for a strightforward specification of priors, where the hyperparameters for each Mat\'ern covariance $\Sigma_S$ and $\Sigma_\ell$ each take a vague inverse-Gamma prior.
+Specifically:
+
+\begin{align}
+\Sigma_\ell^{i,j} = \tau_\ell^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_\ell}} + \frac{5d^2}{3\theta_\ell}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_\ell}}\right),\\
+\Sigma_S^{i,j} = \tau_S^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_S}} + \frac{5d^2}{3\theta_S}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_S}}\right),
+\end{align}
+
+where $d=||x_i-x_j||^2$.
+From here, we estimate $\tau_\ell, \theta_\ell$ with maximum likelihood, and conditional on $\Sigma_\varepsilon$, we specify priors for $\tau_S, \theta_S$ following \cite{sauer2023active}:
+
+priors here
 
@@
 
-Here we describe the PCA-GP model.  We discuss how we use the predicted curves that we got previously as training data.
+Here we describe the PCA-GP model.  We discuss how we use the predicted curves that we got previously as training data. \cite{higdon2008computer, higdon2010estcosmo}
 

### 2024-06-03 fitting/run_sim.R

@@ -1,0 +1,79 @@
+###############################################################################
+# This script fits the Bayesian hierarchical model for simulated data under
+# two different functions: f1 and f2
+###############################################################################
+
+library(dgp.hm)
+
+# Simulated toy example from DPC paper
+x <- seq(0, 4, by=0.1)
+n <- length(x)
+r <- 50 # number of replicates
+
+# parameters for f1 and f2
+m1 <- runif(1, 0.5, 1.5)
+m2 <- runif(1, 0.6, 1.4)
+u1 <- runif(1, 1.5, 2.5)
+u2 <- runif(1, 0.6, 1.4)
+
+# functions f1 and f2 to generate sims
+f1 <- function(x, m1, u1, sd=0.1){
+  w <- sqrt(25 - (u1/2)^2)
+  return(m1*exp(-u1*x/2)*cos(w*x)-m1*x/5 + rnorm(length(x), sd=sd))
+}
+
+f2 <- function(x, m2, u2, sd=0.15){
+  return(exp(-m2*(x-3)^2) + exp(-u2*(x-1)^2) - 0.05*sin(x-1.9) + rnorm(length(x), sd=sd))
+}
+
+# Get n function evaluations for f1 and f2 
+y1s <- y2s <- matrix(nrow = r, ncol = n)
+for (i in 1:r) y1s[i,] <- f1(x, m1=m1, u1=u1)
+for (i in 1:r) y2s[i,] <- f2(x, m2=m2, u2=u2)
+
+# Get average and sigma
+y_avg1 <- colMeans(y1s)
+y_avg2 <- colMeans(y2s)
+Sigma1 = diag(0.1^2, n)
+Sigma2 = diag(0.15^2, n)
+
+# Plot these realizations
+matplot(x, t(y1s), ylab="f1(x)", type="l")
+matplot(x, t(y2s), ylab="f2(x)", type="l")
+
+# Fit DGP model to f1
+fit <- dgp.hm::fit_two_layer_hm(x, y_avg1, Sigma_hat = Sigma1, nmcmc = 7500)
+plot(fit)
+fit <- trim(fit, 2500, 5)
+plot(fit)
+fit <- est_true(fit)
+
+# plot estimated function alongside data and avg
+xx <- seq(0,4,length=100)
+matplot(x, t(y1s), type="l", col="gray")
+lines(xx, f1(xx, m1=m1,u1=u1,sd = 0))
+lines(x, y_avg1, col="red", lty=2)
[…]

### 2024-06-04 fitting/run_sim.R
change boxplots, 50 replicates each

@@ -59,7 +59,6 @@ lines(x, fit$lb, col="blue")
 legend(x = "topright", legend = c("data","truth", "wt avg", "95% UQ"),
        col = c("gray","black","red","blue"), lty = c(1,1,2,1))
 
-
 # Fit DGP model to f2
 fit <- dgp.hm::fit_two_layer_hm(x, y_avg2, Sigma_hat = Sigma2, nmcmc = 10000)
 plot(fit)
@@ -70,10 +69,59 @@ fit <- est_true(fit)
 # plot estimated function alongside data and avg
 xx <- seq(0,4,length=100)
 matplot(x, t(y2s), type="l", col="gray")
-lines(xx, f2(xx, m2=m2,u2=u2,sd = 0))
+lines(x, f2(x, m2=m2,u2=u2,sd = 0))
 lines(x, y_avg2, col="red", lty=2)
 lines(x, fit$m, col="blue")
 lines(x, fit$ub, col="blue")
 lines(x, fit$lb, col="blue")
 legend(x = "topright", legend = c("data","truth", "wt avg", "95% UQ"),
        col = c("gray","black","red","blue"), lty = c(1,1,2,1))
+
+# repeat for r replicates (don't plot these)
+mses1 <- c()
+for (j in 1:r) {
+  print(j)
+  m1 <- runif(1, 0.5, 1.5)
+  u1 <- runif(1, 1.5, 2.5)
+  y1s <- matrix(nrow = r, ncol = n)
+  for (i in 1:r) y1s[i,] <- f1(x, m1=m1, u1=u1)
+  y_avg1 <- colMeans(y1s)
+  var_y1 <- mean(apply(y1s, 2, var))
+  Sigma1 = diag(var_y1, n)
+  fit <- dgp.hm::fit_two_layer_hm(x, y_avg1, Sigma_hat = Sigma1, nmcmc = 7500)
+  fit <- trim(fit, 2500, 5)
+  fit <- est_true(fit)
+  mses1[j] <- mean((fit$m - f1(x, m1=m1,u1=u1,sd = 0))^2)
+}
+
+mean(mses1)
+
+# Boxplot
+boxplot(mses1, ylim=c(0,.018), main="mses1")
+grid(nx = NULL, ny = NULL,
+     col = "#ebebeb", lwd = 2, lty=1)
[…]

### 2024-06-04 fitting/run_sim.R
add both models and all settings

@@ -8,69 +8,76 @@ library(dgp.hm)
 # Simulated toy example from DPC paper
 x <- seq(0, 4, by=0.1)
 n <- length(x)
-r <- 50 # number of replicates
+model <- 1 # set to 1 for "function 1", 2 for "function 2"
+setting <- 1 #set to 1 for "Setting A", 2 for "B", 3 for "C"
+if(!(model %in% 1:2)) stop("model should be 1 or 2")
+if(!(setting %in% 1:3)) stop("setting should be 1, 2, or 3")
+r <- 5 # number of replicates
 
 # parameters for f1 and f2
 m1 <- runif(1, 0.5, 1.5)
 m2 <- runif(1, 0.6, 1.4)
 u1 <- runif(1, 1.5, 2.5)
 u2 <- runif(1, 0.6, 1.4)
+if(model == 1){
+  if(setting == 1) sdvec <- rep(0.1, n)
+  if(setting == 2) sdvec <- 1/(10*(x+1))
+  if(setting == 3){
+    sdvec <- 1/(10*(x+1))
+    sdvec <- ifelse(x<=0.4, 5e-4, sdvec)
+    sdvec <- ifelse(x>0.4 & x<3, 0.75*sdvec, sdvec)
+    sdvec <- ifelse(x>=3, 0.5*sdvec, sdvec)
+  }
+} else {
+  if(setting == 1) sdvec <- rep(0.15, n)
+  if(setting == 2) sdvec <- sqrt(abs(x-2))/10
+  if(setting == 3){
+    sdvec <- sqrt(abs(x-2))/10
+    sdvec <- ifelse(x<=0.5, 5e-4, sdvec)
+    sdvec <- ifelse(x>0.5 & x<2, 0.75*sdvec, sdvec)
+    sdvec <- ifelse(x>=2, 0.5*sdvec, sdvec)
+  }}
+
+plot(sdvec)
 
 # functions f1 and f2 to generate sims
-f1 <- function(x, m1, u1, sd=0.1){
+f1 <- function(x, m1, u1, sd=0){
   w <- sqrt(25 - (u1/2)^2)
   return(m1*exp(-u1*x/2)*cos(w*x)-m1*x/5 + rnorm(length(x), sd=sd))
 }
 
-f2 <- function(x, m2, u2, sd=0.15){
+f2 <- function(x, m2, u2, sd=0){
[…]

### 2024-06-04 fitting/run_sim.R
allow command line args

@@ -5,6 +5,11 @@
 
 library(dgp.hm)
 
+args <- commandArgs(TRUE)
+if(length(args) > 0)
+  for(i in 1:length(args))
+    eval(parse(text=args[[i]]))
+
 # Simulated toy example from DPC paper
 x <- seq(0, 4, by=0.1)
 n <- length(x)

### 2024-06-04 fitting/run_sim.R
allow command line args, correctly

@@ -5,11 +5,6 @@
 
 library(dgp.hm)
 
-args <- commandArgs(TRUE)
-if(length(args) > 0)
-  for(i in 1:length(args))
-    eval(parse(text=args[[i]]))
-
 # Simulated toy example from DPC paper
 x <- seq(0, 4, by=0.1)
 n <- length(x)
@@ -18,6 +13,11 @@ setting <- 1 #set to 1 for "Setting A", 2 for "B", 3 for "C"
 if(!(model %in% 1:2)) stop("model should be 1 or 2")
 if(!(setting %in% 1:3)) stop("setting should be 1, 2, or 3")
 r <- 5 # number of replicates
+
+args <- commandArgs(TRUE)
+if(length(args) > 0)
+  for(i in 1:length(args))
+    eval(parse(text=args[[i]]))
 
 # parameters for f1 and f2
 m1 <- runif(1, 0.5, 1.5)

### 2024-06-04 fitting/run_sim.R
bump replicates to 50

@@ -12,7 +12,7 @@ model <- 1 # set to 1 for "function 1", 2 for "function 2"
 setting <- 1 #set to 1 for "Setting A", 2 for "B", 3 for "C"
 if(!(model %in% 1:2)) stop("model should be 1 or 2")
 if(!(setting %in% 1:3)) stop("setting should be 1, 2, or 3")
-r <- 5 # number of replicates
+r <- 50 # number of replicates
 
 args <- commandArgs(TRUE)
 if(length(args) > 0)
@@ -41,8 +41,10 @@ if(model == 1){
     sdvec <- ifelse(x<=0.5, 5e-4, sdvec)
     sdvec <- ifelse(x>0.5 & x<2, 0.75*sdvec, sdvec)
     sdvec <- ifelse(x>=2, 0.5*sdvec, sdvec)
-  }}
+  }
+}
 
+# make sure the sd plot looks right (compared to appropriate values)
 plot(sdvec)
 
 # functions f1 and f2 to generate sims
@@ -114,7 +116,7 @@ for (j in 1:r) {
 mean(mses)
 
 # Boxplot
-boxplot(mses, ylim=c(0,.018), main="mses")
+boxplot(mses, ylim=c(0,ifelse(model==1,.018,.005)), main="mses")
 grid(nx = NULL, ny = NULL,
      col = "#ebebeb", lwd = 2, lty=1)
 boxplot(mses, add = TRUE)

### 2024-06-04 fitting/collect_sims.R
init commit sim results and collection script

@@ -1,0 +1,22 @@
+###############################################################################
+# This script fits the Bayesian hierarchical model for simulated data under
+# two different functions: f1 and f2
+###############################################################################
+
+library(dgp.hm)
+
+r <- 50
+j <- 1
+
+results <- list()
+
+for(model in 1:2){
+  for(setting in 1:3){
+    result <- read.csv(paste0("results/sims_",model,"_",setting,"_",r,".csv"))$x
+    results[[j]] <- result
+    j <- j + 1
+  }
+}
+
+boxplot(results, names = c("1,A","1,B","1,C","2,A","2,B","2,C"), 
+        ylab="Average MSE", main = "DGP-HM MSE across models/settings")

### 2024-06-04 fitting/run_sim.R
fix f2

@@ -54,7 +54,7 @@ f1 <- function(x, m1, u1, sd=0){
 }
 
 f2 <- function(x, m2, u2, sd=0){
-  return(exp(-m2*(x-3)^2) + exp(-u2*(x-1)^2) - 0.05*sin(x-1.9) + rnorm(length(x), sd=sd))
+  return(exp(-m2*(x-3)^2) + exp(-u2*(x-1)^2) - 0.05*sin(8*(x-1.9)) + rnorm(length(x), sd=sd))
 }
 
 # Get n function evaluations for f1 and f2 

### 2024-06-04 fitting/collect_sims.R
update results for fixed f2; make boxplots to match draft

@@ -20,3 +20,17 @@ for(model in 1:2){
 
 boxplot(results, names = c("1,A","1,B","1,C","2,A","2,B","2,C"), 
         ylab="Average MSE", main = "DGP-HM MSE across models/settings")
+
+# Boxplot for f1
+boxplot(results[1:3], ylim=c(0,.018), main="mses: function 1",
+        names = c("1,A","1,B","1,C"), ylab="MSE")
+grid(nx = NULL, ny = NULL,
+     col = "#ebebeb", lwd = 2, lty=1)
+boxplot(results[1:3], add = TRUE, names=F)
+
+# Boxplot for f2
+boxplot(results[4:6], ylim=c(0,.005), main="mses: function 2",
+        names = c("2,A","2,B","2,C"), ylab="MSE")
+grid(nx = NULL, ny = NULL,
+     col = "#ebebeb", lwd = 2, lty=1)
+boxplot(results[4:6], add = TRUE, names = F)

### 2024-06-04 fitting/collect_sims.R
allow n_obs and n_sims to varyy

@@ -5,14 +5,16 @@
 
 library(dgp.hm)
 
-r <- 50
+r <- 5
+n_sims <- 50
 j <- 1
 
 results <- list()
 
 for(model in 1:2){
   for(setting in 1:3){
-    result <- read.csv(paste0("results/sims_",model,"_",setting,"_",r,".csv"))$x
+    result <- read.csv(paste0("results/sims_",model,"_",setting,"_"
+                              ,r,"_",n_sims,".csv"))$x
     results[[j]] <- result
     j <- j + 1
   }

### 2024-06-04 fitting/run_sim.R

@@ -12,7 +12,8 @@ model <- 1 # set to 1 for "function 1", 2 for "function 2"
 setting <- 1 #set to 1 for "Setting A", 2 for "B", 3 for "C"
 if(!(model %in% 1:2)) stop("model should be 1 or 2")
 if(!(setting %in% 1:3)) stop("setting should be 1, 2, or 3")
-r <- 50 # number of replicates
+r <- 5       # number of function realizations
+n_sims <- 50 # number of sim studies to replicate
 
 args <- commandArgs(TRUE)
 if(length(args) > 0)
@@ -93,7 +94,7 @@ legend(x = "topright", legend = c("data","truth", "wt avg", "95% UQ"),
 
 # repeat for r replicates (don't plot these)
 mses <- c()
-for (j in 1:r) {
+for (j in 1:n_sims) {
   print(j)
   m1 <- runif(1, 0.5, 1.5)
   u1 <- runif(1, 1.5, 2.5)
@@ -121,4 +122,5 @@ grid(nx = NULL, ny = NULL,
      col = "#ebebeb", lwd = 2, lty=1)
 boxplot(mses, add = TRUE)
 
-write.csv(mses, file = paste0("results/sims_",model,"_",setting,"_",r,".csv"))
+write.csv(mses, file = paste0("results/sims_",model,"_",setting,"_"
+                              ,r,"_",n_sims,".csv"))

### 2024-06-04 fitting/collect_sims.R
rerun sims with 5 function realizations

@@ -5,9 +5,9 @@
 
 library(dgp.hm)
 
-r <- 5
-n_sims <- 50
-j <- 1
+r <- 5       # number of function realizations
+n_sims <- 50 # number of sim studies to replicate
+j <- 1       # index to keep track of function/setting pairs
 
 results <- list()
 

### 2024-06-05 paper/main.tex
add in toy sims & image

@@
 Here, we provide results from a simulation study comparing the GP and DGP models, in which case the true curve is known.
-\textcolor{blue}{I can include details for a sim study based on simulated data I construct, or I can use the dataset Dave mentioned getting to use here.}
+Following \cite{moran2024dpc}, we consider two functions (1 and 2) each with three variance specifications (settings ``A", ``B", ``C").
+Our mean squared error (MSE) results are seen in Figure \ref{fig:collect_sims}; our DGP-HM model outperforms these current benchmark methods.
+\textcolor{blue}{We can also include a setting ``D", where the underlying error covariance has spatial dependence (e.g., noise with a Mat'ern covariance).
+Additionally, we can look at linear theory simulations.}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=6in]{collect_sims.png}
+    \caption{Average MSE of our method across all $k$ values. \textcolor{blue}{Add competitors.}}
+    \label{fig:collect_sims}
+\end{figure}
 

### 2024-06-05 fitting/run_pert.R
init commit run_pert.R

@@ -1,0 +1,34 @@
+# Power Spectra Simulations
+
+# From the "linear" folder
+sim <- read.table("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN1/oneh_matterpower.dat")
+plot(log10(sim$V1), scrP(sim$V2,sim$V1), type="l",
+     xlab=expression(log[10](k)), ylab="script P(k)", ylim=c(-4.7,1.1),
+     main = "all 'linear' runs")
+for (run in 2:32) {
+  sim <- read.table(paste0("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN",
+                           run,"/oneh_matterpower.dat"))
+  lines(log10(sim$V1), scrP(sim$V2,sim$V1), col=run)
+}
+
+# From one of the "pow*" folders
+sim2 <- read.table("../CosmicEmu_etc/32 Models Mira Titan/pow.ic/M011/L1300/PM000/analysis/Pow/m011.pk.ini")
+head(sim2)
+pairs(sim2)
+plot(log10(sim2$V1), log10(sim2$V2), main = "orig, no cutoff")
+plot(log10(sim2$V1), scrP(sim2$V2, sim2$V1), main = "emu, no cutoff")
+plot(log10(sim2$V1)[sim2$V1<=1], scrP(sim2$V2, sim2$V1)[sim2$V1<=1], 
+     main = "emu, cutoff at k=1", type="l", xlim=c(-2.25,0),ylim=c(-4.3,-2.6))
+for (run in 12:42) {
+  if(run == 27 | run == 28) next
+  print(run)
+  if(run < 20)  folder <- ""
+  if(run >= 20) folder <- "_2"
+  if(run >= 30) folder <- "_3"
+  if(run >= 40) folder <- "_4"
+  sim2 <- read.table(paste0("../CosmicEmu_etc/32 Models Mira Titan/pow",folder,
+                            ".ic/M0", run, 
+                            "/L1300/PM000/analysis/Pow/m0", run, ".pk.ini"))
+  lines(log10(sim2$V1)[sim2$V1<=1], scrP(sim2$V2, sim2$V1)[sim2$V1<=1], 
+       main = "emu, cutoff at k=1", col=run)
+}

### 2024-06-06 fitting/run_pert.R
load library

@@ -1,5 +1,7 @@
 # Power Spectra Simulations
 
+library(dgp.hm)
+
 # From the "linear" folder
 sim <- read.table("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN1/oneh_matterpower.dat")
 plot(log10(sim$V1), scrP(sim$V2,sim$V1), type="l",
@@ -21,7 +23,6 @@ plot(log10(sim2$V1)[sim2$V1<=1], scrP(sim2$V2, sim2$V1)[sim2$V1<=1],
      main = "emu, cutoff at k=1", type="l", xlim=c(-2.25,0),ylim=c(-4.3,-2.6))
 for (run in 12:42) {
   if(run == 27 | run == 28) next
-  print(run)
   if(run < 20)  folder <- ""
   if(run >= 20) folder <- "_2"
   if(run >= 30) folder <- "_3"

### 2024-06-06 fitting/run_sim.R
add fourth toy example with correlated errors; generalize covariance for settings A,B,C accordingly

@@ -4,16 +4,21 @@
 ###############################################################################
 
 library(dgp.hm)
+library(mvtnorm)
 
 # Simulated toy example from DPC paper
 x <- seq(0, 4, by=0.1)
 n <- length(x)
+
 model <- 1 # set to 1 for "function 1", 2 for "function 2"
-setting <- 1 #set to 1 for "Setting A", 2 for "B", 3 for "C"
+setting <- 4 #set to 1 for "Setting A", 2 for "B", 3 for "C"; 4 for "D"
 if(!(model %in% 1:2)) stop("model should be 1 or 2")
-if(!(setting %in% 1:3)) stop("setting should be 1, 2, or 3")
+if(!(setting %in% 1:4)) stop("setting should be 1, 2, 3, or 4")
+
 r <- 5       # number of function realizations
 n_sims <- 50 # number of sim studies to replicate
+
+use_true_Sigma <- T # if T, use Sigma_true; if F, estimate Sigma_hat
 
 args <- commandArgs(TRUE)
 if(length(args) > 0)
@@ -25,60 +30,102 @@ m1 <- runif(1, 0.5, 1.5)
 m2 <- runif(1, 0.6, 1.4)
 u1 <- runif(1, 1.5, 2.5)
 u2 <- runif(1, 0.6, 1.4)
+
+# specify variance structure for the function
 if(model == 1){
-  if(setting == 1) sdvec <- rep(0.1, n)
-  if(setting == 2) sdvec <- 1/(10*(x+1))
+  if(setting == 1){
+    sdvec <- rep(0.1, n)
+    Sigma_true <- diag(sdvec^2)
+  } 
+  if(setting == 2){
+    sdvec <- 1/(10*(x+1))
+    Sigma_true <- diag(sdvec^2)
+  } 
   if(setting == 3){
     sdvec <- 1/(10*(x+1))
     sdvec <- ifelse(x<=0.4, 5e-4, sdvec)
     sdvec <- ifelse(x>0.4 & x<3, 0.75*sdvec, sdvec)
     sdvec <- ifelse(x>=3, 0.5*sdvec, sdvec)
+    Sigma_true <- diag(sdvec^2)
   }
-} else {
-  if(setting == 1) sdvec <- rep(0.15, n)
-  if(setting == 2) sdvec <- sqrt(abs(x-2))/10
[…]

### 2024-06-06 fitting/collect_sims.R
collect new sim results

@@ -12,7 +12,7 @@ j <- 1       # index to keep track of function/setting pairs
 results <- list()
 
 for(model in 1:2){
-  for(setting in 1:3){
+  for(setting in 1:4){
     result <- read.csv(paste0("results/sims_",model,"_",setting,"_"
                               ,r,"_",n_sims,".csv"))$x
     results[[j]] <- result
@@ -20,19 +20,19 @@ for(model in 1:2){
   }
 }
 
-boxplot(results, names = c("1,A","1,B","1,C","2,A","2,B","2,C"), 
+boxplot(results, names = c("1,A","1,B","1,C","1,D","2,A","2,B","2,C","2,D"), 
         ylab="Average MSE", main = "DGP-HM MSE across models/settings")
 
 # Boxplot for f1
-boxplot(results[1:3], ylim=c(0,.018), main="mses: function 1",
-        names = c("1,A","1,B","1,C"), ylab="MSE")
+boxplot(results[1:4], ylim=c(0,.018), main="mses: function 1",
+        names = c("1,A","1,B","1,C","1,D"), ylab="MSE")
 grid(nx = NULL, ny = NULL,
      col = "#ebebeb", lwd = 2, lty=1)
-boxplot(results[1:3], add = TRUE, names=F)
+boxplot(results[1:4], add = TRUE, names=F)
 
 # Boxplot for f2
-boxplot(results[4:6], ylim=c(0,.005), main="mses: function 2",
-        names = c("2,A","2,B","2,C"), ylab="MSE")
+boxplot(results[5:8], ylim=c(0,.005), main="mses: function 2",
+        names = c("2,A","2,B","2,C","2,D"), ylab="MSE")
 grid(nx = NULL, ny = NULL,
      col = "#ebebeb", lwd = 2, lty=1)
-boxplot(results[4:6], add = TRUE, names = F)
+boxplot(results[5:8], add = TRUE, names = F)

### 2024-06-06 fitting/run_pert.R
add D+ values to match LT and sims

@@ -2,26 +2,27 @@
 
 library(dgp.hm)
 
+# D+ values to scale sims in order to match the linear theory
+Dp <- c(0.007985, 0.006853, 0.007923, 0.007024, 0.007214, 
+        0.008146, 0.006776, 0.007998, 0.007517, 0.006175) 
+        # 0.010293, 0.008178, 0.011243, 0.007489, 0.009507, 
+        # 0.013198, 0.007817, 0.008683, 0.006918, 0.008124, 
+        # 0.006601, 0.006948, 0.009904, 0.006510, 0.007727, 
+        # 0.010022, 0.007219, 0.007437, 0.010621)
+
 # From the "linear" folder
 sim <- read.table("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN1/oneh_matterpower.dat")
 plot(log10(sim$V1), scrP(sim$V2,sim$V1), type="l",
-     xlab=expression(log[10](k)), ylab="script P(k)", ylim=c(-4.7,1.1),
-     main = "all 'linear' runs")
-for (run in 2:32) {
+     xlab=expression(log[10](k)), ylab="script P(k)", ylim=c(-4.7,1.3),
+     main = "LT and simulated spectra")
+for (run in 2:10) {
   sim <- read.table(paste0("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN",
                            run,"/oneh_matterpower.dat"))
   lines(log10(sim$V1), scrP(sim$V2,sim$V1), col=run)
 }
 
 # From one of the "pow*" folders
-sim2 <- read.table("../CosmicEmu_etc/32 Models Mira Titan/pow.ic/M011/L1300/PM000/analysis/Pow/m011.pk.ini")
-head(sim2)
-pairs(sim2)
-plot(log10(sim2$V1), log10(sim2$V2), main = "orig, no cutoff")
-plot(log10(sim2$V1), scrP(sim2$V2, sim2$V1), main = "emu, no cutoff")
-plot(log10(sim2$V1)[sim2$V1<=1], scrP(sim2$V2, sim2$V1)[sim2$V1<=1], 
-     main = "emu, cutoff at k=1", type="l", xlim=c(-2.25,0),ylim=c(-4.3,-2.6))
-for (run in 12:42) {
+for (run in 11:20) {
   if(run == 27 | run == 28) next
   if(run < 20)  folder <- ""
[…]

### 2024-06-06 fitting/run_LT.R
rename run_pert.R to run_LT.R

@@ -1,4 +1,4 @@
-# Power Spectra Simulations
+# Power Spectra Simulations and linear theory (LT)
 
 library(dgp.hm)
 

### 2024-06-06 fitting/run_LT.R
align D+ values

@@ -13,7 +13,7 @@ Dp <- c(0.007985, 0.006853, 0.007923, 0.007024, 0.007214,
 # From the "linear" folder
 sim <- read.table("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN1/oneh_matterpower.dat")
 plot(log10(sim$V1), scrP(sim$V2,sim$V1), type="l",
-     xlab=expression(log[10](k)), ylab="script P(k)", ylim=c(-4.7,1.3),
+     xlab=expression(log[10](k)), ylab="script P(k)", ylim=c(0,1.3), xlim = c(-2.1,0),
      main = "LT and simulated spectra")
 for (run in 2:10) {
   sim <- read.table(paste0("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN",
@@ -32,5 +32,8 @@ for (run in 11:20) {
                             ".ic/M0", run, 
                             "/L1300/PM000/analysis/Pow/m0", run, ".pk.ini"))
   lines(log10(sim2$V1)[sim2$V1<=1], 
-        scrP(sim2$V2/(Dp[run-11])^2, sim2$V1)[sim2$V1<=1], col="red", lwd=2)
+        scrP(sim2$V2/(Dp[run-10])^2, sim2$V1)[sim2$V1<=1], col=run-10, lwd=3, lty=2)
 }
+
+legend(x = "bottom", legend = c("linear theory", "sims"),
+      lty = c(1,2), lwd=c(1,3))

### 2024-08-26 plotting/plot_data_fit.R
update color palette and data png

@@ -59,52 +59,72 @@ fit <- read.csv(paste0("../fitting/results/dgp_", model_name, ".csv"))
 # Plot data types #
 ###################
 
-matplot(log10(k), y_lo, col = "gray", 
+# color-blind friendly color palette
+
+cbcols <- palette.colors(palette = "Okabe-Ito")
+#1 black, 2 orange, 3 skyblue , 4 bluishgreen, 5 yellow 
+#6 blue, 7 vermillion,  8 reddishpurple, 9 gray 
+
+par(mfrow=c(1,2))
+matplot(log10(k), y_lo, col = cbcols[9], 
         type = "l", lty = 1, ylim = range(y_avg),
         xlab=expression(log[10](k)), 
         ylab=expression(paste("\U1D4AB","(k)")))
-lines(log10(k), y_pt, col = "green")
-lines(log10(k), y_hi, col = "red")
+lines(log10(k), y_pt, col = cbcols[4])
+lines(log10(k), y_hi, col = cbcols[8])
 lines(log10(k)[index_list$pert.ix], rep(-1, length(index_list$pert.ix)), 
-      col = "green", lty = 3, lwd=3)
+      col = cbcols[4], lty = 3, lwd=3)
 lines(log10(k)[index_list$lowres.ix], rep(-1.1, length(index_list$lowres.ix)), 
-      col = "gray", lty = 3, lwd=3)
+      col = cbcols[9], lty = 3, lwd=3)
 lines(log10(k)[index_list$highres.ix], rep(-1.2, length(index_list$highres.ix)), 
-      col = "red", lty = 3, lwd=3)
-lines(log10(k), y_avg, col = "orange", lty = 2, lwd = 4)
+      col = cbcols[8], lty = 3, lwd=3)
+lines(log10(k), y_avg, col = cbcols[2], lty = 2, lwd = 4)
 legend(x = "right", legend = c("pert","low res","hi res", "wt avg","ind"),
-       col = c("green","gray","red","orange","black"), lty = c(1,1,1,2,3), 
+       col = cbcols[c(4,9,8,2,1)], lty = c(1,1,1,2,3), 
        lwd = c(1,1,1,4,3))
+
+# Same plot, but only for the low-resolution indices
+matplot(log10(k)[index_list$lowres.ix], y_lo[index_list$lowres.ix,], col = cbcols[9], 
[…]

### 2024-08-26 paper/main.tex
update color palette and data png
use color-blind palette; update images

@@
         David Higdon\thanks{Department of Statistics, Virginia Tech} \and
-        Marco A.R. Ferreira\footnotemark[3]}
+        Marco A.R. Ferreira\footnotemark[3] \and
+        Kelly Moran\thanks{Los Alamos National Laboratory} \and
+        Katrin Heitmann\thanks{Argonne National Laboratory}}
 \date{\today}
@@
     \includegraphics[width=6in]{plot_data.png}
-    \caption{The perturbation theory, low resolution runs, and high resolution run for the first training cosmology.
+    \caption{Left: The perturbation theory, low resolution runs, and high resolution run for the first training cosmology.
 The weighted average is also shown as a dashed line.
-Dotted lines at the bottom indicate where each data type is deemed approximately unbiased.}
+Dotted lines at the bottom indicate where each data type is deemed approximately unbiased.
+Right: The same image, but only where the low resolution runs are approximately unbiased.}
     \label{fig:plot_data}
@@
 
-priors here
+\textcolor{blue}{Annie, please put the newest prior specs here.}
 
@@
 
+\subsection{CAMB Data}
+\label{subsec:camb}
+
+Here, we look over the CAMB data to see how our method performs.
+
 \subsection{Estimating Spectra with Mira-Titan Data}
@@
     \includegraphics[width=6in]{rmse_by_k.png}
-    \caption{Average RMSE of each method across all $k$ values.}
+    \caption{Left: Average RMSE of each prediction method across all $k$ values.
+Right: Average RMSE for each method when using the held-out computer model runs to train the model.
+In each case, the truth is assumed to be the weighted average of the computer model runs.}
     \label{fig:plot_rmse_k}
@@
 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=6in]{pred_1to6.png}
+    \caption{Results of all six predictions for both methods on the held-out cosmologies (weighted average subtracted).
+CosmicEMU is dashed, DGP.HM is dotted, and solid gray lines show results of held-out computer model runs.}
+    \label{fig:plot_pred_1to6}
+\end{figure}
+
 \bibliographystyle{jasa}

### 2024-08-26 plotting/plot_data_fit.R

@@ -116,15 +116,15 @@ legend(x = "bottomright", legend = c("pert","low res","hi res", "wt avg", "UQ"),
        lwd = c(2,1,2,2,2))
 
 # Plot model fit alongside data (posterior mean removed)
-matplot(log10(k), y_lo - fit$m, type="l", lty=1,
+matplot(log10(k), y_lo - fit$m, type="l", lty=3,
         col=cbcols[9], ylim = c(-.02,.02), xlab=expression(log[10](k)), 
         ylab=expression(paste("\U1D4AB","(k), mean removed")))
 abline(h=0, col=cbcols[1],lty=1,lwd=1)
-lines(log10(k), y_pt - fit$m, col=cbcols[4], lwd=3.5, lty=1)
-lines(log10(k), y_hi - fit$m, col=cbcols[8], lwd=3.5, lty=1)
+lines(log10(k), y_pt - fit$m, col=cbcols[4], lwd=3, lty=4)
+lines(log10(k), y_hi - fit$m, col=cbcols[8], lwd=3, lty=2)
 lines(log10(k), y_avg - fit$m , col=cbcols[2], lwd=2, lty=1)
 lines(log10(k), fit$ub - fit$m, col=cbcols[1],lty=2,lwd=2)
 lines(log10(k), fit$lb - fit$m, col=cbcols[1],lty=2,lwd=2)
 legend(x = "bottomright", legend = c("pert","low res","hi res", "wt avg", "UQ"),
-       col = cbcols[c(4,9,8,2,1)], lty = c(3,1,3,2,2,2), 
+       col = cbcols[c(4,9,8,2,1)], lty = c(4,3,2,1,2), 
        lwd = c(2,1,2,2,2))

### 2024-08-26 plotting/plot_pca.R

@@ -21,6 +21,11 @@ if(length(args) > 0)
   for(i in 1:length(args))
     eval(parse(text=args[[i]]))
 
+# color-blind friendly color palette
+cbcols <- palette.colors(palette = "Okabe-Ito")
+#1 black, 2 orange, 3 skyblue , 4 bluishgreen, 5 yellow 
+#6 blue, 7 vermillion,  8 reddishpurple, 9 gray 
+
 if(PDF) pdf(paste0("compare_pred_",n_pc,"_",pwdExp,".pdf"))
 
 # read in predicted power matter spectra for 6 test models
@@ -31,9 +36,14 @@ colnames(etaEmu) <- c("E001", "E002", "E003", "E009", "E010", "M000")
 # this is not emulating, this is estimating (so it should do better)
 convln_pred <- read.csv("../CosmicEmu_etc/z_zero_test.txt", sep = "", header=F)
 
-# record MSEs for each of three approaches
+# record MSEs for each of four approaches
 mses <- mses_emu <- mses_conv_t <- mses_dgp_t <- matrix(NA, 6, 351)
 
+#####################################
+# Plot predictions for both methods #
+#####################################
+
+par(mfrow=c(2,3),oma=c(4,4,1.5,1),mar=c(0,0,0,0))
 # For each of the six test models...
 for(i in 1:6){
   
@@ -68,66 +78,78 @@ for(i in 1:6){
   
   # Make a plot showing each computer model, along with predictions
   # Average is removed to better visualize the variability and bias
-  par(mar=c(4,4.5,1,1), mfrow=c(1,1))
+  # par(mar=c(4,4.5,1,1), mfrow=c(1,1))
   matplot(log10(k), scrP(pk2[,3],k) - meanterm, type="n",
-       ylim = c(-.12,.12),
+       ylim = c(-.07,.07),
        xlab=expression(log[10](k)),
-       ylab='script P')
+       ylab='script P', yaxt=ifelse(i %in% c(2,3,5,6),"n","s"),
+       xaxt=ifelse(i %in% c(1,2,3),"n","s"))
   # plot the low-res runs
   for (j in 3:18) lines(log10(k)[index_list$lowres.ix], 
[…]

### 2024-08-26 paper/main.tex
update sims plots' palette and layout

@@
 Following \cite{moran2024dpc}, we consider two functions (1 and 2) each with three variance specifications (settings ``A", ``B", ``C").
-Our mean squared error (MSE) results are seen in Figure \ref{fig:collect_sims}; our DGP-HM model outperforms these current benchmark methods.
+Our mean squared error (MSE) results are seen in Figure \ref{fig:sims_MSE}; our DGP-HM model outperforms these current benchmark methods.
 \textcolor{blue}{We can also include a setting ``D", where the underlying error covariance has spatial dependence (e.g., noise with a Mat'ern covariance).
@@
     \centering
-    \includegraphics[width=6in]{collect_sims.png}
-    \caption{Average MSE of our method across all $k$ values. \textcolor{blue}{Add competitors.}}
-    \label{fig:collect_sims}
+    \includegraphics[width=6in]{sims_logS.png}
+    \caption{Boxplot of log scores across two different functions and covariance settings (lower is better). For each case, twenty random batches were simulated.}
+    \label{fig:sims_logS}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=6in]{sims_MSE.png}
+    \caption{Boxplot of MSEs across two different functions and covariance settings (lower is better). For each case, twenty random batches were simulated.}    \label{fig:sims_MSE}
 \end{figure}
@@
 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=6in]{rmse_box.png}
-    \caption{Boxplot of all RMSEs of each method across all six test models.
+% \begin{figure}[ht]
+%     \centering
+%     \includegraphics[width=6in]{rmse_box.png}
+%     \caption{Boxplot of all RMSEs of each method across all six test models.
 From top to bottom: box plots for our training method, process convolution training method, Cosmic EMU predictions, and our DGP-PC predictions.}
-    \label{fig:plot_rmse_box}
-\end{figure}
+%     \label{fig:plot_rmse_box}
+% \end{figure}
 

### 2024-08-26 plotting/plot_data_fit.R

@@ -60,7 +60,6 @@ fit <- read.csv(paste0("../fitting/results/dgp_", model_name, ".csv"))
 ###################
 
 # color-blind friendly color palette
-
 cbcols <- palette.colors(palette = "Okabe-Ito")
 #1 black, 2 orange, 3 skyblue , 4 bluishgreen, 5 yellow 
 #6 blue, 7 vermillion,  8 reddishpurple, 9 gray 

### 2024-08-26 sims/collect_sims.R

@@ -7,26 +7,84 @@
 PDF <- F
 if(PDF) pdf("collect_sims.pdf")
 
-par(mfrow = c(1,2))
+# color-blind friendly color palette
+cbcols <- palette.colors(palette = "Okabe-Ito")
+#1 black, 2 orange, 3 skyblue , 4 bluishgreen, 5 yellow 
+#6 blue, 7 vermillion,  8 reddishpurple, 9 gray 
+
+###################################################
+# Get ylim's for MSE/logScore for both r=5 and 20 #
+###################################################
+
+MSE5_range <- MSE20_range <- logS5_range <- logS20_range <- c()
+
+r=5
+for (func in 1:2) {
+  for (setting in 4:5) {
+    filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
+    # Compare MSEs
+    results <- read.csv(filename)[, c(3,5,4)]
+    # NOTE: removed one outlier to not skew the ylim
+    if(func==1&setting==4) results <- results[-which.max(results[,3]),]
+    MSE5_range <- range(MSE5_range, results)
+    results <- read.csv(filename)[, c(7,9,8)]
+    logS5_range <- range(logS5_range, results)
+  }
+}
+
+r=20
+for (func in 1:2) {
+  for (setting in 4:5) {
+    filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
+    # Compare MSEs
+    results <- read.csv(filename)[, c(3,5,4)]
+    MSE20_range <- range(MSE20_range, results)
+    results <- read.csv(filename)[, c(7,9,8)]
+    logS20_range <- range(logS20_range, results)
+  }
+}
+
+####################
+# Boxplots of MSEs #
+####################
+
+par(mfrow=c(2,4),oma=c(4,4,1.5,1),mar=c(0,0,0,0))
 for (r in c(5,20)) {
-  for (func in 1:2) {
-    for (setting in 1:5) {
+  for (setting in 4:5) {
+    for (func in 1:2) {
       filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
       # Compare MSEs
       results <- read.csv(filename)[, c(3,5,4)]
-      boxplot(results, ylab = "MSE",
-              main = paste0("MSEs: Fn ", func, " Stg ", setting, " r=", r),
-              names = c("dgp.hm","hetGP","deepGP"), col = c("green4","gold","gray"))
[…]

### 2024-08-27 paper/main.tex
init commit pred_1to6.png; update paper

@@
 
-We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.
+\textcolor{blue}{Annie: want to add something here? We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.}
+
+Through the synthesis of the DGP model alongside the additional level of hierarchical modeling (HM) to incorporate multiple functional model runs, we call this method DGP.HM.
 
@@
 
-Here, we provide results from a simulation study comparing the GP and DGP models, in which case the true curve is known.
-Following \cite{moran2024dpc}, we consider two functions (1 and 2) each with three variance specifications (settings ``A", ``B", ``C").
-Our mean squared error (MSE) results are seen in Figure \ref{fig:sims_MSE}; our DGP-HM model outperforms these current benchmark methods.
-\textcolor{blue}{We can also include a setting ``D", where the underlying error covariance has spatial dependence (e.g., noise with a Mat'ern covariance).
-Additionally, we can look at linear theory simulations.}
+We compare our DGP.HM model to competing models with a simulation study, in which case the true curve is known.
+Following \cite{moran2024dpc}, we consider two functions (1 and 2) each with two variance specifications (settings ``A", ``B", \textcolor{blue}{, we can consider additional homoskedastic cases in the appendix if necessary}).
+\textcolor{blue}{Put in the details of the two functions, and the two variance specs.} The competing methods we consider are an out-of-the-box DGP fit using the \texttt{deepgp} package \citep{sauer2023active}, a heteroskedastic GP fit using the \textt{hetGP} package \citep{binois2018practical, binois2021hetgp}, as well as a deep process convolution (DPC) approach, which is utilized within Cosmic Emu \citep{moran2023mira}.
+
+For each simulated function and variance specification, we generate either 5 or 20 functional observations and estimate the underlying true function based on the observations.
+The metrics we use to evaluate performance are mean squared error (MSE) and the log score, a proper scoring rule which takes both prediction and intervals into consideration \citep{gneiting2007strictly}.
+
+MSE and log score results are shown in Figures \ref{fig:sims_MSE} and \ref{fig:sims_logS} respectively; our DGP.HM model performs favorably against these current benchmark methods.
 
@@
     \label{fig:plot_fit}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=6in]{plot_warp_M001.png}
+    \caption{Left: MCMC samples of $W$ for model M001 after burnin and thinning. Right: Corresponding 95\% credible intervals for $W$.}
+    \label{fig:plot_warp}
 \end{figure}

### 2024-08-28 paper/main.tex
one typo, comment out missing figure

@@
 Following \cite{moran2024dpc}, we consider two functions (1 and 2) each with two variance specifications (settings ``A", ``B", \textcolor{blue}{, we can consider additional homoskedastic cases in the appendix if necessary}).
-\textcolor{blue}{Put in the details of the two functions, and the two variance specs.} The competing methods we consider are an out-of-the-box DGP fit using the \texttt{deepgp} package \citep{sauer2023active}, a heteroskedastic GP fit using the \textt{hetGP} package \citep{binois2018practical, binois2021hetgp}, as well as a deep process convolution (DPC) approach, which is utilized within Cosmic Emu \citep{moran2023mira}.
+\textcolor{blue}{Put in the details of the two functions, and the two variance specs.} The competing methods we consider are an out-of-the-box DGP fit using the \texttt{deepgp} package \citep{sauer2023active}, a heteroskedastic GP fit using the \texttt{hetGP} package \citep{binois2018practical, binois2021hetgp}, as well as a deep process convolution (DPC) approach, which is utilized within Cosmic Emu \citep{moran2023mira}.
 
@@
 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=6in]{plot_warp_M001.png}
-    \caption{Left: MCMC samples of $W$ for model M001 after burnin and thinning. Right: Corresponding 95\% credible intervals for $W$.}
-    \label{fig:plot_warp}
-\end{figure}
+%\begin{figure}[ht]
+%    \centering
+%    \includegraphics[width=6in]{plot_warp_M001.png}
+%    \caption{Left: MCMC samples of $W$ for model M001 after burnin and thinning. Right: Corresponding 95\% credible intervals for $W$.}
+%    \label{fig:plot_warp}
+%\end{figure}
 

### 2024-08-28 paper/main.tex
update initial details for CAMB/LT runs

@@
 The code to reproduce these results are available at \url{https://github.com/stevewalsh124/dgp.hm}.
-\textcolor{blue}{Please note: this repository is currently private until we hear back regarding permissions for sharing the Mira-Titan data alongside our code.}
+\textcolor{red}{Please note: this repository is currently private until we hear back regarding permissions for sharing the Mira-Titan data alongside our code.}
 
@@
 
-\textcolor{blue}{Annie, please put the newest prior specs here.}
+\textcolor{cyan}{Annie, please put the newest prior specs here.}
 
@@
 
-$$
-\begin{aligned}
+\begin{align}
 \bar Y|S &\sim \mathcal{N}_n(S,\Sigma_\varepsilon) \\
 S &\sim GP\left(\mu, \Sigma_S(x,x')\right)
-\end{aligned}
-$$
+\end{align}
 
@@
 
-$$
-\begin{aligned}
+\begin{align}
 \bar Y|S,W &\sim \mathcal{N}_n(S,\Sigma_\varepsilon) \\
@@
 W &\sim GP\left(0, \Sigma_W(x,x')\right)
-\end{aligned}
-$$
+\end{align}
 
@@
 
-\textcolor{blue}{Annie: want to add something here? We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.}
+\textcolor{cyan}{Annie: want to add something here? We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.}
 
@@
 We compare our DGP.HM model to competing models with a simulation study, in which case the true curve is known.
-Following \cite{moran2024dpc}, we consider two functions (1 and 2) each with two variance specifications (settings ``A", ``B", \textcolor{blue}{, we can consider additional homoskedastic cases in the appendix if necessary}).
-\textcolor{blue}{Put in the details of the two functions, and the two variance specs.} The competing methods we consider are an out-of-the-box DGP fit using the \texttt{deepgp} package \citep{sauer2023active}, a heteroskedastic GP fit using the \texttt{hetGP} package \citep{binois2018practical, binois2021hetgp}, as well as a deep process convolution (DPC) approach, which is utilized within Cosmic Emu \citep{moran2023mira}.
+Following \cite{moran2024dpc}, we consider two functions ($f_1$ and $f_2$) each with two variance specifications (settings ``4" and ``5", \textcolor{magenta}{we can consider additional homoskedastic cases in the appendix if necessary}).
 
-For each simulated function and variance specification, we generate either 5 or 20 functional observations and estimate the underlying true function based on the observations.
+Here, $f_1$ and $f_2$ are defined as follows:
+
+\begin{align}
+  f_1(x) &= m_1 \exp(-u_1x/2) * \cos(wx) - m_1x/5, \quad w=\sqrt{25-[u_1/2]^2} \\
+  f_2(x) &= \exp(-m_2[x-3]^2)+\exp(-u_2[x-1]^2)-0.05\sin(8[x-1.9]), \\ 
+  \quad x &\in [0,4].
+\end{align}
+
+For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are random variables where $m_1 \sim Uniform(0.5,1.5), u_1 \sim Uniform(1.5,2.5),$ and $m_2,u_2 \sim Uniform(0.6,1.4)$.
+For each simulated function and variance specification, we generate 20 values of $m_1, m_2, f_1, f_2$, and conditional on these values, we sample either 5 or 20 functional observations and estimate the underlying true function based on the observations.
+
+The competing methods we consider are an out-of-the-box DGP fit using the \texttt{deepgp} package \citep{sauer2023active}, a heteroskedastic GP fit using the \texttt{hetGP} package \citep{binois2018practical, binois2021hetgp}, as well as a deep process convolution (DPC) approach, which is utilized within Cosmic Emu \citep{moran2023mira}.
+
 The metrics we use to evaluate performance are mean squared error (MSE) and the log score, a proper scoring rule which takes both prediction and intervals into consideration \citep{gneiting2007strictly}.
-
+For both metrics, a lower value indicates better performance.
 MSE and log score results are shown in Figures \ref{fig:sims_MSE} and \ref{fig:sims_logS} respectively; our DGP.HM model performs favorably against these current benchmark methods.
 
-\begin{figure}[ht]
+\begin{figure}[t]
     \centering
@@
 
-Here, we look over the CAMB data to see how our method performs.
+In addition to simulated data, we investigate model fits the for different cosmological datasets (based on linear theory and/or runs from the Code for Anisotropies in the Microwave Background, CAMB).
+\textcolor{magenta}{We have looked at two sets of model runs, "LT" runs where there is an underlying ``true" value alongside 15 computer model runs, or the "CAMB" runs, where we only have a single output.
+The performance from the CAMB runs was better (with a 6-dimensional design space), and also did not require obtaining a posterior mean from multiple model runs.
+When looking at the CAMB runs on the $\Delta$ space, we see it generally is withing $\pm 5\%$ error.}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=6in]{plot_pca_CAMB.png}
+    \caption{Proportion of $\Delta_{emu}$ vs. $\Delta_{holdout}$ for model runs 11-42 (excluding 27-28).}   
+    \label{fig:pca_camb}
+\end{figure}
 

### 2024-08-28 fitting/run_LT.R

@@ -2,7 +2,7 @@
 
 library(dgp.hm)
 
-# pdf("LT_sims_scrP_all.pdf")
+pdf("LT_sims_scrP.pdf")
 
 # All the Runs and Models are in units Mpc/h and h/Mpc.
 # To be able to plot everything in one plot, we have to convert to Mpc or 1/Mpc.
@@ -15,35 +15,38 @@ Dp <- data$D.
 h <- data$Hubble
 
 # Working with runs 1-32 (aka models 11-42), but skip 17-18 (27-28)
-run <- 1
-model <- run + 10
-
-# read in the linear output
-sim <- read.table(paste0("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN",
-                         run,"/oneh_matterpower.dat"))
-k_s <- sim$V1*h[run]
-pk_s <- sim$V2/ (h[run])^3
-plot(log10(k_s), scrP(pk_s, k_s), type="l", col="blue", lwd=4,
-     xlab=expression(log[10](k)), ylab="script P(k)", ylim=c(-0.2,1.4), xlim = c(-2.3,0),
-     main = paste0("LT and simulated spectra, model ", model))
-
-# read in first batch of model runs (model #11)
-pk_ms <- matrix(NA, 886, 15)
-for (i in 0:15) {
-  mod <- read.table(paste0("../CosmicEmu_etc/32 Models Mira Titan/pow.ic/M0", 
-                           model, "/L1300/PM0",if(i < 10){"0"}, i,
-                           "/analysis/Pow/m0", model, ".pk.ini"))
-  k_m <- mod$V1 * h[model-10]
-  pk_ms[,i] <- pk_m <- mod$V2 / (h[model-10])^3 / (Dp[model-10])^2
-  lines(log10(k_m)[k_m <= 1], scrP(pk_m, k_m)[k_m <= 1], lty=3, col="gray")
-}
-
-lines(log10(k_m)[k_m <= 1], scrP(rowMeans(pk_ms), k_m)[k_m <= 1], 
-      col="orange", lwd=2, lty=1)
-
-legend(x = "bottom", legend = c("linear theory", "sims", "sim avg"),
-       lty = c(1, 3, 1), lwd=c(4, 1, 2), col=c("blue","gray","orange"))
-
+for(run in 1:32){
+  model <- run + 10
+  if(model %in% 27:28) next
+  # read in the linear output
[…]

### 2024-08-28 paper/main.tex
remove a few outliers from plot to make the plots fit better

@@
 
-For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are random variables where $m_1 \sim Uniform(0.5,1.5), u_1 \sim Uniform(1.5,2.5),$ and $m_2,u_2 \sim Uniform(0.6,1.4)$.
+For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are random variables where $m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 \sim \text{Uniform}(0.6,1.4)$.
 For each simulated function and variance specification, we generate 20 values of $m_1, m_2, f_1, f_2$, and conditional on these values, we sample either 5 or 20 functional observations and estimate the underlying true function based on the observations.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=6in]{sims_logS.png}
+    \caption{Boxplot of log scores across two different functions and covariance settings (lower is better).
+For each case, the boxplot is constructed from 20 random batches were simulated.
+Each column represents a function/variance specification pair, and each row shows results for batch sizes of either 5 (top) or 20 (bottom).}
+    \label{fig:sims_logS}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=6in]{sims_logS.png}
-    \caption{Boxplot of log scores across two different functions and covariance settings (lower is better). For each case, twenty random batches were simulated.}
-    \label{fig:sims_logS}
-\end{figure}
-
-\begin{figure}[ht]
+\begin{figure}
     \centering
     \includegraphics[width=6in]{sims_MSE.png}
-    \caption{Boxplot of MSEs across two different functions and covariance settings (lower is better). For each case, twenty random batches were simulated.}    \label{fig:sims_MSE}
+    \caption{Boxplot of MSEs across two different functions and covariance settings (lower is better).
+For each case, the boxplot is constructed from 20 random batches were simulated.
+Each column represents a function/variance specification pair, and each row shows results for batch sizes of either 5 (top) or 20 (bottom).}
+    \label{fig:sims_MSE}
 \end{figure}
@@
 The performance from the CAMB runs was better (with a 6-dimensional design space), and also did not require obtaining a posterior mean from multiple model runs.
-When looking at the CAMB runs on the $\Delta$ space, we see it generally is withing $\pm 5\%$ error.}
+When looking at the CAMB runs on the $\Delta$ space, we see it generally is withing $\pm 5\%$ error.
+Regarding these results, compared to the results for the 8-dimensional LT runs which had less impressive predictive performance, I'm not sure if we will include details on one or the other.}
 
-\begin{figure}[ht]
+\begin{figure}
     \centering

### 2024-08-28 sims/collect_sims.R

@@ -28,7 +28,8 @@ for (func in 1:2) {
     if(func==1&setting==4) results <- results[-which.max(results[,3]),]
     MSE5_range <- range(MSE5_range, results)
     results <- read.csv(filename)[, c(7,9,8)]
-    logS5_range <- range(logS5_range, results)
+    # Note: removes three outliers from setting==5, r==5
+    logS5_range <- range(logS5_range, results[results <= 150])
   }
 }
 

### 2024-08-29 paper/main.tex
update paper, include W plot

@@
 
-\textcolor{cyan}{Annie, please put the newest prior specs here.}
+\textcolor{cyan}{Annie, please include the most up-to-date prior specs here.}
 
@@
 Additionally, this model yields the posterior distribution $S|\bar Y \sim N(m, C)$, where $C=\left(\Sigma_S^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1}$ and $m=C\left(\Sigma_\varepsilon^{-1}\bar Y+\Sigma_S^{-1}\mu\right)$.
-
 This distribution provides the posterior mean and corresponding uncertainty for $S$ within a given cosmology.
@@
 Here, $f_1$ and $f_2$ are defined as follows:
-
 \begin{align}
@@
 
-For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are random variables where $m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 \sim \text{Uniform}(0.6,1.4)$.
+For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are random where $m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 \sim \text{Uniform}(0.6,1.4)$.
+
+For the two variance settings, we consider matrices $\Sigma_4$ and $\Sigma_5$ to have a Mat\'ern covariance function with a nugget of $10^{-8}$ and smoothness of 2.5.
+For setting ``4", we have $\tau^2=0.0225$ and $\theta=0.01$.
+For setting ``5", we we have $\tau^2=0.01$ and $\theta=0.05$, and additionally scale this covariance matrix by a vector of values $s$ so that $\Sigma_5 = s M s^T$, where $M$ is a Mat\'ern covariance matrix.
+
 For each simulated function and variance specification, we generate 20 values of $m_1, m_2, f_1, f_2$, and conditional on these values, we sample either 5 or 20 functional observations and estimate the underlying true function based on the observations.
 
-\begin{figure}[h]
+\begin{figure}[t]
     \centering
@@
 
-Estimate one cosmology with 50,000 runs, and use this as a starting point for each of the others
+To being, we initialize our hyperparameter estimates across all cosmologies by modeling the first (M001) with 50,000 MCMC runs, and use these as initial values for all cosmologies.
+Specifically, we save the last MCMC iteration of $W$, $\theta_W$ and $\theta_S$ as $W_0$, $\theta_W0$ and $\theta_S0$ respectively and use each as starting values for all 111 training cosmology fits, where these are now fit with 10,000 MCMC runs (5,000 removed as burn-in and every 5th sample is kept, resulting in 1,000 posterior samples).
 
@@
 
-%\begin{figure}[ht]
-%    \centering
-%    \includegraphics[width=6in]{plot_warp_M001.png}
-%    \caption{Left: MCMC samples of $W$ for model M001 after burnin and thinning. Right: Corresponding 95\% credible intervals for $W$.}
-%    \label{fig:plot_warp}
-%\end{figure}
-
-Images:
-\begin{itemize}
-    \item Illustration of DGP fit (post mean and UQ, could show cosmicEMU as well?)
-    \item estimated warping (or, how all 111 warpings compare)
-\end{itemize}
+\begin{figure}[ht]
+   \centering
+   \includegraphics[width=6in]{plot_warp_M001.png}
+   \caption{Left: MCMC samples of $W$ for model M001 after burnin and thinning.
+Right: Corresponding 95\% credible intervals for $W$, where departure from the identity line represents nonstationary behavior.}
+   \label{fig:plot_warp}
+\end{figure}
 
@@
 
-The different appendices from dissertation as necessary. Can also put details of dataset in supplementary material.
+The different appendices from dissertation as necessary. Can also put details of dataset in supplementary material. \textcolor{magenta}{Fix references as necessary, e.g. Moran et al. 2024).}
 

### 2024-09-20 paper/main.tex
update manuscript

@@
 Estimating and understanding the influence of these parameters on cosmic expansion are areas of active research.
-In addition to presenting the Mira-Titan simulation suite, \cite{moran2023mira} builds off of previous spectrum emulation \citep{lawrence2017mira} to provide the final emulator (Cosmic Emu) based on the full suite of simulations.
+In addition to presenting the Mira-Titan simulation suite, \cite{moran2023mira} builds off of previous spectrum emulation \citep{lawrence2017mira} to provide the final emulator (Cosmic Emu) based on the full suite of 117 batches of simulations.
 
-In this work, we expand on the previous emulation methods and propose a novel synthesis of methods to quantify uncertainty and predict matter power spectra for different cosmologies using deep Gaussian processes (DGPs).
-In order to handle multiple realizations of functional output (from the Mira-Titan simulation suite), we use hierarchical and basis functions modeling within our DGP framework to accurately estimate and predict power spectra for different cosmologies.
+In this work, we expand on the previous emulation approaches and propose a novel synthesis of methods to quantify uncertainty and predict matter power spectra for different cosmologies using deep Gaussian processes (DGPs).
+In order to handle multiple realizations of functional output (from the Mira-Titan simulation suite), we use hierarchical modeling and basis functions within our DGP framework to accurately estimate and predict power spectra for different cosmologies.
 When comparing to Cosmic Emu, we see favorable results for our method.
@@
 Different simulation output exists for different values of a ninth value, the redshift; in this work, we fix this value at 0.
-For each cosmology, we have a set of 18 simulated spectra: an inexpensive spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from lower-resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from a high resolution simulation ($y_h$).
+For each cosmology, we have a batch of 18 simulated spectra: an inexpensive spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from lower-resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from a high resolution simulation ($y_h$).
 
@@
 
+From the multiple low- and high-resolution spectra across all cosmologies, \cite{moran2023mira} obtained precision estimates of these spectra across the $n$ wavenumber values ($p_1,\dots,p_n$) using a log-log regression model, and a multiplier $c$ for the increase in precision from the low- to high- resolution output.
+Given that the perturbation theory spectra is non-stochastic, we treat this as having a precision of $10^5$.
+We leverage these precision estimates within our model at each data type's unbiased wavenumber values to define $\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, $\Lambda_{\ell} = \text{diag}\left(16(p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, $\Lambda_h = \text{diag}\left(c(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$.
+Here, each precision matrix is diagonal and the diagonal elements consist of the precision values or zero, depending on whether the $k$ value is approximately unbiased for that data type.
+With this, we calculate a weighted average for each cosmology: $\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for the weighted average for a particular cosmology.
+
 \begin{figure}[ht]
@@
     \caption{Left: The perturbation theory, low resolution runs, and high resolution run for the first training cosmology.
-The weighted average is also shown as a dashed line.
+The weighted average of these is shown as a dashed line.
 Dotted lines at the bottom indicate where each data type is deemed approximately unbiased.
-Right: The same image, but only where the low resolution runs are approximately unbiased.}
+Right: The same image, but restricted to $k$ values where the low resolution runs are approximately unbiased.}
     \label{fig:plot_data}
@@
 
-From the multiple low- and high-resolution spectra across all cosmologies, \cite{moran2023mira} obtained estimates for the precision of these spectra across the $n$ wavenumber values ($p_1,\dots,p_n$) using a log-log regression model, and a multiplier $c$ for the increase in precision from the low- to high- resolution output.
-Given that the perturbation theory spectra is non-stochastic, we treat this as having a precision of $10^5$.
-We leverage these precision estimates within our model at each data type's unbiased wavenumber values to define $\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, $\Lambda_{\ell} = \text{diag}\left(16(p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, $\Lambda_h = \text{diag}\left(c(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$.
-Here, each precision matrix is diagonal and the diagonal elements consists of the precision values or zero, depending on whether the $k$ value is approximately unbiased for that data type.
-With this, we calculate a weighted average for each cosmology: $\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for the weighted average for a particular cosmology.
-
 We will handle the data in two stages.
-First, we will focus on 111 training cosmologies and obtain the best esimated spectrum for each.
+First, we will focus on 111 training cosmologies and obtain the best estimated spectrum for each.
 Second, we will use these estimates to predict the spectra for six held-out cosmologies.
@@
 
-From plots of the multiple low-resolution spectra, we find they vary smoothly about some mean $S$, suggesting that a dense covariance matrix $\Sigma_\varepsilon$ may outperform a diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$.
+From plots of the multiple low-resolution spectra (e.g., Figure \ref{fig:plot_data}), we find they vary smoothly about some mean $S$, suggesting that a dense covariance matrix $\Sigma_\varepsilon$ (accounting for spatial dependence) may outperform a diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$.
 In previous work, we have modeled multiple candidates for $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and found that a Mat\'ern covariance $\Sigma_\ell$ trained on the low-resolution runs (pre-scaled by the precision values and subtracting a LOESS-smoothed average) performs the best.
@@
 We transform $X$ values to fall within $[0,1]$ and $Y$ so that the mean and variance are 0 and 1, respectively.
-This allows for a strightforward specification of priors, where the hyperparameters for each Mat\'ern covariance $\Sigma_S$ and $\Sigma_\ell$ each take a vague inverse-Gamma prior.
+This allows for a straightforward specification of priors, where the hyperparameters for the Mat\'ern covariance $\Sigma_S$ each take a vague prior.
 Specifically:
@@
 where $d=||x_i-x_j||^2$.
-From here, we estimate $\tau_\ell, \theta_\ell$ with maximum likelihood, and conditional on $\Sigma_\varepsilon$, we specify priors for $\tau_S, \theta_S$ following \cite{sauer2023active}:
-
-\textcolor{cyan}{Annie, please include the most up-to-date prior specs here.}
+From here, we estimate $\tau_\ell, \theta_\ell$ with maximum likelihood, and conditional on $\Sigma_\varepsilon$, we specify priors for $\tau_S, \theta_S$ following \cite{sauer2023active}: \textcolor{cyan}{Annie, please ensure the most up-to-date prior specs are shown above and here.
+Here is what I found from \texttt{fit\$settings} when using \texttt{deepgp v1.1.3}: $\theta_S \sim G(\alpha=1.2, \beta=1)$ and $\tau_S \propto 1/\tau_S^2$.}
 
@@
 
-Upon integrating out $S$, we obtain $\bar Y \sim GP(\mu, \Sigma_S+\Sigma_\varepsilon)$.
-Additionally, this model yields the posterior distribution $S|\bar Y \sim N(m, C)$, where $C=\left(\Sigma_S^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1}$ and $m=C\left(\Sigma_\varepsilon^{-1}\bar Y+\Sigma_S^{-1}\mu\right)$.
+Upon integrating out $S$, we obtain $\bar Y \sim \mathcal{N}_n(\mu, \Sigma_S+\Sigma_\varepsilon)$.
+Additionally, this model yields the posterior distribution $S|\bar Y \sim \mathcal{N}_n(m, C)$, where $C=\left(\Sigma_S^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1}$ and $m=C\left(\Sigma_\varepsilon^{-1}\bar Y+\Sigma_S^{-1}\mu\right)$.
 This distribution provides the posterior mean and corresponding uncertainty for $S$ within a given cosmology. 
-Without this hierarchy, uncertainty would not be properly addressed; the weighted average would simply be nearly interpolated across the relatively dense set of $k$ values.
+% With this hierarchical model, we improve the efficiency of estimating $S|\bar{Y}$.
 
@@
 Considering the Mira-Titan data, some nonstationarity is expected (e.g., due to baryonic acoustic oscillations when $-2 \leq \log_{10}(k) \leq -1$).
-To account for this, we incorporate a latent layer warping $W$ within the hierarchical model to incorporate nonstationary dynamics of the matter power spectra.
+To account for this, we incorporate a latent layer $W$ within the hierarchical model to warp locations of $X$ and incorporate nonstationary dynamics of the matter power spectra.
 When $W$ is a GP, this creates a deep Gaussian process (DGP) model \citep{damianou2013deep}.
@@
 
-We have found that whether the prior mean of $W$ is set to 0 or $X$ \citep[which would indicate stationarity apriori, ][]{schmidt2003bayesian} does not have a substantial impact on the model fit.
-We construct this model in the compositional form, and also use MCMC to obtain full uncertainty for each of the model parameters \citep{sauer2023active}.
+We have found that whether the prior mean of $W$ is set to 0 or $X$ \citep[which would indicate stationarity apriori,][]{schmidt2003bayesian} does not have a substantial impact on the model fit.
+We construct this model in the compositional form, and also use Markov Chain Monte Carlo (MCMC) to estimate and obtain full uncertainty for each of the model parameters \citep{sauer2023active}.
 
-\textcolor{cyan}{Annie: want to add something here? We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.}
+\textcolor{cyan}{Annie: want to add something here?
+For $\theta_w$, we also specify a Gamma prior of the form $\theta_w \sim G(\alpha=1.2, \beta=2)$.
+We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.}
 
@@
 
-We compare our DGP.HM model to competing models with a simulation study, in which case the true curve is known.
+We compare DGP.HM to competing models with a simulation study where the underlying true curve is known.
 Following \cite{moran2024dpc}, we consider two functions ($f_1$ and $f_2$) each with two variance specifications (settings ``4" and ``5", \textcolor{magenta}{we can consider additional homoskedastic cases in the appendix if necessary}).
@@
 For setting ``4", we have $\tau^2=0.0225$ and $\theta=0.01$.
-For setting ``5", we we have $\tau^2=0.01$ and $\theta=0.05$, and additionally scale this covariance matrix by a vector of values $s$ so that $\Sigma_5 = s M s^T$, where $M$ is a Mat\'ern covariance matrix.
+For setting ``5", we we have $\tau^2=0.01$ and $\theta=0.05$, and additionally scale this covariance matrix by a vector of values $s$ so that $\Sigma_5 = s M s^T$ is nonstationary; here $M$ is a Mat\'ern covariance matrix with parameters $\tau^2$ and $\theta$.
 
@@
 For both metrics, a lower value indicates better performance.
-MSE and log score results are shown in Figures \ref{fig:sims_MSE} and \ref{fig:sims_logS} respectively; our DGP.HM model performs favorably against these current benchmark methods.
+Log score and MSE results are shown in Figures \ref{fig:sims_logS} and \ref{fig:sims_MSE} respectively; our DGP.HM model performs favorably against these current benchmark methods.
 
@@
 In addition to simulated data, we investigate model fits the for different cosmological datasets (based on linear theory and/or runs from the Code for Anisotropies in the Microwave Background, CAMB).
-\textcolor{magenta}{We have looked at two sets of model runs, "LT" runs where there is an underlying ``true" value alongside 15 computer model runs, or the "CAMB" runs, where we only have a single output.
+\textcolor{magenta}{We have looked at two sets of model runs, ``LT" runs where there is an underlying ``true" value alongside 15 computer model runs, or the ``CAMB" runs, where we only have a single output.
 The performance from the CAMB runs was better (with a 6-dimensional design space), and also did not require obtaining a posterior mean from multiple model runs.
 When looking at the CAMB runs on the $\Delta$ space, we see it generally is withing $\pm 5\%$ error.
-Regarding these results, compared to the results for the 8-dimensional LT runs which had less impressive predictive performance, I'm not sure if we will include details on one or the other.}
+Regarding these results, compared to the results for the 8-dimensional LT runs which had less impressive predictive performance, I'm open to suggestions on whether to include details on one or the other.}
 
@@
 
-To being, we initialize our hyperparameter estimates across all cosmologies by modeling the first (M001) with 50,000 MCMC runs, and use these as initial values for all cosmologies.
-Specifically, we save the last MCMC iteration of $W$, $\theta_W$ and $\theta_S$ as $W_0$, $\theta_W0$ and $\theta_S0$ respectively and use each as starting values for all 111 training cosmology fits, where these are now fit with 10,000 MCMC runs (5,000 removed as burn-in and every 5th sample is kept, resulting in 1,000 posterior samples).
+To train our hierarchical model, we initialize our hyperparameter estimates across all cosmologies by modeling the first (M001) with 50,000 MCMC runs, and use these as initial values for all cosmologies.
+Specifically, we save the last MCMC iteration of $W$, $\theta_w$ and $\theta_S$ as $W_0$, $\theta_{w0}$ and $\theta_{S0}$ respectively and use each as starting values for all 111 training cosmology fits, where each cosmology is fit with 10,000 MCMC runs (5,000 removed as burn-in and every 5th sample is kept, resulting in 1,000 posterior samples).
 
-Here we can provide figures of the predicted mean and variances for a subset of the actual cosmologies.
+From these samples, we can estimate and quantify the uncertainty within each cosmology's average and latent layer $W$.
 Figure \ref{fig:plot_fit} shows an example of the uncertainty for our posterior mean for the first (of 111) training cosmologies; the posterior mean is subtracted in order to visualize this uncertainty.
+Figure \ref{fig:plot_warp} provides elliptical slice samples of $W$ along with the average warping.
+Some departure from stationarity is estimated in the higher $X$ (i.e., scaled $k$) values.
 
@@
 
-This section details our second contribution - the use of the Bayesian DGP model of the previous section to predict the curve for held-out cosmologies.
+Upon obtaining the posterior means from each of the 111 training cosmologies, this section details our second contribution - the use of the Bayesian DGP.HM model to predict the curve for held-out cosmologies.
 
@@
 
-The different appendices from dissertation as necessary. Can also put details of dataset in supplementary material. \textcolor{magenta}{Fix references as necessary, e.g. Moran et al. 2024).}
+The different appendices from dissertation as necessary. Can also put details of dataset in supplementary material. \textcolor{magenta}{Fix references as necessary, e.g. Moran et al. 2024.}
 

### 2024-09-20 paper/main.tex
update manuscript again

@@
 Upon obtaining the posterior means from each of the 111 training cosmologies, this section details our second contribution - the use of the Bayesian DGP.HM model to predict the curve for held-out cosmologies.
+This is achieved with modeling the weights of different basis functions as GPs.
 
-\subsection{PCA-GP Model}
+\subsection{GP-PC Model}
 \label{subsec:pca}
 
-Here we describe the PCA-GP model. We discuss how we use the predicted curves that we got previously as training data. \cite{higdon2008computer, higdon2010estcosmo}
+Here, we begin with $\boldsymbol\eta$, a $n \times m$ matrix composed of $m=111$ posterior means, each of length $n$.
+To facilitate more efficient estimation, we use singular value decomposition \citep[SVD; e.g.,][]{banerjee2014linear}, where $\boldsymbol\eta = UDV^T$; $U$ and $V$ are orthogonal matrices of size $n \times n$ and  $m\times m$, respectively.
+$D$ is a diagonal matrix containing the singular values.
+For this application, $\boldsymbol\eta$ has rank $r=\text{min}\{m,n\}=111$, which determines the number of non-zero elements in $D$.
+So, we can find an alternative decomposition where $\Sigma$ is a diagonal matrix with the non-zero singular values of $\boldsymbol\eta$, and $U_1$ and $V_1$ are matrices with orthonormal columns of size $n_\eta \times r$ and $m \times r$, respectively.
+
+From here, following modeling strategies from \cite{higdon2008computer, higdon2010estcosmo}, we equivalently decompose $\boldsymbol\eta$ into a principal component (PC) basis matrix $B^* = \frac{1}{\sqrt{m}}U_1\Sigma$ along with its corresponding weights $W^* = \sqrt{m}V_1$.
+Without loss of generality, we can perform this decomposition after an overall mean trend is removed.
+If we use $p_\eta < r$ PCs, this will reduce $B$ and $W$ to be the first $p_\eta$ columns of $B^*$ and $W^*$ respectively and result in an approximation where
+
+\begin{equation}
+    \boldsymbol\eta= UDV^T = U_1\Sigma V_1^T \approx BW^T.
+\end{equation}
+
+Figure \ref{fig:mean_PCs_oneW}, illustrates this decomposition with the estimated mean trend, the principal components contained  in $B$, and one example of how the weight for the first PC will vary dependent on the first parameter $\kappa$.
+
+In the following subsections, we consider different modeling strategies for $W$, which will influence predictions and sensitivity analyses based on $\boldsymbol\eta$.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{mean_PCs_oneW.png}
+    \caption{\textcolor{red}{UPDATE this from GSMF to spectra!!
+The estimated mean trend of the GSMF (left), The principal components contained in $B$ (middle), and an illustration of how the weights for the first PC will change as $\kappa$ varies (right)}.}
+    \label{fig:mean_PCs_oneW}
+\end{figure}
+
+Following \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use GPs to model the principal components' weights; we shall denote this method GP-PC.
+We assume a powered exponential kernel for the Gaussian process and employ \texttt{GPfit} to perform a multi-start gradient based search for GP's hyperparameters \citep{macdonald2015gpfit}.
+That is, the $i$th vector of PC weights $w_i(\theta)$, $i \in \{1,\ldots,p_\eta\}$, will be modeled as a zero mean GP: $w_i(\theta) \sim GP(0, \sigma^2R)$, with
+
+\begin{equation*}
+    R_{ij} = \prod_{k=1}^{p_\theta}\exp\left\{-10^{\beta_k}|x_{ik}-x_{jk}|^\alpha\right\}
+\end{equation*}
+
+We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\theta\}$. \textcolor{red}{Replace all $p_\theta, p_\eta, \kappa$; reread and update this section.}
+
+Here we describe the GP-PC model. We discuss how we use the predicted curves that we got previously as training data. \cite{higdon2008computer, higdon2010estcosmo}
 

### 2024-09-21 paper/main.tex
add mean_PCs_oneW plot; update manuscript

@@
 Log score and MSE results are shown in Figures \ref{fig:sims_logS} and \ref{fig:sims_MSE} respectively; our DGP.HM model performs favorably against these current benchmark methods.
+\textcolor{magenta}{If we mention that DGP.HM timings speed up compared to \texttt{deepgp}, we would also need to mention that \texttt{hetGP} is much faster than both of these.}
 
@@
 
-Here, we begin with $\boldsymbol\eta$, a $n \times m$ matrix composed of $m=111$ posterior means, each of length $n$.
+Here, we begin with a $n \times m$ matrix composed of $m=111$ posterior means, each of length $n$, denoted by $\boldsymbol\eta$.
 To facilitate more efficient estimation, we use singular value decomposition \citep[SVD; e.g.,][]{banerjee2014linear}, where $\boldsymbol\eta = UDV^T$; $U$ and $V$ are orthogonal matrices of size $n \times n$ and  $m\times m$, respectively.
@@
 For this application, $\boldsymbol\eta$ has rank $r=\text{min}\{m,n\}=111$, which determines the number of non-zero elements in $D$.
-So, we can find an alternative decomposition where $\Sigma$ is a diagonal matrix with the non-zero singular values of $\boldsymbol\eta$, and $U_1$ and $V_1$ are matrices with orthonormal columns of size $n_\eta \times r$ and $m \times r$, respectively.
+We find an alternative decomposition where $\Sigma$ is a diagonal matrix with the non-zero singular values of $\boldsymbol\eta$, and $U_1$ and $V_1$ are matrices with orthonormal columns of size $n_\eta \times r$ and $m \times r$, respectively.
 
-From here, following modeling strategies from \cite{higdon2008computer, higdon2010estcosmo}, we equivalently decompose $\boldsymbol\eta$ into a principal component (PC) basis matrix $B^* = \frac{1}{\sqrt{m}}U_1\Sigma$ along with its corresponding weights $W^* = \sqrt{m}V_1$.
+Following \cite{higdon2008computer, higdon2010estcosmo}, we equivalently decompose $\boldsymbol\eta$ into a principal component (PC) basis matrix $B^* = \frac{1}{\sqrt{m}}U_1\Sigma$ along with its corresponding weights $W^* = \sqrt{m}V_1$.
 Without loss of generality, we can perform this decomposition after an overall mean trend is removed.
@@
 
-Figure \ref{fig:mean_PCs_oneW}, illustrates this decomposition with the estimated mean trend, the principal components contained  in $B$, and one example of how the weight for the first PC will vary dependent on the first parameter $\kappa$.
+Figure \ref{fig:mean_PCs_oneW} illustrates this decomposition with the estimated mean trend, the principal components contained in $B$, and one example of how the weight for the first PC will vary dependent on the first of $p_\theta=8$ cosmological parameters contained in $\theta$.
 
@@
     \includegraphics[width=\textwidth]{mean_PCs_oneW.png}
-    \caption{\textcolor{red}{UPDATE this from GSMF to spectra!!
-The estimated mean trend of the GSMF (left), The principal components contained in $B$ (middle), and an illustration of how the weights for the first PC will change as $\kappa$ varies (right)}.}
+    \caption{The estimated mean trend of the 111 posterior means (left), the principal components contained in $B$ (middle), and an illustration of how the weights for the first PC will change as the first cosmological parameter varies (right) \textcolor{red}{This last plot is wonky because it was made with only the 6 test cosmologies, not a large uniform or FF design, Steve can update this}.}
     \label{fig:mean_PCs_oneW}
@@
 
-Following \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use GPs to model the principal components' weights; we shall denote this method GP-PC.
+In alignment with \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use GPs to model the principal components' weights; we shall denote this method GP-PC.
 We assume a powered exponential kernel for the Gaussian process and employ \texttt{GPfit} to perform a multi-start gradient based search for GP's hyperparameters \citep{macdonald2015gpfit}.
@@
 
-We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\theta\}$. \textcolor{red}{Replace all $p_\theta, p_\eta, \kappa$; reread and update this section.}
-
-Here we describe the GP-PC model. We discuss how we use the predicted curves that we got previously as training data. \cite{higdon2008computer, higdon2010estcosmo}
-
-\begin{itemize}
-    \item show breakdown of estimated mean trend, basis functions/PCs, and weights of a particular PC/BF.
-\end{itemize}
+We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\theta\}$. Using these trained GPs from the PCs, we predict the spectra for unobserved cosmologies.
 
@@
 
-Here we show results on held-out cosmologies.  The Cosmic Emu emulator is constructed by (process convolutions, ...). Segue to our method after and compare/contrast with Cosmic Emu.
+Here we show prediction results for 6 held-out cosmologies.
+We compare our method with Cosmic Emu \citep{moran2023mira}, the emulator constructed using the full suite of 111 Mira-Titan model runs.
+Cosmic Emu uses a Bayesian approach where the smooth spectrum for each cosmology is modeled with a process convolution on Brownian motion.
+Nonstationarity is permitted through modeling the bandwidth parameter with a process convolution as well; this composition is referred to as a deep process convolution \citep{moran2024dpc}.
 
-In addition to predicting for held-out cosmologies, we can also predict over a large uniform or full factorial design to obtain estimated main effects for each parameter in $\theta$, interactions amongst the 8 parameters, and variance decomposition.
-Should this be included?
+Both Cosmic Emu and DGP.HM are capable of modeling a true underlying functions based upon multiple functional realizations, but our approach utilizes DGPs within a hierarchical model in lieu of deep process convolutions.
+After obtaining the estimated smooth spectra from the training cosmologies, each method utilizes GPs in order to estimate the weights of the PCs from the estimated spectra.
+Below, we compare prediction performance between Cosmic Emu and DGP.HM for the 6 held-out cosmologies.
+\textcolor{magenta}{In addition, we can also predict over a large uniform or full factorial design to obtain estimated main effects for each cosmological parameter in $\theta$, interactions amongst the $p_\theta=8$ parameters, and decompose the overall variation by main effects, two-way interactions, and so on.
+Should this be included?}
+
+Plots for the average root mean squared error (RMSE) for each prediction method over the 6 cosmologies and each $k$ value are shown in Figure \ref{fig:plot_rmse_k}.
+Comparisons of the spectrum estimation for training cosmologies are shown as well (using RMSE against the weighted average of observations).
+For predictions, DGP.HM achieves the greatest reduction in RMSE in the region where perturbation theory and the low-resolution runs are deemed unbiased.
+For the region where only the high-resolution run is unbiased, Cosmic Emu has a lower average RMSE.
+Over all $k$ values and predictions, DGP.HM has an average RMSE of 0.0089 and Cosmic Emu has an average RMSE of 0.0106.
+Our RMSE is lower at 56\% of $k$ values considered.
+The individual predictions for each method and cosmology are shown in Figure \ref{fig:plot_pred_1to6}.
 
@@
 
-Brief recap, directions for future research: BSS-ANOVA in lieu of DGPs, hydrodynamical sims or TCs as other applications of this methodology.
-A more detailed sensitivity analysis for the parameters, modeling $W$ across different cosmologies, consider modeling for different redshift values other than $z=0$.
+In this work, we propose a novel Bayesian hierarchical model which utilizes DGPs in order to estimate a smooth underlying function from correlated functional observations (as cosmological computer model runs).
+In addition to the compositional form of the DGP, we include a third component modeling the correlated errors for each spectrum.
+With all smooth spectra estimated from the training set of 111 batches of model runs, we can predict spectra for unobserved cosmologies.
+
+Multiple avenues for future research emerge: in addition to estimating dark matter power spectra, batches of hydrodynamical simulations exist, and this method could easily be adapted to handle such datasets.
+Other applications outside of cosmology can be explored, wherever multiple functional outputs exist, perhaps with spatial correlation (for example, observing multiple tropical cyclone forecasts).
+More detailed sensitivity analyses for the cosmological parameters can be explored to gain insights into how each of these parameters interact, and which are most significant.
+From a modeling perspective, changes to the model could include modeling the latent layer $W$ jointly across different cosmologies, as well as considering different redshift values other than $z=0$.
+Leveraging a hybrid model between DGP.HM and Cosmic Emu may also help to obtain accurate estimates on the lower and higher ends of $k$ values, respectively.
+Additionally, Bayesian smoothing spline ANOVA (BSS-ANOVA) or other methods could be compared in addition to DGPs.
 
@@
 
-The different appendices from dissertation as necessary. Can also put details of dataset in supplementary material. \textcolor{magenta}{Fix references as necessary, e.g. Moran et al. 2024.}
+\textcolor{magenta}{The different appendices from dissertation as necessary.
+Can also put details of dataset in supplementary material.} \textcolor{red}{Fix references as necessary, e.g.
+Moran et al.
+2024.}
 

### 2024-09-21 plotting/plot_pca.R

@@ -158,4 +158,15 @@ mean(mses)
 # find proportion of k values where our method has lower MSE
 mean(colMeans(mses) < colMeans(mses_emu))
 
+# Make plot of basis function decomposition
+# run predict.R first to load des_test and aps[[1]]
+# mean_PCs_oneW; hydro_plotting.R has original with GSMF
+par(mfrow=c(1,3), mar=c(4,4,1.8,1))
+plot(apply(eta,1,mean),type="l", xlab=expression(log[10](k)), ylab='script P')
+matplot(bases, type="l", xlab=expression(log[10](k)), ylab = 'deviation from average')
+des_test <- read.csv("../Mira-Titan-IV-Data/design_test.txt",sep="",header = F)
+x1_pred <- des_test[,1]
+plot(x1_pred[order(x1_pred)], aps[[1]]$Y_hat[order(x1_pred)], type="l", 
+     xlab = expression(theta[1]), ylab = expression(W[1]))
+
 if(PDF) dev.off()

### 2024-10-12 paper/main.tex
abstract edits

@@
 
-\title{Deep Gaussian Processes and Hierarchical Modeling with Functional Data: A Case Study in Cosmological Power Spectra}
+\title{Bayesian Deep Gaussian Processes for Correlated Functional Data: 
+        A Case Study in Cosmological Power Spectra}
 \author{Stephen A. Walsh\thanks{Corresponding author: Division of Natural Sciences, 
@@
 \begin{abstract} 
-Understanding the structure of our universe and how matter is spread out and expanding is an area of active research.
-As cosmological surveys grow in complexity, developing surrogate models to efficiently predict matter power spectra is an important research area in cosmology.
-In this work, we synthesize methods from deep Gaussian processes, Bayesian hierarchical modeling, and basis functions in a novel way to estimate and predict matter power spectra (which are functional in nature) for different cosmologies.
-Our method has favorable results compared to the benchmark cosmological emulator (Cosmic Emu), and our code is available in a repository.
+Understanding the structure of our universe and the distribution of matter is an 
+area of active research.  As cosmological surveys grow in complexity, developing 
+surrogate models or emulators to efficiently predict matter power spectra is 
+important.  We are particularly motivated by the Mira Titan Universe simulation
+suite which, for a specified cosmology parameterization, provides multiple response
+curves of various fidelities, including correlated functional realizations.
+In this work, we propose a Bayesian deep Gaussian process (DGP) hierarchical 
+model to synthesize all the information to estimate the underlying matter power
+spectra, while providing effective uncertainty quantification.  This model
+extends previous works on Bayesian DGPs from scaler responses to functional ones.
+We deploy basis function representations of the estimated spectra to train another
+surrogate for the prediction of unobserved cosmologies.  Our method performs well
+compared to the benchmark cosmological emulator (Cosmic Emu).  We provide
+reproducible code in a public git repository.
 \end{abstract}
@@
     \centering
-    \includegraphics[width=\textwidth]{mean_PCs_oneW.png}
+    %\includegraphics[width=\textwidth]{mean_PCs_oneW.png}
     \caption{The estimated mean trend of the 111 posterior means (left), the principal components contained in $B$ (middle), and an illustration of how the weights for the first PC will change as the first cosmological parameter varies (right) \textcolor{red}{This last plot is wonky because it was made with only the 6 test cosmologies, not a large uniform or FF design, Steve can update this}.}

### 2024-10-16 paper/main.tex
revise abstract

@@
 Understanding the structure of our universe and the distribution of matter is an 
-area of active research.  As cosmological surveys grow in complexity, developing 
-surrogate models or emulators to efficiently predict matter power spectra is 
-important.  We are particularly motivated by the Mira Titan Universe simulation
-suite which, for a specified cosmology parameterization, provides multiple response
-curves of various fidelities, including correlated functional realizations.
-In this work, we propose a Bayesian deep Gaussian process (DGP) hierarchical 
-model to synthesize all the information to estimate the underlying matter power
-spectra, while providing effective uncertainty quantification.  This model
-extends previous works on Bayesian DGPs from scaler responses to functional ones.
-We deploy basis function representations of the estimated spectra to train another
-surrogate for the prediction of unobserved cosmologies.  Our method performs well
-compared to the benchmark cosmological emulator (Cosmic Emu).  We provide
-reproducible code in a public git repository.
+area of active research.  As cosmological surveys grow in complexity, the development 
+of emulators to efficiently and effectively predict matter power spectra is essential.  
+We are particularly motivated by the Mira Titan Universe simulation
+suite which, for a specified cosmological parameterization (termed a ``cosmology''), 
+provides multiple response curves of various fidelities, including correlated 
+functional realizations.  Our objective is two-fold.  First, we seek to 
+estimate the underlying true matter power spectra, with appropriate uncertainty 
+quantification (UQ), from all of the provided curves.  To this end, we propose a 
+novel Bayesian deep Gaussian process (DGP) hierarchical model which synthesize 
+all the information and estimates the underlying matter power spectra, while providing 
+effective UQ.  Our model extends previous work on Bayesian DGPs from scaler responses 
+to functional ones.  Second, we aim to leverage simulation results of various 
+cosmologies in order to accurately predict the entire matter power spectra for an 
+unobserved cosmology.  For this we leverage basis function representations 
+of the estimated spectra to train a separate Gaussian process emulator.  
+Our method performs well in synthetic exercises and against the benchmark cosmological 
+emulator (Cosmic Emu).  We provide reproducible code in a public git repository.
 \end{abstract}
+
+\noindent \textbf{Keywords:} surrogate, computer experiment, uncertainty quantification,
+principal components analysis, ???
 

### 2024-10-16 paper/main.tex
some edits to intro

@@
 
-\title{Bayesian Deep Gaussian Processes for Correlated Functional Data: 
+\title{Bayesian Deep Gaussian Processes for Correlated Functional Data: \\
         A Case Study in Cosmological Power Spectra}
@@
 
-To further our understanding of the structure and movement of the universe, researchers utilize numerous cosmological surveys such as the Sloan Digital Sky Survey \citep{york2000sloan} and the upcoming Nancy Grace Roman Space Telescope \citep{Dore2019WFIRST} which are continuously growing in complexity.
-With these tools, cosmologists aim to continue learning more about cosmic acceleration \citep{caldwell2009physics}.
-To this end, one important ingredient to aid in this understanding of our universe is the matter power spectrum.
+%To further our understanding of the structure and movement of the universe, 
+%researchers utilize numerous cosmological surveys such as the Sloan Digital Sky 
+%Survey \citep{york2000sloan} and the upcoming Nancy Grace Roman Space Telescope 
+%\citep{Dore2019WFIRST} which are continuously growing in complexity. With these 
+%tools, cosmologists aim to continue learning more about cosmic acceleration 
+%\citep{caldwell2009physics}. To this end, one important ingredient to aid in this 
+%understanding of our universe is the matter power spectrum. 
+% @STEVE - I don't think these "surveys" are particularly relevant
 
-The matter power spectrum is a fundamental concept in cosmology and describes the distribution of matter as a function of spatial scale.
-It is often represented as a function of wavenumber $k$ (units Mpc$^{-1}$), which is inversely related to spatial scale.
-On large scales (i.e., small $k$ values), cosmic expansion dynamics and the matter power spectrum behave according to linear perturbation theory \citep{pietroni2008flowing, lesgourgues2009non}.
-On smaller scales, non-linear dynamics require the use of computationally intensive simulations; the Coyote Universe \citep{lawrence2010coyote} and more recently the Mira-Titan Universe \citep{moran2023mira} are two simulation suites dedicated to this effort.
+Computer simulation experiments are invaluable tools in the study of cosmology.
+Experiments which simulate the growth and expansion of the universe are growing
+in prevalance and complexity \citep[e.g.,][]{lawrence2010coyote,moran2023mira}
+\textcolor{red}{(more cites here?)}.  
+We are particularly motivated by a simulation
+of the matter power spectrum, which describes the distribution of matter as a 
+function of spatial scale. 
+%It is often represented as a function of wavenumber $k$ (units Mpc$^{-1}$), which is 
+%inversely related to spatial scale. 
+On large scales, power spectrum behave according to linear perturbation 
+theory \citep{pietroni2008flowing, lesgourgues2009non}.  On smaller scales, nonlinear 
+dynamics require the use of computationally intensive simulations, yielding
+correlated functional data.
+%the Coyote Universe \citep{lawrence2010coyote} and the Mira-Titan Universe \citep{moran2023mira} 
+%are two simulation suites dedicated to this effort.  
+From the highest level, our aim is to leverage such simulation data to 
+first estimate and then predict underlying power matter spectra for various 
+cosmological parameterizations.
+
+%The expense of these simulations necessitates statistical ``emulators'' or
+%``surrogate models'' \citep{santner2003design,gramacy2020surrogates}.
 

### 2024-10-19 paper/main.tex
work on introduction

@@
 \footskip=0.5in
-\doublespacing
+%\doublespacing
 
@@
 of emulators to efficiently and effectively predict matter power spectra is essential.  
-We are particularly motivated by the Mira Titan Universe simulation
+We are particularly motivated by the Mira-Titan Universe simulation
 suite which, for a specified cosmological parameterization (termed a ``cosmology''), 
 provides multiple response curves of various fidelities, including correlated 
-functional realizations.  Our objective is two-fold.  First, we seek to 
-estimate the underlying true matter power spectra, with appropriate uncertainty 
+functional realizations.  Our objective is two-fold.  First, we estimate 
+the underlying true matter power spectra, with appropriate uncertainty 
 quantification (UQ), from all of the provided curves.  To this end, we propose a 
@@
 all the information and estimates the underlying matter power spectra, while providing 
-effective UQ.  Our model extends previous work on Bayesian DGPs from scaler responses 
-to functional ones.  Second, we aim to leverage simulation results of various 
+effective UQ.  Our model extends previous work on Bayesian DGPs from scalar responses 
+to functional ones.  Second, we leverage our predicted power spectra from various 
 cosmologies in order to accurately predict the entire matter power spectra for an 
-unobserved cosmology.  For this we leverage basis function representations 
-of the estimated spectra to train a separate Gaussian process emulator.  
+unobserved cosmology.  For this task, we leverage basis function representations 
+of the functional spectra to train a separate Gaussian process emulator.  
 Our method performs well in synthetic exercises and against the benchmark cosmological 
@@
 \noindent \textbf{Keywords:} surrogate, computer experiment, uncertainty quantification,
-principal components analysis, ???
+principal components analysis, Mira-Titan, Cosmic Emu (???)
 
@@
 %understanding of our universe is the matter power spectrum.
-% @STEVE - I don't think these "surveys" are particularly relevant
 
 Computer simulation experiments are invaluable tools in the study of cosmology.
-Experiments which simulate the growth and expansion of the universe are growing
+Experiments which simulate the expansion of the universe are growing
 in prevalance and complexity \citep[e.g.,][]{lawrence2010coyote,moran2023mira}
@@
 
-%The expense of these simulations necessitates statistical ``emulators'' or
-%``surrogate models'' \citep{santner2003design,gramacy2020surrogates}.
+We are particularly motivated by the Mira-Titan simulation suite \citep{moran2023mira}
+which defines ``cosmologies'' according to eight cosmological parameters (more on this
+in Section TBD).  For a particular cosmology, the simulation of the power matter spectra 
+is rather complex.  Although we presume there exists some true matter spectra as a function
+of wavenumber, we are not able to directly observe it.  Rather, the simulation outputs 
+a convoluted patchwork of potential spectra, which vary in fidelity and which are trusted
+only over particular wavenumber ranges.  Our first objective is to synthesize expert knowledge
+and all of this simulation information into an effective estimate of the true underlying
+matter spectrum.
+\textcolor{red}{(@Steve - it might be worth moving something like Figure 1
+here to help in this narration.  It is important that we impress upon the reader why
+it is difficult to estimate the truth.  We can still refer back to the figure in 
+Section 2 and narrate it further.)}
 
-Leveraging these simulations allows for emulation and prediction of the matter power spectrum under varying specifications for eight different cosmological parameters (i.e., different cosmologies).
-Estimating and understanding the influence of these parameters on cosmic expansion are areas of active research.
-In addition to presenting the Mira-Titan simulation suite, \cite{moran2023mira} builds off of previous spectrum emulation \citep{lawrence2017mira} to provide the final emulator (Cosmic Emu) based on the full suite of 117 batches of simulations.
+There are several challenges to estimating the underlying spectra.  Our model
+must: handle multiple correlated functional observations, offer enough 
+flexibility to accommodate the nonstationarity present in power matter spectra 
+(which vary in smoothenss across wavenumbers), provide effective uncertainty
+quantification (UQ), and allow for the incorporation of expert knowledge 
+regarding the wavenumbers over which various outputs are trusted.  To this end, 
+we propose a Bayesian hierarchical model which treats each of functional simulation 
+output as a realization of a deep Gaussian process \citep[DGP;][]{damianou2013deep} 
+centered on the true underlying spectra.  The hierarchical structure naturally
+accommodates the correlated observations.  The depth of the GP offers nonstationary
+flexibility.  The Bayesian framework facilitates UQ and the incorporation of prior
+knowledge.  While Bayesian DGPs have been previously deployed for computer experiments 
+with scalar responses \citep[e.g.,][]{sauer2023active,sauer2023vecchia,ming2023deep}, 
+they have yet to be extended to correlated functional outputs. 
+We will demonstrate profiency on synthetic examples and a low fidelity
+cosmological simulation, before deploying our model on the motivating 
+Mira-Titan simulation.
 
-In this work, we expand on the previous emulation approaches and propose a novel synthesis of methods to quantify uncertainty and predict matter power spectra for different cosmologies using deep Gaussian processes (DGPs).
-In order to handle multiple realizations of functional output (from the Mira-Titan simulation suite), we use hierarchical modeling and basis functions within our DGP framework to accurately estimate and predict power spectra for different cosmologies.
-When comparing to Cosmic Emu, we see favorable results for our method.
-The code to reproduce these results are available at \url{https://github.com/stevewalsh124/dgp.hm}.
-\textcolor{red}{Please note: this repository is currently private until we hear back regarding permissions for sharing the Mira-Titan data alongside our code.}
+Our second objective is to leverage simulation data from a limited set of
+cosmologies in order to predict power matter spectra for unobserved cosmologies.
+Due to the expense and propriatary nature of the Mira-Titan simulation suite
+\textcolor{red}{(do we know how long it takes to run?  Is it really propriatary?)},
+it is infeasible to evaluate the simulation for every possible eight-dimensional
+cosmological configuration which may be of interest.  Instead, we desire
+a statistical ``emulator'' or ``surrogate model'' 
+\citep{santner2003design,gramacy2020surrogates} which will provide quick and effective
+predictions of power matter spectra for any cosmological parameterization.
+\citet{moran2023mira} provide a state-of-the-art emulator for the Mira-Titan
+simulation suite, which was trained on 117 simulations, and is termed 
+the ``Cosmic Emu''.\footnote{\url{https://github.com/lanl/CosmicEmu}}
+We leverage the same training data, in conjunction with our Bayesian hierarchical
+DGP model and basis function representations, to train a Gaussian process surrogate 
+on the principal component weights in order to predict spectra for unobserved
+cosmologies.  \textcolor{red}{(need more cites, like to Daves work)}
+Our method compares favorably to Cosmic Emu on held-out cosmologies.
 
-The remainder of the paper is organized as follows: Section \ref{sec:data} provides more details regarding the Mira-Titan simulation suite.
-Section \ref{sec:hm_fit} describes our model and the training procedure for both toy data as well as the Mira-Titan dataset.
-With this trained model, Section \ref{sec:pred} explains our modeling approach for predictions at held-out cosmologies and compares with other methods.
-Section \ref{sec:disc} concludes with a discussion of our contributions and avenues for future research.
+\textcolor{red}{(Should we introduce the CAMB data here?  I think so, if we
+want to refer to it in the next paragraph.)}
+
+The remainder of the paper is organized as follows.  Section \ref{sec:data} 
+introduces our motivating application, the Mira-Titan simulation suite.  
+Section \ref{sec:hm_fit} describes our hierarchical Bayesian DGP model and 
+validates its performance on synthetic exercises and the CAMB simulation.  
+Leveraging this trained model, Section \ref{sec:pred} details 
+our procedure for predicting at unobserved cosmologies, benchmarking against
+state-of-the-art competitors on the CAMB and Mira-Titan simulations. 
+Section \ref{sec:disc} concludes with a discussion of our contributions and avenues for future research.  We provide reproducible code and an R package for our Bayesian DGP model
+at \url{https://github.com/stevewalsh124/dgp.hm}. 
+\textcolor{red}{Please note: this repository is currently private until we 
+hear back regarding permissions for sharing the Mira-Titan data alongside our code.}
+
+%Leveraging these simulations allows for emulation and prediction of the matter power 
+%spectrum under varying specifications for eight different cosmological parameters 
+%(i.e., different cosmologies). Estimating and understanding the influence of these 
+%parameters on cosmic expansion are areas of active research. In addition to 
+%presenting the Mira-Titan simulation suite, \cite{moran2023mira} builds off 
+%of previous spectrum emulation \citep{lawrence2017mira} to provide the final 
+%emulator (Cosmic Emu) based on the full suite of 117 batches of simulations.
+
+%In this work, we expand on the previous emulation approaches and propose a 
+%novel synthesis of methods to quantify uncertainty and predict matter power 
+%spectra for different cosmologies using deep Gaussian processes (DGPs). In order 
+%to handle multiple realizations of functional output (from the Mira-Titan 
+%simulation suite), we use hierarchical modeling and basis functions within our 
+%DGP framework to accurately estimate and predict power spectra for different 
+%cosmologies.
 

### 2025-04-04 paper/main.tex
move url to footnote?
edits to manuscript; include CAMB/higdon references, initial CAMB image, other small tweaks

@@
 We are particularly motivated by the Mira-Titan simulation suite \citep{moran2023mira}
-which defines ``cosmologies'' according to eight cosmological parameters (more on this
-in Section TBD).  For a particular cosmology, the simulation of the power matter spectra 
+which defines ``cosmologies'' according to eight cosmological parameters (\textcolor{red}{more on this
+in Section TBD}).  For a particular cosmology, the simulation of the power matter spectra 
 is rather complex.  Although we presume there exists some true matter spectra as a function
@@
 matter spectrum.
-\textcolor{red}{(@Steve - it might be worth moving something like Figure 1
-here to help in this narration.  It is important that we impress upon the reader why
-it is difficult to estimate the truth.  We can still refer back to the figure in 
-Section 2 and narrate it further.)}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=6in]{plot_data.png}
+    \caption{Left: The perturbation theory, low resolution runs, and high resolution run 
+             for the first training cosmology. The weighted average of these is shown as a dashed line. 
+             Dotted lines at the bottom indicate where each data type is deemed approximately unbiased. 
+             Right: The same image, but restricted to wavenumber ($k$) values where the low resolution 
+             runs are approximately unbiased.}
+    \label{fig:plot_data}
+\end{figure}
 
@@
 cosmologies in order to predict power matter spectra for unobserved cosmologies.
-Due to the expense and propriatary nature of the Mira-Titan simulation suite
-\textcolor{red}{(do we know how long it takes to run?  Is it really propriatary?)},
+Due to the computationally expensive nature of the Mira-Titan simulation suite
+\textcolor{red}{(do we know how long it takes to run?)},
 it is infeasible to evaluate the simulation for every possible eight-dimensional
@@
 on the principal component weights in order to predict spectra for unobserved
-cosmologies.  \textcolor{red}{(need more cites, like to Daves work)}
+cosmologies \citep{higdon2008computer, higdon2010estcosmo}. 
 Our method compares favorably to Cosmic Emu on held-out cosmologies.
 
-\textcolor{red}{(Should we introduce the CAMB data here?  I think so, if we
-want to refer to it in the next paragraph.)}
+For a comprehensive assessment of our Bayesian hierarchical DGP model, we 
+further study its performance in a simulation study with synthetic data mimicing 
+the Mira-Titan data. Additionally, we also train our model with data output from 
+the Code for Anisotropies in the Microwave Background \citep[CAMB,][]{lewis2011CAMB};
+this model output exhibits similar behavior to Mira-Titan as well, but has the benefit 
+of including a ``true" power spectrum corresponding toeach cosmology.
 
@@
 Section \ref{sec:hm_fit} describes our hierarchical Bayesian DGP model and 
-validates its performance on synthetic exercises and the CAMB simulation.  
+validates its performance on synthetic exercises and the CAMB simulations.  
 Leveraging this trained model, Section \ref{sec:pred} details 
@@
 state-of-the-art competitors on the CAMB and Mira-Titan simulations. 
-Section \ref{sec:disc} concludes with a discussion of our contributions and avenues for future research.  We provide reproducible code and an R package for our Bayesian DGP model
-at \url{https://github.com/stevewalsh124/dgp.hm}. 
+Section \ref{sec:disc} concludes with a discussion of our contributions and avenues for 
+future research.  We provide reproducible code and an R package for our Bayesian DGP model
+in a public git repository.\footnote{\url{https://github.com/stevewalsh124/dgp.hm}}
 \textcolor{red}{Please note: this repository is currently private until we 
@@
 
-The Mira-Titan simulation suite is a dataset consisting of simulated power spectra for 117 different cosmologies.
-Each of these cosmologies is specified by 8 cosmological parameter values, which we will denote as $\theta \in \mathbb{R}^8$.
-Different simulation output exists for different values of a ninth value, the redshift; in this work, we fix this value at 0.
-For each cosmology, we have a batch of 18 simulated spectra: an inexpensive spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from lower-resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from a high resolution simulation ($y_h$).
+The Mira-Titan simulation suite is a dataset consisting of simulated power spectra for 
+117 different cosmologies. Each of these cosmologies is specified by 8 cosmological 
+parameter values, which we will denote as $\theta \in \mathbb{R}^8$. Different simulation 
+output exists for different values of a ninth value, the redshift; in this work, we fix 
+this value at 0. For each cosmology, we have a batch of 18 simulated spectra: an inexpensive 
+spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from 
+lower-resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from 
+a high resolution simulation ($y_h$).
 
 These spectra are represented as a function of wavenumber $k$, and hence is denoted $P(k)$. 
-Following \cite{moran2023mira}, we model on the emulation space $\mathcal{P}(k)=\log_{10}\left(\frac{k^{1.5}P(k)}{2\pi^2}\right)$.
-The output is available for $0.001 \leq k \leq 5$ for $n=351$ different values of $k$, but each data type has differing values where it is deemed approximately unbiased.
-For $k<0.04$, only $y_p$ provides a reliable estimate of the spectra.
-$y_{\ell_r}$ are valid for $0.04 \leq k \leq 0.25$, and $y_h$ is valid for $0.04 \leq k \leq 5$.
-Figure \ref{fig:plot_data} shows an example of the output for a particular cosmology.
+Following \cite{moran2023mira}, we model on the emulation space 
+$\mathcal{P}(k)=\log_{10}\left(\frac{k^{1.5}P(k)}{2\pi^2}\right)$. The output is available 
+for $0.001 \leq k \leq 5$ for $n=351$ different values of $k$, but each data type has differing 
+values where it is deemed approximately unbiased. For $k<0.04$, only $y_p$ provides a reliable 
+estimate of the spectra. $y_{\ell_r}$ are valid for $0.04 \leq k \leq 0.25$, and $y_h$ is valid 
+for $0.04 \leq k \leq 5$. Figure \ref{fig:plot_data} shows an example of the output for a 
+particular cosmology.
 
-From the multiple low- and high-resolution spectra across all cosmologies, \cite{moran2023mira} obtained precision estimates of these spectra across the $n$ wavenumber values ($p_1,\dots,p_n$) using a log-log regression model, and a multiplier $c$ for the increase in precision from the low- to high- resolution output.
-Given that the perturbation theory spectra is non-stochastic, we treat this as having a precision of $10^5$.
-We leverage these precision estimates within our model at each data type's unbiased wavenumber values to define $\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, $\Lambda_{\ell} = \text{diag}\left(16(p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, $\Lambda_h = \text{diag}\left(c(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$.
-Here, each precision matrix is diagonal and the diagonal elements consist of the precision values or zero, depending on whether the $k$ value is approximately unbiased for that data type.
-With this, we calculate a weighted average for each cosmology: $\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for the weighted average for a particular cosmology.
+From the multiple low- and high-resolution spectra across all cosmologies, 
+\cite{moran2023mira} obtained precision estimates of these spectra across the $n$ 
+wavenumber values ($p_1,\dots,p_n$) using a log-log regression model, and a multiplier 
+$c$ for the increase in precision from the low- to high- resolution output. Given 
+that the perturbation theory spectra is non-stochastic, we treat this as having a 
+precision of $10^5$. We leverage these precision estimates within our model at each 
+data type's unbiased wavenumber values to define 
+$\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, 
+$\Lambda_{\ell} = \text{diag}\left(16(p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, 
+$\Lambda_h = \text{diag}\left(c(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$. 
+Here, each precision matrix is diagonal and the diagonal elements consist of the 
+precision values or zero, depending on whether the $k$ value is approximately unbiased 
+for that data type. With this, we calculate a weighted average for each cosmology: 
+$\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, 
+where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for 
+the weighted average for a particular cosmology.
 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=6in]{plot_data.png}
-    \caption{Left: The perturbation theory, low resolution runs, and high resolution run for the first training cosmology.
-The weighted average of these is shown as a dashed line.
-Dotted lines at the bottom indicate where each data type is deemed approximately unbiased.
-Right: The same image, but restricted to $k$ values where the low resolution runs are approximately unbiased.}
-    \label{fig:plot_data}
-\end{figure}
-
-We will handle the data in two stages.
-First, we will focus on 111 training cosmologies and obtain the best estimated spectrum for each.
-Second, we will use these estimates to predict the spectra for six held-out cosmologies.
+We will handle the data in two stages.  First, we will focus on 111 training cosmologies 
+and obtain the best estimated spectrum for each.  Second, we will use these estimates 
+to predict the spectra for six held-out cosmologies.
 
@@
 
-Here we describe our first contribution: the use of a Bayesian hierarchical model to estimate the underlying spectra (and quantify uncertainty) for a particular cosmology.
+Here we describe our first contribution: the use of a Bayesian hierarchical model 
+to estimate the underlying spectra (and quantify uncertainty) for a particular cosmology.  
 We focus solely on one cosmology, independent of the others.
@@
 
-Here, we introduce the Bayesian Gaussian Process (GP) model leveraging the aggregated data $D_n \equiv \{X_n, \bar Y, \Lambda\}$ for a particular cosmology.
-We assume the matter power spectrum $S$ for a given cosmology follows a GP with mean $\mu$ and covariance matrix $\Sigma_S$.
+Here, we introduce the Bayesian Gaussian Process (GP) model leveraging the aggregated 
+data $D_n \equiv \{X_n, \bar Y, \Lambda\}$ for a particular cosmology. We assume 
+the matter power spectrum $S$ for a given cosmology follows a GP with mean $\mu$ 
+and covariance matrix $\Sigma_S$.
 
-From plots of the multiple low-resolution spectra (e.g., Figure \ref{fig:plot_data}), we find they vary smoothly about some mean $S$, suggesting that a dense covariance matrix $\Sigma_\varepsilon$ (accounting for spatial dependence) may outperform a diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$.
-In previous work, we have modeled multiple candidates for $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and found that a Mat\'ern covariance $\Sigma_\ell$ trained on the low-resolution runs (pre-scaled by the precision values and subtracting a LOESS-smoothed average) performs the best.
-We specify this structure within $\Sigma_\varepsilon$ and refer readers to \cite{walsh2023bayesian} for a more detailed exposition.
-This yields $\Sigma_\varepsilon=\left(\Lambda_p^{-1} + \Sigma_\ell + \Lambda_h^{-1}\right)^{-1}$, where $\Lambda_p, \Sigma_\ell^{-1}, \text{ and } \Lambda_h$ are all $n\times n$ and are 0 wherever the corresponding data type is biased.
+From plots of the multiple low-resolution spectra (e.g., Figure \ref{fig:plot_data}), 
+we find they vary smoothly about some mean $S$, suggesting that a dense covariance 
+matrix $\Sigma_\varepsilon$ (accounting for spatial dependence) may outperform a 
+diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$. In previous work, we have 
+modeled multiple candidates for $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and 
+found that a Mat\'ern covariance $\Sigma_\ell$ trained on the low-resolution runs 
+(pre-scaled by the precision values and subtracting a LOESS-smoothed average) performs 
+the best. We specify this structure within $\Sigma_\varepsilon$ and refer readers 
+to \cite{walsh2023bayesian} for a more detailed exposition. This yields 
+$\Sigma_\varepsilon=\left(\Lambda_p^{-1} + \Sigma_\ell + \Lambda_h^{-1}\right)^{-1}$, 
+where $\Lambda_p, \Sigma_\ell^{-1}, \text{ and } \Lambda_h$ are all $n\times n$ 
+and are 0 wherever the corresponding data type is biased.
 
-We transform $X$ values to fall within $[0,1]$ and $Y$ so that the mean and variance are 0 and 1, respectively.
-This allows for a straightforward specification of priors, where the hyperparameters for the Mat\'ern covariance $\Sigma_S$ each take a vague prior.
+We transform $X$ values to fall within $[0,1]$ and $Y$ so that the mean and variance 
+are 0 and 1, respectively. This allows for a straightforward specification of priors, 
+where the hyperparameters for the Mat\'ern covariance $\Sigma_S$ each take a vague prior. 
 Specifically:
@@
 \begin{align}
-\Sigma_\ell^{i,j} = \tau_\ell^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_\ell}} + \frac{5d^2}{3\theta_\ell}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_\ell}}\right),\\
-\Sigma_S^{i,j} = \tau_S^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_S}} + \frac{5d^2}{3\theta_S}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_S}}\right),
+\Sigma_\ell^{i,j} = \tau_\ell^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_\ell}} + 
+  \frac{5d^2}{3\theta_\ell}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_\ell}}\right),\\
+\Sigma_S^{i,j} = \tau_S^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_S}} + 
+  \frac{5d^2}{3\theta_S}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_S}}\right),
 \end{align}
 
-where $d=||x_i-x_j||^2$.
-From here, we estimate $\tau_\ell, \theta_\ell$ with maximum likelihood, and conditional on $\Sigma_\varepsilon$, we specify priors for $\tau_S, \theta_S$ following \cite{sauer2023active}: \textcolor{cyan}{Annie, please ensure the most up-to-date prior specs are shown above and here.
-Here is what I found from \texttt{fit\$settings} when using \texttt{deepgp v1.1.3}: $\theta_S \sim G(\alpha=1.2, \beta=1)$ and $\tau_S \propto 1/\tau_S^2$.}
+where $d=||x_i-x_j||^2$. From here, we estimate $\tau_\ell, \theta_\ell$ with maximum 
+likelihood, and conditional on $\Sigma_\varepsilon$, we specify priors for $\tau_S, \theta_S$ 
+following \cite{sauer2023active}: \textcolor{cyan}{Annie, please ensure the most up-to-date 
+prior specs are shown above and here. Here is what I found from fit\$settings when using 
+deepgp v1.1.3: $\theta_S \sim G(\alpha=1.2, \beta=1)$ and $\tau_S \propto 1/\tau_S^2$.}
 
@@
 
-Therefore, we model $\bar Y|S$ as a GP with mean $S$ and block-diagonal covariance matrix $\Sigma_\varepsilon$:
+Therefore, we model $\bar Y|S$ as a GP with mean $S$ and block-diagonal covariance 
+matrix $\Sigma_\varepsilon$:
 
@@
 Upon integrating out $S$, we obtain $\bar Y \sim \mathcal{N}_n(\mu, \Sigma_S+\Sigma_\varepsilon)$. 
-Additionally, this model yields the posterior distribution $S|\bar Y \sim \mathcal{N}_n(m, C)$, where $C=\left(\Sigma_S^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1}$ and $m=C\left(\Sigma_\varepsilon^{-1}\bar Y+\Sigma_S^{-1}\mu\right)$.
-This distribution provides the posterior mean and corresponding uncertainty for $S$ within a given cosmology. 
+Additionally, this model yields the posterior distribution $S|\bar Y \sim \mathcal{N}_n(m, C)$, 
+where $C=\left(\Sigma_S^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1}$ and 
+$m=C\left(\Sigma_\varepsilon^{-1}\bar Y+\Sigma_S^{-1}\mu\right)$. 
+This distribution provides the posterior mean and corresponding uncertainty for 
+$S$ within a given cosmology. 
 % With this hierarchical model, we improve the efficiency of estimating $S|\bar{Y}$.
@@
 
-Considering the Mira-Titan data, some nonstationarity is expected (e.g., due to baryonic acoustic oscillations when $-2 \leq \log_{10}(k) \leq -1$).
-To account for this, we incorporate a latent layer $W$ within the hierarchical model to warp locations of $X$ and incorporate nonstationary dynamics of the matter power spectra.
+Considering the Mira-Titan data, some nonstationarity is expected (e.g., due to 
+baryonic acoustic oscillations when $-2 \leq \log_{10}(k) \leq -1$). To account 
+for this, we incorporate a latent layer $W$ within the hierarchical model to warp 
+locations of $X$ and incorporate nonstationary dynamics of the matter power spectra. 
 When $W$ is a GP, this creates a deep Gaussian process (DGP) model \citep{damianou2013deep}. 
-Additional latent layers could be considered, although one is sufficient for this dataset \citep{dunlop2018deep}.
+Additional latent layers could be considered, although one is sufficient for this 
+dataset \citep{dunlop2018deep}.
 
@@
 
-We have found that whether the prior mean of $W$ is set to 0 or $X$ \citep[which would indicate stationarity apriori,][]{schmidt2003bayesian} does not have a substantial impact on the model fit.
-We construct this model in the compositional form, and also use Markov Chain Monte Carlo (MCMC) to estimate and obtain full uncertainty for each of the model parameters \citep{sauer2023active}.
+We have found that whether the prior mean of $W$ is set to 0 or $X$ 
+\citep[which would indicate stationarity apriori,][]{schmidt2003bayesian} does not 
+have a substantial impact on the model fit. We construct this model in the 
+compositional form, and also use Markov Chain Monte Carlo (MCMC) to estimate and 
+obtain full uncertainty for each of the model parameters \citep{sauer2023active}.
 
-\textcolor{cyan}{Annie: want to add something here?
-For $\theta_w$, we also specify a Gamma prior of the form $\theta_w \sim G(\alpha=1.2, \beta=2)$.
-We discuss the elliptical slice sampling of $W$, but otherwise it is plugged in to the methods of the previous section with very few changes.}
+\textcolor{cyan}{Annie: want to add something here? For $\theta_w$, we also specify 
+a Gamma prior of the form $\theta_w \sim G(\alpha=1.2, \beta=2)$. We discuss the 
+elliptical slice sampling of $W$, but otherwise it is plugged in to the methods 
+of the previous section with very few changes.}
 
-Through the synthesis of the DGP model alongside the additional level of hierarchical modeling (HM) to incorporate multiple functional model runs, we call this method DGP.HM.
+Through the synthesis of the DGP model alongside the additional level of hierarchical 
+modeling (HM) to incorporate multiple functional model runs, we call this method DGP.HM.
 
@@
 
-We compare DGP.HM to competing models with a simulation study where the underlying true curve is known.
-Following \cite{moran2024dpc}, we consider two functions ($f_1$ and $f_2$) each with two variance specifications (settings ``4" and ``5", \textcolor{magenta}{we can consider additional homoskedastic cases in the appendix if necessary}).
+We compare DGP.HM to competing models with a simulation study where the underlying 
+true curve is known. Following \cite{moran2024dpc}, we consider two functions 
+($f_1$ and $f_2$) each with two variance specifications (settings ``4" and ``5", 
+\textcolor{magenta}{we can consider additional homoskedastic cases in the appendix if necessary}).
 
@@
 
-For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are random where $m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 \sim \text{Uniform}(0.6,1.4)$.
+For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are random where 
+$m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 
+\sim \text{Uniform}(0.6,1.4)$.
 
-For the two variance settings, we consider matrices $\Sigma_4$ and $\Sigma_5$ to have a Mat\'ern covariance function with a nugget of $10^{-8}$ and smoothness of 2.5.
-For setting ``4", we have $\tau^2=0.0225$ and $\theta=0.01$.
-For setting ``5", we we have $\tau^2=0.01$ and $\theta=0.05$, and additionally scale this covariance matrix by a vector of values $s$ so that $\Sigma_5 = s M s^T$ is nonstationary; here $M$ is a Mat\'ern covariance matrix with parameters $\tau^2$ and $\theta$.
+For the two variance settings, we consider matrices $\Sigma_4$ and $\Sigma_5$ to 
+have a Mat\'ern covariance function with a nugget of $10^{-8}$ and smoothness of 2.5. 
+For setting ``4", we have $\tau^2=0.0225$ and $\theta=0.01$. For setting ``5", we have 
+$\tau^2=0.01$ and $\theta=0.05$, and additionally scale this covariance matrix by 
+a vector of values $s$ so that $\Sigma_5 = s M s^T$ is nonstationary; here $M$ is 
+a Mat\'ern covariance matrix with parameters $\tau^2$ and $\theta$.
 
-For each simulated function and variance specification, we generate 20 values of $m_1, m_2, f_1, f_2$, and conditional on these values, we sample either 5 or 20 functional observations and estimate the underlying true function based on the observations.
+For each simulated function and variance specification, we generate 20 values of 
+$m_1, m_2, f_1, f_2$, and conditional on these values, we sample either 5 or 20 
+functional observations and estimate the underlying true function based on the observations.
 
@@
     \includegraphics[width=6in]{sims_logS.png}
-    \caption{Boxplot of log scores across two different functions and covariance settings (lower is better).
-For each case, the boxplot is constructed from 20 random batches were simulated.
-Each column represents a function/variance specification pair, and each row shows results for batch sizes of either 5 (top) or 20 (bottom).}
+    \caption{Boxplot of log scores across two different functions and covariance 
+             settings (lower is better). For each case, the boxplot is constructed 
+             from 20 random batches were simulated. Each column represents a function/variance 
+             specification pair, and each row shows results for batch sizes of either 
+             5 (top) or 20 (bottom).}
     \label{fig:sims_logS}
@@
 
-The competing methods we consider are an out-of-the-box DGP fit using the \texttt{deepgp} package \citep{sauer2023active}, a heteroskedastic GP fit using the \texttt{hetGP} package \citep{binois2018practical, binois2021hetgp}, as well as a deep process convolution (DPC) approach, which is utilized within Cosmic Emu \citep{moran2023mira}.
+The competing methods we consider are an out-of-the-box DGP fit using the 
+\texttt{deepgp} package \citep{sauer2023active}, a heteroskedastic GP fit using 
+the \texttt{hetGP} package \citep{binois2018practical, binois2021hetgp}, as well 
+as a deep process convolution (DPC) approach, which is utilized within Cosmic 
+Emu \citep{moran2023mira}.
 
-The metrics we use to evaluate performance are mean squared error (MSE) and the log score, a proper scoring rule which takes both prediction and intervals into consideration \citep{gneiting2007strictly}.
-For both metrics, a lower value indicates better performance.
-Log score and MSE results are shown in Figures \ref{fig:sims_logS} and \ref{fig:sims_MSE} respectively; our DGP.HM model performs favorably against these current benchmark methods.
-\textcolor{magenta}{If we mention that DGP.HM timings speed up compared to \texttt{deepgp}, we would also need to mention that \texttt{hetGP} is much faster than both of these.}
+The metrics we use to evaluate performance are mean squared error (MSE) and the 
+log score, a proper scoring rule which takes both prediction and intervals into 
+consideration \citep{gneiting2007strictly}. For both metrics, a lower value indicates 
+better performance. Log score and MSE results are shown in Figures \ref{fig:sims_logS} 
+and \ref{fig:sims_MSE} respectively; our DGP.HM model performs favorably against 
+these current benchmark methods. \textcolor{magenta}{If we mention that DGP.HM timings 
+speed up compared to \texttt{deepgp}, we would also need to mention that \texttt{hetGP} 
+is much faster than both of these.}
 
@@
     \includegraphics[width=6in]{sims_MSE.png}
-    \caption{Boxplot of MSEs across two different functions and covariance settings (lower is better).
-For each case, the boxplot is constructed from 20 random batches were simulated.
-Each column represents a function/variance specification pair, and each row shows results for batch sizes of either 5 (top) or 20 (bottom).}
+    \caption{Boxplot of MSEs across two different functions and covariance settings 
+             (lower is better). For each case, the boxplot is constructed from 20 
+             random batches were simulated. Each column represents a function/variance 
+             specification pair, and each row shows results for batch sizes of 
+             either 5 (top) or 20 (bottom).}    
     \label{fig:sims_MSE}
@@
 
-In addition to simulated data, we investigate model fits the for different cosmological datasets (based on linear theory and/or runs from the Code for Anisotropies in the Microwave Background, CAMB).
-\textcolor{magenta}{We have looked at two sets of model runs, ``LT" runs where there is an underlying ``true" value alongside 15 computer model runs, or the ``CAMB" runs, where we only have a single output.
-The performance from the CAMB runs was better (with a 6-dimensional design space), and also did not require obtaining a posterior mean from multiple model runs.
-When looking at the CAMB runs on the $\Delta$ space, we see it generally is withing $\pm 5\%$ error.
-Regarding these results, compared to the results for the 8-dimensional LT runs which had less impressive predictive performance, I'm open to suggestions on whether to include details on one or the other.}
+In addition to simulated data, we investigate model fits for different cosmological 
+datasets from CAMB \citep{lewis2011CAMB}. \textcolor{magenta}{We have looked at two sets of model 
+runs, ``LT" runs where there is an underlying ``true" value alongside 15 computer 
+model runs, or the ``CAMB" runs, where we only have a single output. The performance 
+from the CAMB runs was better (with a 6-dimensional design space), and also did not 
+require obtaining a posterior mean from multiple model runs. When looking at the 
+CAMB runs on the $\Delta$ space, we see it generally is withing $\pm 5\%$ error. 
+Regarding these results, compared to the results for the 8-dimensional LT runs 
+which had less impressive predictive performance, I'm open to suggestions on whether 
+to include details on one or the other.}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{CAMB_fit_model1.jpg}
+    \caption{Example of model fit for the first model from the CAMB data, with LOESS
+             mean term subtracted. \textcolor{red}{Steve: Update this plot to be
+             more clean, use different line types: dotted, dash, solid, etc.}}   
+    \label{fig:fit_camb}
+\end{figure}
 
@@
     \includegraphics[width=6in]{plot_pca_CAMB.png}
-    \caption{Proportion of $\Delta_{emu}$ vs. $\Delta_{holdout}$ for model runs 11-42 (excluding 27-28).}   
+    \caption{Proportion of $\Delta_{emu}$ vs. $\Delta_{holdout}$ for model 
+             runs 11-42 (excluding 27-28).}   
     \label{fig:pca_camb}
@@
 
-To train our hierarchical model, we initialize our hyperparameter estimates across all cosmologies by modeling the first (M001) with 50,000 MCMC runs, and use these as initial values for all cosmologies.
-Specifically, we save the last MCMC iteration of $W$, $\theta_w$ and $\theta_S$ as $W_0$, $\theta_{w0}$ and $\theta_{S0}$ respectively and use each as starting values for all 111 training cosmology fits, where each cosmology is fit with 10,000 MCMC runs (5,000 removed as burn-in and every 5th sample is kept, resulting in 1,000 posterior samples).
+To train our hierarchical model, we initialize our hyperparameter estimates across 
+all cosmologies by modeling the first (M001) with 50,000 MCMC runs, and use these 
+as initial values for all cosmologies. Specifically, we save the last MCMC 
+iteration of $W$, $\theta_w$ and $\theta_S$ as $W_0$, $\theta_{w0}$ and $\theta_{S0}$ 
+respectively and use each as starting values for all 111 training cosmology fits, 
+where each cosmology is fit with 10,000 MCMC runs (5,000 removed as burn-in and 
+every 5th sample is kept, resulting in 1,000 posterior samples).
 
-From these samples, we can estimate and quantify the uncertainty within each cosmology's average and latent layer $W$.
-Figure \ref{fig:plot_fit} shows an example of the uncertainty for our posterior mean for the first (of 111) training cosmologies; the posterior mean is subtracted in order to visualize this uncertainty.
-Figure \ref{fig:plot_warp} provides elliptical slice samples of $W$ along with the average warping.
-Some departure from stationarity is estimated in the higher $X$ (i.e., scaled $k$) values.
+From these samples, we can estimate and quantify the uncertainty within each 
+cosmology's average and latent layer $W$. Figure \ref{fig:plot_fit} shows an 
+example of the uncertainty for our posterior mean for the first (of 111) training 
+cosmologies; the posterior mean is subtracted in order to visualize this uncertainty. 
+Figure \ref{fig:plot_warp} provides elliptical slice samples of $W$ along with the 
+average warping. Some departure from stationarity is estimated in the higher $X$ 
+(i.e., scaled $k$) values.
 
@@
     \includegraphics[width=6in]{plot_fit.png}
-    \caption{95\% credible intervals for the power spectrum of model 1 (posterior mean subtracted).
-Plots of the perturbation theory, low-resolution and high-resolution runs are also shown, alongside the weighted average.}
+    \caption{95\% credible intervals for the power spectrum of model 1 (posterior mean 
+             subtracted). Plots of the perturbation theory, low-resolution and high-resolution 
+             runs are also shown, alongside the weighted average.}
     \label{fig:plot_fit}
@@
    \caption{Left: MCMC samples of $W$ for model M001 after burnin and thinning. 
-Right: Corresponding 95\% credible intervals for $W$, where departure from the identity line represents nonstationary behavior.}
+            Right: Corresponding 95\% credible intervals for $W$, where departure 
+            from the identity line represents nonstationary behavior.}
    \label{fig:plot_warp}
@@
 
-Upon obtaining the posterior means from each of the 111 training cosmologies, this section details our second contribution - the use of the Bayesian DGP.HM model to predict the curve for held-out cosmologies.
-This is achieved with modeling the weights of different basis functions as GPs.
+Upon obtaining the posterior means from each of the 111 training cosmologies, this 
+section details our second contribution - the use of the Bayesian DGP.HM model 
+to predict the curve for held-out cosmologies. This is achieved with modeling the 
+weights of different basis functions as GPs.
 
@@
 
-Here, we begin with a $n \times m$ matrix composed of $m=111$ posterior means, each of length $n$, denoted by $\boldsymbol\eta$.
-To facilitate more efficient estimation, we use singular value decomposition \citep[SVD; e.g.,][]{banerjee2014linear}, where $\boldsymbol\eta = UDV^T$; $U$ and $V$ are orthogonal matrices of size $n \times n$ and  $m\times m$, respectively.
-$D$ is a diagonal matrix containing the singular values.
-For this application, $\boldsymbol\eta$ has rank $r=\text{min}\{m,n\}=111$, which determines the number of non-zero elements in $D$.
-We find an alternative decomposition where $\Sigma$ is a diagonal matrix with the non-zero singular values of $\boldsymbol\eta$, and $U_1$ and $V_1$ are matrices with orthonormal columns of size $n_\eta \times r$ and $m \times r$, respectively.
+Here, we begin with a $n \times m$ matrix composed of $m=111$ posterior means, 
+each of length $n$, denoted by $\boldsymbol\eta$. To facilitate more efficient estimation, 
+we use singular value decomposition \citep[SVD; e.g.,][]{banerjee2014linear}, 
+where $\boldsymbol\eta = UDV^T$; $U$ and $V$ are orthogonal matrices of size $n \times n$ 
+and $m\times m$, respectively. $D$ is a diagonal matrix containing the singular values. 
+For this application, $\boldsymbol\eta$ has rank $r=\text{min}\{m,n\}=111$, which 
+determines the number of non-zero elements in $D$. We find an alternative decomposition 
+where $\Sigma$ is a diagonal matrix with the non-zero singular values of $\boldsymbol\eta$, 
+and $U_1$ and $V_1$ are matrices with orthonormal columns of size $n_\eta \times r$ 
+and $m \times r$, respectively.
 
-Following \cite{higdon2008computer, higdon2010estcosmo}, we equivalently decompose $\boldsymbol\eta$ into a principal component (PC) basis matrix $B^* = \frac{1}{\sqrt{m}}U_1\Sigma$ along with its corresponding weights $W^* = \sqrt{m}V_1$.
-Without loss of generality, we can perform this decomposition after an overall mean trend is removed.
-If we use $p_\eta < r$ PCs, this will reduce $B$ and $W$ to be the first $p_\eta$ columns of $B^*$ and $W^*$ respectively and result in an approximation where
+Following \cite{higdon2008computer, higdon2010estcosmo}, we equivalently decompose 
+$\boldsymbol\eta$ into a principal component (PC) basis matrix 
+$B^* = \frac{1}{\sqrt{m}}U_1\Sigma$ along with its corresponding weights 
+$W^* = \sqrt{m}V_1$. Without loss of generality, we can perform this decomposition 
+after an overall mean trend is removed. If we use $p_\eta < r$ PCs, this will reduce 
+$B$ and $W$ to be the first $p_\eta$ columns of $B^*$ and $W^*$ respectively and 
+result in an approximation where
 
@@
 
-Figure \ref{fig:mean_PCs_oneW} illustrates this decomposition with the estimated mean trend, the principal components contained in $B$, and one example of how the weight for the first PC will vary dependent on the first of $p_\theta=8$ cosmological parameters contained in $\theta$.
+Figure \ref{fig:mean_PCs_oneW} illustrates this decomposition with the estimated 
+mean trend, the principal components contained in $B$, and one example of how the 
+weight for the first PC will vary dependent on the first of $p_\theta=8$ cosmological 
+parameters contained in $\theta$.
 
-In the following subsections, we consider different modeling strategies for $W$, which will influence predictions and sensitivity analyses based on $\boldsymbol\eta$.
+In the following subsections, we consider different modeling strategies for $W$, 
+which will influence predictions and sensitivity analyses based on $\boldsymbol\eta$.
 
@@
     \centering
-    %\includegraphics[width=\textwidth]{mean_PCs_oneW.png}
-    \caption{The estimated mean trend of the 111 posterior means (left), the principal components contained in $B$ (middle), and an illustration of how the weights for the first PC will change as the first cosmological parameter varies (right) \textcolor{red}{This last plot is wonky because it was made with only the 6 test cosmologies, not a large uniform or FF design, Steve can update this}.}
+    \includegraphics[width=\textwidth]{mean_PCs_oneW.png}
+    \caption{The estimated mean trend of the 111 posterior means (left), the principal 
+    components contained in $B$ (middle), and an illustration of how the weights 
+    for the first PC will change as the first cosmological parameter varies (right) 
+    \textcolor{red}{This last plot is wonky because it was made with only the 6 test 
+    cosmologies, not a large uniform or FF design, Steve can update this}.}
     \label{fig:mean_PCs_oneW}
@@
 
-In alignment with \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use GPs to model the principal components' weights; we shall denote this method GP-PC.
-We assume a powered exponential kernel for the Gaussian process and employ \texttt{GPfit} to perform a multi-start gradient based search for GP's hyperparameters \citep{macdonald2015gpfit}.
-That is, the $i$th vector of PC weights $w_i(\theta)$, $i \in \{1,\ldots,p_\eta\}$, will be modeled as a zero mean GP: $w_i(\theta) \sim GP(0, \sigma^2R)$, with
+In alignment with \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use 
+GPs to model the principal components' weights; we shall denote this method GP-PC. 
+We assume a powered exponential kernel for the Gaussian process and employ 
+\texttt{GPfit} to perform a multi-start gradient based search for GP's hyperparameters 
+\citep{macdonald2015gpfit}. That is, the $i$th vector of PC weights $w_i(\theta)$, 
+$i \in \{1,\ldots,p_\eta\}$, will be modeled as a zero mean GP: 
+$w_i(\theta) \sim GP(0, \sigma^2R)$, with
 
@@
 
-We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\theta\}$. Using these trained GPs from the PCs, we predict the spectra for unobserved cosmologies.
+We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\theta\}$. 
+Using these trained GPs from the PCs, we predict the spectra for unobserved cosmologies.
 
@@
 
-Here we show prediction results for 6 held-out cosmologies.
-We compare our method with Cosmic Emu \citep{moran2023mira}, the emulator constructed using the full suite of 111 Mira-Titan model runs.
-Cosmic Emu uses a Bayesian approach where the smooth spectrum for each cosmology is modeled with a process convolution on Brownian motion.
-Nonstationarity is permitted through modeling the bandwidth parameter with a process convolution as well; this composition is referred to as a deep process convolution \citep{moran2024dpc}.
+Here we show prediction results for 6 held-out cosmologies. We compare our method 
+with Cosmic Emu \citep{moran2023mira}, the emulator constructed using the full suite 
+of 111 Mira-Titan model runs. Cosmic Emu uses a Bayesian approach where the smooth 
+spectrum for each cosmology is modeled with a process convolution on Brownian motion. 
+Nonstationarity is permitted through modeling the bandwidth parameter with a process 
+convolution as well; this composition is referred to as a deep process convolution 
+\citep{moran2024dpc}.
 
-Both Cosmic Emu and DGP.HM are capable of modeling a true underlying functions based upon multiple functional realizations, but our approach utilizes DGPs within a hierarchical model in lieu of deep process convolutions.
-After obtaining the estimated smooth spectra from the training cosmologies, each method utilizes GPs in order to estimate the weights of the PCs from the estimated spectra.
-Below, we compare prediction performance between Cosmic Emu and DGP.HM for the 6 held-out cosmologies.
-\textcolor{magenta}{In addition, we can also predict over a large uniform or full factorial design to obtain estimated main effects for each cosmological parameter in $\theta$, interactions amongst the $p_\theta=8$ parameters, and decompose the overall variation by main effects, two-way interactions, and so on.
-Should this be included?}
+Both Cosmic Emu and DGP.HM are capable of modeling a true underlying functions based 
+upon multiple functional realizations, but our approach utilizes DGPs within a 
+hierarchical model in lieu of deep process convolutions. After obtaining the estimated 
+smooth spectra from the training cosmologies, each method utilizes GPs in order to 
+estimate the weights of the PCs from the estimated spectra. Below, we compare prediction 
+performance between Cosmic Emu and DGP.HM for the 6 held-out cosmologies. 
+\textcolor{magenta}{In addition, we can also predict over a large uniform or full 
+factorial design to obtain estimated main effects for each cosmological parameter in 
+$\theta$, interactions amongst the $p_\theta=8$ parameters, and decompose the overall 
+variation by main effects, two-way interactions, and so on. Should this be included?}
 
-Plots for the average root mean squared error (RMSE) for each prediction method over the 6 cosmologies and each $k$ value are shown in Figure \ref{fig:plot_rmse_k}.
-Comparisons of the spectrum estimation for training cosmologies are shown as well (using RMSE against the weighted average of observations).
-For predictions, DGP.HM achieves the greatest reduction in RMSE in the region where perturbation theory and the low-resolution runs are deemed unbiased.
-For the region where only the high-resolution run is unbiased, Cosmic Emu has a lower average RMSE.
-Over all $k$ values and predictions, DGP.HM has an average RMSE of 0.0089 and Cosmic Emu has an average RMSE of 0.0106.
-Our RMSE is lower at 56\% of $k$ values considered.
-The individual predictions for each method and cosmology are shown in Figure \ref{fig:plot_pred_1to6}.
+Plots for the average root mean squared error (RMSE) for each prediction method over the 
+6 cosmologies and each $k$ value are shown in Figure \ref{fig:plot_rmse_k}. Comparisons 
+of the spectrum estimation for training cosmologies are shown as well (using RMSE against 
+the weighted average of observations). For predictions, DGP.HM achieves the greatest 
+reduction in RMSE in the region where perturbation theory and the low-resolution runs 
+are deemed unbiased. For the region where only the high-resolution run is unbiased, 
+Cosmic Emu has a lower average RMSE. Over all $k$ values and predictions, DGP.HM has 
+an average RMSE of 0.0089 and Cosmic Emu has an average RMSE of 0.0106. Our RMSE is 
+lower at 56\% of $k$ values considered. The individual predictions for each method and 
+cosmology are shown in Figure \ref{fig:plot_pred_1to6}.
 
@@
     \caption{Left: Average RMSE of each prediction method across all $k$ values. 
-Right: Average RMSE for each method when using the held-out computer model runs to train the model.
-In each case, the truth is assumed to be the weighted average of the computer model runs.}
+             Right: Average RMSE for each method when using the held-out computer 
+             model runs to train the model. In each case, the truth is assumed to 
+             be the weighted average of the computer model runs.}
     \label{fig:plot_rmse_k}
@@
 %     \caption{Boxplot of all RMSEs of each method across all six test models. 
-From top to bottom: box plots for our training method, process convolution training method, Cosmic EMU predictions, and our DGP-PC predictions.}
+%.             From top to bottom: box plots for our training method, process 
+%              convolution training method, Cosmic EMU predictions, and our DGP-PC predictions.}
 %     \label{fig:plot_rmse_box}
@@
 
-In this work, we propose a novel Bayesian hierarchical model which utilizes DGPs in order to estimate a smooth underlying function from correlated functional observations (as cosmological computer model runs).
-In addition to the compositional form of the DGP, we include a third component modeling the correlated errors for each spectrum.
-With all smooth spectra estimated from the training set of 111 batches of model runs, we can predict spectra for unobserved cosmologies.
+In this work, we propose a novel Bayesian hierarchical model which utilizes DGPs 
+in order to estimate a smooth underlying function from correlated functional observations 
+(as cosmological computer model runs). In addition to the compositional form of the DGP, 
+we include a third component modeling the correlated errors for each spectrum. 
+With all smooth spectra estimated from the training set of 111 batches of model runs, 
+we can predict spectra for unobserved cosmologies.
 
-Multiple avenues for future research emerge: in addition to estimating dark matter power spectra, batches of hydrodynamical simulations exist, and this method could easily be adapted to handle such datasets.
-Other applications outside of cosmology can be explored, wherever multiple functional outputs exist, perhaps with spatial correlation (for example, observing multiple tropical cyclone forecasts).
-More detailed sensitivity analyses for the cosmological parameters can be explored to gain insights into how each of these parameters interact, and which are most significant.
-From a modeling perspective, changes to the model could include modeling the latent layer $W$ jointly across different cosmologies, as well as considering different redshift values other than $z=0$.
-Leveraging a hybrid model between DGP.HM and Cosmic Emu may also help to obtain accurate estimates on the lower and higher ends of $k$ values, respectively.
-Additionally, Bayesian smoothing spline ANOVA (BSS-ANOVA) or other methods could be compared in addition to DGPs.
+Multiple avenues for future research emerge: in addition to estimating dark matter 
+power spectra, batches of hydrodynamical simulations exist, and this method could 
+easily be adapted to handle such datasets. Other applications outside of cosmology 
+can be explored, wherever multiple functional outputs exist, perhaps with spatial 
+correlation (for example, observing multiple tropical cyclone forecasts). More detailed 
+sensitivity analyses for the cosmological parameters can be explored to gain insights 
+into how each of these parameters interact, and which are most significant. From a 
+modeling perspective, changes to the model could include modeling the latent layer 
+$W$ jointly across different cosmologies, as well as considering different redshift 
+values other than $z=0$. Leveraging a hybrid model between DGP.HM and Cosmic Emu 
+may also help to obtain accurate estimates on the lower and higher ends of $k$ values, 
+respectively. Additionally, Bayesian smoothing spline ANOVA (BSS-ANOVA) or other methods 
+could be compared in addition to DGPs.
 
@@
 \textcolor{magenta}{The different appendices from dissertation as necessary. 
-Can also put details of dataset in supplementary material.} \textcolor{red}{Fix references as necessary, e.g.
-Moran et al.
-2024.}
+Can also put details of dataset in supplementary material.} 
+\textcolor{red}{Fix references as necessary, e.g. Moran et al. 2024.}
 
@@
     \includegraphics[width=6in]{pred_1to6.png}
-    \caption{Results of all six predictions for both methods on the held-out cosmologies (weighted average subtracted).
-CosmicEMU is dashed, DGP.HM is dotted, and solid gray lines show results of held-out computer model runs.}
+    \caption{Results of all six predictions for both methods on the held-out 
+    cosmologies (weighted average subtracted). CosmicEMU is dashed, DGP.HM is dotted, 
+    and solid gray lines show results of held-out computer model runs.}
     \label{fig:plot_pred_1to6}

### 2025-04-18 paper/main.tex
include some initial CAMB results; add basic cosmology details and references

@@
         Marco A.R. Ferreira\footnotemark[3] \and
+        Jared Clark\footnotemark[3] \and
         Kelly Moran\thanks{Los Alamos National Laboratory} \and
@@
 We are particularly motivated by the Mira-Titan simulation suite \citep{moran2023mira}
-which defines ``cosmologies'' according to eight cosmological parameters (\textcolor{red}{more on this
-in Section TBD}).  For a particular cosmology, the simulation of the power matter spectra 
+which defines ``cosmologies'' according to eight cosmological parameters (more on this
+in Section \ref{sec:data}).  For a particular cosmology, the simulation of the power matter spectra 
 is rather complex.  Although we presume there exists some true matter spectra as a function
@@
 The Mira-Titan simulation suite is a dataset consisting of simulated power spectra for 
-117 different cosmologies. Each of these cosmologies is specified by 8 cosmological 
-parameter values, which we will denote as $\theta \in \mathbb{R}^8$. Different simulation 
+117 different cosmologies. Each of these cosmologies is specified by eight cosmological 
+parameter values: matter density, baryon density, amplitude of density fluctuations, 
+dimensionless Hubble parameter, spectral index of scalar perturbations,
+dark energy equation of state parameters, and neutrino density. For more details 
+on these parameters and their effects, see \cite{dodelson2020modern, aghanim2020planck, heitmann2016mira}. 
+Parameter ranges for each of these parameters are specified in Table 2.1 of \cite{moran2023mira}.
+For simplicity, we will denote these eight parameters as $\theta \in \mathbb{R}^8$. Different simulation 
 output exists for different values of a ninth value, the redshift; in this work, we fix 
@@
 likelihood, and conditional on $\Sigma_\varepsilon$, we specify priors for $\tau_S, \theta_S$ 
-following \cite{sauer2023active}: \textcolor{cyan}{Annie, please ensure the most up-to-date 
+following \cite{sauer2023active}: \textcolor{red}{Annie, please ensure the most up-to-date 
 prior specs are shown above and here. Here is what I found from fit\$settings when using 
@@
 
-\textcolor{cyan}{Annie: want to add something here? For $\theta_w$, we also specify 
+\textcolor{red}{Annie: want to add something here? For $\theta_w$, we also specify 
 a Gamma prior of the form $\theta_w \sim G(\alpha=1.2, \beta=2)$. We discuss the 
@@
 In addition to simulated data, we investigate model fits for different cosmological 
-datasets from CAMB \citep{lewis2011CAMB}. \textcolor{magenta}{We have looked at two sets of model 
-runs, ``LT" runs where there is an underlying ``true" value alongside 15 computer 
-model runs, or the ``CAMB" runs, where we only have a single output. The performance 
-from the CAMB runs was better (with a 6-dimensional design space), and also did not 
-require obtaining a posterior mean from multiple model runs. When looking at the 
-CAMB runs on the $\Delta$ space, we see it generally is withing $\pm 5\%$ error. 
-Regarding these results, compared to the results for the 8-dimensional LT runs 
-which had less impressive predictive performance, I'm open to suggestions on whether 
-to include details on one or the other.}
+datasets from CAMB \citep{lewis2011CAMB}. These datasets contain power spectra based
+on six-dimensional cosmological parameter settings. Similar to the Mira-Titan suite, the
+CAMB dataset has multiple (sixty-four) batches of power spectra for each cosmology, with each batch 
+containing multiple (fifteen) low-resolution spectra, and one high-resolution sprectum. 
+Unlike the Mira-Titan suite, this dataset favorably contains an infinite-resolution 
+spectrum, which can be treated as the true spectrum for each cosmology,
+allowing us to accurately assess posterior mean estimates and credible intervals.
 
@@
     \caption{Example of model fit for the first model from the CAMB data, with LOESS
-             mean term subtracted. \textcolor{red}{Steve: Update this plot to be
+             mean term subtracted. \textcolor{red}{Currently, we can compare the 
+             orange lines (posterior mean and 95\% credible interval) to the blue line,
+             representing the infinite-resolution ``true" spectrum. Steve: Update this plot to be
              more clean, use different line types: dotted, dash, solid, etc.}}   
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=6in]{plot_pca_CAMB.png}
-    \caption{Proportion of $\Delta_{emu}$ vs. $\Delta_{holdout}$ for model 
-             runs 11-42 (excluding 27-28).}   
-    \label{fig:pca_camb}
-\end{figure}
+We use 32 CAMB batches to train the DGP.HM model and reserve the other 32 as a test set.
+To imitate the role of the perturbation theory output of Mira-Titan, we anchor estimates of the CAMB
+power spectrum at the infinite-resolution for $\log_{10}(k) < -1.7$, and then rely 
+strictly on the low- and high-resolution runs for $k$ above this. We take a weighted
+average of the CAMB runs and assume heteroskedastic errors in a similar fashion, now with
+no spatial dependence (compare low-resolution runs of Figure \ref{fig:fit_camb} 
+with Figure \ref{fig:plot_data}).
 
@@
 
-\subsection{GP-PC Model}
+\subsection{Functional Prediction Model}
 \label{subsec:pca}
@@
 
+% This is old: switched to SEPIA
+% We assume a powered exponential kernel for the Gaussian process and employ 
+% \texttt{GPfit} to perform a multi-start gradient based search for GP's hyperparameters 
+% \citep{macdonald2015gpfit}. 
 In alignment with \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use 
-GPs to model the principal components' weights; we shall denote this method GP-PC. 
-We assume a powered exponential kernel for the Gaussian process and employ 
-\texttt{GPfit} to perform a multi-start gradient based search for GP's hyperparameters 
-\citep{macdonald2015gpfit}. That is, the $i$th vector of PC weights $w_i(\theta)$, 
+GPs to model the principal components' weights. For direct comparison alongside
+\cite{moran2023mira}, we employ a Python implementation of this approach called 
+SEPIA \citep{gattiker2020sepia}. \textcolor{red}{Please note: currently show GPfit 
+predictions for Mira-Titan Data, NOT SEPIA. The CAMB results are based on SEPIA.}
+That is, the $i$th vector of PC weights $w_i(\theta)$, 
 $i \in \{1,\ldots,p_\eta\}$, will be modeled as a zero mean GP: 
@@
 
-We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\theta\}$. 
+\textcolor{red}{We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\theta\}$. 
+The above with $R_{ij}$ is based on GPfit; will update this to reflect SEPIA settings later.}
 Using these trained GPs from the PCs, we predict the spectra for unobserved cosmologies.
+
+\subsection{Predicting Spectra for CAMB Data}
+\label{subsec:camb_pred}
+
+From our 32 posterior means using DGP.HM, these are input into SEPIA in order to predict power spectra
+for the 32 test cosmologies. With the infinite-resolution output, we have a sense of the ``best-case" 
+predictions we can have (i.e., if we trained with the 32 true power spectra, as opposed to our 
+estimated posterior means). In Figure \ref{fig:pca_preds_v_camb}, we see comparisons of the heldout 
+infinite resolution alongside two predictions: one based on DGP.HM and the ``best-case" scenario using the
+infinite-resolution spectra of the training set. There is a high level of agreement between the pairs
+of predictions across cosmologies. Upon comparing MSEs for each of these prediction scenarios 
+(Figure \ref{fig:mse_camb}, most are very similar across the two groups (with the two most difficult 
+predictions being the exceptions), suggesting that overall DGP.HM is estimating the true spectra 
+for each batch well.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{preds_diff1_4_8.jpg}
+    \caption{Predictions for 16 of the 32 test cosmologies after centering and scaling. 
+             The interpolated infinite resolution run is shown in black, the prediction 
+             (trained on the infinite resolution) is in blue and the prediction from 
+             the DGP.HM posterior mean is in red.}   
+    \label{fig:pca_preds_v_camb}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=3in]{mse_dot.jpg}
+    \caption{Comparison of MSE for CAMB predictions, when training with DGP.HM posterior 
+             means (left), and training with the ``true" infinite-resolution spectra (right).}   
+    \label{fig:mse_camb}
+\end{figure}
 

### 2025-04-22 paper/main.tex
minor edits in abstract

@@
         Math and Technology, Elms College, {\tt walshst@elms.edu}} \and 
-        Annie S. Booth\thanks{Department of Statistics, North Carolina State University} \and
-        David Higdon\thanks{Department of Statistics, Virginia Tech} \and
-        Marco A.R. Ferreira\footnotemark[3] \and
-        Jared Clark\footnotemark[3] \and
+        Annie S. Booth\thanks{Department of Statistics, Virginia Tech} \and
+        David Higdon\footnotemark[2] \and
+        Marco A.R. Ferreira\footnotemark[2] \and
+        Jared Clark\footnotemark[2] \and
         Kelly Moran\thanks{Los Alamos National Laboratory} \and
@@
 quantification (UQ), from all of the provided curves.  To this end, we propose a 
-novel Bayesian deep Gaussian process (DGP) hierarchical model which synthesize 
-all the information and estimates the underlying matter power spectra, while providing 
-effective UQ.  Our model extends previous work on Bayesian DGPs from scalar responses 
-to functional ones.  Second, we leverage our predicted power spectra from various 
-cosmologies in order to accurately predict the entire matter power spectra for an 
-unobserved cosmology.  For this task, we leverage basis function representations 
+novel Bayesian deep Gaussian process (DGP) hierarchical model which synthesizes 
+all the simulation information to estimate the underlying matter power spectra
+while providing effective UQ.  Our model extends previous work on Bayesian DGPs 
+from scalar responses to functional ones.  Second, we leverage our predicted 
+power spectra from various cosmologies in order to accurately predict the entire 
+matter power spectra for an 
+unobserved cosmology.  For this task, we use basis function representations 
 of the functional spectra to train a separate Gaussian process emulator.  
 Our method performs well in synthetic exercises and against the benchmark cosmological 
-emulator (Cosmic Emu).  We provide reproducible code in a public git repository.
+emulator (Cosmic Emu).
 \end{abstract}

### 2025-04-22 paper/main.tex
Annie through section 1

@@
 
-\noindent \textbf{Keywords:} surrogate, computer experiment, uncertainty quantification,
-principal components analysis, Mira-Titan, Cosmic Emu (???)
+\noindent \textbf{Keywords:} computer experiment, Cosmic Emu, 
+principal components analysis, Mira-Titan, surrogate, uncertainty quantification
 
@@
 in prevalance and complexity \citep[e.g.,][]{lawrence2010coyote,moran2023mira}
-\textcolor{red}{(more cites here?)}.  
+\textcolor{red}{(more cites here? possibly those commented out above?)}.  
 We are particularly motivated by a simulation
@@
 %are two simulation suites dedicated to this effort.  
-From the highest level, our aim is to leverage such simulation data to 
+Our overarching goal is to leverage such simulation data to 
 first estimate and then predict underlying power matter spectra for various 
@@
 
-We are particularly motivated by the Mira-Titan simulation suite \citep{moran2023mira}
-which defines ``cosmologies'' according to eight cosmological parameters (more on this
+The Mira-Titan simulation suite \citep{moran2023mira}
+defines ``cosmologies'' according to eight cosmological parameters (more on this
 in Section \ref{sec:data}).  For a particular cosmology, the simulation of the power matter spectra 
@@
 only over particular wavenumber ranges.  Our first objective is to synthesize expert knowledge
-and all of this simulation information into an effective estimate of the true underlying
-matter spectrum.
+and all available simulation information into an effective estimate of the true underlying
+matter spectrum.  \textcolor{red}{(In this paragraph, we need to narrate Figure 1.  The
+figure provides a useful visual, but it needs to be talked through.  Comment on how
+the weighted average takes into account the appropriate ranges.  Explain why the weighted
+average on its own is not an effective estimate.  Motivate why we need a more complex model.
+This is also a good place to introduce wavenumber as $k$.  Is there a particular reason
+we plot it on the log scale?)}
 
@@
     \centering
-    \includegraphics[width=6in]{plot_data.png}
-    \caption{Left: The perturbation theory, low resolution runs, and high resolution run 
-             for the first training cosmology. The weighted average of these is shown as a dashed line. 
-             Dotted lines at the bottom indicate where each data type is deemed approximately unbiased. 
-             Right: The same image, but restricted to wavenumber ($k$) values where the low resolution 
-             runs are approximately unbiased.}
+    \includegraphics[width=\linewidth, trim=0 13 0 45, clip=TRUE]{plot_data.png}
+    \caption{{\it Left:} The perturbation theory (solid green), low resolution runs (solid grey), 
+    and high resolution run  (solid pink) for a particular cosmology. 
+    The weighted average of these is shown in dashed orange. Dotted lines at the bottom 
+    indicate where each data type is deemed approximately unbiased. 
+    {\it Right:} Same, but restricted to wavenumber ($k$) values where the low resolution 
+    runs are approximately unbiased.
+    \textcolor{red}{(Either make one legend for both graphs or ditch the legend 
+    altogether and narrate the colors/line types in the caption.  If legend stays, "ind" needs
+    to go.  Pink line is too hard to see.  Make gray lines thinner and pink thicker?
+    Since later we will use a centered version, we could add a 3rd panel here that shows the
+    same data centered, so the reader can understand the connection to the plots later.)}}
     \label{fig:plot_data}
@@
 There are several challenges to estimating the underlying spectra.  Our model
-must: handle multiple correlated functional observations, offer enough 
+must: handle multiple correlated functional observations (like the low resolution
+runs shown in Figure \ref{fig:plot_data}), offer enough 
 flexibility to accommodate the nonstationarity present in power matter spectra 
@@
 regarding the wavenumbers over which various outputs are trusted.  To this end, 
-we propose a Bayesian hierarchical model which treats each of functional simulation 
-output as a realization of a deep Gaussian process \citep[DGP;][]{damianou2013deep} 
+we propose a Bayesian hierarchical model which treats functional simulation 
+outputs as realizations of a deep Gaussian process \citep[DGP;][]{damianou2013deep} 
 centered on the true underlying spectra.  The hierarchical structure naturally
@@
 simulation suite, which was trained on 117 simulations, and is termed 
-the ``Cosmic Emu''.\footnote{\url{https://github.com/lanl/CosmicEmu}}
+the ``Cosmic Emu.''\footnote{\url{https://github.com/lanl/CosmicEmu}}
 We leverage the same training data, in conjunction with our Bayesian hierarchical
 DGP model and basis function representations, to train a Gaussian process surrogate 
-on the principal component weights in order to predict spectra for unobserved
+on principal component weights in order to predict spectra for unobserved
 cosmologies \citep{higdon2008computer, higdon2010estcosmo}. 
@@
 
-For a comprehensive assessment of our Bayesian hierarchical DGP model, we 
-further study its performance in a simulation study with synthetic data mimicing 
-the Mira-Titan data. Additionally, we also train our model with data output from 
-the Code for Anisotropies in the Microwave Background \citep[CAMB,][]{lewis2011CAMB};
-this model output exhibits similar behavior to Mira-Titan as well, but has the benefit 
-of including a ``true" power spectrum corresponding toeach cosmology.
+For a comprehensive assessment of our Bayesian hierarchical DGP, we benchmark
+its performance in two additional settings: with synthetic data mimicing 
+the Mira-Titan data and with real-world data from 
+the Code for Anisotropies in the Microwave Background \citep[CAMB,][]{lewis2011CAMB}.
+The CAMB model also exhibits similar behavior to Mira-Titan, but crucially
+provides a ``true" power spectrum corresponding to each cosmology.
 
@@
 introduces our motivating application, the Mira-Titan simulation suite.  
-Section \ref{sec:hm_fit} describes our hierarchical Bayesian DGP model and 
+Section \ref{sec:hm_fit} describes our hierarchical Bayesian DGP and 
 validates its performance on synthetic exercises and the CAMB simulations.  
@@
 Section \ref{sec:disc} concludes with a discussion of our contributions and avenues for 
-future research.  We provide reproducible code and an R package for our Bayesian DGP model
-in a public git repository.\footnote{\url{https://github.com/stevewalsh124/dgp.hm}}
+future research.  We provide reproducible code and an {\sf R} package for our 
+Bayesian DGP model in a public git repository.\footnote{\url{https://github.com/stevewalsh124/dgp.hm}}
 \textcolor{red}{Please note: this repository is currently private until we 

### 2025-04-23 paper/main.tex
move figure 1?
working on section 2

@@
 \usepackage{graphicx}
+\usepackage{xcolor} % only for coloring text while revising
 
@@
 and all available simulation information into an effective estimate of the true underlying
-matter spectrum.  \textcolor{red}{(In this paragraph, we need to narrate Figure 1.  The
-figure provides a useful visual, but it needs to be talked through.  Comment on how
-the weighted average takes into account the appropriate ranges.  Explain why the weighted
-average on its own is not an effective estimate.  Motivate why we need a more complex model.
-This is also a good place to introduce wavenumber as $k$.  Is there a particular reason
-we plot it on the log scale?)}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\linewidth, trim=0 13 0 45, clip=TRUE]{plot_data.png}
-    \caption{{\it Left:} The perturbation theory (solid green), low resolution runs (solid grey), 
-    and high resolution run  (solid pink) for a particular cosmology. 
-    The weighted average of these is shown in dashed orange. Dotted lines at the bottom 
-    indicate where each data type is deemed approximately unbiased. 
-    {\it Right:} Same, but restricted to wavenumber ($k$) values where the low resolution 
-    runs are approximately unbiased.
-    \textcolor{red}{(Either make one legend for both graphs or ditch the legend 
-    altogether and narrate the colors/line types in the caption.  If legend stays, "ind" needs
-    to go.  Pink line is too hard to see.  Make gray lines thinner and pink thicker?
-    Since later we will use a centered version, we could add a 3rd panel here that shows the
-    same data centered, so the reader can understand the connection to the plots later.)}}
-    \label{fig:plot_data}
-\end{figure}
+matter spectrum.
 
@@
 Bayesian DGP model in a public git repository.\footnote{\url{https://github.com/stevewalsh124/dgp.hm}}
-\textcolor{red}{Please note: this repository is currently private until we 
+\textcolor{cyan}{Please note: this repository is currently private until we 
 hear back regarding permissions for sharing the Mira-Titan data alongside our code.}
@@
 
-\section{Data}
+\section{Mira-Titan}
 \label{sec:data}
@@
 117 different cosmologies. Each of these cosmologies is specified by eight cosmological 
-parameter values: matter density, baryon density, amplitude of density fluctuations, 
+parameters: matter density, baryon density, amplitude of density fluctuations, 
 dimensionless Hubble parameter, spectral index of scalar perturbations,
@@
 on these parameters and their effects, see \cite{dodelson2020modern, aghanim2020planck, heitmann2016mira}. 
-Parameter ranges for each of these parameters are specified in Table 2.1 of \cite{moran2023mira}.
-For simplicity, we will denote these eight parameters as $\theta \in \mathbb{R}^8$. Different simulation 
-output exists for different values of a ninth value, the redshift; in this work, we fix 
-this value at 0. For each cosmology, we have a batch of 18 simulated spectra: an inexpensive 
+Parameter ranges are specified in Table 2.1 of \cite{moran2023mira}.
+For simplicity, we will denote these eight parameters as $\theta \in \mathbb{R}^8$. 
+There is an optional ninth parameter, the redshift, which we fix at 0.
+%Different simulation output exists for different values of a ninth value, 
+%the redshift; in this work, we fix 
+%this value at 0. 
+For each cosmology, Mira-Titan contains a batch of 18 simulated spectra: an inexpensive 
 spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from 
-lower-resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from 
+lower resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from 
 a high resolution simulation ($y_h$).
 
-These spectra are represented as a function of wavenumber $k$, and hence is denoted $P(k)$. 
-Following \cite{moran2023mira}, we model on the emulation space 
-$\mathcal{P}(k)=\log_{10}\left(\frac{k^{1.5}P(k)}{2\pi^2}\right)$. The output is available 
-for $0.001 \leq k \leq 5$ for $n=351$ different values of $k$, but each data type has differing 
-values where it is deemed approximately unbiased. For $k<0.04$, only $y_p$ provides a reliable 
-estimate of the spectra. $y_{\ell_r}$ are valid for $0.04 \leq k \leq 0.25$, and $y_h$ is valid 
-for $0.04 \leq k \leq 5$. Figure \ref{fig:plot_data} shows an example of the output for a 
-particular cosmology.
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\linewidth, trim=0 13 0 45, clip=TRUE]{plot_data.png}
+    \caption{{\it Left:} The perturbation theory (solid green), low resolution runs (solid grey), 
+    and high resolution run  (solid pink) for a particular cosmology. 
+    The weighted average of these is shown in dashed orange. Dotted lines at the bottom 
+    indicate where each data type is deemed approximately unbiased. 
+    {\it Right:} Same, but restricted to wavenumber ($k$) values where the low resolution 
+    runs are approximately unbiased.
+    \textcolor{red}{(Either make one legend for both graphs or ditch the legend 
+    altogether and narrate the colors/line types in the caption.  If legend stays, "ind" needs
+    to go.  Pink line is too hard to see.  Make gray lines thinner and pink thicker?
+    Since later we will use a centered version, we could add a 3rd panel here that shows the
+    same data centered, so the reader can understand the connection to the plots later.)}}
+    \label{fig:plot_data}
+\end{figure}
+
+These spectra are represented as a function of wavenumber $k$.  Following 
+\cite{moran2023mira}, we model on the emulation space 
+$\mathcal{P}(k)=\log_{10}\left(\frac{k^{1.5}P(k)}{2\pi^2}\right)$, where $P(k)$ represents the
+original spectra. Output is available for $n=351$ values of $k$, spanning
+$0.001 \leq k \leq 5$, but each data type has unique wavenumber ranges
+where it is deemed approximately unbiased. For $k<0.04$, only $y_p$ provides a reliable 
+estimate of the true spectra. $y_{\ell_r}$ are valid for $0.04 \leq k \leq 0.25$, and 
+$y_h$ is valid for $0.04 \leq k \leq 5$.  Figure \ref{fig:plot_data} shows an example 
+of the output for a particular cosmology. \textcolor{red}{(Add some narration here?  Something
+that doesn't repeat the caption but complements it.)}
 
@@
 wavenumber values ($p_1,\dots,p_n$) using a log-log regression model, and a multiplier 
-$c$ for the increase in precision from the low- to high- resolution output. Given 
-that the perturbation theory spectra is non-stochastic, we treat this as having a 
+$c$ for the increase in precision from the low- to high- resolution output. 
+\textcolor{red}{(I'm missing a bit of context here - why did we need the precision
+estimates in the first place?  Can we show a graph of $p_1,\dots, p_n$ to illustrate
+how these change over $k$ values?)}
+Given that the perturbation theory spectra is non-stochastic, we treat this as having a 
 precision of $10^5$. We leverage these precision estimates within our model at each 
@@
 $\Lambda_h = \text{diag}\left(c(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$. 
+\textcolor{red}{(Unless we are really short on space, could we show these in
+actual "bmatrix" form?  Or visually somehow?)}
 Here, each precision matrix is diagonal and the diagonal elements consist of the 
@@
 where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for 
-the weighted average for a particular cosmology.
+the weighted average for a particular cosmology. \textcolor{red}{(Revisit the figure and
+comment on the weighted average.  Explain why the weighted
+average on its own is not an effective estimate.  Motivate why we need a more complex model.)}
 
-We will handle the data in two stages.  First, we will focus on 111 training cosmologies 
-and obtain the best estimated spectrum for each.  Second, we will use these estimates 
-to predict the spectra for six held-out cosmologies.
+We consider a training-testing split with 111 cosmologies used for estimation and 6 
+cosmologies held out for testing.  Our ultimate objective is to effectively predict
+the underlying power matter spectra for the held-out cosmologies based on their 
+8-dimensional input configuration (Section \ref{sec:pred}).  But first, we must 
+obtain effective estimates of the underlying spectrum for each training 
+cosmology (Section \ref{sec:hm_fit}).
 

### 2025-04-25 paper/main.tex
update plots, include suggested edits

@@
 Experiments which simulate the expansion of the universe are growing
-in prevalance and complexity \citep[e.g.,][]{lawrence2010coyote,moran2023mira}
-\textcolor{red}{(more cites here? possibly those commented out above?)}.  
+in prevalance and complexity \citep[e.g.,][]{lawrence2010coyote,derose2019aemulus,
+nishimichi2019dark,angulo2021bacco,euclid2021euclid,moran2023mira}.  
 We are particularly motivated by a simulation
@@
 Due to the computationally expensive nature of the Mira-Titan simulation suite
-\textcolor{red}{(do we know how long it takes to run?)},
+(one batch of simulations can take multiple weeks to run on a supercomputer),
 it is infeasible to evaluate the simulation for every possible eight-dimensional
@@
     indicate where each data type is deemed approximately unbiased. 
-    {\it Right:} Same, but restricted to wavenumber ($k$) values where the low resolution 
+    {\it Middle:} Same as left, but restricted to wavenumber ($k$) values where the low resolution 
     runs are approximately unbiased.
-    \textcolor{red}{(Either make one legend for both graphs or ditch the legend 
-    altogether and narrate the colors/line types in the caption.  If legend stays, "ind" needs
-    to go.  Pink line is too hard to see.  Make gray lines thinner and pink thicker?
-    Since later we will use a centered version, we could add a 3rd panel here that shows the
-    same data centered, so the reader can understand the connection to the plots later.)}}
+    {\it Right:} Same as left, but with estimated posterior mean subtracted.
+    \textcolor{red}{Rerun on Windows for $\mathcal{P}$}}
     \label{fig:plot_data}
@@
 $y_h$ is valid for $0.04 \leq k \leq 5$.  Figure \ref{fig:plot_data} shows an example 
-of the output for a particular cosmology. \textcolor{red}{(Add some narration here?  Something
-that doesn't repeat the caption but complements it.)}
+of the output for a particular cosmology. When we focus on $k$ values where $y_{\ell_r}$ 
+is valid, we see each low-resolution and the high-resolution run oscillate around 
+a mean term, which is relatively smooth, compared to these realizations. When such a 
+mean term is subtracted, we will obtain a plot like the right panel in 
+Figure \ref{fig:plot_data}.
 
@@
 $c$ for the increase in precision from the low- to high- resolution output. 
-\textcolor{red}{(I'm missing a bit of context here - why did we need the precision
-estimates in the first place?  Can we show a graph of $p_1,\dots, p_n$ to illustrate
-how these change over $k$ values?)}
+We can easily incorporate these estimates in our model to account for nonstationarity 
+(i.e., non-constant variance) of the runs. We see as $k$ increases, the runs' precision 
+increases in the middle panel of Figure \ref{fig:plot_data}). 
 Given that the perturbation theory spectra is non-stochastic, we treat this as having a 
 precision of $10^5$. We leverage these precision estimates within our model at each 
-data type's unbiased wavenumber values to define 
-$\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, 
-$\Lambda_{\ell} = \text{diag}\left(16(p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, 
-$\Lambda_h = \text{diag}\left(c(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$. 
-\textcolor{red}{(Unless we are really short on space, could we show these in
-actual "bmatrix" form?  Or visually somehow?)}
-Here, each precision matrix is diagonal and the diagonal elements consist of the 
-precision values or zero, depending on whether the $k$ value is approximately unbiased 
-for that data type. With this, we calculate a weighted average for each cosmology: 
+data type's unbiased wavenumber values to define the following diagonal precision matrix:
+% $\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, 
+% $\Lambda_{\ell} = \text{diag}\left(16\times(p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, 
+% $\Lambda_h = \text{diag}\left(c\times(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$.
+
+\[
+\Lambda = 
+\begin{bmatrix}
+\Lambda_p & 0 & 0 \\
+0 & \Lambda_\ell & 0 \\
+0 & 0 & \Lambda_h
+\end{bmatrix}
+=
+\begin{bmatrix}
+\operatorname{diag}(10^5, \ldots, 10^5) &  &  \\
+ & \operatorname{diag}(16\times(p_1, \ldots, p_n)) &  \\
+ &  & \operatorname{diag}(c\times(p_1, \ldots, p_n))
+\end{bmatrix}
+\]
+
+\[
+\text{where } 
+\begin{cases}
+\Lambda_p \text{ for } k < 0.04 \\
+\Lambda_\ell \text{ for } 0.04 \leq k < 0.25 \\
+\Lambda_h \text{ for } 0.25 \leq k \leq 5
+\end{cases}
+\]
+
+Above, each matrix element is either the corresponding precision value or zero, 
+depending on whether the $k$ value is approximately unbiased for that data type. 
+With this, we calculate a weighted average for each cosmology: 
 $\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, 
 where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for 
-the weighted average for a particular cosmology. \textcolor{red}{(Revisit the figure and
-comment on the weighted average.  Explain why the weighted
-average on its own is not an effective estimate.  Motivate why we need a more complex model.)}
+the weighted average for a particular cosmology. On its own, the weighted average
+has too much variation baked in from the limited number of model runs. There is also
+no spatial dependence accounting for the correlated errors seen in Figure \ref{fig:plot_data}.
+To account for these modeling challenges, we need a more complex model.
 
@@
     \centering
-    \includegraphics[width=6in]{CAMB_fit_model1.jpg}
-    \caption{Example of model fit for the first model from the CAMB data, with LOESS
-             mean term subtracted. \textcolor{red}{Currently, we can compare the 
-             orange lines (posterior mean and 95\% credible interval) to the blue line,
-             representing the infinite-resolution ``true" spectrum. Steve: Update this plot to be
-             more clean, use different line types: dotted, dash, solid, etc.}}   
+    \includegraphics[width=\textwidth]{CAMB_fit_model1.jpg}
+    \caption{Example of DGP.HM model fit for the first batch of model runs from the CAMB data, 
+             with LOESS mean term subtracted. We can compare the orange lines 
+             (posterior mean and dotted 95\% credible interval) to the light blue line,
+             representing the infinite-resolution ``true" spectrum.}   
     \label{fig:fit_camb}
@@
 strictly on the low- and high-resolution runs for $k$ above this. We take a weighted
-average of the CAMB runs and assume heteroskedastic errors in a similar fashion, now with
-no spatial dependence (compare low-resolution runs of Figure \ref{fig:fit_camb} 
+average of the CAMB runs and assume heteroskedastic errors in a similar fashion. For the CAMB
+data, lengthscale estimates are consistently around $3\times 10^{-7}$, evidence that we
+can simplify our covariance structure to contain no spatial dependence 
+(compare low-resolution runs of Figure \ref{fig:fit_camb} 
 with Figure \ref{fig:plot_data}).

### 2025-04-25 fitting/run_CAMB.R

@@ -17,10 +17,6 @@ library(zoo) # rollmean
 library(Matrix) # bdiag
 library(pracma) #interp1
 
-PDF = T
-if(PDF) pdf("results/run_CAMB.pdf")
-par(mfrow=c(2,1))
-
 # keep track of coverages for CAMB based on DGP fit
 coverages = c()
 cov_mat = list()
@@ -50,7 +46,7 @@ for(model in 1:32){ # integer 1-32
   # upper bound to cut off x values (orig scale, not log10)
   x_ub = 0.2
   
-  # read in and plot all n_lo=15 low-res (L1300) runs (adjusted by h and h^3)
+  # read in all n_lo=15 low-res (L1300) runs (adjusted by h and h^3)
   n_lo = 15
   lr = read.table(paste0("../CosmicEmu_etc/pows/M0",if(model<10){"0"},model,
                          "/L1300/PM001/output/m0",
@@ -59,18 +55,12 @@ for(model in 1:32){ # integer 1-32
   y_lo = matrix(NA, length(x_lo), n_lo)
   y_lo[,1] = scrP(lr[which(h*lr[,1]<=x_ub),2]/h^3/g^2, 10^x_lo)
   
-  plot(x_lo, y_lo[,1],
-       col="gray", type="l", main = model, xlim=c(-3,0), ylim=c(-2,2),
-       ylab = "script P (adj by h,g)",
-       xlab = "log10(k) (adj by h)")
-  
   for (i in 2:n_lo) {
     if(model==8 & i==9) next
     lr = read.table(paste0("../CosmicEmu_etc/pows/M0",if(model<10){"0"},model,
                            "/L1300/PM0",if(i<10){"0"},i,"/output/m0",
                            if(model<10){"0"},model,".pk.ini"))
     y_lo[,i] = scrP(lr[which(h*lr[,1]<=x_ub),2]/h^3/g^2, 10^x_lo)
-    lines(x_lo, y_lo[,i], col="gray")
   }
   
   # only model 8 is missing the 9th run
@@ -79,21 +69,18 @@ for(model in 1:32){ # integer 1-32
     n_lo = 14
   }
   
-  # read in and plot the high-res (L2100) run (adjusted by h and h^3)
+  # read in the high-res (L2100) run (adjusted by h and h^3)
   hr = read.table(paste0("../CosmicEmu_etc/pows/M0",if(model<10){"0"},model,
[…]

### 2025-04-25 plotting/plot_data_CAMB.R

@@ -1,0 +1,63 @@
+
+###############################################################################
+# This script fits the Bayesian hierarchical model for a particular CAMB model
+# and writes the results to a csv file.
+#
+# Command line arguments:
+#    model: number of particular model (1-32)
+#    deep: indicator for whether to fit a DGP or GP (1 = deep, 0 = not deep)
+# 
+# Outputs:
+#    writes csv file of predicted values to the results folder
+#
+###############################################################################
+
+library(dgp.hm)
+library(zoo) # rollmean
+library(Matrix) # bdiag
+library(pracma) #interp1
+
+load(file = "../fitting/results/CAMB/fit_1camb_image.rda")
+
+# # Plot the first batch of model runs for CAMB
+# plot(x_lo, y_lo[,1],
+#      col="gray", type="l", main = model, xlim=c(-3,0), ylim=c(-2,2),
+#      ylab = "script P (adj by h,g)",
+#      xlab = "log10(k) (adj by h)")
+# for (i in 2:n_lo) {
+#   if(model==8 & i==9) next
+#   lr = read.table(paste0("../CosmicEmu_etc/pows/M0",if(model<10){"0"},model,
+#                          "/L1300/PM0",if(i<10){"0"},i,"/output/m0",
+#                          if(model<10){"0"},model,".pk.ini"))
+#   y_lo[,i] = scrP(lr[which(h*lr[,1]<=x_ub),2]/h^3/g^2, 10^x_lo)
+#   lines(x_lo, y_lo[,i], col="gray")
+# }
+# lines(x_hi, y_hi, col="red")#, lwd=2)
+# lines(x_camb, y_camb, col="blue1",type="l", lwd=3, lty=3)
+# legend("bottomright", c("camb","lr","hr"), lty=1, col=c("blue","grey","red"))
+
+
+# color-blind friendly color palette
+cbcols <- palette.colors(palette = "Okabe-Ito")
+#1 black, 2 orange, 3 skyblue , 4 bluishgreen, 5 yellow 
+#6 blue, 7 vermillion,  8 reddishpurple, 9 gray 
+
+# Plot showing dgp.hm fit compared with camb
+# Plot with all of [-2.5, -2.2] camb
+# CAMB_fit_model1
[…]

### 2025-04-25 plotting/plot_data_fit.R

@@ -64,35 +64,44 @@ cbcols <- palette.colors(palette = "Okabe-Ito")
 #1 black, 2 orange, 3 skyblue , 4 bluishgreen, 5 yellow 
 #6 blue, 7 vermillion,  8 reddishpurple, 9 gray 
 
-par(mfrow=c(1,2))
+par(mfrow=c(1,3))
 matplot(log10(k), y_lo, col = cbcols[9], 
-        type = "l", lty = 1, ylim = range(y_avg),
+        type = "l", lty = 1, ylim = range(y_avg), lwd=0.5,
         xlab=expression(log[10](k)), 
         ylab=expression(paste("\U1D4AB","(k)")))
 lines(log10(k), y_pt, col = cbcols[4])
-lines(log10(k), y_hi, col = cbcols[8])
+lines(log10(k), y_hi, col = cbcols[8], lwd=2)
 lines(log10(k)[index_list$pert.ix], rep(-1, length(index_list$pert.ix)), 
       col = cbcols[4], lty = 3, lwd=3)
 lines(log10(k)[index_list$lowres.ix], rep(-1.1, length(index_list$lowres.ix)), 
       col = cbcols[9], lty = 3, lwd=3)
 lines(log10(k)[index_list$highres.ix], rep(-1.2, length(index_list$highres.ix)), 
       col = cbcols[8], lty = 3, lwd=3)
-lines(log10(k), y_avg, col = cbcols[2], lty = 2, lwd = 4)
-legend(x = "right", legend = c("pert","low res","hi res", "wt avg","ind"),
-       col = cbcols[c(4,9,8,2,1)], lty = c(1,1,1,2,3), 
-       lwd = c(1,1,1,4,3))
+lines(log10(k), y_avg, col = cbcols[2], lty = 2, lwd = 3)
 
-# Same plot, but only for the low-resolution indices
+# Same as first plot, but only for the low-resolution indices
 matplot(log10(k)[index_list$lowres.ix], y_lo[index_list$lowres.ix,], col = cbcols[9], 
-        type = "l", lty = 1, ylim = range(y_avg[index_list$lowres.ix]),
+        type = "l", lty = 1, ylim = range(y_avg[index_list$lowres.ix])+c(-.05,.05), lwd=0.5,
         xlab=expression(log[10](k)), 
         ylab=expression(paste("\U1D4AB","(k)")))
[…]

### 2025-04-29 paper/main.tex
more revisions

@@
 \usepackage{graphicx}
+\usepackage{bbm} % for indicator function
 \usepackage{xcolor} % only for coloring text while revising
+
+\usepackage{tikz} % for pretty diagonal matrix
+\usetikzlibrary{matrix,decorations.pathreplacing,calc}
 
@@
 of wavenumber, we are not able to directly observe it.  Rather, the simulation outputs 
-a convoluted patchwork of potential spectra, which vary in fidelity and which are trusted
+a convoluted patchwork of potential spectra, which vary in fidelity and are trusted
 only over particular wavenumber ranges.  Our first objective is to synthesize expert knowledge
@@
 There are several challenges to estimating the underlying spectra.  Our model
-must: handle multiple correlated functional observations (like the low resolution
-runs shown in Figure \ref{fig:plot_data}), offer enough 
+must: handle multiple correlated functional observations, offer enough 
 flexibility to accommodate the nonstationarity present in power matter spectra 
@@
 they have yet to be extended to correlated functional outputs. 
-We will demonstrate profiency on synthetic examples and a low fidelity
-cosmological simulation, before deploying our model on the motivating 
-Mira-Titan simulation.
+To demonstrate the profiency of our Bayesian hierarchical DGP, 
+we benchmark its performance in two unique settings: with synthetic data 
+mimicing the Mira-Titan data and with real-world data from 
+the Code for Anisotropies in the Microwave Background \citep[CAMB;][]{lewis2011CAMB}.
+The CAMB model also exhibits similar behavior to Mira-Titan, but crucially
+provides a ``true" power spectrum corresponding to each cosmology.
 
@@
 Our method compares favorably to Cosmic Emu on held-out cosmologies.
-
-For a comprehensive assessment of our Bayesian hierarchical DGP, we benchmark
-its performance in two additional settings: with synthetic data mimicing 
-the Mira-Titan data and with real-world data from 
-the Code for Anisotropies in the Microwave Background \citep[CAMB,][]{lewis2011CAMB}.
-The CAMB model also exhibits similar behavior to Mira-Titan, but crucially
-provides a ``true" power spectrum corresponding to each cosmology.
 
@@
     {\it Right:} Same as left, but with estimated posterior mean subtracted.
-    \textcolor{red}{Rerun on Windows for $\mathcal{P}$}}
+    \textcolor{red}{Rerun on Windows for $\mathcal{P}$.  Make axis/labels and legend 
+    a bit larger so they are easier to read.  Am I tripping or are the horizontal lines
+    in the left panel not lined up correctly?}}
     \label{fig:plot_data}
@@
 of the output for a particular cosmology. When we focus on $k$ values where $y_{\ell_r}$ 
-is valid, we see each low-resolution and the high-resolution run oscillate around 
-a mean term, which is relatively smooth, compared to these realizations. When such a 
+is valid (center panel), we see each low-resolution and the high-resolution run oscillate around 
+a mean term, which is relatively smooth, compared to these realizations. When this
 mean term is subtracted, we will obtain a plot like the right panel in 
-Figure \ref{fig:plot_data}.
+Figure \ref{fig:plot_data} \textcolor{red}{(This plot goes back to the original range,
+would it be prettier and easier for the reader to follow if we kept it to the 
+range of the middle panel here?)}
 
-From the multiple low- and high-resolution spectra across all cosmologies, 
-\cite{moran2023mira} obtained precision estimates of these spectra across the $n$ 
-wavenumber values ($p_1,\dots,p_n$) using a log-log regression model, and a multiplier 
+Notice, in the center panel of Figure \ref{fig:plot_data}, the precision of each
+spectra increases as $k$ increases.  The curves get progressively tighter from left
+to right.  We would like to conveniently wrap this domain-specific knowledge inside
+our Bayesian framework later.  To this end, \cite{moran2023mira} 
+used the multiple low- and high-resolution spectra across all cosmologies to
+obtained precision estimates of these spectra across the $n$ 
+wavenumber values ($p_1,\dots,p_n$) using a log-log regression model and a multiplier 
 $c$ for the increase in precision from the low- to high- resolution output.
-We can easily incorporate these estimates in our model to account for nonstationarity 
-(i.e., non-constant variance) of the runs. We see as $k$ increases, the runs' precision 
-increases in the middle panel of Figure \ref{fig:plot_data}). 
-Given that the perturbation theory spectra is non-stochastic, we treat this as having a 
-precision of $10^5$. We leverage these precision estimates within our model at each 
-data type's unbiased wavenumber values to define the following diagonal precision matrix:
+To streamline notation, we denote three unique diagonal matrices:
+\[
+\Lambda_p = 
+   \begin{tikzpicture}[decoration={brace,amplitude=5pt},baseline=(current bounding box.west)]
+     \matrix (magic) [matrix of math nodes,left delimiter=(,right delimiter=)] {
+      10^5 \\
+      & \ddots \\
+      & & 10^5 \\
+      & & & 0 \\
+      & & & & 0 \\
+      & & & & & \ddots \\
+      & & & & & & 0\\
+     };
+     \draw[decorate] (magic-1-1.north) -- (magic-3-3.north east) node[above=5pt,midway,sloped] {$k < 0.04$};
+   \end{tikzpicture}
+\quad
+\Lambda_l = 
+   \begin{tikzpicture}[decoration={brace,amplitude=5pt},baseline=(current bounding box.west)]
+     \matrix (magic) [matrix of math nodes,left delimiter=(,right delimiter=)] {
+      0 \\
+      & \ddots \\
+      & & 0 \\
+      & & & 10p_1 \\
+      & & & & \ddots \\
+      & & & & & 10p_n ? \\
+      & & & & & & 0 \\
+      & & & & & & & \ddots \\
+      & & & & & & & & 0 \\
+     };
+     \draw[decorate] (magic-4-4.north) -- (magic-6-6.north east) node[above=5pt,midway,sloped] {$0.04\leq k < 0.25$};
+   \end{tikzpicture}
+\]
+\[
+\Lambda_h = 
+   \begin{tikzpicture}[decoration={brace,amplitude=5pt},baseline=(current bounding box.west)]
+     \matrix (magic) [matrix of math nodes,left delimiter=(,right delimiter=)] {
+      0 \\
+      & \ddots \\
+      & & 0 \\
+      & & & cp_1 \\
+      & & & & \ddots \\
+      & & & & & cp_n \\
+     };
+     \draw[decorate] (magic-4-4.north) -- (magic-6-6.north east) node[above=5pt,midway,sloped] {$0.25\leq k \leq 5$};
+   \end{tikzpicture}
+\]
+\textcolor{red}{(@Steve - I am still quite confused here.  Are the above matrices helpful?
+I'm confused by $p_i$.  Isn't there some overlap
+between the low and high res runs?  The piecewise form you had doesn't have any overlap.  You
+had the large $\Lambda$ as the concatenation of the matrices, but then you also had
+it as a sum.  Perhaps we should talk this out sometime.  There has to be an easier 
+way to convey how we calculate the weighted average.  Perhaps we could represent 
+this better with indicator functions?  
+Or as a function of $k_i$?  The equations below are just some ideas.)}
+\[
+\Lambda_p(k) = 10^5 \mathbbm{I}_{\{k < 0.04\}}
+\]
+where $\mathbbm{I}$ is the indicator function.
+\[
+\Lambda(k_i) = \begin{cases}
+    10^5 \text{ for } k_i < 0.04 \text{  (perturbation theory)}\\
+    16p_i \text{ for } 0.04 \leq k_i < 0.25 \text{  (low resolution)}\\
+    cp_i \text{ for } 0.25 \leq k_i \leq 5 \text{  (high resolution)}
+    \end{cases},
+\]
+where we use the sufficiently high precision of $10^5$ to impose the deterministic
+nature of the perturbation theory output.  \textcolor{red}{(Then comment on how 
+we can put this all into one big diagonal matrix that we multiply to our y values?)}
+
 % $\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, 
@@
 
-\[
-\Lambda = 
-\begin{bmatrix}
-\Lambda_p & 0 & 0 \\
-0 & \Lambda_\ell & 0 \\
-0 & 0 & \Lambda_h
-\end{bmatrix}
-=
-\begin{bmatrix}
-\operatorname{diag}(10^5, \ldots, 10^5) &  &  \\
- & \operatorname{diag}(16\times(p_1, \ldots, p_n)) &  \\
- &  & \operatorname{diag}(c\times(p_1, \ldots, p_n))
-\end{bmatrix}
-\]
+%\[
+%\Lambda = 
+%\begin{bmatrix}
+%\Lambda_p & 0 & 0 \\
+%0 & \Lambda_\ell & 0 \\
+%0 & 0 & \Lambda_h
+%\end{bmatrix}
+%=
+%\begin{bmatrix}
+%\operatorname{diag}(10^5, \ldots, 10^5) &  &  \\
+% & \operatorname{diag}(16\times(p_1, \ldots, p_n)) &  \\
+% &  & \operatorname{diag}(c\times(p_1, \ldots, p_n))
+%\end{bmatrix}
+%\]
 
-\[
-\text{where } 
-\begin{cases}
-\Lambda_p \text{ for } k < 0.04 \\
-\Lambda_\ell \text{ for } 0.04 \leq k < 0.25 \\
-\Lambda_h \text{ for } 0.25 \leq k \leq 5
-\end{cases}
-\]
-
+\textcolor{red}{(This part will need some revising, it is still unclear to me
+how the weighted average is calculated.  Although I know in reality it is super simple.}
 Above, each matrix element is either the corresponding precision value or zero, 
@@
 Here we describe our first contribution: the use of a Bayesian hierarchical model 
-to estimate the underlying spectra (and quantify uncertainty) for a particular cosmology.  
-We focus solely on one cosmology, independent of the others.
+to estimate the underlying spectra and quantify uncertainty for a particular cosmology,
+independent of the others.  Let $D_n = \{X_n, \bar{Y}, \Lambda\}$ represent the aggregated 
+data for a particular cosmology.  Let $S$ represent the true (but unknown) underlying
+power matter spectra from which $D_n$ was generated.
+\textcolor{red}{(What is X?  Why is $\bar{Y}$ capital here, but it wasn't earlier?
+Since $\Lambda$ is fixed and doesn't change, I do not think we need to consider it part
+of the data.)}
 
@@
 
-Here, we introduce the Bayesian Gaussian Process (GP) model leveraging the aggregated 
-data $D_n \equiv \{X_n, \bar Y, \Lambda\}$ for a particular cosmology. We assume 
+Here, we introduce the Bayesian Gaussian Process (GP) model We assume 
 the matter power spectrum $S$ for a given cosmology follows a GP with mean $\mu$ 

### 2025-05-02 paper/main.tex
update draft, add script P to images

@@
 
+% after \linewidth:, trim=0 13 0 45, clip=TRUE
 \begin{figure}[ht]
     \centering 
-    \includegraphics[width=\linewidth, trim=0 13 0 45, clip=TRUE]{plot_data.png}
+    \includegraphics[width=\linewidth]{plot_data.png}
     \caption{{\it Left:} The perturbation theory (solid green), low resolution runs (solid grey), 
@@
     runs are approximately unbiased.
-    {\it Right:} Same as left, but with estimated posterior mean subtracted.
-    \textcolor{red}{Rerun on Windows for $\mathcal{P}$.  Make axis/labels and legend 
-    a bit larger so they are easier to read.  Am I tripping or are the horizontal lines
-    in the left panel not lined up correctly?}}
+    {\it Right:} Same as middle, but with estimated posterior mean subtracted.}
     \label{fig:plot_data}
@@
 mean term is subtracted, we will obtain a plot like the right panel in 
-Figure \ref{fig:plot_data} \textcolor{red}{(This plot goes back to the original range,
-would it be prettier and easier for the reader to follow if we kept it to the 
-range of the middle panel here?)}
+Figure \ref{fig:plot_data}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{CAMB_fit_model1.jpg}
+    \includegraphics[width=\textwidth]{CAMB_fit_model1.png}
     \caption{Example of DGP.HM model fit for the first batch of model runs from the CAMB data, 

### 2025-05-02 plotting/plot_data_CAMB.R

@@ -45,10 +45,13 @@ cbcols <- palette.colors(palette = "Okabe-Ito")
 # Plot showing dgp.hm fit compared with camb
 # Plot with all of [-2.5, -2.2] camb
 # CAMB_fit_model1
+png("../paper/CAMB_fit_model1.png", width = 11, height = 4.5, 
+    units = "in", res=300)
+par(mar=c(4.5, 4.5, 1.5, 1.5) + 0.1)
 plot(x, y_cambi - loess_fit$fitted, type="l", col=cbcols[3],
      ylim = range(y_cambi - loess_fit$fitted)+c(-.025,.025),
      xlab=expression(paste(log[10](k),", transformed to [0,1]")), 
-     ylab=expression(paste("\U1D4AB","(k), centered")))
+     ylab=expression(paste("\U1D4AB","  (k), centered")), cex.lab = 1.35)
      # main = paste(model, "cover", coverages[model]))
 abline(h=0, col="black", lty=2)
 for (i in 1:n_lo) lines(x, y_loi[,i] - loess_fit$fitted, col=cbcols[9], lwd=0.4)
@@ -59,5 +62,6 @@ lines(x, fit$ub - loess_fit$fitted, col=cbcols[2], lwd=2, lty=3)
 lines(x, fit$lb - loess_fit$fitted, col=cbcols[2], lwd=2, lty=3)
 lines(x, y_cambi - loess_fit$fitted, col=cbcols[3], lwd=2)
 legend("topright", c("inf-res","low-res","hi-res","wt avg","dgp.hm"), 
-       lty=c(1,1,1,2,1), cex=0.85, bty="n",
+       lty=c(1,1,1,2,1), cex=1, bty="n",
        col=c(cbcols[3],cbcols[9],cbcols[8],cbcols[4],cbcols[2]), lwd=c(2,0.5,1.5,2,2))
+dev.off()

### 2025-05-02 plotting/plot_data_fit.R

@@ -64,11 +64,12 @@ cbcols <- palette.colors(palette = "Okabe-Ito")
 #1 black, 2 orange, 3 skyblue , 4 bluishgreen, 5 yellow 
 #6 blue, 7 vermillion,  8 reddishpurple, 9 gray 
 
-par(mfrow=c(1,3))
+# plot_data.png
+par(mfrow=c(1,3), mar=c( 5.6, 4.6, 4.6, 2.1))
 matplot(log10(k), y_lo, col = cbcols[9], 
         type = "l", lty = 1, ylim = range(y_avg), lwd=0.5,
-        xlab=expression(log[10](k)), 
-        ylab=expression(paste("\U1D4AB","(k)")))
+        xlab=expression(log[10](k)), cex.axis=1.75, cex.lab=1.75,
+        ylab=expression(paste("\U1D4AB","  (k)")))
 lines(log10(k), y_pt, col = cbcols[4])
 lines(log10(k), y_hi, col = cbcols[8], lwd=2)
 lines(log10(k)[index_list$pert.ix], rep(-1, length(index_list$pert.ix)), 
@@ -82,8 +83,9 @@ lines(log10(k), y_avg, col = cbcols[2], lty = 2, lwd = 3)
 # Same as first plot, but only for the low-resolution indices
 matplot(log10(k)[index_list$lowres.ix], y_lo[index_list$lowres.ix,], col = cbcols[9], 
         type = "l", lty = 1, ylim = range(y_avg[index_list$lowres.ix])+c(-.05,.05), lwd=0.5,
-        xlab=expression(log[10](k)), 
-        ylab=expression(paste("\U1D4AB","(k)")))
+        xlab=expression(log[10](k)), cex.axis=1.75, cex.lab=1.75,
+        ylab=expression(paste("\U1D4AB","  (k)")))
+lines(log10(k), y_pt, col = cbcols[4])
 lines(log10(k)[index_list$lowres.ix], y_hi[index_list$lowres.ix], col = cbcols[8], lwd=2)
 lines(log10(k)[index_list$lowres.ix], y_avg[index_list$lowres.ix], 
       col = cbcols[2], lty = 2, lwd = 3)
@@ -94,10 +96,10 @@ legend(x = "topleft", legend = c("pert","low res","hi res", "wt avg"),
 )
 
 # same as first plot, but posterior mean subtracted
[…]

### 2025-05-06 paper/main.tex
annie done in section 2

@@
 Notice, in the center panel of Figure \ref{fig:plot_data}, the precision of each
-spectra increases as $k$ increases.  The curves get progressively tighter from left
-to right.  We would like to conveniently wrap this domain-specific knowledge inside
-our Bayesian framework later.  To this end, \cite{moran2023mira} 
+low/high resolution spectra increases as $k$ increases.  The curves get progressively 
+tighter from left to right.  We would like to conveniently wrap this domain-specific 
+knowledge inside our Bayesian framework later.  To this end, \cite{moran2023mira} 
 used the multiple low- and high-resolution spectra across all cosmologies to
-obtained precision estimates of these spectra across the $n$ 
-wavenumber values ($p_1,\dots,p_n$) using a log-log regression model and a multiplier 
-$c$ for the increase in precision from the low- to high- resolution output.
-To streamline notation, we denote three unique diagonal matrices:
+obtained precision estimates, $p_1,\dots, p_n$, across the $n$ 
+wavenumber values using a log-log regression model.
+% and a multiplier $c$ for the increase in precision from the 
+% low- to high- resolution output.
+We use these estimates to define three diagonal precision matrices, 
+one for each data type (perturbation theory, low resolution, and high resolution).
+Specifically, denote $\Lambda_p$, $\Lambda_\ell$, and $\Lambda_h$ as the
+diagonal matrices with elements
 \[
-\Lambda_p = 
-   \begin{tikzpicture}[decoration={brace,amplitude=5pt},baseline=(current bounding box.west)]
-     \matrix (magic) [matrix of math nodes,left delimiter=(,right delimiter=)] {
-      10^5 \\
-      & \ddots \\
-      & & 10^5 \\
-      & & & 0 \\
-      & & & & 0 \\
-      & & & & & \ddots \\
-      & & & & & & 0\\
-     };
-     \draw[decorate] (magic-1-1.north) -- (magic-3-3.north east) node[above=5pt,midway,sloped] {$k < 0.04$};
-   \end{tikzpicture}
+\Lambda_p^{(ii)} = \begin{cases}
+    10^5 &\text{for } k_i < 0.04 \\
+    0  &\text{otherwise}\\
+    \end{cases}
 \quad
-\Lambda_l = 
-   \begin{tikzpicture}[decoration={brace,amplitude=5pt},baseline=(current bounding box.west)]
-     \matrix (magic) [matrix of math nodes,left delimiter=(,right delimiter=)] {
-      0 \\
-      & \ddots \\
-      & & 0 \\
-      & & & 10p_1 \\
-      & & & & \ddots \\
-      & & & & & 10p_n ? \\
-      & & & & & & 0 \\
-      & & & & & & & \ddots \\
-      & & & & & & & & 0 \\
-     };
-     \draw[decorate] (magic-4-4.north) -- (magic-6-6.north east) node[above=5pt,midway,sloped] {$0.04\leq k < 0.25$};
-   \end{tikzpicture}
+\Lambda_\ell^{(ii)} = \begin{cases}
+    16p_i &\text{for } 0.04 \leq k_i < 0.25 \\
+    0  &\text{otherwise}\\
+    \end{cases}
+\quad
+\Lambda_h^{(ii)} = \begin{cases}
+    cp_i &\text{for } 0.04 \leq k_i < 5 \\
+    0  &\text{otherwise}\\
+    \end{cases}
 \]
-\[
-\Lambda_h = 
-   \begin{tikzpicture}[decoration={brace,amplitude=5pt},baseline=(current bounding box.west)]
-     \matrix (magic) [matrix of math nodes,left delimiter=(,right delimiter=)] {
-      0 \\
-      & \ddots \\
-      & & 0 \\
-      & & & cp_1 \\
-      & & & & \ddots \\
-      & & & & & cp_n \\
-     };
-     \draw[decorate] (magic-4-4.north) -- (magic-6-6.north east) node[above=5pt,midway,sloped] {$0.25\leq k \leq 5$};
-   \end{tikzpicture}
-\]
-\textcolor{red}{(@Steve - I am still quite confused here.  Are the above matrices helpful?
-I'm confused by $p_i$.  Isn't there some overlap
-between the low and high res runs?  The piecewise form you had doesn't have any overlap.  You
-had the large $\Lambda$ as the concatenation of the matrices, but then you also had
-it as a sum.  Perhaps we should talk this out sometime.  There has to be an easier 
-way to convey how we calculate the weighted average.  Perhaps we could represent 
-this better with indicator functions?  
-Or as a function of $k_i$?  The equations below are just some ideas.)}
-\[
-\Lambda_p(k) = 10^5 \mathbbm{I}_{\{k < 0.04\}}
-\]
-where $\mathbbm{I}$ is the indicator function.
-\[
-\Lambda(k_i) = \begin{cases}
-    10^5 \text{ for } k_i < 0.04 \text{  (perturbation theory)}\\
-    16p_i \text{ for } 0.04 \leq k_i < 0.25 \text{  (low resolution)}\\
-    cp_i \text{ for } 0.25 \leq k_i \leq 5 \text{  (high resolution)}
-    \end{cases},
-\]
-where we use the sufficiently high precision of $10^5$ to impose the deterministic
-nature of the perturbation theory output.  \textcolor{red}{(Then comment on how 
-we can put this all into one big diagonal matrix that we multiply to our y values?)}
-
-% $\Lambda_p = \text{diag}\left((10^5)_{k < 0.04}\right)$, 
-% $\Lambda_{\ell} = \text{diag}\left(16\times(p_1,\dots,p_n)_{0.04 \leq k < 0.25}\right)$, 
-% $\Lambda_h = \text{diag}\left(c\times(p_1,\dots,p_n)_{0.04 \leq k \leq 5}\right)$.
-
-%\[
-%\Lambda = 
-%\begin{bmatrix}
-%\Lambda_p & 0 & 0 \\
-%0 & \Lambda_\ell & 0 \\
-%0 & 0 & \Lambda_h
-%\end{bmatrix}
-%=
-%\begin{bmatrix}
-%\operatorname{diag}(10^5, \ldots, 10^5) &  &  \\
-% & \operatorname{diag}(16\times(p_1, \ldots, p_n)) &  \\
-% &  & \operatorname{diag}(c\times(p_1, \ldots, p_n))
-%\end{bmatrix}
-%\]
-
-\textcolor{red}{(This part will need some revising, it is still unclear to me
-how the weighted average is calculated.  Although I know in reality it is super simple.}
-Above, each matrix element is either the corresponding precision value or zero, 
-depending on whether the $k$ value is approximately unbiased for that data type. 
-With this, we calculate a weighted average for each cosmology: 
+for $i=1,\dots, n$.  In $\Lambda_p$, we use the sufficiently high precision 
+of $10^5$ to impose the deterministic nature of the perturbation theory output.  
+\textcolor{red}{(@Steve - what is $c$
+exactly?)}  We then calculate a weighted average for each cosmology: 
 $\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, 
-where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ is the precision matrix for 
-the weighted average for a particular cosmology. On its own, the weighted average
-has too much variation baked in from the limited number of model runs. There is also
-no spatial dependence accounting for the correlated errors seen in Figure \ref{fig:plot_data}.
+where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ and 
+$\bar{y}_\ell = \frac{1}{16}\sum_{r=1}^{16} y_{\ell_r}$. In Figure \ref{fig:plot_data},
+the weighted average is shown in dashed orange.  As an estimate of the true
+underlying spectra, the weighted average fluctuates too drastically due 
+to the limited number of model runs.  It fails to account for the
+spatial dependence inherent in the smooth but stochastic spectra.
 To account for these modeling challenges, we need a more complex model.
@@
 obtain effective estimates of the underlying spectrum for each training 
-cosmology (Section \ref{sec:hm_fit}).
+cosmology (Section \ref{sec:hm_fit}). \textcolor{red}{(@Steve - maybe this is a
+good place to introduce the ``hold out'' and ``within sample'' terms?)}
 

### 2025-05-09 paper/main.tex
updates to section 2-3

@@
 used the multiple low- and high-resolution spectra across all cosmologies to
-obtained precision estimates, $p_1,\dots, p_n$, across the $n$ 
-wavenumber values using a log-log regression model.
-% and a multiplier $c$ for the increase in precision from the 
-% low- to high- resolution output.
-We use these estimates to define three diagonal precision matrices, 
-one for each data type (perturbation theory, low resolution, and high resolution).
-Specifically, denote $\Lambda_p$, $\Lambda_\ell$, and $\Lambda_h$ as the
-diagonal matrices with elements
+obtain precision estimates, $p_1,\dots, p_n$, across the $n$ 
+wavenumber values using a log-log regression model. A multiplier $c$ for 
+the increase in precision from the low- to high- resolution output is also 
+estimated with this regression model. We use these estimates to define three 
+diagonal precision matrices, one for each data type (perturbation theory, 
+low resolution average, and high resolution). Specifically, denote $\Lambda_p$, 
+$\Lambda_\ell$, and $\Lambda_h$ as the diagonal matrices with elements
 \[
@@
 of $10^5$ to impose the deterministic nature of the perturbation theory output.  
-\textcolor{red}{(@Steve - what is $c$
-exactly?)}  We then calculate a weighted average for each cosmology: 
+We then calculate a weighted average for each cosmology: 
 $\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, 
@@
 to estimate the underlying spectra and quantify uncertainty for a particular cosmology,
-independent of the others.  Let $D_n = \{X_n, \bar{Y}, \Lambda\}$ represent the aggregated 
+independent of the others. In the previous seciton, we used lowercase $\bar{y}$ to denote 
+the observed weighted average; here, we use $\bar{Y}$ to indicate the corresponding 
+random variable within the statistical model. Specifically, our input $X$ is the 
+$n$ values of $\log_{10}(k)$ and the output is the weighted average $\bar{Y}$. 
+Let $D = \{X, \bar{Y}, \Lambda\}$ represent the aggregated 
 data for a particular cosmology.  Let $S$ represent the true (but unknown) underlying
-power matter spectra from which $D_n$ was generated.
-\textcolor{red}{(What is X?  Why is $\bar{Y}$ capital here, but it wasn't earlier?
-Since $\Lambda$ is fixed and doesn't change, I do not think we need to consider it part
-of the data.)}
+power matter spectra from which $D$ was generated.
+\textcolor{red}{(@Annie You suggested removing $\Lambda$ from $D$; is fixed and 
+doesn't change here, but it would if $X$ or $Y$ changed. Also, we use $\Lambda$ to pre-scale the data)}
 
@@
 
-Here, we introduce the Bayesian Gaussian Process (GP) model We assume 
+Here, we introduce the Bayesian Gaussian Process (GP) model. We assume 
 the matter power spectrum $S$ for a given cosmology follows a GP with mean $\mu$ 
-and covariance matrix $\Sigma_S$.
+and covariance matrix $\Sigma_S$. From the data, we obtain $\bar Y$, which is a 
+noisy version of $S$.
 
@@
 
-We transform $X$ values to fall within $[0,1]$ and $Y$ so that the mean and variance 
+Therefore, we model $\bar Y|S$ as a GP with mean $S$ and block-diagonal covariance 
+matrix $\Sigma_\varepsilon$:
+
+\begin{align}
+\bar Y|S &\sim GP\left(S,\Sigma_\varepsilon\right) \\
+S &\sim GP\left(\mu, \Sigma_S(x,x')\right)
+\end{align}
+
+We transform $X$ values to fall within $[0,1]$ and $\bar Y$ so that the mean and variance 
 are 0 and 1, respectively. This allows for a straightforward specification of priors, 
@@
 
-Therefore, we model $\bar Y|S$ as a GP with mean $S$ and block-diagonal covariance 
-matrix $\Sigma_\varepsilon$:
-
-\begin{align}
-\bar Y|S &\sim \mathcal{N}_n(S,\Sigma_\varepsilon) \\
-S &\sim GP\left(\mu, \Sigma_S(x,x')\right)
-\end{align}
-
-Upon integrating out $S$, we obtain $\bar Y \sim \mathcal{N}_n(\mu, \Sigma_S+\Sigma_\varepsilon)$. 
+A key property of GPs is that any finite set of $n$ observations from the GP is a 
+multivariate normal distribution. Therefore, upon integrating out $S$, we obtain 
+$\bar{Y} \sim \mathcal{N}_n(\mu, \Sigma_S+\Sigma_\varepsilon)$ for the $n$ observed X locations. 
 Additionally, this model yields the posterior distribution $S|\bar Y \sim \mathcal{N}_n(m, C)$, 

### 2025-05-23 paper/main.tex
annie working on revisions

@@
 is rather complex.  Although we presume there exists some true matter spectra as a function
-of wavenumber, we are not able to directly observe it.  Rather, the simulation outputs 
-a convoluted patchwork of potential spectra, which vary in fidelity and are trusted
-only over particular wavenumber ranges.  Our first objective is to synthesize expert knowledge
+of wavenumber, we are not able to directly observe it.  Rather, the simulation returns
+a convoluted patchwork of potential spectra, which vary in fidelity and are only trusted
+over particular wavenumber ranges.  Our first objective is to synthesize expert knowledge
 and all available simulation information into an effective estimate of the true underlying
@@
 centered on the true underlying spectra.  The hierarchical structure naturally
-accommodates the correlated observations.  The depth of the GP offers nonstationary
+accommodates the correlated observations.  The depth of the DGP offers nonstationary
 flexibility.  The Bayesian framework facilitates UQ and the incorporation of prior
@@
 Parameter ranges are specified in Table 2.1 of \cite{moran2023mira}.
-For simplicity, we will denote these eight parameters as $\theta \in \mathbb{R}^8$. 
 There is an optional ninth parameter, the redshift, which we fix at 0.
@@
 %this value at 0.
-For each cosmology, Mira-Titan contains a batch of 18 simulated spectra: an inexpensive 
-spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from 
-lower resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from 
-a high resolution simulation ($y_h$).
 
@@
 
+For each cosmology, Mira-Titan contains a batch of 18 simulated spectra: an inexpensive 
+spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from 
+lower resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from 
+a high resolution simulation ($y_h$). 
 These spectra are represented as a function of wavenumber $k$.  Following 
@@
 
-Notice, in the center panel of Figure \ref{fig:plot_data}, the precision of each
+Notice, in the center/right panels of Figure \ref{fig:plot_data}, the precision of each
 low/high resolution spectra increases as $k$ increases.  The curves get progressively 
@@
 low resolution average, and high resolution). Specifically, denote $\Lambda_p$, 
-$\Lambda_\ell$, and $\Lambda_h$ as the diagonal matrices with elements
+$\Lambda_\ell$, and $\Lambda_h$ as the $n\times n$ diagonal matrices with elements
 \[
@@
 of $10^5$ to impose the deterministic nature of the perturbation theory output.
+In $\Lambda_\ell$, the multiplication by 16 accounts for the sample size of the
+low resolution runs.
 We then calculate a weighted average for each cosmology: 
-$\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_h y_h + \Lambda_{\ell} \bar{y}_\ell)$, 
+$\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_{\ell} \bar{y}_\ell + \Lambda_h y_h)$, 
 where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ and 
@@
 spatial dependence inherent in the smooth but stochastic spectra.
-To account for these modeling challenges, we need a more complex model.
+To account for these modeling challenges, we need a more complex model which can
+also provide UQ.
 
@@
 
+For simplicity, we will denote these eight parameters as $\theta \in \mathbb{R}^8$.
+
 Upon obtaining the posterior means from each of the 111 training cosmologies, this 

### 2025-05-23 paper/main.tex
more revisions

@@
 the weighted average is shown in dashed orange.  As an estimate of the true
-underlying spectra, the weighted average fluctuates too drastically due 
-to the limited number of model runs.  It fails to account for the
-spatial dependence inherent in the smooth but stochastic spectra.
-To account for these modeling challenges, we need a more complex model which can
-also provide UQ.
+underlying spectrum, the weighted average fluctuates too drastically due 
+to the limited amount of data.  It fails to account for the
+spatial dependence inherent in the smooth but stochastic spectra,
+which is pivotal to effectively inferring the true spectrum.
 
-We consider a training-testing split with 111 cosmologies used for estimation and 6 
-cosmologies held out for testing.  Our ultimate objective is to effectively predict
-the underlying power matter spectra for the held-out cosmologies based on their 
-8-dimensional input configuration (Section \ref{sec:pred}).  But first, we must 
-obtain effective estimates of the underlying spectrum for each training 
-cosmology (Section \ref{sec:hm_fit}). \textcolor{red}{(@Steve - maybe this is a
-good place to introduce the ``hold out'' and ``within sample'' terms?)}
+%We consider a training-testing split with 111 cosmologies used for estimation and 6 
+%cosmologies held out for testing.  Our ultimate objective is to effectively predict
+%the underlying power matter spectra for the held-out cosmologies based on their 
+%8-dimensional input configuration (Section \ref{sec:pred}).  But first, we must 
+%obtain effective estimates of the underlying spectrum for each training 
+%cosmology (Section \ref{sec:hm_fit}). \textcolor{red}{(@Steve - maybe this is a
+%good place to introduce the ``hold out'' and ``within sample'' terms?)}
 
@@
 Here we describe our first contribution: the use of a Bayesian hierarchical model 
-to estimate the underlying spectra and quantify uncertainty for a particular cosmology,
-independent of the others. In the previous seciton, we used lowercase $\bar{y}$ to denote 
-the observed weighted average; here, we use $\bar{Y}$ to indicate the corresponding 
-random variable within the statistical model. Specifically, our input $X$ is the 
-$n$ values of $\log_{10}(k)$ and the output is the weighted average $\bar{Y}$. 
-Let $D = \{X, \bar{Y}, \Lambda\}$ represent the aggregated 
-data for a particular cosmology.  Let $S$ represent the true (but unknown) underlying
-power matter spectra from which $D$ was generated.
-\textcolor{red}{(@Annie You suggested removing $\Lambda$ from $D$; is fixed and 
-doesn't change here, but it would if $X$ or $Y$ changed. Also, we use $\Lambda$ to pre-scale the data)}
+to estimate the underlying spectra and quantify uncertainty for a particular cosmology.
+Let input $X$ represent the $n$-vector of $\log_{10}(k)$ values.
+In the previous section, we used lowercase $\bar{y}$ to denote 
+the observed weighted average; now, we use $\bar{Y}$ to indicate the corresponding 
+random variable within the statistical model.  $\bar{Y}$ is typically pre-processed
+in a domain-specific manner, such as that defined in Section \ref{sec:data}.
+Together, $X$ and $\bar{Y}$ comprise the training data for a particular cosmology.
+We will reserve specific integration 
+of expert knowledge for the CAMB and Mira-Titan suites, such as this pre-processing,
+for Sections \ref{subsec:camb} and \ref{subsec:mira_fit}, respectively.
 
-\subsection{Gaussian Process Model}
+\subsection{Deep Gaussian Process Hierarchical Model}
 
@@
 and covariance matrix $\Sigma_S$. From the data, we obtain $\bar Y$, which is a 
-noisy version of $S$.
+noisy version of $S$.  Let $S$ represent the true (but unknown) underlying
+power matter spectra from which $D$ was generated.
 
@@
 % With this hierarchical model, we improve the efficiency of estimating $S|\bar{Y}$.
-
-\subsection{Deep Gaussian Process Model}
 

### 2025-05-23 paper/main.tex
work on section 3

@@
 to estimate the underlying spectra and quantify uncertainty for a particular cosmology.
-Let input $X$ represent the $n$-vector of $\log_{10}(k)$ values.
+Let input $X$ represent the vector with elements $x_i = \log_{10}(k_i)$ for $i=1,\dots,n$.
 In the previous section, we used lowercase $\bar{y}$ to denote 
@@
 
-\subsection{Deep Gaussian Process Hierarchical Model}
+\subsection{Model Specification}
 
-Here, we introduce the Bayesian Gaussian Process (GP) model. We assume 
-the matter power spectrum $S$ for a given cosmology follows a GP with mean $\mu$ 
-and covariance matrix $\Sigma_S$. From the data, we obtain $\bar Y$, which is a 
-noisy version of $S$.  Let $S$ represent the true (but unknown) underlying
-power matter spectra from which $D$ was generated.
+Let $S$ represent the true (but unknown) underlying power matter spectra for
+a particular cosmology.  We consider $\bar{Y}$ a noisy realization of $S$.
+Specifically, we assume $\bar{Y}$ is a random realization of a Gaussian 
+process centered at $S$ with some covariance, i.e. 
+$\bar{Y}|S \sim \mathrm{GP}(S, \Sigma_\varepsilon)$.  While previous works
+have considered diagonal $\Sigma_\varepsilon$ \citep{moran2024dpc}, we find it
+essential to account for the smoothly varying spatial dependence of the
+low resolution runs (as visualized in Figure \ref{fig:plot_data}).  For
+this, we embrace variations of the Mat\'ern kernel \citep{stein1999interpolation},
+tailored to each scenario (more on this in later subsections).
 
-From plots of the multiple low-resolution spectra (e.g., Figure \ref{fig:plot_data}), 
-we find they vary smoothly about some mean $S$, suggesting that a dense covariance 
-matrix $\Sigma_\varepsilon$ (accounting for spatial dependence) may outperform a 
-diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$. In previous work, we have 
-modeled multiple candidates for $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and 
-found that a Mat\'ern covariance $\Sigma_\ell$ trained on the low-resolution runs 
-(pre-scaled by the precision values and subtracting a LOESS-smoothed average) performs 
-the best. We specify this structure within $\Sigma_\varepsilon$ and refer readers 
-to \cite{walsh2023bayesian} for a more detailed exposition. This yields 
-$\Sigma_\varepsilon=\left(\Lambda_p^{-1} + \Sigma_\ell + \Lambda_h^{-1}\right)^{-1}$, 
-where $\Lambda_p, \Sigma_\ell^{-1}, \text{ and } \Lambda_h$ are all $n\times n$ 
-and are 0 wherever the corresponding data type is biased.
+We next place a zero-mean Gaussian process prior on $S$, i.e.,
+$S\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(\cdot)\right)$, again with a 
+spatial Mat\`ern kernel.  
+\textcolor{red}{(@Steve - is a zero mean the right thing?  If not, what is the mean?)}
+Standard practice would be to use locations $X$ as the inputs to kernel $\Sigma_S(\cdot)$.
+Yet we have found that prior to lack sufficient flexibility for our motivating
+application.  In the Mira-Titan setting, some nonstationarity is expected due to 
+baryonic acoustic oscillations when $-2 \leq \log_{10}(k) \leq -1$. To account 
+for this, we incorporate an additional latent Gaussian process, $W$, which accommodates
+nonstationarity by warping $X$.  Ultimately, this yields the following hierarchical model:
+\begin{equation}\label{eq:dgphm}
+\begin{aligned}
+\bar{Y}|S,W &\sim \mathrm{GP}(S,\Sigma_\varepsilon) \\
+S|W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(W)\right) \\
+W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_W(X)\right)
+\end{aligned}
+\end{equation}
+The functional composition of GPs in the latter two lines forms a ``deep Gaussian process'' 
+\citep{damianou2013deep,dunlop2018deep}. Additional latent layers could be considered, 
+although we find one is sufficient.
 
-Therefore, we model $\bar Y|S$ as a GP with mean $S$ and block-diagonal covariance 
-matrix $\Sigma_\varepsilon$:
+\[
+\Sigma_S(X)^{i,j} = \tau^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta}} + 
+  \frac{5d^2}{3\theta}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta}}\right)
+\quad\textrm{for}\quad 
+d=||x_i-x_j||^2
+\]
+Hyperparameters $\tau^2$ and $\theta$ govern the correlation scale and decay respectively.
 
-\begin{align}
-\bar Y|S &\sim GP\left(S,\Sigma_\varepsilon\right) \\
-S &\sim GP\left(\mu, \Sigma_S(x,x')\right)
-\end{align}
+We have found that whether the prior mean of $W$ is set to 0 or $X$ 
+\citep[which would indicate stationarity apriori,][]{schmidt2003bayesian} does not 
+have a substantial impact on the model fit. We construct this model in the 
+compositional form, and also use Markov Chain Monte Carlo (MCMC) to estimate and 
+obtain full uncertainty for each of the model parameters \citep{sauer2023active}.
 
@@
 
-\begin{align}
-\Sigma_\ell^{i,j} = \tau_\ell^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_\ell}} + 
-  \frac{5d^2}{3\theta_\ell}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_\ell}}\right),\\
-\Sigma_S^{i,j} = \tau_S^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_S}} + 
-  \frac{5d^2}{3\theta_S}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_S}}\right),
-\end{align}
-
-where $d=||x_i-x_j||^2$. From here, we estimate $\tau_\ell, \theta_\ell$ with maximum 
+where. From here, we estimate $\tau_\ell, \theta_\ell$ with maximum 
 likelihood, and conditional on $\Sigma_\varepsilon$, we specify priors for $\tau_S, \theta_S$ 
@@
 
+\subsection{Model Inference}
+
 A key property of GPs is that any finite set of $n$ observations from the GP is a 
@@
 % With this hierarchical model, we improve the efficiency of estimating $S|\bar{Y}$.
-
-Considering the Mira-Titan data, some nonstationarity is expected (e.g., due to 
-baryonic acoustic oscillations when $-2 \leq \log_{10}(k) \leq -1$). To account 
-for this, we incorporate a latent layer $W$ within the hierarchical model to warp 
-locations of $X$ and incorporate nonstationary dynamics of the matter power spectra. 
-When $W$ is a GP, this creates a deep Gaussian process (DGP) model \citep{damianou2013deep}. 
-Additional latent layers could be considered, although one is sufficient for this 
-dataset \citep{dunlop2018deep}.
-
-\begin{align}
-\bar Y|S,W &\sim \mathcal{N}_n(S,\Sigma_\varepsilon) \\
-S|W &\sim GP\left(\mu, \Sigma_S(w,w')\right) \\
-W &\sim GP\left(0, \Sigma_W(x,x')\right)
-\end{align}
-
-We have found that whether the prior mean of $W$ is set to 0 or $X$ 
-\citep[which would indicate stationarity apriori,][]{schmidt2003bayesian} does not 
-have a substantial impact on the model fit. We construct this model in the 
-compositional form, and also use Markov Chain Monte Carlo (MCMC) to estimate and 
-obtain full uncertainty for each of the model parameters \citep{sauer2023active}.
 
@@
 \label{subsec:mira_fit}
+
+From plots of the multiple low-resolution spectra (e.g., Figure \ref{fig:plot_data}), 
+we find they vary smoothly about some mean $S$, suggesting that a dense covariance 
+matrix $\Sigma_\varepsilon$ (accounting for spatial dependence) may outperform a 
+diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$. In previous work, we have 
+modeled multiple candidates for $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and 
+found that a Mat\'ern covariance $\Sigma_\ell$ trained on the low-resolution runs 
+(pre-scaled by the precision values and subtracting a LOESS-smoothed average) performs 
+the best. We specify this structure within $\Sigma_\varepsilon$ and refer readers 
+to \cite{walsh2023bayesian} for a more detailed exposition. This yields 
+$\Sigma_\varepsilon=\left(\Lambda_p^{-1} + \Sigma_\ell + \Lambda_h^{-1}\right)^{-1}$, 
+where $\Lambda_p, \Sigma_\ell^{-1}, \text{ and } \Lambda_h$ are all $n\times n$ 
+and are 0 wherever the corresponding data type is biased.
 

### 2025-05-23 paper/main.tex
more work in section 3

@@
 $S\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(\cdot)\right)$, again with a 
-spatial Mat\`ern kernel.  
+spatial kernel.  
 \textcolor{red}{(@Steve - is a zero mean the right thing?  If not, what is the mean?)}
@@
 \end{equation}
+$W$ is also an $n$-vector whose entries $w_i$ correspond to warped versions of each $x_i$.
 The functional composition of GPs in the latter two lines forms a ``deep Gaussian process'' 
@@
 
-\[
-\Sigma_S(X)^{i,j} = \tau^2  \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta}} + 
-  \frac{5d^2}{3\theta}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta}}\right)
+We use the Mat\`ern kernels,
+\begin{align}
+&\Sigma_S(W)^{i,j} = \tau^2 \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_S}} + 
+  \frac{5d^2}{3\theta_S}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_S}}\right)
 \quad\textrm{for}\quad 
-d=||x_i-x_j||^2
-\]
-Hyperparameters $\tau^2$ and $\theta$ govern the correlation scale and decay respectively.
+d=||w_i-w_j||^2, \;\;\textrm{and} \\
+&\Sigma_W(X)^{i,j} = \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_W}} + 
+  \frac{5d^2}{3\theta_W}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_W}}\right)
+\quad\textrm{for}\quad 
+d=||x_i-x_j||^2.
+\end{align}
+On the $S$ layer, hyperparameters $\tau^2$ and $\theta_S$ govern the 
+correlation scale and lengthscale respectively.  On the $W$ layer, we allow for a
+unique lengthscale $\theta_W$ but enforce unit scale following \citet{sauer2023active}.
 
-We have found that whether the prior mean of $W$ is set to 0 or $X$ 
+We have found that setting the prior mean of $W$ to $X$ instead of the $\mathbf{0}$
+in Eq.~\ref{eq:dgphm},
 \citep[which would indicate stationarity apriori,][]{schmidt2003bayesian} does not 
@@
 obtain full uncertainty for each of the model parameters \citep{sauer2023active}.
+\textcolor{red}{(Need more details on inference, in next section...)}
 

### 2025-05-23 paper/main.tex
start 3.2

@@
 \begin{aligned}
-\bar{Y}|S,W &\sim \mathrm{GP}(S,\Sigma_\varepsilon) \\
+\bar{Y}|S &\sim \mathrm{GP}(S,\Sigma_\varepsilon) \\
 S|W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(W)\right) \\
@@
 We use the Mat\`ern kernels,
-\begin{align}
+\[
+\begin{aligned}
 &\Sigma_S(W)^{i,j} = \tau^2 \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_S}} + 
@@
 d=||x_i-x_j||^2.
-\end{align}
+\end{aligned}
+\]
 On the $S$ layer, hyperparameters $\tau^2$ and $\theta_S$ govern the 
@@
 unique lengthscale $\theta_W$ but enforce unit scale following \citet{sauer2023active}.
+We considered setting the prior mean of $W$ to $X$ instead of $\mathbf{0}$
+\citep[which would indicate stationarity apriori;][]{schmidt2003bayesian}, but found
+it had a neglible effect on the model fits.
 
-We have found that setting the prior mean of $W$ to $X$ instead of the $\mathbf{0}$
-in Eq.~\ref{eq:dgphm},
-\citep[which would indicate stationarity apriori,][]{schmidt2003bayesian} does not 
-have a substantial impact on the model fit. We construct this model in the 
-compositional form, and also use Markov Chain Monte Carlo (MCMC) to estimate and 
-obtain full uncertainty for each of the model parameters \citep{sauer2023active}.
-\textcolor{red}{(Need more details on inference, in next section...)}
+\subsection{Bayesian Inference}
 
@@
 deepgp v1.1.3: $\theta_S \sim G(\alpha=1.2, \beta=1)$ and $\tau_S \propto 1/\tau_S^2$.}
-
-% $$
-% \mathbf{\Sigma_\varepsilon} =
-% \begin{bmatrix}
-% \Lambda_p^{-1} & \mathbf{0} & \mathbf{0} \\
-% \mathbf{0} & \mathbf{\Sigma}_{\ell} & \mathbf{0} \\
-% \mathbf{0} & \mathbf{0} & \Lambda_h^{-1}
-% \end{bmatrix}
-% $$
-
-\subsection{Model Inference}
 

### 2025-05-23 paper/main.tex
annie done through 3.2

@@
 process centered at $S$ with some covariance, i.e. 
-$\bar{Y}|S \sim \mathrm{GP}(S, \Sigma_\varepsilon)$.  While previous works
+$\bar{Y}\mid S \sim \mathrm{GP}(S, \Sigma_\varepsilon)$.  While previous works
 have considered diagonal $\Sigma_\varepsilon$ \citep{moran2024dpc}, we find it
@@
 \begin{aligned}
-\bar{Y}|S &\sim \mathrm{GP}(S,\Sigma_\varepsilon) \\
-S|W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(W)\right) \\
+\bar{Y}\mid S,\Sigma_\varepsilon &\sim \mathrm{GP}(S,\Sigma_\varepsilon) \\
+S\mid W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(W)\right) \\
 W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_W(X)\right)
@@
 \]
-On the $S$ layer, hyperparameters $\tau^2$ and $\theta_S$ govern the 
-correlation scale and lengthscale respectively.  On the $W$ layer, we allow for a
+In the $S$ layer, hyperparameters $\tau^2$ and $\theta_S$ govern the 
+correlation scale and lengthscale respectively.  In the $W$ layer, we allow for a
 unique lengthscale $\theta_W$ but enforce unit scale following \citet{sauer2023active}.
@@
 
-We transform $X$ values to fall within $[0,1]$ and $\bar Y$ so that the mean and variance 
-are 0 and 1, respectively. This allows for a straightforward specification of priors, 
-where the hyperparameters for the Mat\'ern covariance $\Sigma_S$ each take a vague prior. 
-Specifically:
+While our inferential goal is to obtain the posterior distribution of 
+$S\mid\bar{Y},\Sigma_\varepsilon$, this requires inferring latent quantity $W$.  
+We first integrate over $S$ to condense our hierarchical model of Eq.~\ref{eq:dgphm} into
+\begin{align}
+\label{eq:likelihood}
+\bar{Y} \mid W, \Sigma_\varepsilon &\sim \textrm{GP}(\mathbf{0}, \Sigma_S(W) + \Sigma_\varepsilon) \\
+\label{eq:prior}
+W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_W(X)\right)
+\end{align}
+Details are provided in Appendix \textcolor{red}{???}. We then use elliptical slice
+sampling \citep[ESS;][]{murray2010elliptical} to generate posterior samples of $W$.  
+Proposals are generated from prior samples from Eq.~(\ref{eq:prior}), and acceptances 
+involve the ratio of likelihoods from Eq.~\ref{eq:likelihood}.
 
-where. From here, we estimate $\tau_\ell, \theta_\ell$ with maximum 
-likelihood, and conditional on $\Sigma_\varepsilon$, we specify priors for $\tau_S, \theta_S$ 
-following \cite{sauer2023active}: \textcolor{red}{Annie, please ensure the most up-to-date 
-prior specs are shown above and here. Here is what I found from fit\$settings when using 
-deepgp v1.1.3: $\theta_S \sim G(\alpha=1.2, \beta=1)$ and $\tau_S \propto 1/\tau_S^2$.}
+We also embrace a fully-Bayesian treatment of unknown kernel hyperparameters 
+$\{\tau^2, \theta_S, \theta_W\}$.  We prescale $X$ to $[0, 1]$
+and $\bar{Y}$ to zero mean and unit variance.  On this scale, we adopt the default 
+prior specifications provided in the {\tt deepgp} {\sf R}-package \citep{deepgp}.
+\textcolor{red}{(@Steve - check the version number of the package in this reference.  Also,
+I'm not sure what we really do with $\tau^2$.  In my package, we assign it a reference prior and
+integrate it out.  I don't think we can do that here since we have the $\Sigma_\varepsilon$.)}
+%\textcolor{red}{Annie, please ensure the most up-to-date 
+%prior specs are shown above and here. Here is what I found from fit\$settings when using 
+%deepgp v1.1.3: $\theta_S \sim G(\alpha=1.2, \beta=1)$ and $\tau_S \propto 1/\tau_S^2$.}
+We integrate Metropolis-Hastings sampling of these parameters in a Gibbs framework with ESS
+sampling of $W$ following the scheme of \citep{sauer2023active}.
 
-A key property of GPs is that any finite set of $n$ observations from the GP is a 
-multivariate normal distribution. Therefore, upon integrating out $S$, we obtain 
-$\bar{Y} \sim \mathcal{N}_n(\mu, \Sigma_S+\Sigma_\varepsilon)$ for the $n$ observed X locations. 
-Additionally, this model yields the posterior distribution $S|\bar Y \sim \mathcal{N}_n(m, C)$, 
-where $C=\left(\Sigma_S^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1}$ and 
-$m=C\left(\Sigma_\varepsilon^{-1}\bar Y+\Sigma_S^{-1}\mu\right)$. 
-This distribution provides the posterior mean and corresponding uncertainty for 
-$S$ within a given cosmology. 
-% With this hierarchical model, we improve the efficiency of estimating $S|\bar{Y}$.
-
-\textcolor{red}{Annie: want to add something here? For $\theta_w$, we also specify 
-a Gamma prior of the form $\theta_w \sim G(\alpha=1.2, \beta=2)$. We discuss the 
-elliptical slice sampling of $W$, but otherwise it is plugged in to the methods 
-of the previous section with very few changes.}
-
-Through the synthesis of the DGP model alongside the additional level of hierarchical 
-modeling (HM) to incorporate multiple functional model runs, we call this method DGP.HM.
+Given burned-in posterior samples $\{W^{(t)}, \tau^{2(t)}, \theta_S^{(t)}, \theta_W^{(t)}\}$
+for $t\in\mathcal{T}$, we may obtain the
+posterior distribution of the true power matter spectra for a particular cosmology as
+\[
+S^{(t)}\mid\bar{Y}, \Sigma_\varepsilon \sim \mathcal{N}_n(m^{(t)}, C^{(t)})
+\quad\textrm{where}\quad
+\begin{array}{rl}
+C^{(t)}&=\left(\Sigma_S(W^{(t)})^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1} \\
+m^{(t)}&=C^{(t)}\left(\Sigma_\varepsilon^{-1}\bar{Y}+\Sigma_S(W^{(t)})^{-1}\mu\right).
+\end{array}
+\] 
+\textcolor{red}{(@Steve - provide details of this derivation.  Also what is $\mu$?  Are we 
+going with zero here?  Add details of how we combine these things across all the $t$ samples.)}
+We denote our inference procedure for the hierarchical model of Eq.~\ref{eq:dgphm} as
+``DGP.HM'' given its deep Gaussian process foundation with additional hierarchical modeling (HM)
+structure to incorporate multiple functional outputs.
+\textcolor{red}{(Is it too late to change this?  Maybe DGP-FCO for functional correlated output?)}
 
@@
 and are 0 wherever the corresponding data type is biased.
+
+From here, we estimate $\tau_\ell, \theta_\ell$ with maximum 
+likelihood, and conditional on $\Sigma_\varepsilon$,
 

### 2025-06-04 sims/functions.R
add 6th setting for sim study

@@ -41,6 +41,11 @@ get_Sigma_true <- function(x, n, func = 1, setting = 1) {
       Sigma <- deepgp:::Matern(distmat = D, tau2 = 0.1^2, 
                                theta = 5e-2, g = 1e-8, v = 2.5)
       Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
+    } else if(setting == 6) {
+      D <- deepgp::sq_dist(x)
+      Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
+                               theta = 1e-1, g = 1e-8, v = 2.5)
+      Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
     }
   } else if(func == 2) {
     if(setting == 1) {
@@ -63,6 +68,11 @@ get_Sigma_true <- function(x, n, func = 1, setting = 1) {
       D <- deepgp::sq_dist(x)
       Sigma <- deepgp:::Matern(distmat = D, tau2 = 0.1^2, 
                                theta = 5e-2, g = 1e-8, v = 2.5)
+      Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
+    } else if(setting == 6) {
+      D <- deepgp::sq_dist(x)
+      Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
+                               theta = 1e-1, g = 1e-8, v = 2.5)
       Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
     }
   }

### 2025-06-04 sims/run_sim.R

@@ -36,7 +36,7 @@ if(length(args) > 0)
   for(i in 1:length(args))
     eval(parse(text=args[[i]]))
 if(!(func %in% 1:2)) stop("func should be 1 or 2")
-if(!(setting %in% 1:5)) stop("setting should be 1, 2, 3, 4, or 5")
+if(!(setting %in% 1:6)) stop("setting should be 1, 2, 3, 4, 5, or 6")
 
 set.seed(seed)
 source("functions.R") # generates random model parameters upon sourcing
@@ -49,6 +49,7 @@ vis <- FALSE # should plots be generated
 x <- seq(0, 4, by = 0.1)
 n <- length(x)
 if(setting==5) sddtrue <- 1/exp(seq(0,2,length=length(x)))
+if(setting==6) sddtrue <- 1/1.5^(seq(0,2,length=length(x)))
 Sigma_true <- get_Sigma_true(x, n, func, setting)
 if(vis) image(Sigma_true) # make sure sd plot looks right
 
@@ -113,7 +114,7 @@ if(setting %in% 1:3) {
                                     log(tau2hat_0)))
   Sigma_hat <- deepgp:::Matern(dx, params_hat$tau2_hat, params_hat$theta_hat,
                                g = 1e-8, v = 2.5)
-} else if(setting==5) {
+} else if(setting %in% 5:6) {
   # Get initial tau2 estimate
   tau2hat_0 <- mean(apply(Yzm, 2, var))
   
@@ -245,7 +246,6 @@ for(i in 1:1){ # if you only have one function
 }
 
 # Run DPC model
-library(dpc)
 n_u=30; n_v=10
 dpc_out = dpc(Y=Y_list, vars=vars_list, n_u=n_u, n_v=n_v, 
               nmcmc=15000, burn=5000,

### 2025-06-04 sims/collect_sims.R
add new sim setting; improve boxplots

@@ -20,7 +20,7 @@ MSE5_range <- MSE20_range <- logS5_range <- logS20_range <- c()
 
 r=5
 for (func in 1:2) {
-  for (setting in 4:5) {
+  for (setting in 4:6) {
     filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
     # Compare MSEs
     results <- read.csv(filename)[, c(3,6,5,4)]
@@ -33,9 +33,9 @@ for (func in 1:2) {
   }
 }
 
-r=20
+r=15
 for (func in 1:2) {
-  for (setting in 4:5) {
+  for (setting in 4:7) {
     filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
     # Compare MSEs
     results <- read.csv(filename)[, c(3,6,5,4)]
@@ -49,44 +49,77 @@ for (func in 1:2) {
 # Boxplots of MSEs #
 ####################
 
-par(mfrow=c(2,4),oma=c(4,4,1.5,1),mar=c(0,0,0,0))
-for (r in c(5,20)) {
-  for (setting in 4:5) {
+par(mfrow=c(4,4), oma=c(4,4,1.5,1), mar=c(0,0,0,0))
+
+# Step 1: Precompute y-limits for each setting (row)
+row_ylims <- list()
+for (setting in 4:7) {
+  all_mses <- NULL
+  for (r in c(5,15)) {
     for (func in 1:2) {
       filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
-      # Compare MSEs
+      results <- read.csv(filename)[, c(3,6,5,4)]
+      all_mses <- c(all_mses, unlist(results))
+    }
+  }
+  row_ylims[[as.character(setting)]] <- range(all_mses, na.rm = TRUE)
+}
+
+# Step 2: Plot with per-row ylim
+for (setting in 4:7) {
+  row_ylim <- row_ylims[[as.character(setting)]]
+  for (r in c(5,15)) {
+    for (func in 1:2) {
+      filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
       results <- read.csv(filename)[, c(3,6,5,4)]
       boxplot(results, ylab = "MSE", axes=F,
               names = c("dgp.hm","dpc","hetGP","deepGP"), col = cbcols[c(4,5,7,9)],
-              ylim = if(r==5){MSE5_range}else{MSE20_range})
+              ylim = row_ylim)
[…]

### 2025-06-04 sims/functions.R

@@ -46,6 +46,11 @@ get_Sigma_true <- function(x, n, func = 1, setting = 1) {
       Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
                                theta = 1e-1, g = 1e-8, v = 2.5)
       Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
+    } else if(setting == 7) {
+      D <- deepgp::sq_dist(x)
+      Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
+                               theta = 5e-2, g = 1e-8, v = 2.5)
+      Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
     }
   } else if(func == 2) {
     if(setting == 1) {
@@ -73,6 +78,11 @@ get_Sigma_true <- function(x, n, func = 1, setting = 1) {
       D <- deepgp::sq_dist(x)
       Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
                                theta = 1e-1, g = 1e-8, v = 2.5)
+      Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
+    } else if(setting == 7) {
+      D <- deepgp::sq_dist(x)
+      Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
+                               theta = 5e-2, g = 1e-8, v = 2.5)
       Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
     }
   }

### 2025-06-04 sims/run_sim.R

@@ -36,7 +36,7 @@ if(length(args) > 0)
   for(i in 1:length(args))
     eval(parse(text=args[[i]]))
 if(!(func %in% 1:2)) stop("func should be 1 or 2")
-if(!(setting %in% 1:6)) stop("setting should be 1, 2, 3, 4, 5, or 6")
+if(!(setting %in% 1:7)) stop("setting should be 1, 2, 3, 4, 5, 6, or 7")
 
 set.seed(seed)
 source("functions.R") # generates random model parameters upon sourcing
@@ -49,7 +49,7 @@ vis <- FALSE # should plots be generated
 x <- seq(0, 4, by = 0.1)
 n <- length(x)
 if(setting==5) sddtrue <- 1/exp(seq(0,2,length=length(x)))
-if(setting==6) sddtrue <- 1/1.5^(seq(0,2,length=length(x)))
+if(setting %in% 6:7) sddtrue <- 1/1.5^(seq(0,2,length=length(x)))
 Sigma_true <- get_Sigma_true(x, n, func, setting)
 if(vis) image(Sigma_true) # make sure sd plot looks right
 
@@ -114,7 +114,7 @@ if(setting %in% 1:3) {
                                     log(tau2hat_0)))
   Sigma_hat <- deepgp:::Matern(dx, params_hat$tau2_hat, params_hat$theta_hat,
                                g = 1e-8, v = 2.5)
-} else if(setting %in% 5:6) {
+} else if(setting %in% 5:7) {
   # Get initial tau2 estimate
   tau2hat_0 <- mean(apply(Yzm, 2, var))
   

### 2025-06-05 paper/main.tex
some tweaks, initial add of first appendix

@@
 
-We next place a zero-mean Gaussian process prior on $S$, i.e.,
-$S\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(\cdot)\right)$, again with a 
-spatial kernel.  
-\textcolor{red}{(@Steve - is a zero mean the right thing?  If not, what is the mean?)}
-Standard practice would be to use locations $X$ as the inputs to kernel $\Sigma_S(\cdot)$.
+We next place a Gaussian process prior on $S$, i.e.,
+$S\sim \mathrm{GP}\left(\boldsymbol{\mu}, \Sigma_S(\cdot)\right)$, again with a 
+spatial kernel. Here, we set $\boldsymbol{\mu}=\mathbf{0}$, although the modeling can 
+be carried out with any selection of $\boldsymbol{\mu}$. Standard practice would 
+be to use locations $X$ as the inputs to kernel $\Sigma_S(\cdot)$.
 Yet we have found that prior to lack sufficient flexibility for our motivating
@@
 \bar{Y}\mid S,\Sigma_\varepsilon &\sim \mathrm{GP}(S,\Sigma_\varepsilon) \\
-S\mid W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(W)\right) \\
+S\mid W &\sim \mathrm{GP}\left(\boldsymbol{\mu}, \Sigma_S(W)\right) \\
 W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_W(X)\right)
@@
 \label{eq:likelihood}
-\bar{Y} \mid W, \Sigma_\varepsilon &\sim \textrm{GP}(\mathbf{0}, \Sigma_S(W) + \Sigma_\varepsilon) \\
+\bar{Y} \mid W, \Sigma_\varepsilon &\sim \textrm{GP}(\boldsymbol{\mu}, \Sigma_S(W) + \Sigma_\varepsilon) \\
 \label{eq:prior}
@@
 C^{(t)}&=\left(\Sigma_S(W^{(t)})^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1} \\
-m^{(t)}&=C^{(t)}\left(\Sigma_\varepsilon^{-1}\bar{Y}+\Sigma_S(W^{(t)})^{-1}\mu\right).
+m^{(t)}&=C^{(t)}\left(\Sigma_\varepsilon^{-1}\bar{Y}+\Sigma_S(W^{(t)})^{-1}\boldsymbol{\mu}\right).
 \end{array}
@@
 
-\section{Appendix}
-\label{sec:apdx}
+\appendix
+\renewcommand{\thesection}{Appendix \Alph{section}}
+
+\section{Integrated Likelihood}
+\label{sec:apdx_int_lik}
+Here, we establish that under the model:
+
+$$
+\begin{aligned}
+Y_i|S,W &\sim N\big(S, \Sigma_\varepsilon(w,w')\big), \quad i \in \{1,\dots, r\}\\
+S|W &\sim N\big(m, \Sigma_S(w,w')\big)\\
+W &\sim N\big(0, \Sigma_W(x,x')\big) \\
+\end{aligned}
+$$
+
+We can obtain the integrated likelihood $p(Y_i|W)=\int p(Y_i|S,W)p(S|W) dS$, with the following calculations:
+$$
+\begin{aligned}
+p(Y_i|W) &= \int p(Y_i|S,W)p(S|W)dS \\
+&\propto \int\exp\left(-\frac{1}{2} \left[Y_i^T\Sigma_\varepsilon^{-1} Y_i - 2Y_i^T\Sigma_\varepsilon^{-1} S + S^T \Sigma_\varepsilon^{-1} S + S^T \Sigma_S^{-1} S -2S^T \Sigma_S^{-1} m + m^T \Sigma_S^{-1} m \right]\right)dS \\
+\end{aligned}
+$$
+
+If we focus only on terms containing $S$, we have 
+$$
+\begin{aligned}
+    \exp\Big(-\frac{1}{2} \big[S^T &(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1}) S -2S^T (\Sigma_S^{-1} m + \Sigma_\varepsilon^{-1} Y_i) \big]\Big) \\
+&= \exp\left(-\frac{1}{2} \left[S^T (\Sigma_\varepsilon^{-1}+\Sigma_S^{-1}) S -2S^T (\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}(\Sigma_S^{-1} m + \Sigma_\varepsilon^{-1} Y_i) \right]\right)
+\end{aligned}
+$$
+
+To complete the square, we add and subtract $(\Sigma_S^{-1} m + \Sigma_\varepsilon^{-1} Y_i)^T(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}(\Sigma_S^{-1} m + \Sigma_\varepsilon^{-1} Y_i)$, which then gives the kernel for $S \sim N\left((\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}(\Sigma_S^{-1} m + \Sigma_\varepsilon^{-1} Y_i), (\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1} \right)$.
+This removes the integral and leaves us with the following terms for $Y_i|W$:
+$$
+\begin{aligned}
+&\exp\Big(-\frac{1}{2}\big[Y_i^T\Sigma_\varepsilon^{-1} Y + m^T\Sigma_S^{-1} m - (\Sigma_S^{-1} m + \Sigma_\varepsilon^{-1} Y_i)^T(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1} (\Sigma_S^{-1} m + \Sigma_\varepsilon^{-1} Y_i)\big]\Big) \\
+&= \exp\Big(-\frac{1}{2}\big[Y_i^T\big(\Sigma_\varepsilon^{-1} - \Sigma_\varepsilon^{-1}(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}\Sigma_\varepsilon^{-1}\big)Y_i -2Y_i^T \Sigma_\varepsilon^{-1}(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}\Sigma_S^{-1} m \\
+&\qquad \qquad \qquad + m^T (\Sigma_S^{-1} - \Sigma_S^{-1}(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}\Sigma_S^{-1}) m\big]\Big) \\
+&= \exp\Big(-\frac{1}{2}\big((Y_i-m)^T(\Sigma_\varepsilon+\Sigma_S)^{-1}(Y_i-m) \big)\Big)
+\end{aligned}
+$$
+
+With the last equality resulting from the use of the Sherman–Morrison–Woodbury formula, yielding $\big(\Sigma_\varepsilon^{-1} - \Sigma_\varepsilon^{-1}(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}\Sigma_\varepsilon^{-1} = (\Sigma_S^{-1} - \Sigma_S^{-1}(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}\Sigma_S^{-1}) = (\Sigma_\varepsilon+\Sigma_S)^{-1}$.
+We also use the fact that $\Sigma_\varepsilon^{-1}(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}\Sigma_S^{-1} = (\Sigma_\varepsilon + \Sigma_S)^{-1}$, which holds since it can be shown that $[\Sigma_S(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})\Sigma_\varepsilon](\Sigma_\varepsilon+\Sigma_S)=(\Sigma_\varepsilon+\Sigma_S)(\Sigma_\varepsilon+\Sigma_S)$.
+
+Therefore, we have that $Y_i|W \sim N\big(m, \Sigma_S(w,w')+\Sigma_\varepsilon(w,w')\big)$.
+
+\section{Mira-Titan Predictive Plots}
+\label{sec:apdx_mira_plots}
 

### 2025-06-06 paper/main.tex
add appendices, change theta to psi for design, general updates to manuscript

--- 2025-05-23 paper/main.tex
@@
 \end{align}
-Details are provided in Appendix \textcolor{red}{???}. We then use elliptical slice
+Details are provided in \ref{sec:apdx_int_lik}. We then use elliptical slice
 sampling \citep[ESS;][]{murray2010elliptical} to generate posterior samples of $W$.  
@@
 prior specifications provided in the {\tt deepgp} {\sf R}-package \citep{deepgp}.
-\textcolor{red}{(@Steve - check the version number of the package in this reference.  Also,
-I'm not sure what we really do with $\tau^2$.  In my package, we assign it a reference prior and
-integrate it out.  I don't think we can do that here since we have the $\Sigma_\varepsilon$.)}
+\textcolor{red}{(@Annie - from our email 5/29, can you comment on how we handle $\tau^2$/
+cite Bobby's book/etc? Thanks!)}
 %\textcolor{red}{Annie, please ensure the most up-to-date 
@@
 C^{(t)}&=\left(\Sigma_S(W^{(t)})^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1} \\
-m^{(t)}&=C^{(t)}\left(\Sigma_\varepsilon^{-1}\bar{Y}+\Sigma_S(W^{(t)})^{-1}\mu\right).
+m^{(t)}&=C^{(t)}\left(\Sigma_\varepsilon^{-1}\bar{Y}+\Sigma_S(W^{(t)})^{-1}\boldsymbol{\mu}\right).
 \end{array}
 \]
-\textcolor{red}{(@Steve - provide details of this derivation.  Also what is $\mu$?  Are we 
-going with zero here?  Add details of how we combine these things across all the $t$ samples.)}
+
+These results come from Bayes' Theorem, with details of this derivation provided 
+in \ref{sec:apdx_SgivenY}. In our application, we find success setting 
+$\boldsymbol{\mu}=\boldsymbol{0}$. With all posterior draws $S^{(t)}$, we compute
+the average of these to obtain our posterior mean for a particular cosmology.
 We denote our inference procedure for the hierarchical model of Eq.~\ref{eq:dgphm} as
@@
 structure to incorporate multiple functional outputs.
-\textcolor{red}{(Is it too late to change this?  Maybe DGP-FCO for functional correlated output?)}
 
@@
 
-For simplicity, we will denote these eight parameters as $\theta \in \mathbb{R}^8$.
+For simplicity, we will denote these eight parameters as $\psi \in \mathbb{R}^8$.
 
@@
 mean trend, the principal components contained in $B$, and one example of how the 
-weight for the first PC will vary dependent on the first of $p_\theta=8$ cosmological 
-parameters contained in $\theta$.
+weight for the first PC will vary dependent on the first of $p_\psi=8$ cosmological 
+parameters contained in $\psi$.
 
@@
 predictions for Mira-Titan Data, NOT SEPIA. The CAMB results are based on SEPIA.}
-That is, the $i$th vector of PC weights $w_i(\theta)$, 
+That is, the $i$th vector of PC weights $w_i(\psi)$, 
 $i \in \{1,\ldots,p_\eta\}$, will be modeled as a zero mean GP: 
-$w_i(\theta) \sim GP(0, \sigma^2R)$, with
+$w_i(\psi) \sim GP(0, \sigma^2R)$, with
 
 \begin{equation*}
-    R_{ij} = \prod_{k=1}^{p_\theta}\exp\left\{-10^{\beta_k}|x_{ik}-x_{jk}|^\alpha\right\}
+    R_{ij} = \prod_{k=1}^{p_\psi}\exp\left\{-10^{\beta_k}|x_{ik}-x_{jk}|^\alpha\right\}
 \end{equation*}
 
-\textcolor{red}{We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\theta\}$. 
+\textcolor{red}{We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\psi\}$. 
 The above with $R_{ij}$ is based on GPfit; will update this to reflect SEPIA settings later.}
@@
 factorial design to obtain estimated main effects for each cosmological parameter in 
-$\theta$, interactions amongst the $p_\theta=8$ parameters, and decompose the overall 
+$\psi$, interactions amongst the $p_\psi=8$ parameters, and decompose the overall 
 variation by main effects, two-way interactions, and so on. Should this be included?}
@@
 
-\section{Appendix}
-\label{sec:apdx}
+\appendix
+\renewcommand{\thesection}{Appendix \Alph{section}}
+
+\section{Integrated Likelihood}
+\label{sec:apdx_int_lik}
+Here, we establish that under the model:
+
+$$
+\begin{aligned}
+Y_i|S,W &\sim N\big(S, \Sigma_\varepsilon\big), \quad i \in \{1,\dots, r\}\\
+S|W &\sim N\big(\boldsymbol{\mu}, \Sigma_S(w,w')\big)\\
+W &\sim N\big(\mathbf{0}, \Sigma_W(x,x')\big) \\
+\end{aligned}
+$$
+
+We can obtain the integrated likelihood $p(Y_i|W)=\int p(Y_i|S,W)p(S|W) dS$, 
+with the following calculations:
+$$
+\begin{aligned}
+p(Y_i|W) &= \int p(Y_i|S,W)p(S|W)dS \\
+&\propto \int\exp\left(-\frac{1}{2} \left[Y_i^T\Sigma_\varepsilon^{-1} Y_i 
+- 2Y_i^T\Sigma_\varepsilon^{-1} S + S^T \Sigma_\varepsilon^{-1} S + S^T \Sigma_S^{-1} S 
+-2S^T \Sigma_S^{-1} \boldsymbol{\mu} 
++ \boldsymbol{\mu}^T \Sigma_S^{-1} \boldsymbol{\mu} \right]\right)dS \\
+\end{aligned}
+$$
+
+If we focus only on terms containing $S$, we have 
+$$
+\begin{aligned}
+    \exp\Big(-\frac{1}{2} \big[S^T &(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1}) S 
+    -2S^T (\Sigma_S^{-1} \boldsymbol{\mu} + \Sigma_\varepsilon^{-1} Y_i) \big]\Big) \\
+&= \exp\left(-\frac{1}{2} \left[S^T (\Sigma_\varepsilon^{-1}+\Sigma_S^{-1}) S 
+-2S^T (\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})(\Sigma_\varepsilon^{-1}
++\Sigma_S^{-1})^{-1}(\Sigma_S^{-1} \boldsymbol{\mu} + \Sigma_\varepsilon^{-1} Y_i) \right]\right)
+\end{aligned}
+$$
+
+To complete the square, we add and subtract $(\Sigma_S^{-1} \boldsymbol{\mu} 
++ \Sigma_\varepsilon^{-1} Y_i)^T(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}
+(\Sigma_S^{-1} \boldsymbol{\mu} + \Sigma_\varepsilon^{-1} Y_i)$, 
+which then gives the kernel for $S \sim N\left((\Sigma_\varepsilon^{-1}
++\Sigma_S^{-1})^{-1}(\Sigma_S^{-1} \boldsymbol{\mu} + \Sigma_\varepsilon^{-1} Y_i), 
+(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1} \right)$. 
+This removes the integral and leaves us with the following terms for $Y_i|W$:
+$$
+\begin{aligned}
+&\exp\Big(-\frac{1}{2}\big[Y_i^T\Sigma_\varepsilon^{-1} Y + 
+\boldsymbol{\mu}^T\Sigma_S^{-1} \boldsymbol{\mu} - (\Sigma_S^{-1} \boldsymbol{\mu} + 
+\Sigma_\varepsilon^{-1} Y_i)^T(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1} 
+(\Sigma_S^{-1} \boldsymbol{\mu} + \Sigma_\varepsilon^{-1} Y_i)\big]\Big) \\
+&= \exp\Big(-\frac{1}{2}\big[Y_i^T\big(\Sigma_\varepsilon^{-1} - 
+\Sigma_\varepsilon^{-1}(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}\Sigma_\varepsilon^{-1}\big)Y_i 
+-2Y_i^T \Sigma_\varepsilon^{-1}(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}\Sigma_S^{-1} \boldsymbol{\mu} \\
+&\qquad \qquad \qquad + \boldsymbol{\mu}^T (\Sigma_S^{-1} - \Sigma_S^{-1}(\Sigma_\varepsilon^{-1}
++\Sigma_S^{-1})^{-1}\Sigma_S^{-1}) \boldsymbol{\mu}\big]\Big) \\
+&= \exp\Big(-\frac{1}{2}\big((Y_i-\boldsymbol{\mu})^T(\Sigma_\varepsilon+\Sigma_S)^{-1}
+(Y_i-\boldsymbol{\mu}) \big)\Big)
+\end{aligned}
+$$
+
+With the last equality resulting from the use of the Sherman–Morrison–Woodbury formula, 
+yielding $\big(\Sigma_\varepsilon^{-1} - \Sigma_\varepsilon^{-1}(\Sigma_\varepsilon^{-1}+
+\Sigma_S^{-1})^{-1}\Sigma_\varepsilon^{-1} = (\Sigma_S^{-1} - \Sigma_S^{-1}
+(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})^{-1}\Sigma_S^{-1}) = (\Sigma_\varepsilon+\Sigma_S)^{-1}$. 
+We also use the fact that $\Sigma_\varepsilon^{-1}(\Sigma_\varepsilon^{-1}+
+\Sigma_S^{-1})^{-1}\Sigma_S^{-1} = (\Sigma_\varepsilon + \Sigma_S)^{-1}$, 
+which holds since it can be shown that 
+$[\Sigma_S(\Sigma_\varepsilon^{-1}+\Sigma_S^{-1})\Sigma_\varepsilon](\Sigma_\varepsilon+\Sigma_S)
+=(\Sigma_\varepsilon+\Sigma_S)(\Sigma_\varepsilon+\Sigma_S)$.
+
+Therefore, we have that $Y_i|W \sim N\big(\boldsymbol{\mu}, \Sigma_S(w,w')+\Sigma_\varepsilon(w,w')\big)$.
+
+\section{Conditional Distribution of Power Spectrum, Given Weighted Average}
+\label{sec:apdx_SgivenY}
+
+From our model specification, we have the following:
+
+\[
+S \sim \mathcal{N}(\boldsymbol{\mu}, \Sigma_S)
+\]
+\[
+\bar{Y} \mid S \sim \mathcal{N}(S, \Sigma_\epsilon)
+\]
+
+Then, to find the distribution of $S|\bar{Y}$, we apply Bayes' Theorem below, 
+where $\propto$ indicates proportionality:
+\[
+P(S \mid \bar{Y}) \propto P(\bar{Y} \mid S) P(S)
+\]
+
+\[
+\propto \exp\left(-\tfrac{1}{2} (\bar{Y} - S)^T \Sigma_\epsilon^{-1} (\bar{Y} - S) \right) 
+\cdot \exp\left(-\tfrac{1}{2} (S - \boldsymbol{\mu})^T \Sigma_S^{-1} (S - \boldsymbol{\mu})\right)
+\]
+
+\[
+\propto \exp\left(-\tfrac{1}{2} \left[ \bar{Y}^T \Sigma_\epsilon^{-1} \bar{Y} 
+- 2 S^T \Sigma_\epsilon^{-1} \bar{Y} + S^T \Sigma_\epsilon^{-1} S + S^T \Sigma_S^{-1} S 
+- 2 S^T \Sigma_S^{-1} \boldsymbol{\mu} + \boldsymbol{\mu}^T \Sigma_S^{-1} \boldsymbol{\mu} \right] \right)
+\]
+
+\[
+\propto \exp\left( -\tfrac{1}{2} \left[ S^T (\Sigma_\epsilon^{-1} + \Sigma_S^{-1}) S 
+- 2 S^T (\Sigma_\epsilon^{-1} \bar{Y} + \Sigma_S^{-1} \boldsymbol{\mu}) \right] \right)
+\]
+
+From properties of the multivariate normal distribution \citep[e.g., Equation 7.1 of][]{hoff2009first},
+we can establish that the mean and covariance of $S|\bar{Y}$ will be $\boldsymbol{m}$
+and $C$ respectively, where we can solve for these values with:
+
+\[
+C^{-1} = \Sigma_\epsilon^{-1} + \Sigma_S^{-1}
+\]
+\[
+C^{-1} \boldsymbol{m} = \Sigma_\epsilon^{-1} \bar{Y} + \Sigma_S^{-1} \boldsymbol{\mu}
+\]
+
+Then,
+\[
+C = (\Sigma_\epsilon^{-1} + \Sigma_S^{-1})^{-1}
+\]
+\[
+\boldsymbol{m} = \left( \Sigma_S^{-1} + \Sigma_\epsilon^{-1} \right)^{-1} 
+\left( \Sigma_\epsilon^{-1} \bar{Y} + \Sigma_S^{-1} \boldsymbol{\mu} \right)
+= C \cdot \left( \Sigma_\epsilon^{-1} \bar{Y} + \Sigma_S^{-1} \boldsymbol{\mu} \right)
+\]
+
+Therefore, we have the result that $S|\bar{Y} \sim \mathcal{N}(\boldsymbol{m}, C)$.
+
+\section{Mira-Titan Predictive Plots}
+\label{sec:apdx_mira_plots}
 

### 2025-06-06 sims/collect_sims.R
include newest sim setting

@@ -20,7 +20,7 @@ MSE5_range <- MSE20_range <- logS5_range <- logS20_range <- c()
 
 r=5
 for (func in 1:2) {
-  for (setting in 4:6) {
+  for (setting in 4:7) {
     filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
     # Compare MSEs
     results <- read.csv(filename)[, c(3,6,5,4)]

### 2025-06-06 paper/main.tex
update r options to 5 and 15

@@
 For each simulated function and variance specification, we generate 20 values of 
-$m_1, m_2, f_1, f_2$, and conditional on these values, we sample either 5 or 20 
+$m_1, m_2, f_1, f_2$, and conditional on these values, we sample either 5 or 15 
 functional observations and estimate the underlying true function based on the observations.
@@
              specification pair, and each row shows results for batch sizes of either 
-             5 (top) or 20 (bottom).}
+             5 (top) or 15 (bottom).}
     \label{fig:sims_logS}
@@
              specification pair, and each row shows results for batch sizes of 
-             either 5 (top) or 20 (bottom).}    
+             either 5 (top) or 15 (bottom).}    
     \label{fig:sims_MSE}

### 2025-06-08 paper/main.tex
some edits to Section 3

@@
 of the output for a particular cosmology. When we focus on $k$ values where $y_{\ell_r}$ 
-is valid (center panel), we see each low-resolution and the high-resolution run oscillate around 
-a mean term, which is relatively smooth, compared to these realizations. When this
+is valid (center panel), we see both low-resolution and high-resolution runs oscillate around 
+a mean term, which is relatively smooth compared to these realizations. When this
 mean term is subtracted, we will obtain a plot like the right panel in 
@@
 the observed weighted average; now, we use $\bar{Y}$ to indicate the corresponding 
-random variable within the statistical model.  $\bar{Y}$ is typically pre-processed
-in a domain-specific manner, such as that defined in Section \ref{sec:data}.
-Together, $X$ and $\bar{Y}$ comprise the training data for a particular cosmology.
-We will reserve specific integration 
-of expert knowledge for the CAMB and Mira-Titan suites, such as this pre-processing,
+random variable within the statistical model.  Together, $X$ and $\bar{Y}$ comprise the 
+training data for a particular cosmology. $\bar{Y}$ is typically pre-processed
+based on domain expertise, such as that defined in Section \ref{sec:data}.
+We will reserve details of domain-specific pre-processing for the CAMB and Mira-Titan suites
 for Sections \ref{subsec:camb} and \ref{subsec:mira_fit}, respectively.
@@
 
-Let $S$ represent the true (but unknown) underlying power matter spectra for
+Let $S$ represent the true (but unknown) underlying power matter spectrum for
 a particular cosmology.  We consider $\bar{Y}$ a noisy realization of $S$.
 Specifically, we assume $\bar{Y}$ is a random realization of a Gaussian 
-process centered at $S$ with some covariance, i.e. 
+process centered at $S$ with some covariance, i.e., 
 $\bar{Y}\mid S \sim \mathrm{GP}(S, \Sigma_\varepsilon)$.  While previous works
@@
 tailored to each scenario (more on this in later subsections).
+We consider $\Sigma_\epsilon$ as known, given domain-matter expertise. (right)?
+\textcolor{red}{(Do we actually tailor this?  I thought it was consistent across the board.)}
 
 We next place a Gaussian process prior on $S$, i.e.,
-$S\sim \mathrm{GP}\left(\boldsymbol{\mu}, \Sigma_S(\cdot)\right)$, again with a 
-spatial kernel. Here, we set $\boldsymbol{\mu}=\mathbf{0}$, although the modeling can 
-be carried out with any selection of $\boldsymbol{\mu}$. Standard practice would 
-be to use locations $X$ as the inputs to kernel $\Sigma_S(\cdot)$.
-Yet we have found that prior to lack sufficient flexibility for our motivating
+$S\sim \mathrm{GP}\left(\boldsymbol{\mu}_S, \Sigma_S(\cdot)\right)$, again with a 
+Mat\'ern kernel. Here, we set $\boldsymbol{\mu}_S=\mathbf{0}$, although our model
+is applicable to any selection of $\boldsymbol{\mu}_S$. Standard practice would 
+use locations $X$ as the inputs to kernel $\Sigma_S(\cdot)$, yet we have found that 
+prior to lack sufficient flexibility for our motivating
 application.  In the Mira-Titan setting, some nonstationarity is expected due to 
@@
 for this, we incorporate an additional latent Gaussian process, $W$, which accommodates
-nonstationarity by warping $X$.  Ultimately, this yields the following hierarchical model:
+nonstationarity by warping $X$ via $W \sim \mathrm{GP}\left(\boldsymbol{\mu}_W, \Sigma_W(X)\right)$.
+$W$ is also an $n$-vector whose entries $w_i$ correspond to warped versions of each $x_i$.
+Again, we set $\boldsymbol{\mu}_W = \mathbf{0}$ without loss of generatlity (we have
+found $\boldsymbol{\mu}_W = X$ works similarly).
+Ultimately, this yields the following hierarchical model:
 \begin{equation}\label{eq:dgphm}
@@
 \bar{Y}\mid S,\Sigma_\varepsilon &\sim \mathrm{GP}(S,\Sigma_\varepsilon) \\
-S\mid W &\sim \mathrm{GP}\left(\boldsymbol{\mu}, \Sigma_S(W)\right) \\
-W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_W(X)\right)
+S\mid W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(W)\right) \\
+W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_W(X)\right).
 \end{aligned}
 \end{equation}
-$W$ is also an $n$-vector whose entries $w_i$ correspond to warped versions of each $x_i$.
 The functional composition of GPs in the latter two lines forms a ``deep Gaussian process'' 
@@
 unique lengthscale $\theta_W$ but enforce unit scale following \citet{sauer2023active}.
-We considered setting the prior mean of $W$ to $X$ instead of $\mathbf{0}$
-\citep[which would indicate stationarity apriori;][]{schmidt2003bayesian}, but found
-it had a neglible effect on the model fits.
 
@@
 
-While our inferential goal is to obtain the posterior distribution of 
-$S\mid\bar{Y},\Sigma_\varepsilon$, this requires inferring latent quantity $W$.  
-We first integrate over $S$ to condense our hierarchical model of Eq.~\ref{eq:dgphm} into
+Our inferential goal is to obtain the posterior distribution of 
+$S\mid\bar{Y},\Sigma_\varepsilon$, but this requires inferring latent quantity $W$.  
+We first integrate over $S$ to condense our hierarchical model of Eq.~(\ref{eq:dgphm}) into
 \begin{align}
 \label{eq:likelihood}
-\bar{Y} \mid W, \Sigma_\varepsilon &\sim \textrm{GP}(\boldsymbol{\mu}, \Sigma_S(W) + \Sigma_\varepsilon) \\
+\bar{Y} \mid W, \Sigma_\varepsilon &\sim \textrm{GP}(\mathbf{0}, \Sigma_S(W) + \Sigma_\varepsilon) \\
 \label{eq:prior}
@@
 \end{align}
-Details are provided in \ref{sec:apdx_int_lik}. We then use elliptical slice
-sampling \citep[ESS;][]{murray2010elliptical} to generate posterior samples of $W$.  
-Proposals are generated from prior samples from Eq.~(\ref{eq:prior}), and acceptances 
-involve the ratio of likelihoods from Eq.~\ref{eq:likelihood}.
+Details are provided in \ref{sec:apdx_int_lik}. The incorporation of $\Sigma_\epsilon$ in the
+Gaussian likelihood of this outer layer is an essential upgrade to previous DGP inferential schemes.
+This covariance incorporates knowledge of the smoothly varying dynamics from which $\bar{Y}$ was
+generated.
 
-We also embrace a fully-Bayesian treatment of unknown kernel hyperparameters 
-$\{\tau^2, \theta_S, \theta_W\}$.  We prescale $X$ to $[0, 1]$
-and $\bar{Y}$ to zero mean and unit variance.  On this scale, we adopt the default 
-prior specifications provided in the {\tt deepgp} {\sf R}-package \citep{deepgp}.
-\textcolor{red}{(@Annie - from our email 5/29, can you comment on how we handle $\tau^2$/
-cite Bobby's book/etc? Thanks!)}
-%\textcolor{red}{Annie, please ensure the most up-to-date 
-%prior specs are shown above and here. Here is what I found from fit\$settings when using 
-%deepgp v1.1.3: $\theta_S \sim G(\alpha=1.2, \beta=1)$ and $\tau_S \propto 1/\tau_S^2$.}
-We integrate Metropolis-Hastings sampling of these parameters in a Gibbs framework with ESS
-sampling of $W$ following the scheme of \citep{sauer2023active}.
+We embrace fully-Bayesian inference of latent $W$ and kernel hyperparameters $\{\tau^2, \theta_S, \theta_W\}$.
+Specifically, we use elliptical slice sampling \citep[ESS;][]{murray2010elliptical} to 
+generate posterior samples of $W$, using Eq.~(\ref{eq:prior}) for proposal generation and 
+Eq.~(\ref{eq:likelihood}) for likelihood-based acceptance.  
+Under a reference prior $\pi(\tau^2)\propto\frac{1}{\tau^2}$,
+we integrate $\tau^2$ from the likelihood of Eq.~(\ref{eq:likelihood}).  The resulting marginal 
+posterior is equivalent to that with the plug-in estimator 
+$\hat{\tau}^2 = \dots$ \citep[][Chapter 5]{gramacy2020surrogates}.
+\textcolor{red}{(With $\Sigma_\epsilon$ on there, we can't actually do this.. unless we divide
+$\Sigma_\epsilon$ by $\tau^2$...)}.
+For $\theta_S$ and $\theta_W$, after pre-scaling, we adopt the default prior specifications 
+provided in the {\tt deepgp} {\sf R}-package \citep{deepgp}.
+We then integrate Metropolis-Hastings sampling of these parameters in a Gibbs framework 
+with the ESS sampling of $W$ \citep{sauer2023active}.
 
@@
 These results come from Bayes' Theorem, with details of this derivation provided 
-in \ref{sec:apdx_SgivenY}. In our application, we find success setting 
-$\boldsymbol{\mu}=\boldsymbol{0}$. With all posterior draws $S^{(t)}$, we compute
+in \ref{sec:apdx_SgivenY}. With all posterior draws $S^{(t)}$, we compute
 the average of these to obtain our posterior mean for a particular cosmology.
@@
 structure to incorporate multiple functional outputs.
+\textcolor{red}{(How do you get UQ though?)}
 

### 2025-06-10 dgp.hm/R/fit.R
chg outer and calc_tau2 to FALSE

@@ -98,8 +98,8 @@ gibbs_one_layer_hm <- function(x, y, nmcmc, verb, initial, settings, v,
     # Sample lengthscale (theta), also get MLE estimate of tau2
     samp <- sample_theta_hm(y, dx, theta[j - 1], alpha = settings$alpha$theta, 
                             beta = settings$beta$theta, l = settings$l, 
-                            u = settings$u, outer = TRUE, ll_prev = ll, v = v, 
-                            calc_tau2 = TRUE, Sigma_hat = Sigma_hat)
+                            u = settings$u, outer = FALSE, ll_prev = ll, v = v, 
+                            calc_tau2 = FALSE, Sigma_hat = Sigma_hat)
     theta[j] <- samp$theta
     ll <- samp$ll
     ll_store[j] <- ll
@@ -134,8 +134,8 @@ gibbs_two_layer_hm <- function(x, y, nmcmc, D, verb, initial, settings, v,
     samp <- sample_theta_hm(y, dw, theta_y[j - 1], 
                             alpha = settings$alpha$theta_y, 
                             beta = settings$beta$theta_y, l = settings$l, 
-                            u = settings$u, outer = TRUE, ll_prev = ll_outer, 
-                            v = v, calc_tau2 = TRUE, Sigma_hat = Sigma_hat)
+                            u = settings$u, outer = FALSE, ll_prev = ll_outer, 
+                            v = v, calc_tau2 = FALSE, Sigma_hat = Sigma_hat)
     theta_y[j] <- samp$theta
     ll_outer <- samp$ll
     if (is.null(samp$tau2)) tau2[j] <- tau2[j - 1] else tau2[j] <- samp$tau2

### 2025-06-10 dgp.hm/R/fit.R
REMOVE TAU2 option, force at 1 everywhere

@@ -20,7 +20,7 @@
 #' @export
 
 fit_one_layer_hm <- function(x, y, nmcmc = 10000, verb = TRUE, theta_0 = 0.1, 
-                             settings = NULL, v = 2.5, Sigma_hat = NULL) {
+                             settings = NULL, v = 2.5, Sigma_hat) {
   tic <- proc.time()[[3]]
   if (nmcmc <= 1) stop("nmcmc must be greater than 1")
   
@@ -28,7 +28,7 @@ fit_one_layer_hm <- function(x, y, nmcmc = 10000, verb = TRUE, theta_0 = 0.1,
   if (is.numeric(x)) x <- as.matrix(x)
   test <- deepgp:::check_inputs(x, y, 0) # do not need to check nugget
   settings <- deepgp:::check_settings(settings, layers = 1, noisy = TRUE) # TRUE/FALSE??
-  initial <- list(theta = theta_0, tau2 = 1)
+  initial <- list(theta = theta_0) # force tau2 = 1
   if (!(v %in% c(0.5, 1.5, 2.5))) 
     stop("v must be one of 0.5, 1.5, or 2.5")
   if (nrow(Sigma_hat) != nrow(x))
@@ -52,13 +52,13 @@ fit_one_layer_hm <- function(x, y, nmcmc = 10000, verb = TRUE, theta_0 = 0.1,
 fit_two_layer_hm <- function (x, y, D = ifelse(is.matrix(x), ncol(x), 1), 
                               nmcmc = 10000, verb = TRUE, w_0 = NULL, 
                               theta_y_0 = 0.1, theta_w_0 = 0.1, settings = NULL, 
-                              v = 2.5, Sigma_hat = NULL) {
+                              v = 2.5, Sigma_hat) {
   
   tic <- proc.time()[[3]]
   if (is.numeric(x)) x <- as.matrix(x)
   test <- deepgp:::check_inputs(x, y, 0) # do not need to check nugget
   settings <- deepgp:::check_settings(settings, layers = 2, noisy = TRUE) # TRUE/FALSE??
-  initial <- list(w = w_0, theta_y = theta_y_0, theta_w = theta_w_0, tau2 = 1)
+  initial <- list(w = w_0, theta_y = theta_y_0, theta_w = theta_w_0) # force tau2 = 1
   initial <- deepgp:::check_initialization(initial, layers = 2, x = x, D = D)
   if (!(v %in% c(0.5, 1.5, 2.5))) 
     stop("v must be one of 0.5, 1.5, or 2.5")
[…]

### 2025-06-10 dgp.hm/R/trim.R

@@ -25,7 +25,6 @@ trim.gphm <- function(object, burn, thin = 1) {
   
   object$nmcmc <- length(indx)
   object$theta <- object$theta[indx, drop = FALSE]
-  object$tau2 <- object$tau2[indx, drop = FALSE]
   object$ll <- object$ll[indx, drop = FALSE]
   
   toc <- proc.time()[[3]]
@@ -52,7 +51,6 @@ trim.dgp2hm <- function(object, burn, thin = 1) {
   object$theta_y <- object$theta_y[indx, drop = FALSE]
   object$theta_w <- object$theta_w[indx, , drop = FALSE]
   object$w <- as.list(object$w[indx])
-  object$tau2 <- object$tau2[indx, drop = FALSE]
   object$ll <- object$ll[indx, drop = FALSE]
   
   toc <- proc.time()[[3]]

### 2025-06-10 fitting/run_CAMB.R
smooth precs for weighted avg

@@ -108,16 +108,16 @@ for(model in 1:32){ # integer 1-32
   # Get indices for where each data product is deemed unbiased
   lo_ind <- which(x < log10(x_ub) & x >= min(x_lo))
   hi_ind <- which(x < log10(x_ub) & x >= min(x_lo))
-  camb_ind <- which(x < min(x_lo))
+  camb_ind <- which(x <= min(x_lo))
   hi_only <- hi_ind[which(!(hi_ind %in% lo_ind))]
   
   # Smooth the precision information (to remove steps)
-  prec_lo = rollmean(ifelse(1:length(x) %in% lo_ind, prec, 0),
-                     k = 2, fill = "extend")
-  prec_hi = rollmean(ifelse(1:length(x) %in% hi_ind, prec, 0) * prec_adj,
-                     k = 2, fill = "extend")
-  prec_camb = rollmean(ifelse(1:length(x) %in% camb_ind, 1e6, 0),
-                       k = 2, fill = "extend")
+  prec_lo = smooth.spline(rollmean(ifelse(1:length(x) %in% lo_ind, prec, 0),
+                     k = 3, fill = "extend"))$y
+  prec_hi = smooth.spline(rollmean(ifelse(1:length(x) %in% hi_ind, prec, 0) * prec_adj,
+                     k = 3, fill = "extend"))$y
+  prec_camb = smooth.spline(rollmean(ifelse(1:length(x) %in% camb_ind, 1e6, 0),
+                       k = 3, fill = "extend"))$y
   prec_avg = (prec_lo*n_lo+prec_hi+prec_camb)
   
   # Scale inputs ----------------------------------------------------------------

### 2025-06-11 fitting/run_CAMB.R
smooth precisions for weighted average

@@ -101,8 +101,8 @@ for(model in 1:32){ # integer 1-32
   var_lo <- apply(y_lo, 1, var)
   lm_var <- lm(log(var_lo) ~ x_lo)# + I(x_lo^2))
 
-  # Combine 1e6 prec (for camb) with log-log precs for lo and hi
-  prec = c(rep(1e6, n_camb_only), as.numeric(1/exp(lm_var$fitted.values)))
+  # Combine 1e8 prec (for camb) with log-log precs for lo and hi
+  prec = c(rep(1e8, n_camb_only), as.numeric(1/exp(lm_var$fitted.values)))
   prec_adj = 3.725626
   
   # Get indices for where each data product is deemed unbiased
@@ -112,13 +112,13 @@ for(model in 1:32){ # integer 1-32
   hi_only <- hi_ind[which(!(hi_ind %in% lo_ind))]
   
   # Smooth the precision information (to remove steps)
-  prec_lo = smooth.spline(rollmean(ifelse(1:length(x) %in% lo_ind, prec, 0),
-                     k = 3, fill = "extend"))$y
-  prec_hi = smooth.spline(rollmean(ifelse(1:length(x) %in% hi_ind, prec, 0) * prec_adj,
-                     k = 3, fill = "extend"))$y
-  prec_camb = smooth.spline(rollmean(ifelse(1:length(x) %in% camb_ind, 1e6, 0),
-                       k = 3, fill = "extend"))$y
-  prec_avg = (prec_lo*n_lo+prec_hi+prec_camb)
+  prec_lo = pmax(0, smooth.spline(rollmean(ifelse(1:length(x) %in% lo_ind, prec, 0),
+                                           k = 3, fill = "extend"))$y)
+  prec_hi = pmax(0, smooth.spline(rollmean(ifelse(1:length(x) %in% hi_ind, prec, 0) * prec_adj,
+                                           k = 3, fill = "extend"))$y)
+  prec_camb = pmax(0, smooth.spline(rollmean(ifelse(1:length(x) %in% camb_ind, 1e8, 0),
+                                             k = 3, fill = "extend"))$y)
+  prec_avg = (prec_lo*n_lo + prec_hi + prec_camb)
   
   # Scale inputs ----------------------------------------------------------------
   
@@ -167,7 +167,7 @@ for(model in 1:32){ # integer 1-32
   #                               params$theta_hat, 1e-8, 2.5)
   # 
[…]

### 2025-06-11 dgp.hm/R/fit.R
update fit for only 1D, monotonic warping

@@ -19,13 +19,15 @@
 # fit_one_layer_hm ------------------------------------------------------------
 #' @export
 
-fit_one_layer_hm <- function(x, y, nmcmc = 10000, verb = TRUE, theta_0 = 0.1, 
-                             settings = NULL, v = 2.5, Sigma_hat) {
+fit_one_layer_hm <- function(x, y, Sigma_hat, nmcmc = 10000, verb = TRUE, theta_0 = 0.1, 
+                             settings = NULL, v = 2.5) {
   tic <- proc.time()[[3]]
   if (nmcmc <= 1) stop("nmcmc must be greater than 1")
   
   # Check inputs
-  if (is.numeric(x)) x <- as.matrix(x)
+  if (!is.matrix(x)) {
+    x <- matrix(x, ncol = 1)
+  } else if (ncol(x) != 1) stop("x must be one-dimensional")
   test <- deepgp:::check_inputs(x, y, 0) # do not need to check nugget
   settings <- deepgp:::check_settings(settings, layers = 1, noisy = TRUE) # TRUE/FALSE??
   initial <- list(theta = theta_0) # force tau2 = 1
@@ -49,26 +51,29 @@ fit_one_layer_hm <- function(x, y, nmcmc = 10000, verb = TRUE, theta_0 = 0.1,
 # fit_two_layer_hm ------------------------------------------------------------
 #' @export
 
-fit_two_layer_hm <- function (x, y, D = ifelse(is.matrix(x), ncol(x), 1), 
-                              nmcmc = 10000, verb = TRUE, w_0 = NULL, 
-                              theta_y_0 = 0.1, theta_w_0 = 0.1, settings = NULL, 
-                              v = 2.5, Sigma_hat) {
+fit_two_layer_hm <- function (x, y, Sigma_hat, nmcmc = 10000, verb = TRUE, 
+                              w_0 = NULL, theta_y_0 = 0.1, theta_w_0 = 0.1, 
+                              settings = NULL, v = 2.5, x_grid = NULL) {
   
   tic <- proc.time()[[3]]
-  if (is.numeric(x)) x <- as.matrix(x)
+  if (!is.matrix(x)) {
+    x <- matrix(x, ncol = 1)
+  } else if (ncol(x) != 1) stop("x must be one-dimensional")
[…]

### 2025-06-11 dgp.hm/R/misc.R

@@ -10,8 +10,7 @@
 # Optimize Matern -------------------------------------------------------------
 #' @export
 
-opt_matern <- function (dx, y, sdd, init = c(0.1, 0.1), n_multi = 100) 
-{
+opt_matern <- function (dx, y, sdd, init = c(0.1, 0.1), n_multi = 100) {
   out <- optim(init, nl_matern, dx = dx, y = y, sdd = sdd)
   best_par <- out$par
   best_ll <- out$value
@@ -83,4 +82,29 @@ moore_penrose <- function(H, tolpower = -10) {
   inverse.values[i] <- 1 / H.eigen$values[i]
   H.MP <- H.eigen$vectors %*% diag(inverse.values) %*% t(H.eigen$vectors)
   return(H.MP)
+}
+
+# monowarp_ref ----------------------------------------------------------------
+
+monowarp_ref <- function(x, xg, wg, index) {
+  # x: matrix of input locations for returned w
+  # xg: matrix of grid locations
+  # wg: UNWARPED w values at xg locations
+  # Returns warped values of w at x locations
+  if (!is.matrix(x)) x <- matrix(x, ncol = 1)
+  if (!is.matrix(xg)) xg <- matrix(xg, ncol = 1)
+  if (!is.matrix(wg)) wg <- matrix(wg, ncol = 1)
+  if (!is.matrix(index)) index <- matrix(index, ncol = 1)
+  
+  w <- matrix(nrow = nrow(x), ncol = ncol(wg))
+  for (i in 1:ncol(wg)) {
+    r_orig <- range(wg[, i])
+    wg[, i] <- exp(wg[, i])
+    wg[, i] <- cumsum(wg[, i])
+    r_new <- range(wg[, i])
+    wg[, i] <- (wg[, i] - r_new[1])/(r_new[2] - r_new[1]) # first scale to [0, 1]
+    wg[, i] <- wg[, i]*(r_orig[2] - r_orig[1]) + r_orig[1] # then scale to original range
+    w[, i] <- fo_approx(xg[, i], wg[, i], x[, i], index[, i])
+  }
+  return(w)
 }

### 2025-06-11 dgp.hm/R/fit.R
finish monowarp updates

@@ -12,9 +12,11 @@
 #
 # Updates from deepgp ---------------------------------------------------------
 #   Sigma_hat is added to Sigma(X) in outer likelihood
+#   one dimension only
+#   Matern kernel only
+#   scale tau2 is fixed at 1
 #   No longer a nugget option (g = 1e-8 used for numerical stability)
 #   No longer a prior_mean of x option
-#   Matern kernel only
 
 # fit_one_layer_hm ------------------------------------------------------------
 #' @export

### 2025-06-11 dgp.hm/R/plot.R

@@ -8,7 +8,7 @@ plot.gphm <- function(x) {
   oldpar <- par(no.readonly = TRUE)
   on.exit(par(oldpar))
   
-  Dx <- ncol(x$x)
+  if (ncol(x$x) != 1) stop("ncol(x) must equal 1")
   
   # Trace plots
   par(mfrow = c(1, 2), mar = c(5, 4, 2, 2))
@@ -29,18 +29,16 @@ plot.dgp2hm <- function(x) {
   oldpar <- par(no.readonly = TRUE)
   on.exit(par(oldpar))
   
-  Dx <- ncol(x$x)
-  D <- ncol(x$w[[1]])
+  if (ncol(x$x) != 1) stop("ncol(x) must equal 1")
   
   # Trace plots
-  par(mfrow = c(1, D + 2), mar = c(5, 4, 2, 2))
+  par(mfrow = c(1, 3), mar = c(5, 4, 2, 2))
   plot(x$ll, type = 'l', ylab = 'outer logl', xlab = 'Iteration',
        main = 'Trace Plot of outer logl')
   plot(x$theta_y, type = 'l', ylab = 'theta_y', xlab = 'Iteration',
        main = 'Trace Plot of theta_y')
-  for (i in 1:D)
-    plot(x$theta_w[, i], type = 'l', ylab = 'theta_w', xlab = 'Iteration',
-         main = paste0('Trace Plot of theta_w [', i, ']'))
+  plot(x$theta_w, type = 'l', ylab = 'theta_w', xlab = 'Iteration',
+         main = paste0('Trace Plot of theta_w'))
   
   # Hidden layer plot
   indx <- floor(seq(from = 1, to = x$nmcmc, length = 100))
@@ -48,11 +46,11 @@ plot.dgp2hm <- function(x) {
   col <- heat.colors(100 + 10) # add ten to avoid using colors that are too light
   par(mfrow = c(1, 1), mar = c(4, 4, 2, 2))
   o <- order(x$x)
-  plot(x$x[o], x$w[[indx[1]]][o] - mean(x$w[[indx[1]]]), type = 'l', xlab = 'X', 
+  plot(x$x[o], x$w[o, indx[1]] - mean(x$w[, indx[1]]), type = 'l', xlab = 'X', 
        ylab = 'W', col = col[1], main = paste0('MCMC samples of X to W'), 
[…]

### 2025-06-11 dgp.hm/R/predict.R

@@ -16,7 +16,11 @@ est_true <- function(fit, tolpower = -10, return_all = FALSE) {
   deep <- (class(fit) == "dgp2hm")
   
   # get squared distance matrix
-  if (!deep) d2 <- deepgp::sq_dist(fit$x)
+  if (deep) { # pre-calculate grid indices
+    grid_index <- deepgp:::fo_approx_init(fit$x_grid, fit$x)
+  } else { # pre-calculate pairwise distances for x
+    d2 <- deepgp::sq_dist(fit$x)
+  }
   
   # From Section 3.3.2: Bayesian UQ Algorithm (Walsh dissertation 2023)
   # Get Sigma_epsilon (covariance matrix for correlated errors) and its inverse
@@ -38,7 +42,8 @@ est_true <- function(fit, tolpower = -10, return_all = FALSE) {
     # construct the covariance matrix for the true spectrum (Sigma_S)
     # cov mtx of [S]; based on a matern covariance function
     if (deep) {
-      d2 <- deepgp:::sq_dist(fit$w[[i]])
+      w <- monowarp_ref(fit$x, fit$x_grid, fit$w_grid[, i], grid_index)
+      d2 <- deepgp:::sq_dist(w)
       theta <- fit$theta_y[i]
     } else theta <- fit$theta[i]
     S <- deepgp:::Matern(d2, 1, theta, 1e-8, fit$v)

### 2025-06-11 dgp.hm/R/trim.R

@@ -49,8 +49,8 @@ trim.dgp2hm <- function(object, burn, thin = 1) {
   
   object$nmcmc <- length(indx)
   object$theta_y <- object$theta_y[indx, drop = FALSE]
-  object$theta_w <- object$theta_w[indx, , drop = FALSE]
-  object$w <- as.list(object$w[indx])
+  object$theta_w <- object$theta_w[indx, drop = FALSE]
+  object$w <- object$w[, indx, drop = FALSE]
   object$ll <- object$ll[indx, drop = FALSE]
   
   toc <- proc.time()[[3]]

### 2025-06-11 dgp.hm/R/fit.R
fixing checks

@@ -64,7 +64,7 @@ fit_two_layer_hm <- function (x, y, Sigma_hat, nmcmc = 10000, verb = TRUE,
   test <- deepgp:::check_inputs(x, y, 0) # do not need to check nugget
   settings <- deepgp:::check_settings(settings, layers = 2, noisy = TRUE) # TRUE/FALSE??
   initial <- list(w = w_0, theta_y = theta_y_0, theta_w = theta_w_0) # force tau2 = 1
-  initial <- deepgp:::check_initialization(initial, layers = 2, x = x, D = D)
+  initial <- deepgp:::check_initialization(initial, layers = 2, x = x, D = 1)
   if (is.null(x_grid)) 
     x_grid <- matrix(seq(0, 1, length = 50), ncol = 1)
   if (!(v %in% c(0.5, 1.5, 2.5))) 
@@ -223,7 +223,7 @@ logl_hm <- function(y, dx, theta, v, Sigma_hat) {
   n <- length(y)
   K <- deepgp:::Matern(dx, 1, theta, 1e-8, v) + Sigma_hat
   id <- deepgp:::invdet(K)
-  quadterm <- t(out_vec) %*% id$Mi %*% out_vec
+  quadterm <- t(y) %*% id$Mi %*% y
   ll <- (-0.5)*id$ldet - 0.5*quadterm
   return(ll)
 }

### 2025-06-11 dgp.hm/R/misc.R

@@ -7,6 +7,13 @@
 #   nl_matern
 #   moore_penrose
 
+# Imported Functions ----------------------------------------------------------
+#' @importFrom mvtnorm rmvnorm
+#' @importFrom deepgp sq_dist
+#' @importFrom stats quantile dgamma optim runif
+#' @importFrom grDevices heat.colors
+#' @importFrom graphics lines par
+
 # Optimize Matern -------------------------------------------------------------
 #' @export
 
@@ -104,7 +111,7 @@ monowarp_ref <- function(x, xg, wg, index) {
     r_new <- range(wg[, i])
     wg[, i] <- (wg[, i] - r_new[1])/(r_new[2] - r_new[1]) # first scale to [0, 1]
     wg[, i] <- wg[, i]*(r_orig[2] - r_orig[1]) + r_orig[1] # then scale to original range
-    w[, i] <- fo_approx(xg[, i], wg[, i], x[, i], index[, i])
+    w[, i] <- deepgp:::fo_approx(xg[, i], wg[, i], x[, i], index[, i])
   }
   return(w)
 }

### 2025-06-11 dgp.hm/R/plot.R

@@ -3,7 +3,7 @@
 #' @rdname plot
 #' @export
 
-plot.gphm <- function(x) {
+plot.gphm <- function(x, ...) {
   
   oldpar <- par(no.readonly = TRUE)
   on.exit(par(oldpar))
@@ -23,7 +23,7 @@ plot.gphm <- function(x) {
 #' @rdname plot
 #' @export
 
-plot.dgp2hm <- function(x) {
+plot.dgp2hm <- function(x, ...) {
   
   # save and restore par settings 
   oldpar <- par(no.readonly = TRUE)

### 2025-06-11 dgp.hm/R/predict.R

@@ -43,7 +43,7 @@ est_true <- function(fit, tolpower = -10, return_all = FALSE) {
     # cov mtx of [S]; based on a matern covariance function
     if (deep) {
       w <- monowarp_ref(fit$x, fit$x_grid, fit$w_grid[, i], grid_index)
-      d2 <- deepgp:::sq_dist(w)
+      d2 <- deepgp::sq_dist(w)
       theta <- fit$theta_y[i]
     } else theta <- fit$theta[i]
     S <- deepgp:::Matern(d2, 1, theta, 1e-8, fit$v)

### 2025-06-11 dgp.hm/R/fit.R
done with monowarp (fixed a couple more bugs)

@@ -152,7 +152,7 @@ gibbs_two_layer_hm <- function(x, y, x_grid, nmcmc, verb, initial, settings, v,
     theta_w[j] <- samp$theta
     
     # Sample hidden Gaussian layer (w)
-    samp <- sample_w_hm(y, w_grid[, j - 1], dw, x, x_grid, dx_grid,
+    samp <- sample_w_hm(y, w_grid[, j - 1], dw, x, x_grid, dx_grid, grid_index,
                         theta_y[j], theta_w[j], ll_prev = ll_outer, v = v, 
                         Sigma_hat = Sigma_hat)
     w_grid[, j] <- samp$w_grid

### 2025-06-11 dgp.hm/R/misc.R

@@ -104,14 +104,13 @@ monowarp_ref <- function(x, xg, wg, index) {
   if (!is.matrix(index)) index <- matrix(index, ncol = 1)
   
   w <- matrix(nrow = nrow(x), ncol = ncol(wg))
-  for (i in 1:ncol(wg)) {
-    r_orig <- range(wg[, i])
-    wg[, i] <- exp(wg[, i])
-    wg[, i] <- cumsum(wg[, i])
-    r_new <- range(wg[, i])
-    wg[, i] <- (wg[, i] - r_new[1])/(r_new[2] - r_new[1]) # first scale to [0, 1]
-    wg[, i] <- wg[, i]*(r_orig[2] - r_orig[1]) + r_orig[1] # then scale to original range
-    w[, i] <- deepgp:::fo_approx(xg[, i], wg[, i], x[, i], index[, i])
-  }
+  r_orig <- range(wg)
+  wg <- exp(wg)
+  wg <- cumsum(wg)
+  r_new <- range(wg)
+  wg <- (wg - r_new[1])/(r_new[2] - r_new[1]) # first scale to [0, 1]
+  wg <- wg*(r_orig[2] - r_orig[1]) + r_orig[1] # then scale to original range
+  w <- deepgp:::fo_approx(xg, wg, x, index)
+
   return(w)
 }

### 2025-06-11 dgp.hm/R/plot.R

@@ -41,16 +41,19 @@ plot.dgp2hm <- function(x, ...) {
          main = paste0('Trace Plot of theta_w'))
   
   # Hidden layer plot
+  grid_index <- deepgp:::fo_approx_init(x$x_grid, x$x)
   indx <- floor(seq(from = 1, to = x$nmcmc, length = 100))
   if (indx[1] == 0) indx[1] <- 1
   col <- heat.colors(100 + 10) # add ten to avoid using colors that are too light
   par(mfrow = c(1, 1), mar = c(4, 4, 2, 2))
   o <- order(x$x)
-  plot(x$x[o], x$w[o, indx[1]] - mean(x$w[, indx[1]]), type = 'l', xlab = 'X', 
+  w <- monowarp_ref(x$x, x$x_grid, x$w_grid[, indx[1]], grid_index)
+  plot(x$x[o], w[o] - mean(w), type = 'l', xlab = 'X', 
        ylab = 'W', col = col[1], main = paste0('MCMC samples of X to W'), 
        ylim = c(min(x$w[, indx]), max(x$w[, indx])))
   for (j in 2:length(indx)) {
-    lines(x$x[o], x$w[o, indx[j]] - mean(x$w[, indx[j]]), col = col[j])
+    w <- monowarp_ref(x$x, x$x_grid, x$w_grid[, indx[j]], grid_index)
+    lines(x$x[o], w[o] - mean(w), col = col[j])
   }
 
 }

### 2025-06-11 paper/main.tex
update some comments, include initial SEPIA cov fn info

@@
 tailored to each scenario (more on this in later subsections).
-We consider $\Sigma_\epsilon$ as known, given domain-matter expertise. (right)?
-\textcolor{red}{(Do we actually tailor this?  I thought it was consistent across the board.)}
+We consider $\Sigma_\varepsilon$ as known, given domain-matter expertise.
+\textcolor{red}{(Here, we are fixing $\Sigma_\varepsilon$ upon estimating the Matern 
+parameters and utilizing the precision estimates (domain-matter expertise from 
+log-log regression model)?  The estimate of $\Sigma_\varepsilon$ is tailored for
+each cosmology, as the Matern estimates are computed uniquely for each.)}
 
@@
 in \ref{sec:apdx_SgivenY}. With all posterior draws $S^{(t)}$, we compute
-the average of these to obtain our posterior mean for a particular cosmology.
+the average of these to obtain our posterior mean for a particular cosmology. Additionally,
+we quantify uncertainty by finding the 2.5th and 97.5th percentiles of the posterior 
+draws at each index to obtain a 95\% credible interval.
 We denote our inference procedure for the hierarchical model of Eq.~\ref{eq:dgphm} as
@@
 structure to incorporate multiple functional outputs.
-\textcolor{red}{(How do you get UQ though?)}
 
@@
 \cite{moran2023mira}, we employ a Python implementation of this approach called 
-SEPIA \citep{gattiker2020sepia}. \textcolor{red}{Please note: currently show GPfit 
-predictions for Mira-Titan Data, NOT SEPIA. The CAMB results are based on SEPIA.}
-That is, the $i$th vector of PC weights $w_i(\psi)$, 
-$i \in \{1,\ldots,p_\eta\}$, will be modeled as a zero mean GP: 
-$w_i(\psi) \sim GP(0, \sigma^2R)$, with
+SEPIA \citep{gattiker2020sepia}. That is, the $i$th vector of PC weights $w_i(\psi)$, 
+$i \in \{1,\ldots,p_\eta\}$, is modeled as a zero-mean Gaussian process with a 
+squared exponential (i.e., Gaussian) covariance function:
+\[
+w_i(\psi) \sim GP\left(0,\ \frac{1}{\lambda_{z_i}} R^{(i)} + \frac{1}{\lambda_{s_i}} I \right),
+\]
+with covariance function
+\[
+R^{(i)}_{uv} = \exp\left(-\sum_{k=1}^{p_\psi} \beta_{ik} (x_{uk} - x_{vk})^2 \right),
+\]
+where $\beta_{ik}$ is the inverse squared lengthscale for the $k$th input dimension, $\lambda_{z_i}$ is the marginal precision for $w_i(\psi)$, and $\lambda_{s_i}$ is the nugget precision (set to zero if no nugget is used).
 
-\begin{equation*}
-    R_{ij} = \prod_{k=1}^{p_\psi}\exp\left\{-10^{\beta_k}|x_{ik}-x_{jk}|^\alpha\right\}
-\end{equation*}
-
-\textcolor{red}{We fix $\alpha=1.95$ and estimate each $\beta_k$ with $k \in \{1,...,p_\psi\}$. 
-The above with $R_{ij}$ is based on GPfit; will update this to reflect SEPIA settings later.}
 Using these trained GPs from the PCs, we predict the spectra for unobserved cosmologies.

### 2025-06-11 pca/get_weights_CAMB_64.R
change k_ci to uniform grid

@@ -47,7 +47,7 @@ plot(log10(k_c), scrP(pk_c, k_c), type="l",
      main = "held-out CAMB runs (adjusted by h)")
 
 # create k values of CAMB for interpolation
-k_ci <- 10^(seq(-1.7, 0.7, by=0.015))
+k_ci <- 10^read.csv("../fitting/results/CAMB/xcamb_int.csv")$x
 
 # interpolate to common k grid (k_ci)
 pk_ci <- approx(k_c, scrP(pk_c, k_c), k_ci)$y

### 2025-06-13 paper/main.tex
include Sigma_epsilon details, more info on CAMB, *HM to *FCO, etc.

@@
 low resolution runs (as visualized in Figure \ref{fig:plot_data}).  For
-this, we embrace variations of the Mat\'ern kernel \citep{stein1999interpolation},
-tailored to each scenario (more on this in later subsections).
-We consider $\Sigma_\varepsilon$ as known, given domain-matter expertise.
-\textcolor{red}{(Here, we are fixing $\Sigma_\varepsilon$ upon estimating the Matern 
-parameters and utilizing the precision estimates (domain-matter expertise from 
-log-log regression model)?  The estimate of $\Sigma_\varepsilon$ is tailored for
-each cosmology, as the Matern estimates are computed uniquely for each.)}
+this, we specify a dense covariance matrix $\Sigma_\varepsilon$ that simulataneously
+allows us to model the correlated errors, as well as build domain-matter expertise
+into the prior. To this end, we embrace variations of the Mat\'ern kernel 
+\citep{stein1999interpolation}, tailored to each cosmology (more on this in later 
+subsections).
 
@@
 
-We use the Mat\`ern kernels,
+We use the Mat\`ern kernel \textcolor{red}{@Annie, might need to add/remove $\tau^2$ 
+in one or both of these kernels},
 \[
@@
 \quad\textrm{for}\quad 
-d=||w_i-w_j||^2, \;\;\textrm{and} \\
-&\Sigma_W(X)^{i,j} = \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_W}} + 
-  \frac{5d^2}{3\theta_W}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_W}}\right)
-\quad\textrm{for}\quad 
-d=||x_i-x_j||^2.
+d=||w_i-w_j||^2. 
+% \;\;\textrm{and} \\
+% &\Sigma_W(X)^{i,j} = \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_W}} + 
+%   \frac{5d^2}{3\theta_W}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_W}}\right)
+% \quad\textrm{for}\quad 
+% d=||x_i-x_j||^2,
 \end{aligned}
 \]
-In the $S$ layer, hyperparameters $\tau^2$ and $\theta_S$ govern the 
-correlation scale and lengthscale respectively.  In the $W$ layer, we allow for a
-unique lengthscale $\theta_W$ but enforce unit scale following \citet{sauer2023active}.
+In the $S$ layer, hyperparameters $\tau^2$ and $\theta_S$ govern the correlation 
+scale and lengthscale respectively.  In the $W$ layer, $\Sigma_W(X)^{i,j}$ is defined 
+analogously for squared distances $d=||x_i-x_j||^2$ and we allow for a unique 
+lengthscale $\theta_W$ but enforce unit scale following \citet{sauer2023active}.
 
@@
 C^{(t)}&=\left(\Sigma_S(W^{(t)})^{-1}+\Sigma_\varepsilon^{-1}\right)^{-1} \\
-m^{(t)}&=C^{(t)}\left(\Sigma_\varepsilon^{-1}\bar{Y}+\Sigma_S(W^{(t)})^{-1}\boldsymbol{\mu}\right).
+m^{(t)}&=C^{(t)}\left(\Sigma_\varepsilon^{-1}\bar{Y}\right).
+% "+\Sigma_S(W^{(t)})^{-1}\boldsymbol{\mu}" removed from m^t when mu=0
 \end{array}
@@
 
-These results come from Bayes' Theorem, with details of this derivation provided 
-in \ref{sec:apdx_SgivenY}. With all posterior draws $S^{(t)}$, we compute
+These results come from Bayes' Theorem and reflect that $\boldsymbol{\mu}_S=\boldsymbol{0}$. 
+These results will generalize to any value for $\boldsymbol{\mu}_S$, with details of this 
+general derivation provided in \ref{sec:apdx_SgivenY}. With all posterior draws $S^{(t)}$, we compute
 the average of these to obtain our posterior mean for a particular cosmology. Additionally,
@@
 We denote our inference procedure for the hierarchical model of Eq.~\ref{eq:dgphm} as
-``DGP.HM'' given its deep Gaussian process foundation with additional hierarchical modeling (HM)
-structure to incorporate multiple functional outputs.
+``DGP.FCO'' given its deep Gaussian process foundation with additional hierarchical modeling
+structure to incorporate multiple functional correlated outputs (FCO).
 
@@
 
-We compare DGP.HM to competing models with a simulation study where the underlying 
+We compare DGP.FCO to competing models with a simulation study where the underlying 
 true curve is known. Following \cite{moran2024dpc}, we consider two functions 
-($f_1$ and $f_2$) each with two variance specifications (settings ``4" and ``5", 
+($f_1$ and $f_2$) each with two variance specifications (settings ``A" and ``B", 
 \textcolor{magenta}{we can consider additional homoskedastic cases in the appendix if necessary}).
@@
 
-For the two variance settings, we consider matrices $\Sigma_4$ and $\Sigma_5$ to 
+For the two variance settings, we consider matrices $\Sigma_A$ and $\Sigma_B$ to 
 have a Mat\'ern covariance function with a nugget of $10^{-8}$ and smoothness of 2.5. 
-For setting ``4", we have $\tau^2=0.0225$ and $\theta=0.01$. For setting ``5", we have 
-$\tau^2=0.01$ and $\theta=0.05$, and additionally scale this covariance matrix by 
-a vector of values $s$ so that $\Sigma_5 = s M s^T$ is nonstationary; here $M$ is 
+For setting ``A", we have $\tau^2=0.0225$ and $\theta=0.01$. For setting ``B", we have 
+$\tau^2=0.1$ and $\theta=0.05$, and additionally scale this covariance matrix by 
+a vector of precision values $s$ so that $\Sigma_B = s M s^T$ is nonstationary; here $M$ is 
 a Mat\'ern covariance matrix with parameters $\tau^2$ and $\theta$.
@@
 consideration \citep{gneiting2007strictly}. For both metrics, a lower value indicates 
-better performance. Log score and MSE results are shown in Figures \ref{fig:sims_logS} 
-and \ref{fig:sims_MSE} respectively; our DGP.HM model performs favorably against 
-these current benchmark methods. \textcolor{magenta}{If we mention that DGP.HM timings 
-speed up compared to \texttt{deepgp}, we would also need to mention that \texttt{hetGP} 
-is much faster than both of these.}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=6in]{sims_MSE.png}
-    \caption{Boxplot of MSEs across two different functions and covariance settings 
-             (lower is better). For each case, the boxplot is constructed from 20 
-             random batches were simulated. Each column represents a function/variance 
-             specification pair, and each row shows results for batch sizes of 
-             either 5 (top) or 15 (bottom).}    
-    \label{fig:sims_MSE}
-\end{figure}
+better performance. Log score are shown in Figure \ref{fig:sims_logS}; our DGP.FCO 
+model performs favorably against each of these current benchmark methods by
+accounting for spatial dependence within our model, resulting in more accurate UQ.
+MSE results, which do not address UQ, were comparable across competing methods; these results
+can be found in Figure \ref{fig:sims_MSE}. \textcolor{magenta}{If we mention that DGP.FCO timings 
+speed up compared to \texttt{deepgp} and DPC, we would also need to mention that \texttt{hetGP} 
+is much faster than all of these.}
 
@@
     \includegraphics[width=\textwidth]{CAMB_fit_model1.png}
-    \caption{Example of DGP.HM model fit for the first batch of model runs from the CAMB data, 
+    \caption{Example of DGP.FCO model fit for the first batch of model runs from the CAMB data, 
              with LOESS mean term subtracted. We can compare the orange lines 
@@
 
-We use 32 CAMB batches to train the DGP.HM model and reserve the other 32 as a test set.
+We use 32 CAMB batches to train the DGP.FCO model and reserve the other 32 as a test set.
 To imitate the role of the perturbation theory output of Mira-Titan, we anchor estimates of the CAMB
@@
 strictly on the low- and high-resolution runs for $k$ above this. We take a weighted
-average of the CAMB runs and assume heteroskedastic errors in a similar fashion. For the CAMB
-data, lengthscale estimates are consistently around $3\times 10^{-7}$, evidence that we
-can simplify our covariance structure to contain no spatial dependence 
-(compare low-resolution runs of Figure \ref{fig:fit_camb} 
-with Figure \ref{fig:plot_data}).
+average of the CAMB runs and assume heteroskedastic errors in a similar fashion.
+
+We use a log-log regression model to obtain precision estimates $\Lambda_\ell$ and $\Lambda_h$
+for the low- and high-resolution runs respectively, as outlined in Section \ref{sec:data}. 
+Similarly, we set $\Lambda_i=10^8$ to illustrate high confidence in the infinite-resolution 
+run for low $k$ values and incorporate these estimates within $\Sigma_\varepsilon$. 
+With the precision information, we then estimate the Mat\'ern parameters for 
+$\Sigma_\varepsilon$. For the CAMB data, lengthscale estimates are consistently 
+around $3\times 10^{-7}$, evidence that we can simplify our covariance structure 
+to contain no spatial dependence (compare low-resolution runs of Figure \ref{fig:fit_camb} 
+with Figure \ref{fig:plot_data}). We therefore specify $\Sigma_\varepsilon = 
+(15\Lambda_\ell+\Lambda_h+\Lambda_i)^{-1}$ for the CAMB data.
+
+After fitting the DGP.FCO model, an illustration of the posterior mean and its UQ can
+be seen in Figure \ref{fig:fit_camb}. Notice that there are peaks where both the high-resolution
+run and the weighted average are biased, but our posterior mean adjusts appropriately 
+and has a more reliable estimate for the infinite-resolution run (the ``truth" in this setting). 
+UQ is handled well, as precision increases with increasing $k$, which is captured in our
+95\% credible intervals.
 
@@
 matrix $\Sigma_\varepsilon$ (accounting for spatial dependence) may outperform a 
-diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$. In previous work, we have 
-modeled multiple candidates for $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and 
-found that a Mat\'ern covariance $\Sigma_\ell$ trained on the low-resolution runs 
-(pre-scaled by the precision values and subtracting a LOESS-smoothed average) performs 
-the best. We specify this structure within $\Sigma_\varepsilon$ and refer readers 
+diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$. From Section \ref{sec:data},
+we can incorporate domain knowledge from $\Lambda_p, \Lambda_\ell,$ and $\Lambda_h$ 
+within our model. In previous work, we have modeled multiple candidates for 
+$\Sigma_\varepsilon$ \citep{walsh2023bayesian} and found that a Mat\'ern covariance 
+$\Sigma_\ell$ trained on the low-resolution runs (pre-scaled by the precision 
+values and subtracting a LOESS-smoothed average) performs the best. 
+We specify this structure within $\Sigma_\varepsilon$ and refer readers 
 to \cite{walsh2023bayesian} for a more detailed exposition. This yields 
@@
 where $\Lambda_p, \Sigma_\ell^{-1}, \text{ and } \Lambda_h$ are all $n\times n$ 
-and are 0 wherever the corresponding data type is biased.
-
-From here, we estimate $\tau_\ell, \theta_\ell$ with maximum 
-likelihood, and conditional on $\Sigma_\varepsilon$,
+and are 0 wherever the corresponding data type is biased. From here, we estimate 
+$\tau_\ell, \theta_\ell$ with maximum likelihood, and input these within 
+$\Sigma_\varepsilon$, which we use as an input to our model.
 
@@
 
-For simplicity, we will denote these eight parameters as $\psi \in \mathbb{R}^8$.
-
 Upon obtaining the posterior means from each of the 111 training cosmologies, this 
-section details our second contribution - the use of the Bayesian DGP.HM model 
+section details our second contribution - the use of the Bayesian DGP.FCO model 
 to predict the curve for held-out cosmologies. This is achieved with modeling the 
@@
 
+Our goal is to predict the power spectrum for a held-out cosmology. 
+For simplicity, we will denote these eight parameters for a given cosmology (see 
+Section \ref{sec:data}) as $\psi \in \mathbb{R}^8$. 
 Figure \ref{fig:mean_PCs_oneW} illustrates this decomposition with the estimated 
@@
 
-From our 32 posterior means using DGP.HM, these are input into SEPIA in order to predict power spectra
+From our 32 posterior means using DGP.FCO, these are input into SEPIA in order to predict power spectra
 for the 32 test cosmologies. With the infinite-resolution output, we have a sense of the ``best-case" 
@@
 estimated posterior means). In Figure \ref{fig:pca_preds_v_camb}, we see comparisons of the heldout 
-infinite resolution alongside two predictions: one based on DGP.HM and the ``best-case" scenario using the
+infinite resolution alongside two predictions: one based on DGP.FCO and the ``best-case" scenario using the
 infinite-resolution spectra of the training set. There is a high level of agreement between the pairs
@@
 (Figure \ref{fig:mse_camb}, most are very similar across the two groups (with the two most difficult 
-predictions being the exceptions), suggesting that overall DGP.HM is estimating the true spectra 
+predictions being the exceptions), suggesting that overall DGP.FCO is estimating the true spectra 
 for each batch well.
@@
              (trained on the infinite resolution) is in blue and the prediction from 
-             the DGP.HM posterior mean is in red.}   
+             the DGP.FCO posterior mean is in red.}   
     \label{fig:pca_preds_v_camb}
@@
     \includegraphics[width=3in]{mse_dot.jpg}
-    \caption{Comparison of MSE for CAMB predictions, when training with DGP.HM posterior 
+    \caption{Comparison of MSE for CAMB predictions, when training with DGP.FCO posterior 
              means (left), and training with the ``true" infinite-resolution spectra (right).}   
@@
 
-Both Cosmic Emu and DGP.HM are capable of modeling a true underlying functions based 
+Both Cosmic Emu and DGP.FCO are capable of modeling a true underlying functions based 
 upon multiple functional realizations, but our approach utilizes DGPs within a 
@@
 estimate the weights of the PCs from the estimated spectra. Below, we compare prediction 
-performance between Cosmic Emu and DGP.HM for the 6 held-out cosmologies. 
+performance between Cosmic Emu and DGP.FCO for the 6 held-out cosmologies. 
 \textcolor{magenta}{In addition, we can also predict over a large uniform or full 
@@
 of the spectrum estimation for training cosmologies are shown as well (using RMSE against 
-the weighted average of observations). For predictions, DGP.HM achieves the greatest 
+the weighted average of observations). For predictions, DGP.FCO achieves the greatest 
 reduction in RMSE in the region where perturbation theory and the low-resolution runs 
 are deemed unbiased. For the region where only the high-resolution run is unbiased, 
-Cosmic Emu has a lower average RMSE. Over all $k$ values and predictions, DGP.HM has 
+Cosmic Emu has a lower average RMSE. Over all $k$ values and predictions, DGP.FCO has 
 an average RMSE of 0.0089 and Cosmic Emu has an average RMSE of 0.0106. Our RMSE is 
@@
 $W$ jointly across different cosmologies, as well as considering different redshift 
-values other than $z=0$. Leveraging a hybrid model between DGP.HM and Cosmic Emu 
+values other than $z=0$. Leveraging a hybrid model between DGP.FCO and Cosmic Emu 
 may also help to obtain accurate estimates on the lower and higher ends of $k$ values, 
@@
 could be compared in addition to DGPs.
+
+\section{Acknowledgments}
+\label{sec:ack}
+
+\textcolor{red}{This work was in part supported by the U.S. Department
+of Energy, Office of Science, Office of Advanced Scientific Computing Research, 
+Scientific Discovery through Advanced Computing (SciDAC) program through 
+Grant 420453 "Enabling Cosmic Discoveries in the Exascale Era."}
+
+The authors are pleased to acknowledge Advanced Research Computing at Virginia Tech 
+(\url{https://arc.vt.edu/}) as well as the Shared Computing Cluster (SCC) administered
+by Boston University's Research Computing Services (\url{www.bu.edu/tech/support/research/})
+for providing computational resources and technical support that have contributed 
+to the results reported within this paper.
 
@@
     \caption{Results of all six predictions for both methods on the held-out 
-    cosmologies (weighted average subtracted). CosmicEMU is dashed, DGP.HM is dotted, 
+    cosmologies (weighted average subtracted). CosmicEMU is dashed, DGP.FCO is dotted, 
     and solid gray lines show results of held-out computer model runs.}
     \label{fig:plot_pred_1to6}
+\end{figure}
+
+\section{Simulation Study Additional Results}
+\label{sec:apdx_sims}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{sims_MSE.png}
+    \caption{Boxplot of MSEs across two different functions and covariance settings 
+             (lower is better). For each case, the boxplot is constructed from 20 
+             random batches were simulated. Each column represents a function/variance 
+             specification pair, and each row shows results for batch sizes of 
+             either 5 (top) or 15 (bottom).}    
+    \label{fig:sims_MSE}
 \end{figure}

### 2025-06-16 paper/main.tex
more edits to sections 1 and 2

@@
 \usepackage{xcolor} % only for coloring text while revising
-
-\usepackage{tikz} % for pretty diagonal matrix
-\usetikzlibrary{matrix,decorations.pathreplacing,calc}
 
@@
 while providing effective UQ.  Our model extends previous work on Bayesian DGPs 
-from scalar responses to functional ones.  Second, we leverage our predicted 
+from scalar responses to correlated functional outputs.  Second, we leverage our predicted 
 power spectra from various cosmologies in order to accurately predict the entire 
@@
 defines ``cosmologies'' according to eight cosmological parameters (more on this
-in Section \ref{sec:data}).  For a particular cosmology, the simulation of the power matter spectra 
-is rather complex.  Although we presume there exists some true matter spectra as a function
+in Section \ref{sec:data}).  For a particular cosmology, the simulation of the power matter spectrum 
+is rather complex.  Although we presume there exists some true matter spectrum as a function
 of wavenumber, we are not able to directly observe it.  Rather, the simulation returns
@@
 
-There are several challenges to estimating the underlying spectra.  Our model
+There are several challenges to estimating the underlying spectrum.  Our model
 must: handle multiple correlated functional observations, offer enough 
 flexibility to accommodate the nonstationarity present in power matter spectra 
-(which vary in smoothenss across wavenumbers), provide effective uncertainty
+(which vary in smoothness across wavenumbers), provide effective uncertainty
 quantification (UQ), and allow for the incorporation of expert knowledge 
@@
 we propose a Bayesian hierarchical model which treats functional simulation 
-outputs as realizations of a deep Gaussian process \citep[DGP;][]{damianou2013deep} 
-centered on the true underlying spectra.  The hierarchical structure naturally
-accommodates the correlated observations.  The depth of the DGP offers nonstationary
+outputs as realizations of a Gaussian process (GP) 
+centered on the true underlying spectrum, which itself is a realization of a deep Gaussian process 
+\citep[DGP;][]{damianou2013deep}.  The GP accommodates correlated observations
+through its covariance structure.  The depth of the DGP offers nonstationary
 flexibility.  The Bayesian framework facilitates UQ and the incorporation of prior
@@
 the Code for Anisotropies in the Microwave Background \citep[CAMB;][]{lewis2011CAMB}.
-The CAMB model also exhibits similar behavior to Mira-Titan, but crucially
-provides a ``true" power spectrum corresponding to each cosmology.
+The CAMB model exhibits similar behavior to Mira-Titan but crucially
+provides a ``true" power spectrum for each cosmology.
 
@@
 We leverage the same training data, in conjunction with our Bayesian hierarchical
-DGP model and basis function representations, to train a Gaussian process surrogate 
+DGP model and basis function representations, to train a GP surrogate 
 on principal component weights in order to predict spectra for unobserved
@@
 Bayesian DGP model in a public git repository.\footnote{\url{https://github.com/stevewalsh124/dgp.hm}}
-\textcolor{cyan}{Please note: this repository is currently private until we 
-hear back regarding permissions for sharing the Mira-Titan data alongside our code.}
+%\textcolor{cyan}{Please note: this repository is currently private until we 
+%hear back regarding permissions for sharing the Mira-Titan data alongside our code.}
 
@@
 
-The Mira-Titan simulation suite is a dataset consisting of simulated power spectra for 
-117 different cosmologies. Each of these cosmologies is specified by eight cosmological 
+The Mira-Titan simulation dataset consists of simulated power spectra for 
+117 unique cosmologies. Each of these cosmologies is specified by eight cosmological 
 parameters: matter density, baryon density, amplitude of density fluctuations, 
@@
     runs are approximately unbiased.
-    {\it Right:} Same as middle, but with estimated posterior mean subtracted.}
+    {\it Right:} Same as middle, but with estimated posterior mean subtracted.
+    \textcolor{red}{(@Steve - please fix the legend so it is not going over the lines)}}
     \label{fig:plot_data}
@@
 spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from 
-lower resolutions simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from 
+low resolution simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from 
 a high resolution simulation ($y_h$). 
@@
 of the output for a particular cosmology. When we focus on $k$ values where $y_{\ell_r}$ 
-is valid (center panel), we see both low-resolution and high-resolution runs oscillate around 
-a mean term, which is relatively smooth compared to these realizations. When this
-mean term is subtracted, we will obtain a plot like the right panel in 
-Figure \ref{fig:plot_data}.
+is valid (center panel), the spatially correlated errors of the low-resolution and high-resolution runs 
+becomes apparent.
 
@@
 the increase in precision from the low- to high- resolution output is also 
-estimated with this regression model. We use these estimates to define three 
+estimated with this regression model. 
+\textcolor{red}{(What is the c value?  Can we report it?)}
+We use these estimates to define three 
 diagonal precision matrices, one for each data type (perturbation theory, 
@@
 the weighted average is shown in dashed orange.  As an estimate of the true
-underlying spectrum, the weighted average fluctuates too drastically due 
-to the limited amount of data.  It fails to account for the
-spatial dependence inherent in the smooth but stochastic spectra,
-which is pivotal to effectively inferring the true spectrum.
+underlying spectrum, the weighted average fluctuates too drastically.  
+This is especially apparent after centering (right panel).
+\textcolor{red}{(@Steve - how are we centering?)}
+Although the weighted average incorporates all available observatons and 
+expert knowledge regarding the precisions, it fails to account for the spatial 
+dependence inherent in the smooth but stochastic spectra, which is pivotal to 
+effectively inferring the true spectrum.
 

### 2025-06-16 paper/main.tex
through 3.1

@@
 the observed weighted average; now, we use $\bar{Y}$ to indicate the corresponding 
-random variable within the statistical model.  Together, $X$ and $\bar{Y}$ comprise the 
-training data for a particular cosmology. $\bar{Y}$ is typically pre-processed
-based on domain expertise, such as that defined in Section \ref{sec:data}.
-We will reserve details of domain-specific pre-processing for the CAMB and Mira-Titan suites
-for Sections \ref{subsec:camb} and \ref{subsec:mira_fit}, respectively.
+random variable within the statistical model.  
+%Together, $X$ and $\bar{Y}$ comprise the 
+%training data for a particular cosmology. 
+$\bar{Y}$ is typically pre-processed
+based on domain expertise.  Pre-processing for Mira-Titan was described in Section 
+\ref{sec:data}; we will detail domain-specific pre-processing 
+for the CAMB suite in Section \ref{subsec:camb}.
 
@@
 have considered diagonal $\Sigma_\varepsilon$ \citep{moran2024dpc}, we find it
-essential to account for the smoothly varying spatial dependence of the
-low resolution runs (as visualized in Figure \ref{fig:plot_data}).  For
-this, we specify a dense covariance matrix $\Sigma_\varepsilon$ that simulataneously
-allows us to model the correlated errors, as well as build domain-matter expertise
+essential to account for the smoothly varying spatial dependence
+(as visualized in Figure \ref{fig:plot_data}).  We thus specify a dense 
+covariance matrix $\Sigma_\varepsilon$ that simulataneously
+allows us to model the correlated errors and build domain-matter expertise
 into the prior. To this end, we embrace variations of the Mat\'ern kernel 
@@
 baryonic acoustic oscillations when $-2 \leq \log_{10}(k) \leq -1$. To account 
-for this, we incorporate an additional latent Gaussian process, $W$, which accommodates
-nonstationarity by warping $X$ via $W \sim \mathrm{GP}\left(\boldsymbol{\mu}_W, \Sigma_W(X)\right)$.
+for this, we incorporate a latent monotonic Gaussian process \citep[monoGP;][]{barnett2025monotonic},
+which accommodates nonstationarity by warping $X$ via 
+$W \sim \mathrm{monoGP}\left(\boldsymbol{\mu}_W, \Sigma_W(X)\right)$.
 $W$ is also an $n$-vector whose entries $w_i$ correspond to warped versions of each $x_i$.
@@
 S\mid W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_S(W)\right) \\
-W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_W(X)\right).
+W &\sim \mathrm{monoGP}\left(\mathbf{0}, \Sigma_W(X)\right).
 \end{aligned}
@@
 
-We use the Mat\`ern kernel \textcolor{red}{@Annie, might need to add/remove $\tau^2$ 
-in one or both of these kernels},
+For both $\Sigma_S$ and $\Sigma_W$, we use Mat\`ern kernels with unit scale and
+unique lengthscales,
 \[
-\begin{aligned}
-&\Sigma_S(W)^{i,j} = \tau^2 \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_S}} + 
-  \frac{5d^2}{3\theta_S}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_S}}\right)
-\quad\textrm{for}\quad 
-d=||w_i-w_j||^2. 
-% \;\;\textrm{and} \\
-% &\Sigma_W(X)^{i,j} = \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta_W}} + 
-%   \frac{5d^2}{3\theta_W}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta_W}}\right)
-% \quad\textrm{for}\quad 
-% d=||x_i-x_j||^2,
-\end{aligned}
+\begin{array}{l}
+\Sigma_S(W)^{i,j} = k\left(||w_i - w_j||^2, \theta_S\right) \\[7pt]
+\Sigma_W(X)^{i,j} = k\left(||x_i - x_j||^2, \theta_W\right)
+\end{array}
+\quad\textrm{where}\quad
+k(d, \theta) = \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta}} + 
+  \frac{5d^2}{3\theta}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta}}\right).
 \]
-In the $S$ layer, hyperparameters $\tau^2$ and $\theta_S$ govern the correlation 
-scale and lengthscale respectively.  In the $W$ layer, $\Sigma_W(X)^{i,j}$ is defined 
-analogously for squared distances $d=||x_i-x_j||^2$ and we allow for a unique 
-lengthscale $\theta_W$ but enforce unit scale following \citet{sauer2023active}.
+In our one-dimensional setting, we embrace unit scale on $S$ to preserve identifiability
+with $\theta_S$ \textcolor{red}{(@Steve - can you cite that paper here?)}.
+Unit scale on latent $W$ was recommended by \citet{sauer2023active} for similar reasons.
 
@@
 
-We embrace fully-Bayesian inference of latent $W$ and kernel hyperparameters $\{\tau^2, \theta_S, \theta_W\}$.
+We embrace fully-Bayesian inference of latent $W$ and kernel hyperparameters $\{\theta_S, \theta_W\}$.
 Specifically, we use elliptical slice sampling \citep[ESS;][]{murray2010elliptical} to 

### 2025-06-16 paper/main.tex
working through section 3

@@
 \]
-In our one-dimensional setting, we embrace unit scale on $S$ to preserve identifiability
-with $\theta_S$ \textcolor{red}{(@Steve - can you cite that paper here?)}.
+In our one-dimensional setting, we embrace unit scale on $S$ to preserve parsimony 
+and identifiability with $\theta_S$ \textcolor{red}{(@Steve - can you cite that paper here?)}.
 Unit scale on latent $W$ was recommended by \citet{sauer2023active} for similar reasons.
@@
 \label{eq:prior}
-W &\sim \mathrm{GP}\left(\mathbf{0}, \Sigma_W(X)\right)
+W &\sim \mathrm{monoGP}\left(\mathbf{0}, \Sigma_W(X)\right)
 \end{align}
-Details are provided in \ref{sec:apdx_int_lik}. The incorporation of $\Sigma_\epsilon$ in the
+Details, including generalizations for $\boldsymbol{\mu}_S\neq\mathbf{0}$, are provided in \ref{sec:apdx_int_lik}. 
+The incorporation of $\Sigma_\epsilon$ in the
 Gaussian likelihood of this outer layer is an essential upgrade to previous DGP inferential schemes.
@@
 Eq.~(\ref{eq:likelihood}) for likelihood-based acceptance.  
-Under a reference prior $\pi(\tau^2)\propto\frac{1}{\tau^2}$,
-we integrate $\tau^2$ from the likelihood of Eq.~(\ref{eq:likelihood}).  The resulting marginal 
-posterior is equivalent to that with the plug-in estimator 
-$\hat{\tau}^2 = \dots$ \citep[][Chapter 5]{gramacy2020surrogates}.
-\textcolor{red}{(With $\Sigma_\epsilon$ on there, we can't actually do this.. unless we divide
-$\Sigma_\epsilon$ by $\tau^2$...)}.
 For $\theta_S$ and $\theta_W$, after pre-scaling, we adopt the default prior specifications 

### 2025-06-16 sims/collect_sims.R
rm unused sim settings, adjust plotting

@@ -20,14 +20,14 @@ MSE5_range <- MSE20_range <- logS5_range <- logS20_range <- c()
 
 r=5
 for (func in 1:2) {
-  for (setting in 4:7) {
+  for (setting in 4:5) {
     filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
     # Compare MSEs
-    results <- read.csv(filename)[, c(3,6,5,4)]
+    results <- read.csv(filename)[, c(2,5,4,3)]
     # NOTE: removed one outlier to not skew the ylim
     if(func==1&setting==4) results <- results[-which.max(results[,3]),]
     MSE5_range <- range(MSE5_range, results)
-    results <- read.csv(filename)[, c(8,11,10,9)]
+    results <- read.csv(filename)[, c(6,9,8,7)]
     # Note: removes three outliers from setting==5, r==5
     logS5_range <- range(logS5_range, results[results <= 150])
   }
@@ -35,12 +35,12 @@ for (func in 1:2) {
 
 r=15
 for (func in 1:2) {
-  for (setting in 4:7) {
+  for (setting in 4:5) {
     filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
     # Compare MSEs
-    results <- read.csv(filename)[, c(3,6,5,4)]
+    results <- read.csv(filename)[, c(2,5,4,3)]
     MSE20_range <- range(MSE20_range, results)
-    results <- read.csv(filename)[, c(8,11,10,9)]
+    results <- read.csv(filename)[, c(6,9,8,7)]
     logS20_range <- range(logS20_range, results)
   }
 }
@@ -49,16 +49,17 @@ for (func in 1:2) {
 # Boxplots of MSEs #
 ####################
 
-par(mfrow=c(4,4), oma=c(4,4,1.5,1), mar=c(0,0,0,0))
+par(mfrow=c(2,4), oma=c(4,4,1.5,1), mar=c(0,0,0,0))
 
 # Step 1: Precompute y-limits for each setting (row)
 row_ylims <- list()
-for (setting in 4:7) {
+for (setting in 4:5) {
   all_mses <- NULL
   for (r in c(5,15)) {
+    print(r)
     for (func in 1:2) {
       filename <- paste0("results/sims_", func, "_", setting, "_", r, ".csv")
-      results <- read.csv(filename)[, c(3,6,5,4)]
[…]

### 2025-06-16 sims/run_sim.R

@@ -6,15 +6,15 @@
 # Important Arguments:
 #   seed: random seed (must be set before sourcing functions.R)
 #   func: 1 or 2
-#   setting: 1, 2, 3, or 4 (indicates noise settings A-D from DPC paper)
+#   setting: 1, 2, 3, 4, or 5 (1:3 indicates noise settings from DPC paper, and 
+#                           4:5 indicates settings A & B for the DGP-FCO paper)
 #
 # Output:
 #   data frame of MSE and CRPS values for each model is either written to 
 #   ".csv" in the "results" folder or appended to existing file
 #   
 # Included Models:
-#   dgp: the Bayesian hierarchical DGP model 
-#   dgp_r: the dgp model, but use Sigma_hat/sqrt(r)
+#   dgp: the Bayesian hierarchical DGP-FCO model 
 #   deepgp: original DGP from "deepgp" package
 #   hetgp: heteroskedastic GP from "hetGP" package 
 
@@ -36,7 +36,7 @@ if(length(args) > 0)
   for(i in 1:length(args))
     eval(parse(text=args[[i]]))
 if(!(func %in% 1:2)) stop("func should be 1 or 2")
-if(!(setting %in% 1:7)) stop("setting should be 1, 2, 3, 4, 5, 6, or 7")
+if(!(setting %in% 1:5)) stop("setting should be 1, 2, 3, 4, or 5")
 
 set.seed(seed)
 source("functions.R") # generates random model parameters upon sourcing
@@ -48,8 +48,8 @@ vis <- FALSE # should plots be generated
 
 x <- seq(0, 4, by = 0.1)
 n <- length(x)
-if(setting==5) sddtrue <- 1/exp(seq(0,2,length=length(x)))
-if(setting %in% 6:7) sddtrue <- 1/1.5^(seq(0,2,length=length(x)))
+if(setting == 4) sddtrue <- rep(1,n)
+if(setting == 5) sddtrue <- 1/1.5^(seq(0,2,length=length(x)))
 Sigma_true <- get_Sigma_true(x, n, func, setting)
 if(vis) image(Sigma_true) # make sure sd plot looks right
 
@@ -77,13 +77,13 @@ for(i in 1:r) {
 }
 if(vis) plot(x_all, y_all)
 
-# dgp.hm model ----------------------------------------------------------------
[…]

### 2025-06-16 sims/functions.R
rm unused sim settings

@@ -38,16 +38,6 @@ get_Sigma_true <- function(x, n, func = 1, setting = 1) {
                                     theta = 1e-2, g = 1e-8, v = 2.5)
     } else if(setting == 5) {
       D <- deepgp::sq_dist(x)
-      Sigma <- deepgp:::Matern(distmat = D, tau2 = 0.1^2, 
-                               theta = 5e-2, g = 1e-8, v = 2.5)
-      Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
-    } else if(setting == 6) {
-      D <- deepgp::sq_dist(x)
-      Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
-                               theta = 1e-1, g = 1e-8, v = 2.5)
-      Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
-    } else if(setting == 7) {
-      D <- deepgp::sq_dist(x)
       Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
                                theta = 5e-2, g = 1e-8, v = 2.5)
       Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
@@ -70,16 +60,6 @@ get_Sigma_true <- function(x, n, func = 1, setting = 1) {
       Sigma_true <- deepgp:::Matern(distmat = D, tau2 = 0.15^2, 
                                    theta = 1e-2, g = 1e-8, v = 2.5)
     } else if(setting == 5) {
-      D <- deepgp::sq_dist(x)
-      Sigma <- deepgp:::Matern(distmat = D, tau2 = 0.1^2, 
-                               theta = 5e-2, g = 1e-8, v = 2.5)
-      Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
-    } else if(setting == 6) {
-      D <- deepgp::sq_dist(x)
-      Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
-                               theta = 1e-1, g = 1e-8, v = 2.5)
-      Sigma_true <- diag(sddtrue) %*% Sigma %*% diag(sddtrue)
-    } else if(setting == 7) {
       D <- deepgp::sq_dist(x)
       Sigma <- deepgp:::Matern(distmat = D, tau2 = .1, 
                                theta = 5e-2, g = 1e-8, v = 2.5)

### 2025-06-16 paper/main.tex
through 3.2

@@
 Here we describe our first contribution: the use of a Bayesian hierarchical model 
-to estimate the underlying spectra and quantify uncertainty for a particular cosmology.
+to estimate the underlying spectrum and quantify uncertainty for a particular cosmology.
 Let input $X$ represent the vector with elements $x_i = \log_{10}(k_i)$ for $i=1,\dots,n$.
@@
 based on domain expertise.  Pre-processing for Mira-Titan was described in Section 
-\ref{sec:data}; we will detail domain-specific pre-processing 
+\ref{sec:data}; we will detail pre-processing 
 for the CAMB suite in Section \ref{subsec:camb}.
@@
 $W$ is also an $n$-vector whose entries $w_i$ correspond to warped versions of each $x_i$.
-Again, we set $\boldsymbol{\mu}_W = \mathbf{0}$ without loss of generatlity (we have
+Again, we set $\boldsymbol{\mu}_W = \mathbf{0}$ without loss of generality (we have
 found $\boldsymbol{\mu}_W = X$ works similarly).
@@
 Our inferential goal is to obtain the posterior distribution of 
-$S\mid\bar{Y},\Sigma_\varepsilon$, but this requires inferring latent quantity $W$.  
+$S\mid\bar{Y},\Sigma_\varepsilon$, but this requires inferring latent $W$.  
 We first integrate over $S$ to condense our hierarchical model of Eq.~(\ref{eq:dgphm}) into
@@
 \end{align}
-Details, including generalizations for $\boldsymbol{\mu}_S\neq\mathbf{0}$, are provided in \ref{sec:apdx_int_lik}. 
+A detailed derivation, including generalizations for 
+$\boldsymbol{\mu}_S\neq\mathbf{0}$, is provided in \ref{sec:apdx_int_lik}. 
 The incorporation of $\Sigma_\epsilon$ in the
@@
 
-Given burned-in posterior samples $\{W^{(t)}, \tau^{2(t)}, \theta_S^{(t)}, \theta_W^{(t)}\}$
-for $t\in\mathcal{T}$, we may obtain the
-posterior distribution of the true power matter spectra for a particular cosmology as
+Given burned-in posterior samples $\{W^{(t)}$ and $\theta_S^{(t)}\}$
+for $t\in\mathcal{T}$ ($\theta_W^{(t)}$ is only used to sample $W^{(t)}$), 
+we may leverage Bayes' Theorem to obtain the
+posterior distribution of the true power matter spectrum for a particular cosmology as
 \[
@@
 \] 
-
-These results come from Bayes' Theorem and reflect that $\boldsymbol{\mu}_S=\boldsymbol{0}$. 
-These results will generalize to any value for $\boldsymbol{\mu}_S$, with details of this 
-general derivation provided in \ref{sec:apdx_SgivenY}. With all posterior draws $S^{(t)}$, we compute
-the average of these to obtain our posterior mean for a particular cosmology. Additionally,
-we quantify uncertainty by finding the 2.5th and 97.5th percentiles of the posterior 
-draws at each index to obtain a 95\% credible interval.
-We denote our inference procedure for the hierarchical model of Eq.~\ref{eq:dgphm} as
-``DGP.FCO'' given its deep Gaussian process foundation with additional hierarchical modeling
-structure to incorporate multiple functional correlated outputs (FCO).
+This form is simplified for $\boldsymbol{\mu}_S=\mathbf{0}$; a general derivation 
+is provided in \ref{sec:apdx_SgivenY}.  This closed-form enables direct posterior
+sampling of $S^{(t)}$.  We draw samples accordingly and compute their average 
+to obtain our posterior mean for a particular cosmology. 
+Additionally, we quantify uncertainty by finding the 2.5th and 97.5th percentiles 
+of the posterior draws at each index to obtain a 95\% credible interval.
+We denote our inference procedure for the hierarchical model of Eq.~(\ref{eq:dgphm}) as
+``DGP.FCO'' given its deep Gaussian process foundation with additional hierarchical
+structure to incorporate functional correlated outputs (FCO).
 

### 2025-06-17 paper/main.tex
comments in section 3

@@
 We compare DGP.FCO to competing models with a simulation study where the underlying 
-true curve is known. Following \cite{moran2024dpc}, we consider two functions 
-($f_1$ and $f_2$) each with two variance specifications (settings ``A" and ``B", 
-\textcolor{magenta}{we can consider additional homoskedastic cases in the appendix if necessary}).
-
-Here, $f_1$ and $f_2$ are defined as follows:
-\begin{align}
-  f_1(x) &= m_1 \exp(-u_1x/2) * \cos(wx) - m_1x/5, \quad w=\sqrt{25-[u_1/2]^2} \\
-  f_2(x) &= \exp(-m_2[x-3]^2)+\exp(-u_2[x-1]^2)-0.05\sin(8[x-1.9]), \\ 
-  \quad x &\in [0,4].
-\end{align}
-
-For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are random where 
+true curve is known. Following \cite{moran2024dpc}, we consider two test functions 
+defined as
+\[
+\begin{aligned}
+  f_1(x) &= m_1 \exp\left(-\frac{u_1x}{2}\right) * \cos\left(x\sqrt{25-\left(\frac{u_1}{2}\right)^2}\right) - \frac{m_1x}{5}\\
+  f_2(x) &= \exp\left(-m_2(x-3)^2\right)+\exp\left(-u_2(x-1)^2\right)-0.05\sin\left(8(x-1.9)\right.
+\end{aligned}
+\]
+We consider $x\in[0, 4]$.  For additional variation, the values of $m_1, m_2, u_1,$ and 
+$u_2$ are randomly selected from
 $m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 
@@
 a Mat\'ern covariance matrix with parameters $\tau^2$ and $\theta$.
+\textcolor{red}{(@Steve - this whole paragraph needs re-write.  Need to start saying
+how you draw $r$-many realizations from a GP with some mean (?) and some spatially varying
+covariance.  Then use something like this:
+\[
+\begin{aligned}
+\Sigma_A &= 0.0225 * k(||x_i - x_j||^2, 0.01) \\
+\Sigma_B &= s^\top Ms \quad\textrm{where}\quad M = 0.1 * k(||x_i - x_j||^2, 0.05)
+\quad\textrm{and}\quad s = ???
+\end{aligned}
+\]
+where you say $k$ is defined according to the previous equation.  Also, what is s?}
 
-For each simulated function and variance specification, we generate 20 values of 
-$m_1, m_2, f_1, f_2$, and conditional on these values, we sample either 5 or 15 
-functional observations and estimate the underlying true function based on the observations.
+For each simulated function and variance specification, we consider 20 Monte Carlo repetitions
+with re-randomized $m_1, m_2, u_1, u_2$.  We consider $r = 5$ and $r = 15$.
+\textcolor{red}{@Steve - what is $n$?  How dense is the grid?  I think we need a visual
+of this scenario.  Perhaps two panels, one with an $f_1$ and one with an $f_2$.
+Show the true function.  Then show $r = 5$ realizations.  Use points to highlight
+the training observations.  That way for the other methods, you can comment on how
+they treat the points as all the same, not coming from 5 smooth curves.}
 
@@
     \includegraphics[width=6in]{sims_logS.png}
-    \caption{Boxplot of log scores across two different functions and covariance 
-             settings (lower is better). For each case, the boxplot is constructed 
-             from 20 random batches were simulated. Each column represents a function/variance 
-             specification pair, and each row shows results for batch sizes of either 
-             5 (top) or 15 (bottom).}
+    \caption{Boxplot of log scores (20 repetitions, lower is better) for each 
+             simulated scenario.  Each column represents a function/covariance 
+             pair, and each row shows results for batch sizes of either 
+             5 (top) or 15 (bottom) realizations. \textcolor{red}{(This figure needs udpating.)}}
     \label{fig:sims_logS}
@@
 as a deep process convolution (DPC) approach, which is utilized within Cosmic 
-Emu \citep{moran2023mira}.
+Emu \citep{moran2023mira}. \textcolor{red}{(@Steve - given the new figure, comment
+on how these methods view the points as iid noise/heteroskedastic noise/etc.
+Also need to comment on how $\Sigma_\epsilon$ is estimated for our method.  And
+that there is no pre-processing of the weighted average, right?)}
 
@@
 MSE results, which do not address UQ, were comparable across competing methods; these results
-can be found in Figure \ref{fig:sims_MSE}. \textcolor{magenta}{If we mention that DGP.FCO timings 
-speed up compared to \texttt{deepgp} and DPC, we would also need to mention that \texttt{hetGP} 
-is much faster than all of these.}
+can be found in Figure \ref{fig:sims_MSE}.
+\textcolor{red}{(@Steve - this paragraph needs revising now that the figure is moved
+to the appendix.)}
 
@@
 
-In addition to simulated data, we investigate model fits for different cosmological 
+We next investigate model fits to several cosmological 
 datasets from CAMB \citep{lewis2011CAMB}. These datasets contain power spectra based
@@
 CAMB dataset has multiple (sixty-four) batches of power spectra for each cosmology, with each batch 
-containing multiple (fifteen) low-resolution spectra, and one high-resolution sprectum. 
-Unlike the Mira-Titan suite, this dataset favorably contains an infinite-resolution 
+containing multiple (fifteen) low-resolution spectra, and one high-resolution sprectum
+\textcolor{red}{(@Steve - do you mean there are 64 cosmologies?)} 
+Unlike the Mira-Titan suite, this dataset contains an infinite-resolution 
 spectrum, which can be treated as the true spectrum for each cosmology,
-allowing us to accurately assess posterior mean estimates and credible intervals.
+allowing us to assess the accuracy of posterior mean estimates and effectiveness
+of credible intervals. \textcolor{red}{(@Steve - refer to Figure 3 as a visual.  Comment on the
+true power spectrum and the low/high res runs.)}
 
@@
     \caption{Example of DGP.FCO model fit for the first batch of model runs from the CAMB data, 
-             with LOESS mean term subtracted. We can compare the orange lines 
+             with LOESS mean term subtracted. \textcolor{red}{(LOESS mean of what?  We need
+             to officially say how things are centered.)} We can compare the orange lines 
              (posterior mean and dotted 95\% credible interval) to the light blue line,
@@
 We use 32 CAMB batches to train the DGP.FCO model and reserve the other 32 as a test set.
+\textcolor{red}{(@Steve - I'm so confused.  We are not talking about PCA here.  Each
+cosmology stands on its own.  We use the low/high res runs for training, and the true
+spectrum for testing.)}
 To imitate the role of the perturbation theory output of Mira-Titan, we anchor estimates of the CAMB
 power spectrum at the infinite-resolution for $\log_{10}(k) < -1.7$, and then rely 
-strictly on the low- and high-resolution runs for $k$ above this. We take a weighted
+strictly on the low- and high-resolution runs for $k$ above this. 
+\textcolor{red}{(Refer to the figure here as well to help explain?)}
+We take a weighted
 average of the CAMB runs and assume heteroskedastic errors in a similar fashion. 
+\textcolor{red}{(Need details on how the weighted average is calculated.  Can
+refer to previous section, but need to be specific.)}
 
-We use a log-log regression model to obtain precision estimates $\Lambda_\ell$ and $\Lambda_h$
+To estimate $\Sigma_\epsilon$, we first use a log-log regression model to obtain 
+precision estimates $\Lambda_\ell$ and $\Lambda_h$
 for the low- and high-resolution runs respectively, as outlined in Section \ref{sec:data}. 
-Similarly, we set $\Lambda_i=10^8$ to illustrate high confidence in the infinite-resolution 
+Similarly, we set $\Lambda_i=10^8$ \textcolor{red}{(this notation is no longer right.
+Also why switch from $10^5$ to $10^8$?)} 
+to illustrate high confidence in the infinite-resolution 
 run for low $k$ values and incorporate these estimates within $\Sigma_\varepsilon$. 
 With the precision information, we then estimate the Mat\'ern parameters for 
-$\Sigma_\varepsilon$. For the CAMB data, lengthscale estimates are consistently 
+$\Sigma_\varepsilon$. \textcolor{red}{(Need more specifics.  Are we using a $\tau^2$ now?
+5/2 still?  Maximum likelihood estimation?)}
+For the CAMB data, lengthscale estimates are consistently 
 around $3\times 10^{-7}$, evidence that we can simplify our covariance structure 
 to contain no spatial dependence (compare low-resolution runs of Figure \ref{fig:fit_camb} 
-with Figure \ref{fig:plot_data}). We therefore specify $\Sigma_\varepsilon = 
+with Figure \ref{fig:plot_data} \textcolor{red}{ - I don't understand what we are supposed
+to see from this comparison.)}). We therefore specify $\Sigma_\varepsilon = 
 (15\Lambda_\ell+\Lambda_h+\Lambda_i)^{-1}$ for the CAMB data.
+
+\textcolor{red}{(We need to fix the entire previous paragraph. How about we keep it short
+and sweet and say we considered a Matern kernel, but found that the spatial correlation
+decayed so quickly ($\theta\rightarrow 0$) that spatial dependence was not necessary.)}
 
@@
 
+\textcolor{red}{(Where is the MSE/logscore across the 64 batches????  Are we comparing
+to other methods?  As it stands, this is barely a "benchmark".  It is more of a "proof
+of concept".)}
+
 \subsection{Estimating Spectra with Mira-Titan Data}
@@
 
+\textcolor{red}{(@Steve - we've done a lot of edits up to this point.  Can
+you revise this text now to fit in with the rest?  Need to refer to how the 
+weighted average is calculated.
+And the purpose of $\Sigma_\epsilon$.  But this is the first time Mira-Titan
+specific details of $\Sigma_\epsilon$ will be mentioned.)}
 From plots of the multiple low-resolution spectra (e.g., Figure \ref{fig:plot_data}), 
@@
              subtracted). Plots of the perturbation theory, low-resolution and high-resolution 
-             runs are also shown, alongside the weighted average.}
+             runs are also shown, alongside the weighted average.
+             \textcolor{red}{(This plot needs to be made fresh (not using the package
+             function.  Adjust y-axis.  We don't need the yellow/red coloring.)}}
     \label{fig:plot_fit}
@@
 \end{figure}
+
+\textcolor{red}{(This whole subsection is missing a point.  What is the point?
+We don't know the truth, but what can we glean from this?  What does the warping
+tell us?  Talk about the UQ...)}
 

### 2025-06-17 paper/main.tex
oops, comment was on wrong figure

@@
 and sweet and say we considered a Matern kernel, but found that the spatial correlation
-decayed so quickly ($\theta\rightarrow 0$) that spatial dependence was not necessary.)}
+decayed so quickly ($\theta\approx 0$) that spatial dependence was not necessary.)}
 
@@
              subtracted). Plots of the perturbation theory, low-resolution and high-resolution 
-             runs are also shown, alongside the weighted average.
-             \textcolor{red}{(This plot needs to be made fresh (not using the package
-             function.  Adjust y-axis.  We don't need the yellow/red coloring.)}}
+             runs are also shown, alongside the weighted average.}
     \label{fig:plot_fit}
@@
             Right: Corresponding 95\% credible intervals for $W$, where departure 
-            from the identity line represents nonstationary behavior.}
+            from the identity line represents nonstationary behavior.
+            \textcolor{red}{(This plot needs to be made fresh (not using the package
+             function.  Adjust y-axis.  We don't need the yellow/red coloring.)}}
    \label{fig:plot_warp}

### 2025-06-17 paper/main.tex
update manuscript; adjust plot for sims

--- 2025-06-16 paper/main.tex
@@
 obtain precision estimates, $p_1,\dots, p_n$, across the $n$ 
-wavenumber values using a log-log regression model. A multiplier $c$ for 
+wavenumber values using a log-log regression model. A multiplier $c\approx 3.73$ for 
 the increase in precision from the low- to high- resolution output is also 
 estimated with this regression model. 
-\textcolor{red}{(What is the c value?  Can we report it?)}
 We use these estimates to define three 
@@
 \subsection{Model Specification}
+\label{sec:mod_spec}
 
@@
 In our one-dimensional setting, we embrace unit scale on $S$ to preserve parsimony 
-and identifiability with $\theta_S$ \textcolor{red}{(@Steve - can you cite that paper here?)}.
+and identifiability with $\theta_S$ \citep{zhang2004inconsistent}.
 Unit scale on latent $W$ was recommended by \citet{sauer2023active} for similar reasons.
@@
 We compare DGP.FCO to competing models with a simulation study where the underlying 
-true curve is known. Following \cite{moran2024dpc}, we consider two functions 
-($f_1$ and $f_2$) each with two variance specifications (settings ``A" and ``B", 
-\textcolor{magenta}{we can consider additional homoskedastic cases in the appendix if necessary}).
+true curve is known. Following \cite{moran2024dpc}, we consider two test functions 
+defined as
+\[
+\begin{aligned}
+  f_1(x) &= m_1 \exp\left(-\frac{u_1x}{2}\right) * 
+  \cos\left(x\sqrt{25-\left(\frac{u_1}{2}\right)^2}\right) - \frac{m_1x}{5}\\
+  f_2(x) &= \exp\left(-m_2(x-3)^2\right) + 
+  \exp\left(-u_2(x-1)^2\right)-0.05\sin\left(8(x-1.9))\right.
+\end{aligned}
+\]
+For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are randomly selected from
+$m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 
+\sim \text{Uniform}(0.6,1.4)$. The outputs of $f_1$ and $f_2$ (conditional on $m_1,u_1$ and 
+$m_2,u_2$), determine the true function values for each simulation.
 
-Here, $f_1$ and $f_2$ are defined as follows:
-\begin{align}
-  f_1(x) &= m_1 \exp(-u_1x/2) * \cos(wx) - m_1x/5, \quad w=\sqrt{25-[u_1/2]^2} \\
-  f_2(x) &= \exp(-m_2[x-3]^2)+\exp(-u_2[x-1]^2)-0.05\sin(8[x-1.9]), \\ 
-  \quad x &\in [0,4].
-\end{align}
+For each of the two true functions, we consider two variance settings $A$ and $B$, where 
+matrices $\Sigma_A$ and $\Sigma_B$ have a Mat\'ern covariance function with a nugget of 
+$10^{-8}$ and smoothness of 2.5. We obtain function realizations at $x=\{0, 0.1, \dots, 
+3.9, 4.0\}$, so that each function draw contains $n=41$ observations. We draw $r$-many 
+realizations from a GP with some mean ($f_1$ or $f_2$) and some spatially varying covariance
+($\Sigma_A$ or $\Sigma_B$), where:
 
-For additional variation, the values of $m_1, m_2, u_1,$ and $u_2$ are random where 
-$m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 
-\sim \text{Uniform}(0.6,1.4)$.
+\[
+\begin{aligned}
+\Sigma_A &= 0.0225 * k(||x_i - x_j||^2, 0.01), \\
+\Sigma_B &= s^\top Ms \quad\textrm{where}\quad M = 0.1 * k(||x_i - x_j||^2, 0.05)
+\quad\textrm{and}\quad s = 1.5^{-x/2},
+\end{aligned}
+\]
 
-For the two variance settings, we consider matrices $\Sigma_A$ and $\Sigma_B$ to 
-have a Mat\'ern covariance function with a nugget of $10^{-8}$ and smoothness of 2.5. 
-For setting ``A", we have $\tau^2=0.0225$ and $\theta=0.01$. For setting ``B", we have 
-$\tau^2=0.1$ and $\theta=0.05$, and additionally scale this covariance matrix by 
-a vector of precision values $s$ so that $\Sigma_B = s M s^T$ is nonstationary; here $M$ is 
-a Mat\'ern covariance matrix with parameters $\tau^2$ and $\theta$.
+where $k(\cdot,\cdot)$ is defined in Section \ref{sec:mod_spec}.  Within $\Sigma_B$, $s$ introduces
+nonstationarity into the function realizations, more accurately matching the behavior of the 
+CAMB and Mira-Titan motivating datasets.
 
-For each simulated function and variance specification, we generate 20 values of 
-$m_1, m_2, f_1, f_2$, and conditional on these values, we sample either 5 or 15 
-functional observations and estimate the underlying true function based on the observations.
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{run_sim_f1A_f2B.png}
+    \caption{Illustration of simulation output, each with $r=5$ function realizations. The true function
+             is shown in solid green, while the gray lines and points indicate the realizations. 
+             Left panel: function $f_1$, variance setting $A$. 
+             Right panel: function $f_2$, variance setting $B$.}   
+    \label{fig:run_sim_f1A_f2B}
+\end{figure}
+
+For each simulated function and variance specification, we consider 20 Monte Carlo repetitions
+with re-randomized $m_1, m_2, u_1, u_2$.  We consider $r = 5$ and $r = 15$. Examples of such
+simulation output are shown in Figure \ref{fig:run_sim_f1A_f2B}.
 
@@
 as a deep process convolution (DPC) approach, which is utilized within Cosmic 
-Emu \citep{moran2023mira}.
+Emu \citep{moran2023mira}. These competing methods view the inputs as noisy observations
+(i.e., the points in Figure \ref{fig:run_sim_f1A_f2B}) with no spatial dependence. In
+contrast, our method uses the $r$ functional realizations to estimate parameters
+within $\Sigma_\epsilon$ that account for the spatial dependence.
 
-The metrics we use to evaluate performance are mean squared error (MSE) and the 
+The metric we highlight to evaluate performance of each method is the
 log score, a proper scoring rule which takes both prediction and intervals into 
-consideration \citep{gneiting2007strictly}. For both metrics, a lower value indicates 
+consideration \citep{gneiting2007strictly}. A lower value indicates 
 better performance. Log score are shown in Figure \ref{fig:sims_logS}; our DGP.FCO 
@@
 accounting for spatial dependence within our model, resulting in more accurate UQ.
-MSE results, which do not address UQ, were comparable across competing methods; these results
-can be found in Figure \ref{fig:sims_MSE}. \textcolor{magenta}{If we mention that DGP.FCO timings 
-speed up compared to \texttt{deepgp} and DPC, we would also need to mention that \texttt{hetGP} 
-is much faster than all of these.}
+Mean squared error (MSE) results, which do not address UQ, were comparable across 
+competing methods; these results can be found in Figure \ref{fig:sims_MSE}.
 
@@
 
-In addition to simulated data, we investigate model fits for different cosmological 
+We next investigate model fits to several cosmological 
 datasets from CAMB \citep{lewis2011CAMB}. These datasets contain power spectra based
 on six-dimensional cosmological parameter settings. Similar to the Mira-Titan suite, the
-CAMB dataset has multiple (sixty-four) batches of power spectra for each cosmology, with each batch 
-containing multiple (fifteen) low-resolution spectra, and one high-resolution sprectum. 
-Unlike the Mira-Titan suite, this dataset favorably contains an infinite-resolution 
+CAMB dataset contains data for sixty-four cosmologies, with each containing multiple 
+(fifteen) low-resolution spectra, and one high-resolution spectrum. 
+Unlike the Mira-Titan suite, this dataset additionally contains an infinite-resolution 
 spectrum, which can be treated as the true spectrum for each cosmology,
-allowing us to accurately assess posterior mean estimates and credible intervals.
+allowing us to assess the accuracy of posterior mean estimates and effectiveness
+of credible intervals. An illustration of the first cosmology from CAMB is shown
+in Figure \ref{fig:fit_camb}, where the infinite-resolution spectrum is generally
+much smoother than the low- and high-resolution counterparts.
 
@@
     \caption{Example of DGP.FCO model fit for the first batch of model runs from the CAMB data, 
-             with LOESS mean term subtracted. We can compare the orange lines 
-             (posterior mean and dotted 95\% credible interval) to the light blue line,
-             representing the infinite-resolution ``true" spectrum.}   
+             with a LOESS mean of the weighted average subtracted to emphasize details. We can 
+             compare the orange lines (posterior mean and dotted 95\% credible interval) to the 
+             light blue line, representing the infinite-resolution ``true" spectrum.}   
     \label{fig:fit_camb}
@@
 
-We use 32 CAMB batches to train the DGP.FCO model and reserve the other 32 as a test set.
-To imitate the role of the perturbation theory output of Mira-Titan, we anchor estimates of the CAMB
-power spectrum at the infinite-resolution for $\log_{10}(k) < -1.7$, and then rely 
-strictly on the low- and high-resolution runs for $k$ above this. We take a weighted
-average of the CAMB runs and assume heteroskedastic errors in a similar fashion.
+Our goal here is to train the DGP.FCO model and accurately estimate the true spectrum for each
+cosmology. To imitate the role of the perturbation theory output of Mira-Titan, we anchor 
+estimates of the CAMB power spectrum at the infinite-resolution for $\log_{10}(k) < -1.7$, and 
+then rely strictly on the low- and high-resolution runs for $k$ above this (note the transition
+area in Figure \ref{fig:fit_camb} around the transformed $k$ value of 0.2). We first use a log-log 
+regression model to obtain precision estimates $\Lambda_\ell$ and $\Lambda_h$
+for the low- and high-resolution runs respectively, as outlined in Section \ref{sec:data}. 
+Similarly, we set $\Lambda_c^{(ii)}=10^8$ when $\log_{10}(k_i) < -1.7$ ($0$ otherwise) for $i = 1,\dots,n$.
+\textcolor{red}{(why switch from $10^5$ to $10^8$?)} 
+to illustrate high confidence in the infinite-resolution run for low $k$ values.
+We then take a weighted average of the CAMB runs in a similar fashion
+to Mira-Titan data as described in Section \ref{sec:data}: $\bar y = \Lambda^{-1}(\Lambda_c y_c + 
+\Lambda_{\ell} \bar{y}_\ell + \Lambda_h y_h)$, where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ 
+and $\bar{y}_\ell = \frac{1}{15}\sum_{r=1}^{15} y_{\ell_r}$.
 
-We use a log-log regression model to obtain precision estimates $\Lambda_\ell$ and $\Lambda_h$
-for the low- and high-resolution runs respectively, as outlined in Section \ref{sec:data}. 
-Similarly, we set $\Lambda_i=10^8$ to illustrate high confidence in the infinite-resolution 
-run for low $k$ values and incorporate these estimates within $\Sigma_\varepsilon$. 
-With the precision information, we then estimate the Mat\'ern parameters for 
-$\Sigma_\varepsilon$. For the CAMB data, lengthscale estimates are consistently 
-around $3\times 10^{-7}$, evidence that we can simplify our covariance structure 
-to contain no spatial dependence (compare low-resolution runs of Figure \ref{fig:fit_camb} 
-with Figure \ref{fig:plot_data}). We therefore specify $\Sigma_\varepsilon = 
-(15\Lambda_\ell+\Lambda_h+\Lambda_i)^{-1}$ for the CAMB data.
+With weighted average and precision information, we considered a Mat\'ern kernel for 
+$\Sigma_\varepsilon$. 
+\textcolor{red}{(Yes, using maximum likelihood estimation with smoothness set to $5/2$, 
+and we use $\tau^2$ and $\theta$ estimates here, in the sims, and for Mira-Titan; 
+these estimates are coming out of the ``opt\_matern" function we created.)}
+For the CAMB data, spatial correlation decays very quickly
+($\theta\approx 0$ for CAMB, compare the smoothly varying low-resolution runs of 
+Figure \ref{fig:fit_camb} with the jagged low-resolution CAMB runs in Figure \ref{fig:plot_data}). 
+We therefore assume no spatial dependence for CAMB and specify $\Sigma_\varepsilon = 
+(\Lambda_c+\Lambda_\ell+\Lambda_h)^{-1}$.
 
@@
 
+\textcolor{red}{As of now, we are not comparing to DPC/hetGP/deepgp here, so no
+MSE or log scores. We do have MSEs of CAMB predictions when training on these posterior means
+vs training on the infinite-resolution runs; this comes up in Figures \ref{fig:pca_preds_v_camb} 
+and \ref{fig:mse_camb} later.}
+
 \subsection{Estimating Spectra with Mira-Titan Data}
@@
 we find they vary smoothly about some mean $S$, suggesting that a dense covariance 
-matrix $\Sigma_\varepsilon$ (accounting for spatial dependence) may outperform a 
-diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$. From Section \ref{sec:data},
-we can incorporate domain knowledge from $\Lambda_p, \Lambda_\ell,$ and $\Lambda_h$ 
-within our model. In previous work, we have modeled multiple candidates for 
+matrix $\Sigma_\varepsilon$ (accounting for domain knowledge and spatial dependence) 
+may outperform a diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$. From Section 
+\ref{sec:data}, we can incorporate domain knowledge from $\Lambda_p, \Lambda_\ell,$ 
+and $\Lambda_h$ within our model, and calculate the weighted average for each cosmology: 
+$\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_{\ell} \bar{y}_\ell + \Lambda_h y_h)$, 
+where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$. 
+In previous work, we have modeled multiple candidates for 
 $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and found that a Mat\'ern covariance 
@@
 cosmologies; the posterior mean is subtracted in order to visualize this uncertainty. 
+The UQ from the DGP.FCO model accounts for varying precisions across differing $k$ 
+values, as well as the spatial correlation within the functional realizations.
 Figure \ref{fig:plot_warp} provides elliptical slice samples of $W$ along with the 
-average warping. Some departure from stationarity is estimated in the higher $X$ 
-(i.e., scaled $k$) values.
+average warping. From these samples, we see some departure from stationarity 
+estimated in the higher $X$ (i.e., scaled $k$) values.
 
@@
 \]
-where $\beta_{ik}$ is the inverse squared lengthscale for the $k$th input dimension, $\lambda_{z_i}$ is the marginal precision for $w_i(\psi)$, and $\lambda_{s_i}$ is the nugget precision (set to zero if no nugget is used).
+where $\beta_{ik}$ is the inverse squared lengthscale for the $k$th input dimension, 
+$\lambda_{z_i}$ is the marginal precision for $w_i(\psi)$, and $\lambda_{s_i}$ is the nugget 
+precision (set to zero if no nugget is used).
 

### 2025-06-17 sims/run_sim.R

@@ -49,7 +49,7 @@ vis <- FALSE # should plots be generated
 x <- seq(0, 4, by = 0.1)
 n <- length(x)
 if(setting == 4) sddtrue <- rep(1,n)
-if(setting == 5) sddtrue <- 1/1.5^(seq(0,2,length=length(x)))
+if(setting == 5) sddtrue <- 1/1.5^(seq(0, 2, length=n))
 Sigma_true <- get_Sigma_true(x, n, func, setting)
 if(vis) image(Sigma_true) # make sure sd plot looks right
 
@@ -59,7 +59,7 @@ if(func == 1) {
 } else {
   for (i in 1:r) Y[i,] <- f2(x, m2 = m2, u2 = u2, Sigma = Sigma_true)
 }
-if(vis) matplot(x, t(Y), ylab = "f(x)", type = "l")
+if(vis) matplot(x, t(Y), ylab = "f(x)", type = "l", col = "#999999", lwd=2)
 
 if(func == 1) y_true <- f1(x, m1 = m1, u1 = u1, Sigma = diag(1e-300,n,n))
 if(func == 2) y_true <- f2(x, m2 = m2, u2 = u2, Sigma = diag(1e-300,n,n))
@@ -75,7 +75,8 @@ for(i in 1:r) {
   x_all <- c(x_all, x)
   y_all <- c(y_all, Y[i, ])
 }
-if(vis) plot(x_all, y_all)
+if(vis) points(x_all, y_all, col = rgb(0, 0, 0, alpha = 75, maxColorValue = 255))
+if(vis) lines(x, y_true, lwd=2, col = "#009E73")
 
 # dgp.fco model ---------------------------------------------------------------
 

### 2025-06-18 paper/main.tex
through 3.3

@@
 
-For both $\Sigma_S$ and $\Sigma_W$, we use Mat\`ern kernels with unit scale and
-unique lengthscales,
-\[
+For both $\Sigma_S$ and $\Sigma_W$, we use Mat\`ern kernels with smoothness of $2.5$,
+unit scale, and unique lengthscales:
+\begin{equation}\label{eq:cov}
 \begin{array}{l}
-\Sigma_S(W)^{i,j} = k\left(||w_i - w_j||^2, \theta_S\right) \\[7pt]
-\Sigma_W(X)^{i,j} = k\left(||x_i - x_j||^2, \theta_W\right)
+\Sigma_S(W)^{i,j} = K\left(||w_i - w_j||^2, \theta_S\right) \\[7pt]
+\Sigma_W(X)^{i,j} = K\left(||x_i - x_j||^2, \theta_W\right)
 \end{array}
 \quad\textrm{where}\quad
-k(d, \theta) = \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta}} + 
+K(d, \theta) = \left( 1 + \frac{\sqrt{5}d}{\sqrt{\theta}} + 
   \frac{5d^2}{3\theta}\right) \exp\left(-\frac{\sqrt{5}d}{\sqrt{\theta}}\right).
-\]
+\end{equation}
 In our one-dimensional setting, we embrace unit scale on $S$ to preserve parsimony 
@@
 
-Given burned-in posterior samples $\{W^{(t)}$ and $\theta_S^{(t)}\}$
+Given burned-in posterior samples $W^{(t)}$ and $\theta_S^{(t)}$
 for $t\in\mathcal{T}$ ($\theta_W^{(t)}$ is only used to sample $W^{(t)}$), 
@@
 \[
-S^{(t)}\mid\bar{Y}, \Sigma_\varepsilon \sim \mathcal{N}_n(m^{(t)}, C^{(t)})
+S^{(t)}\mid\bar{Y}, \Sigma_\varepsilon \sim \mathrm{GP}(m^{(t)}, C^{(t)})
 \quad\textrm{where}\quad
@@
 is provided in \ref{sec:apdx_SgivenY}.  This closed-form enables direct posterior
-sampling of $S^{(t)}$.  We draw samples accordingly and compute their average 
+sampling of $S^{(t)}$ at discrete locations $x_i$ for $i=1,\dots, n$.  
+We draw samples accordingly and compute their average 
 to obtain our posterior mean for a particular cosmology. 
@@
   f_2(x) &= \exp\left(-m_2(x-3)^2\right) + 
-  \exp\left(-u_2(x-1)^2\right)-0.05\sin\left(8(x-1.9))\right.
+  \exp\left(-u_2(x-1)^2\right)-0.05\sin\left(8(x-1.9)\right).
 \end{aligned}
@@
 $m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 
-\sim \text{Uniform}(0.6,1.4)$. The outputs of $f_1$ and $f_2$ (conditional on $m_1,u_1$ and 
+\sim \text{Uniform}(0.6,1.4)$. The outputs of $f_1$ and $f_2$ (conditional on $m_1/u_1$ and 
 $m_2,u_2$), determine the true function values for each simulation.
@@
 For each of the two true functions, we consider two variance settings $A$ and $B$, where 
-matrices $\Sigma_A$ and $\Sigma_B$ have a Mat\'ern covariance function with a nugget of 
-$10^{-8}$ and smoothness of 2.5. We obtain function realizations at $x=\{0, 0.1, \dots, 
+matrices $\Sigma_A$ and $\Sigma_B$ have a Mat\'ern kernel with smoothness of 2.5,
+as defined in Eq.~(\ref{eq:cov}). 
+We obtain function realizations at $x=\{0, 0.1, \dots, 
 3.9, 4.0\}$, so that each function draw contains $n=41$ observations. We draw $r$-many 
-realizations from a GP with some mean ($f_1$ or $f_2$) and some spatially varying covariance
-($\Sigma_A$ or $\Sigma_B$), where:
+realizations from a GP with mean $f_1$ or $f_2$ and spatially varying covariance
+$\Sigma_A$ or $\Sigma_B$, where:
 
@@
 \begin{aligned}
-\Sigma_A &= 0.0225 * k(||x_i - x_j||^2, 0.01), \\
-\Sigma_B &= s^\top Ms \quad\textrm{where}\quad M = 0.1 * k(||x_i - x_j||^2, 0.05)
-\quad\textrm{and}\quad s = 1.5^{-x/2},
+\Sigma_A &= 0.0225 * K(||x_i - x_j||^2, 0.01), \\
+\Sigma_B &= s^\top Ms \quad\textrm{where}\quad M = 0.1 * K(||x_i - x_j||^2, 0.05)
+\quad\textrm{and}\quad s = 1.5^{-x/2}.
 \end{aligned}
 \]
-
-where $k(\cdot,\cdot)$ is defined in Section \ref{sec:mod_spec}.  Within $\Sigma_B$, $s$ introduces
-nonstationarity into the function realizations, more accurately matching the behavior of the 
-CAMB and Mira-Titan motivating datasets.
+Observations are kept noise-free, but we do add a jitter of $10^{-8}$ to the diagonal of $\Sigma_A$ 
+and $\Sigma_B$ for numerical stability.  Within $\Sigma_B$, $s$ introduces a degree of
+nonstationarity, more effectively mimicing the behavior of the 
+CAMB and Mira-Titan datasets.  Examples of simulation output are showin in 
+Figure \ref{fig:run_sim_f1A_f2B}.
 
@@
              is shown in solid green, while the gray lines and points indicate the realizations. 
-             Left panel: function $f_1$, variance setting $A$. 
-             Right panel: function $f_2$, variance setting $B$.}   
+             {\it Left:} $f_1$ with variance setting $A$. 
+             {\it Right:} $f_2$ with variance setting $B$.}   
     \label{fig:run_sim_f1A_f2B}
 \end{figure}
-
-For each simulated function and variance specification, we consider 20 Monte Carlo repetitions
-with re-randomized $m_1, m_2, u_1, u_2$.  We consider $r = 5$ and $r = 15$. Examples of such
-simulation output are shown in Figure \ref{fig:run_sim_f1A_f2B}.
 
@@
              simulated scenario.  Each column represents a function/covariance 
-             pair, and each row shows results for batch sizes of either 
-             5 (top) or 15 (bottom) realizations. \textcolor{red}{(This figure needs udpating.)}}
+             pair, and each row shows results for $r=5$ (top) or $r=15$ (bottom). 
+             \textcolor{red}{(This figure needs udpating.)}}
     \label{fig:sims_logS}
@@
 
-The competing methods we consider are an out-of-the-box DGP fit using the 
-\texttt{deepgp} package \citep{sauer2023active}, a heteroskedastic GP fit using 
-the \texttt{hetGP} package \citep{binois2018practical, binois2021hetgp}, as well 
-as a deep process convolution (DPC) approach, which is utilized within Cosmic 
-Emu \citep{moran2023mira}. These competing methods view the inputs as noisy observations
-(i.e., the points in Figure \ref{fig:run_sim_f1A_f2B}) with no spatial dependence. In
-contrast, our method uses the $r$ functional realizations to estimate parameters
+For each simulated function and variance specification, we conduct 20 Monte Carlo repetitions
+with re-randomized $m_1, u_1, m_2, u_2$.  We consider $r = 5$ and $r = 15$.
+We compare to an out-of-the-box DGP from the 
+\texttt{deepgp} package \citep{sauer2023active}, a heteroskedastic GP from
+the \texttt{hetGP} package \citep{binois2018practical, binois2021hetgp}, and
+a deep process convolution (DPC) approach, which is utilized within Cosmic 
+Emu \citep{moran2023mira}.  These competing methods view the training data as 
+noisy observations with no spatial dependence (i.e., they view the gray dots of
+Figure \ref{fig:run_sim_f1A_f2B} without the lines). In
+contrast, our method leverages the fact that the observations represent $r$-many
+smooth functional realizations in order to estimate parameters
 within $\Sigma_\epsilon$ that account for the spatial dependence.
+\textcolor{red}{(@Steve - we have not provided any details on 
+the estimation of $\Sigma_\epsilon$ here.  Can you add some?)}
 
-The metric we highlight to evaluate performance of each method is the
-log score, a proper scoring rule which takes both prediction and intervals into 
-consideration \citep{gneiting2007strictly}. A lower value indicates 
-better performance. Log score are shown in Figure \ref{fig:sims_logS}; our DGP.FCO 
-model performs favorably against each of these current benchmark methods by
-accounting for spatial dependence within our model, resulting in more accurate UQ.
-Mean squared error (MSE) results, which do not address UQ, were comparable across 
-competing methods; these results can be found in Figure \ref{fig:sims_MSE}.
+We evaluate performance of each method using log score, a proper scoring 
+rule which takes both accuracy and UQ into 
+consideration \citep[][lower is better]{gneiting2007strictly}.
+Results are shown in Figure \ref{fig:sims_logS}.  Our DGP.FCO 
+model performs favorably across the board; accounting for the spatial 
+dependence between the observations results in more effective UQ.
+We relegate mean squared error (MSE) results to \ref{sec:apdx_sims}, because
+MSE neglects UQ and was comparable across competing methods.
 

### 2025-06-18 paper/main.tex
done with 3.4

@@
 
-\subsection{CAMB Data}
+\subsection{CAMB Estimation}
 \label{subsec:camb}
 
-We next investigate model fits to several cosmological 
-datasets from CAMB \citep{lewis2011CAMB}. These datasets contain power spectra based
-on six-dimensional cosmological parameter settings. Similar to the Mira-Titan suite, the
-CAMB dataset contains data for sixty-four cosmologies, with each containing multiple 
-(fifteen) low-resolution spectra, and one high-resolution spectrum. 
-Unlike the Mira-Titan suite, this dataset additionally contains an infinite-resolution 
-spectrum, which can be treated as the true spectrum for each cosmology,
-allowing us to assess the accuracy of posterior mean estimates and effectiveness
-of credible intervals. An illustration of the first cosmology from CAMB is shown
-in Figure \ref{fig:fit_camb}, where the infinite-resolution spectrum is generally
-much smoother than the low- and high-resolution counterparts.
+The CAMB \citep{lewis2011CAMB} dataset contains power spectra based
+on six cosmological parameter settings. It includes sixty-four cosmologies,
+each containing fifteen low-resolution spectra ($y_{l_r}$) and one high-resolution spectrum ($y_h$).
+Each spectrum is observed along the same grid of wavenumbers ($k_i$ for $i=1,\dots,n$). 
+Unlike the Mira-Titan suite, this dataset includes an infinite-resolution 
+spectrum ($y_c)$ which can be treated as the true spectrum for each cosmology,
+allowing us to assess model performance. An illustration of the first cosmology 
+from CAMB is shown in Figure \ref{fig:fit_camb}.  The ``true'' infinite-resolution 
+spectrum is notably smoother than its low- and high-resolution counterparts.
 
@@
              compare the orange lines (posterior mean and dotted 95\% credible interval) to the 
-             light blue line, representing the infinite-resolution ``true" spectrum.}   
+             light blue line, representing the infinite-resolution ``true" spectrum.
+             \textcolor{red}{(@Steve - can you leave the x-axis on the usual scale?
+             So -1.7 is directly visible.)}}   
     \label{fig:fit_camb}
@@
 
-Our goal here is to train the DGP.FCO model and accurately estimate the true spectrum for each
-cosmology. To imitate the role of the perturbation theory output of Mira-Titan, we anchor 
-estimates of the CAMB power spectrum at the infinite-resolution for $\log_{10}(k) < -1.7$, and 
-then rely strictly on the low- and high-resolution runs for $k$ above this (note the transition
-area in Figure \ref{fig:fit_camb} around the transformed $k$ value of 0.2). We first use a log-log 
-regression model to obtain precision estimates $\Lambda_\ell$ and $\Lambda_h$
+To imitate the role of the perturbation theory output of Mira-Titan, we anchor 
+estimates of the CAMB power spectrum at the infinite-resolution for 
+$\log_{10}(k) < -1.7$, and then rely strictly on the low- and high-resolution 
+runs for $k$ above this (note the transition
+in Figure \ref{fig:fit_camb} at $\log_{10}(k) = -1.7$). To compute $\bar{y}$ and $\Sigma_\epsilon$,
+we first use a log-log regression model to obtain precision estimates $\Lambda_\ell$ and $\Lambda_h$
 for the low- and high-resolution runs respectively, as outlined in Section \ref{sec:data}. 
-Similarly, we set $\Lambda_c^{(ii)}=10^8$ when $\log_{10}(k_i) < -1.7$ ($0$ otherwise) for $i = 1,\dots,n$.
-\textcolor{red}{(why switch from $10^5$ to $10^8$?)} 
-to illustrate high confidence in the infinite-resolution run for low $k$ values.
-We then take a weighted average of the CAMB runs in a similar fashion
-to Mira-Titan data as described in Section \ref{sec:data}: $\bar y = \Lambda^{-1}(\Lambda_c y_c + 
-\Lambda_{\ell} \bar{y}_\ell + \Lambda_h y_h)$, where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$ 
+To anchor estimates at low resolutions, we define the diagonal matrix $\Lambda_c$ with 
+$\Lambda_c^{(ii)}=10^8$ for $\log_{10}(k_i) < -1.7$ ($0$ otherwise).
+We then compute the weighted average as: $\bar y = \Lambda^{-1}(\Lambda_c y_c + 
+\Lambda_{\ell} \bar{y}_\ell + \Lambda_h y_h)$, where $\Lambda = \Lambda_c + \Lambda_\ell + \Lambda_h$ 
 and $\bar{y}_\ell = \frac{1}{15}\sum_{r=1}^{15} y_{\ell_r}$.
-
-With weighted average and precision information, we considered a Mat\'ern kernel for 
-$\Sigma_\varepsilon$. 
-\textcolor{red}{(Yes, using maximum likelihood estimation with smoothness set to $5/2$, 
-and we use $\tau^2$ and $\theta$ estimates here, in the sims, and for Mira-Titan; 
-these estimates are coming out of the ``opt\_matern" function we created.)}
 For the CAMB data, spatial correlation decays very quickly
-($\theta\approx 0$ for CAMB, compare the smoothly varying low-resolution runs of 
-Figure \ref{fig:fit_camb} with the jagged low-resolution CAMB runs in Figure \ref{fig:plot_data}). 
-We therefore assume no spatial dependence for CAMB and specify $\Sigma_\varepsilon = 
+(compare the smoothly varying low-resolution runs of Mira-Titan in Figure \ref{fig:plot_data} 
+with the jagged low-resolution runs of CAMB in Figure \ref{fig:fit_camb}). 
+Although we considered a Mat\'ern kernel for $\Sigma_\epsilon$ here, we found the covariance
+to decay so quickly ($\theta\approx 0$) that a diagonal covariance would suffice.
+Therefore, we drop spatial dependence and specify $\Sigma_\varepsilon = 
 (\Lambda_c+\Lambda_\ell+\Lambda_h)^{-1}$.
 
-After fitting the DGP.FCO model, an illustration of the posterior mean and its UQ can
-be seen in Figure \ref{fig:fit_camb}. Notice that there are peaks where both the high-resolution
-run and the weighted average are biased, but our posterior mean adjusts appropriately 
-and has a more reliable estimate for the infinite-resolution run (the ``truth" in this setting). 
-UQ is handled well, as precision increases with increasing $k$, which is captured in our
-95\% credible intervals.
+We fit our DGP.FCO model to the low- and high-resolution runs of each CAMB cosmology for
+later use in Section \ref{sec:pred}.  For now, we provide an illustration of the DGP.FCO
+prediction (solid yellow) and credible intervals (dashed yellow) for a single 
+CAMB cosmology in Figure \ref{fig:fit_camb}.  Our posterior mean provides an effective estimate
+of the ``true'' infinite-resolution run.  It aptly smooths over the peaks where 
+both the high-resolution run and the weighted average significantly deviate from the truth
+($\log_{10}(k)\approx 0.4,0.7$ - \textcolor{red}{@Steve - please adjust these numbers after 
+changing the x-axis of the figure}). Our model also provides effective UQ, appropriately shrinking
+credible intervals as precision increases for larger $k$.
 
-\textcolor{red}{As of now, we are not comparing to DPC/hetGP/deepgp here, so no
-MSE or log scores. We do have MSEs of CAMB predictions when training on these posterior means
-vs training on the infinite-resolution runs; this comes up in Figures \ref{fig:pca_preds_v_camb} 
-and \ref{fig:mse_camb} later.}
-
-\subsection{Estimating Spectra with Mira-Titan Data}
+\subsection{Mira-Titan Estimation}
 \label{subsec:mira_fit}

### 2025-06-18 plotting/plot_data_fit.R
update images for manuscript

@@ -11,6 +11,8 @@
 library(dgp.hm)
 library(zoo)
 
+JPG = FALSE
+
 # Get model data
 model <- 1
 if (model <= 111) {
@@ -65,6 +67,7 @@ cbcols <- palette.colors(palette = "Okabe-Ito")
 #6 blue, 7 vermillion,  8 reddishpurple, 9 gray 
 
 # plot_data.png
+if(JPG) jpeg("../paper/plot_data.jpeg", width = 12, height = 4, units = "in", res = 300)
 par(mfrow=c(1,3), mar=c( 5.6, 4.6, 4.6, 2.1))
 matplot(log10(k), y_lo, col = cbcols[9], 
         type = "l", lty = 1, ylim = range(y_avg), lwd=0.5,
@@ -104,7 +107,7 @@ lines(log10(k), y_pt-fit$m, col = cbcols[4])
 lines(log10(k), y_hi-fit$m, col = cbcols[8], lwd=2)
 lines(log10(k), y_avg-fit$m, 
       col = cbcols[2], lty = 2, lwd = 3)
-
+if(JPG) dev.off()
 
 ######################
 # Plot fitted values #
@@ -126,6 +129,7 @@ legend(x = "bottomright", legend = c("pert","low res","hi res", "wt avg", "UQ"),
        lwd = c(2,1,2,2,2))
 
 # plot_fit.png
+if(JPG) jpeg("../paper/plot_fit.jpeg", width = 12, height = 4, units = "in", res = 300)
 # Plot model fit alongside data (posterior mean removed)
 matplot(log10(k), y_lo - fit$m, type="l", lty=3,
         col=cbcols[9], ylim = c(-.02,.02), xlab=expression(log[10](k)), 
@@ -138,7 +142,8 @@ lines(log10(k), fit$ub - fit$m, col=cbcols[1],lty=2,lwd=2)
 lines(log10(k), fit$lb - fit$m, col=cbcols[1],lty=2,lwd=2)
 legend(x = "bottomright", legend = c("pert","low res","hi res", "wt avg", "UQ"),
        col = cbcols[c(4,9,8,2,1)], lty = c(4,3,2,1,2), 
-       lwd = c(2,1,2,2,2), cex=0.67, bty="n")
+       lwd = c(2,1,2,2,2), cex=0.9, bty="n")
+if(JPG) dev.off()
 
 # plot samples and average warping
 plot.warp <- function(fit, wl = 1, wh = length(fit$x), ref.scale = 1, ...){

### 2025-06-18 paper/main.tex
done with Section 3

@@
 $\Lambda_\ell$, and $\Lambda_h$ as the $n\times n$ diagonal matrices with elements
-\[
+\begin{equation}\label{eq:lambda}
 \Lambda_p^{(ii)} = \begin{cases}
@@
     \end{cases}
-\]
+\end{equation}
 for $i=1,\dots, n$.  In $\Lambda_p$, we use the sufficiently high precision 
@@
 
-From plots of the multiple low-resolution spectra (e.g., Figure \ref{fig:plot_data}), 
-we find they vary smoothly about some mean $S$, suggesting that a dense covariance 
-matrix $\Sigma_\varepsilon$ (accounting for domain knowledge and spatial dependence) 
-may outperform a diagonal covariance $\Lambda_\ell^{-1}$ for $\bar Y$. From Section 
-\ref{sec:data}, we can incorporate domain knowledge from $\Lambda_p, \Lambda_\ell,$ 
-and $\Lambda_h$ within our model, and calculate the weighted average for each cosmology: 
+Here, we revisit the Mira-Titan suite described in Section \ref{sec:data} to provide
+final details for our estimation of the true (but unknown) matter power spectrum for
+each cosmology.  Recall, with $\Lambda_p$, $\Lambda_\ell$, and $\Lambda_h$ as 
+defined in Eq.~(\ref{eq:lambda}), we compute
 $\bar y = \Lambda^{-1}(\Lambda_p y_p + \Lambda_{\ell} \bar{y}_\ell + \Lambda_h y_h)$, 
 where $\Lambda = \Lambda_p + \Lambda_\ell + \Lambda_h$.
-In previous work, we have modeled multiple candidates for 
+Turning to $\Sigma_\epsilon$, since the low-resolution spectra 
+vary smoothly about $S$ (e.g., Figure \ref{fig:plot_data}), we prefer a dense 
+covariance ($\Sigma_\ell$) for wavelengths $0.04 \leq k < 0.25$.
+In previous work, we considered multiple candidates for 
 $\Sigma_\varepsilon$ \citep{walsh2023bayesian} and found that a Mat\'ern covariance 
-$\Sigma_\ell$ trained on the low-resolution runs (pre-scaled by the precision 
+for $\Sigma_\ell$ trained on the low-resolution runs (pre-scaled by the precision 
 values and subtracting a LOESS-smoothed average) performs the best. 
@@
 to \cite{walsh2023bayesian} for a more detailed exposition. This yields 
-$\Sigma_\varepsilon=\left(\Lambda_p^{-1} + \Sigma_\ell + \Lambda_h^{-1}\right)^{-1}$, 
-where $\Lambda_p, \Sigma_\ell^{-1}, \text{ and } \Lambda_h$ are all $n\times n$ 
-and are 0 wherever the corresponding data type is biased. From here, we estimate 
-$\tau_\ell, \theta_\ell$ with maximum likelihood, and input these within 
-$\Sigma_\varepsilon$, which we use as an input to our model.
+$\Sigma_\varepsilon=\left(\Lambda_p^{-1} + \Sigma_\ell + \Lambda_h^{-1}\right)^{-1}$.
 
-To train our hierarchical model, we initialize our hyperparameter estimates across 
-all cosmologies by modeling the first (M001) with 50,000 MCMC runs, and use these 
-as initial values for all cosmologies. Specifically, we save the last MCMC 
-iteration of $W$, $\theta_w$ and $\theta_S$ as $W_0$, $\theta_{w0}$ and $\theta_{S0}$ 
-respectively and use each as starting values for all 111 training cosmology fits, 
-where each cosmology is fit with 10,000 MCMC runs (5,000 removed as burn-in and 
-every 5th sample is kept, resulting in 1,000 posterior samples).
-
-From these samples, we can estimate and quantify the uncertainty within each 
-cosmology's average and latent layer $W$. Figure \ref{fig:plot_fit} shows an 
-example of the uncertainty for our posterior mean for the first (of 111) training 
-cosmologies; the posterior mean is subtracted in order to visualize this uncertainty. 
-The UQ from the DGP.FCO model accounts for varying precisions across differing $k$ 
-values, as well as the spatial correlation within the functional realizations.
-Figure \ref{fig:plot_warp} provides elliptical slice samples of $W$ along with the 
-average warping. From these samples, we see some departure from stationarity 
-estimated in the higher $X$ (i.e., scaled $k$) values.
+We then fit our DGP.FCO model on each of 111 training cosmologies, reserving 6 cosmologies
+for later hold-out testing.  To speed-up burn-in of MCMC chains, we conduct
+50,000 MCMC iterations for the first cosmology, then initialize the chains for all 
+other cosmologies with these burned-in samples, so fewer iterations are needed for the
+remaining 110 cosmologies.  Ultimately, we conduct 10,000 iterations for each of these,
+removing 5,000 for burn-in, then thinning by 5.  Figure \ref{fig:plot_fit} 
+visualizes the DGP.FCO fit for the first training cosmology.  We center the plot
+by subtracting the posterior mean in order to more easily visualize the credible
+interval bounds.  The UQ from the DGP.FCO model accounts for varying precisions 
+across differing $k$ values, as well as the spatial correlation within the 
+functional realizations.  To provide insight into the nonstationarity of the power
+matter spectrum, Figure \ref{fig:plot_warp} shows the elliptical slice samples of $W$.
+These warpings depart from the identiy mapping around $x=0.4$ \textcolor{red}{(@Steve - 
+please change the x-axis to match the actual x-axis values, then adjust this number)}, 
+stretching inputs where dynamics are shifting more rapidly for larger $k$ values.
 
@@
     \includegraphics[width=6in]{plot_fit.png}
-    \caption{95\% credible intervals for the power spectrum of model 1 (posterior mean 
-             subtracted). Plots of the perturbation theory, low-resolution and high-resolution 
-             runs are also shown, alongside the weighted average.}
+    \caption{95\% credible intervals for the power spectrum of the first Mira-Titan
+    cosmology (centered by the predicted posterior mean). Perturbation theory, 
+    low-resolution, and high-resolution runs are shown for context.
+    \textcolor{red}{(@Steve - combine this plot with Figure 6.  The two panels of Figure 6
+    are redundant and unecessary (yet oddly, they don't match?).  Have this Mira-Titan 
+    plot take up 2/3 of the space, then have just the right panel from Figure 6 take 
+    up the remaining 1/3.  Update the figure caption and referencing accordingly.)}}
     \label{fig:plot_fit}
@@
             from the identity line represents nonstationary behavior.
-            \textcolor{red}{(This plot needs to be made fresh (not using the package
-             function.  Adjust y-axis.  We don't need the yellow/red coloring.)}}
+            \textcolor{red}{(remove this figure, combine right panel with previous)}}
    \label{fig:plot_warp}

### 2025-06-18 paper/main.tex
passing back to Steve

@@
 
-Upon obtaining the posterior means from each of the 111 training cosmologies, this 
-section details our second contribution - the use of the Bayesian DGP.FCO model 
-to predict the curve for held-out cosmologies. This is achieved with modeling the 
-weights of different basis functions as GPs.
+Here we detail our second contribution - using the Bayesian DGP.FCO model predictions
+from training cosmologies to predict the matter power spectra for held-out cosmologies. 
+%This is achieved with modeling the weights of different basis functions as GPs.
 

### 2025-06-18 plotting/plot_data_CAMB.R
update images

@@ -17,6 +17,8 @@ library(zoo) # rollmean
 library(Matrix) # bdiag
 library(pracma) #interp1
 
+JPG = FALSE
+
 load(file = "../fitting/results/CAMB/fit_1camb_image.rda")
 
 # # Plot the first batch of model runs for CAMB
@@ -45,23 +47,24 @@ cbcols <- palette.colors(palette = "Okabe-Ito")
 # Plot showing dgp.hm fit compared with camb
 # Plot with all of [-2.5, -2.2] camb
 # CAMB_fit_model1
-png("../paper/CAMB_fit_model1.png", width = 11, height = 4.5, 
+if(JPG) jpeg("../paper/CAMB_fit_model1.jpeg", width = 11, height = 4.5, 
     units = "in", res=300)
 par(mar=c(4.5, 4.5, 1.5, 1.5) + 0.1)
-plot(x, y_cambi - loess_fit$fitted, type="l", col=cbcols[3],
+plot(x*(b-a)+a, y_cambi - loess_fit$fitted, type="l", col=cbcols[3],
      ylim = range(y_cambi - loess_fit$fitted)+c(-.025,.025),
-     xlab=expression(paste(log[10](k),", transformed to [0,1]")), 
+     xlab=expression(paste(log[10](k))), 
      ylab=expression(paste("\U1D4AB","  (k), centered")), cex.lab = 1.35)
      # main = paste(model, "cover", coverages[model]))
 abline(h=0, col="black", lty=2)
-for (i in 1:n_lo) lines(x, y_loi[,i] - loess_fit$fitted, col=cbcols[9], lwd=0.4)
-lines(x, y_hii - loess_fit$fitted, col=cbcols[8], lwd=1.5)
-lines(x, y_avg - loess_fit$fitted, col=cbcols[4], lwd=2, lty=2)
-lines(x, fit$m - loess_fit$fitted , col=cbcols[2], lwd=2)
-lines(x, fit$ub - loess_fit$fitted, col=cbcols[2], lwd=2, lty=3)
-lines(x, fit$lb - loess_fit$fitted, col=cbcols[2], lwd=2, lty=3)
-lines(x, y_cambi - loess_fit$fitted, col=cbcols[3], lwd=2)
[…]

### 2025-06-18 plotting/plot_data_fit.R
subtract loess smoothed wt avg

@@ -98,14 +98,16 @@ legend(x = "topleft", legend = c("pert","low res","hi res", "wt avg"),
        inset = 0.01, bty="n"  # nudges legend inward
 )
 
-# same as first plot, but posterior mean subtracted
-matplot(log10(k), y_lo - fit$m, col = cbcols[9], xlim = c(log10(0.04),log10(0.25)),
+loess_fit <- loess(y_avg ~ x, span = 0.15)
+
+# same as first plot, but LOESS-smoothed weighted average subtracted
+matplot(log10(k), y_lo - loess_fit$fitted, col = cbcols[9], xlim = c(log10(0.04),log10(0.25)),
         type = "l", lty = 1, ylim = c(-.02, .02), lwd=0.5,
         xlab=expression(log[10](k)), cex.axis=1.75, cex.lab=1.75,
         ylab=expression(paste("\U1D4AB","  (k)", ", centered")))
-lines(log10(k), y_pt-fit$m, col = cbcols[4])
-lines(log10(k), y_hi-fit$m, col = cbcols[8], lwd=2)
-lines(log10(k), y_avg-fit$m, 
+lines(log10(k), y_pt-loess_fit$fitted, col = cbcols[4])
+lines(log10(k), y_hi-loess_fit$fitted, col = cbcols[8], lwd=2)
+lines(log10(k), y_avg-loess_fit$fitted, 
       col = cbcols[2], lty = 2, lwd = 3)
 if(JPG) dev.off()
 

### 2025-06-19 dgp.hm/R/fit.R
add storage of w in monowarped version

@@ -117,6 +117,7 @@ gibbs_two_layer_hm <- function(x, y, x_grid, nmcmc, verb, initial, settings, v,
   
   ng <- nrow(x_grid)
   dx_grid <- deepgp::sq_dist(x_grid)
+  grid_index <- deepgp:::fo_approx_init(x_grid, x)
   
   theta_y <- vector(length = nmcmc)
   theta_y[1] <- initial$theta_y
@@ -124,13 +125,13 @@ gibbs_two_layer_hm <- function(x, y, x_grid, nmcmc, verb, initial, settings, v,
   theta_w[1] <- initial$theta_w
   w_grid <- matrix(nrow = ng, ncol = nmcmc)
   w_grid[, 1] <- deepgp:::fo_approx(x, initial$w, x_grid) # snap initial$w to grid
+  w <- matrix(nrow = nrow(x), ncol = nmcmc)
+  w[, 1] <- monowarp_ref(x, x_grid, w_grid[, 1], grid_index)
   ll_store <- vector(length = nmcmc)
   ll_store[1] <- NA
   ll_outer <- NULL
   
-  grid_index <- deepgp:::fo_approx_init(x_grid, x)
-  w <- monowarp_ref(x, x_grid, w_grid[, 1], grid_index)
-  dw <- deepgp::sq_dist(w)
+  dw <- deepgp::sq_dist(w[, 1])
   
   for (j in 2:nmcmc) {
     if (verb & (j %% 500 == 0)) cat(j, "\n")
@@ -156,21 +157,23 @@ gibbs_two_layer_hm <- function(x, y, x_grid, nmcmc, verb, initial, settings, v,
                         theta_y[j], theta_w[j], ll_prev = ll_outer, v = v, 
                         Sigma_hat = Sigma_hat)
     w_grid[, j] <- samp$w_grid
+    w[, j] <- samp$w
     dw <- samp$dw
     ll_outer <- samp$ll
     ll_store[j] <- ll_outer
   }
   
-  return(list(theta_y = theta_y, theta_w = theta_w, w_grid = w_grid, ll = ll_store))
+  return(list(theta_y = theta_y, theta_w = theta_w, w = w, w_grid = w_grid, 
+              ll = ll_store))
 }
 
 # sample_w_hm -----------------------------------------------------------------
 
-sample_w_hm <- function(y, w_grid_t, dw_t, x, x_grid, dx_grid, grid_index, 
[…]

### 2025-06-19 dgp.hm/R/plot.R

@@ -41,19 +41,13 @@ plot.dgp2hm <- function(x, ...) {
          main = paste0('Trace Plot of theta_w'))
   
   # Hidden layer plot
-  grid_index <- deepgp:::fo_approx_init(x$x_grid, x$x)
   indx <- floor(seq(from = 1, to = x$nmcmc, length = 100))
   if (indx[1] == 0) indx[1] <- 1
   col <- heat.colors(100 + 10) # add ten to avoid using colors that are too light
   par(mfrow = c(1, 1), mar = c(4, 4, 2, 2))
   o <- order(x$x)
-  w <- monowarp_ref(x$x, x$x_grid, x$w_grid[, indx[1]], grid_index)
-  plot(x$x[o], w[o] - mean(w), type = 'l', xlab = 'X', 
-       ylab = 'W', col = col[1], main = paste0('MCMC samples of X to W'), 
-       ylim = c(min(x$w[, indx]), max(x$w[, indx])))
-  for (j in 2:length(indx)) {
-    w <- monowarp_ref(x$x, x$x_grid, x$w_grid[, indx[j]], grid_index)
-    lines(x$x[o], w[o] - mean(w), col = col[j])
-  }
-
+  w <- x$w[o, indx]
+  for (i in 1:length(indx)) w[, i] <- w[, i] - mean(w[, i])
+  matplot(x$x, w, type = 'l', xlab = 'X', ylab = 'W', col = col[1], 
+          main = paste0('MCMC samples of X to W'), ylim = c(min(w), max(w)))
 }

### 2025-06-19 dgp.hm/R/predict.R

@@ -14,13 +14,7 @@ est_true <- function(fit, tolpower = -10, return_all = FALSE) {
     stop("only applicable to gphm or dgp2hm classes")
   
   deep <- (class(fit) == "dgp2hm")
-  
-  # get squared distance matrix
-  if (deep) { # pre-calculate grid indices
-    grid_index <- deepgp:::fo_approx_init(fit$x_grid, fit$x)
-  } else { # pre-calculate pairwise distances for x
-    d2 <- deepgp::sq_dist(fit$x)
-  }
+  if (!deep) d2 <- deepgp::sq_dist(fit$x)
   
   # From Section 3.3.2: Bayesian UQ Algorithm (Walsh dissertation 2023)
   # Get Sigma_epsilon (covariance matrix for correlated errors) and its inverse
@@ -42,7 +36,7 @@ est_true <- function(fit, tolpower = -10, return_all = FALSE) {
     # construct the covariance matrix for the true spectrum (Sigma_S)
     # cov mtx of [S]; based on a matern covariance function
     if (deep) {
-      w <- monowarp_ref(fit$x, fit$x_grid, fit$w_grid[, i], grid_index)
+      w <- fit$w[, i]
       d2 <- deepgp::sq_dist(w)
       theta <- fit$theta_y[i]
     } else theta <- fit$theta[i]

### 2025-06-19 dgp.hm/R/trim.R

@@ -51,6 +51,7 @@ trim.dgp2hm <- function(object, burn, thin = 1) {
   object$theta_y <- object$theta_y[indx, drop = FALSE]
   object$theta_w <- object$theta_w[indx, drop = FALSE]
   object$w <- object$w[, indx, drop = FALSE]
+  object$w_grid <- object$w_grid[, indx, drop = FALSE]
   object$ll <- object$ll[indx, drop = FALSE]
   
   toc <- proc.time()[[3]]

### 2025-06-19 dgp.hm/R/fit.R
reduce reliance on deepgp package

@@ -30,9 +30,9 @@ fit_one_layer_hm <- function(x, y, Sigma_hat, nmcmc = 10000, verb = TRUE, theta_
   if (!is.matrix(x)) {
     x <- matrix(x, ncol = 1)
   } else if (ncol(x) != 1) stop("x must be one-dimensional")
-  test <- deepgp:::check_inputs(x, y, 0) # do not need to check nugget
-  settings <- deepgp:::check_settings(settings, layers = 1, noisy = TRUE) # TRUE/FALSE??
-  initial <- list(theta = theta_0) # force tau2 = 1
+  if (nrow(x) != length(y)) stop("dimensions of x and y do not match")
+  settings <- check_prior_settings(settings, layers = 1)
+  initial <- list(theta = theta_0)
   if (!(v %in% c(0.5, 1.5, 2.5))) 
     stop("v must be one of 0.5, 1.5, or 2.5")
   if (nrow(Sigma_hat) != nrow(x))
@@ -61,10 +61,11 @@ fit_two_layer_hm <- function (x, y, Sigma_hat, nmcmc = 10000, verb = TRUE,
   if (!is.matrix(x)) {
     x <- matrix(x, ncol = 1)
   } else if (ncol(x) != 1) stop("x must be one-dimensional")
-  test <- deepgp:::check_inputs(x, y, 0) # do not need to check nugget
-  settings <- deepgp:::check_settings(settings, layers = 2, noisy = TRUE) # TRUE/FALSE??
-  initial <- list(w = w_0, theta_y = theta_y_0, theta_w = theta_w_0) # force tau2 = 1
-  initial <- deepgp:::check_initialization(initial, layers = 2, x = x, D = 1)
+  if (nrow(x) != length(y)) stop("dimensions of x and y do not match")
+  settings <- check_prior_settings(settings, layers = 2)
+  initial <- list(w = ifelse(is.null(w_0), x, w_0), 
+                  theta_y = theta_y_0, 
+                  theta_w = theta_w_0)
   if (is.null(x_grid)) 
     x_grid <- matrix(seq(0, 1, length = 50), ncol = 1)
   if (!(v %in% c(0.5, 1.5, 2.5))) 
@@ -98,8 +99,8 @@ gibbs_one_layer_hm <- function(x, y, nmcmc, verb, initial, settings, v,
   
   for (j in 2:nmcmc) {
     if (verb & (j %% 500 == 0)) cat(j, "\n")
[…]

### 2025-06-19 dgp.hm/R/misc.R

@@ -114,3 +114,23 @@ monowarp_ref <- function(x, xg, wg, index) {
 
   return(w)
 }
+
+# Check prior settings --------------------------------------------------------
+
+check_prior_settings <- function(settings, layers = 1) {
+  
+  if (is.null(settings$l)) settings$l <- 1
+  if (is.null(settings$u)) settings$u <- 2
+
+  if (layers == 1) {
+    if (is.null(settings$theta$alpha)) settings$theta$alpha <- 1.2
+    if (is.null(settings$theta$beta)) settings$theta$beta <- 4
+  } else if (layers == 2) {
+    if (is.null(settings$theta_w$alpha)) settings$theta_w$alpha <- 1.2
+    if (is.null(settings$theta_w$beta)) settings$theta_w$beta <- 2
+    if (is.null(settings$theta_y$alpha)) settings$theta_y$alpha <- 1.2
+    if (is.null(settings$theta_y$beta)) settings$theta_y$beta <- 1
+  }
+  
+  return(settings)
+}

### 2025-06-19 dgp.hm/R/plot.R

@@ -43,11 +43,10 @@ plot.dgp2hm <- function(x, ...) {
   # Hidden layer plot
   indx <- floor(seq(from = 1, to = x$nmcmc, length = 100))
   if (indx[1] == 0) indx[1] <- 1
-  col <- heat.colors(100 + 10) # add ten to avoid using colors that are too light
   par(mfrow = c(1, 1), mar = c(4, 4, 2, 2))
   o <- order(x$x)
   w <- x$w[o, indx]
   for (i in 1:length(indx)) w[, i] <- w[, i] - mean(w[, i])
-  matplot(x$x, w, type = 'l', xlab = 'X', ylab = 'W', col = col[1], 
+  matplot(x$x, w, type = 'l', xlab = 'X', ylab = 'W', col = "grey",
           main = paste0('MCMC samples of X to W'), ylim = c(min(w), max(w)))
 }

### 2025-06-19 dgp.hm/R/fit.R
run_fit ready to roll

@@ -63,9 +63,8 @@ fit_two_layer_hm <- function (x, y, Sigma_hat, nmcmc = 10000, verb = TRUE,
   } else if (ncol(x) != 1) stop("x must be one-dimensional")
   if (nrow(x) != length(y)) stop("dimensions of x and y do not match")
   settings <- check_prior_settings(settings, layers = 2)
-  initial <- list(w = ifelse(is.null(w_0), x, w_0), 
-                  theta_y = theta_y_0, 
-                  theta_w = theta_w_0)
+  if (is.null(w_0)) w_0 <- x
+  initial <- list(w = w_0, theta_y = theta_y_0, theta_w = theta_w_0)
   if (is.null(x_grid)) 
     x_grid <- matrix(seq(0, 1, length = 50), ncol = 1)
   if (!(v %in% c(0.5, 1.5, 2.5))) 

### 2025-06-19 fitting/initialize.R

@@ -116,9 +116,9 @@ fit <- fit_two_layer_hm(x, y_avg, nmcmc = 50000, Sigma_hat = Sigma_hat)
 fit <- trim(fit, 49000, 10)
 
 # Unscale results before storing
-w0 <- c(fit$w[[length(fit$w)]])
-params0 <- data.frame(theta_y0 = fit$theta_y[length(fit$theta_y)],
-                      theta_w0 = fit$theta_w[length(fit$theta_w)])
+w0 <- c(fit$w[, fit$nmcmc])
+params0 <- data.frame(theta_y0 = fit$theta_y[fit$nmcmc],
+                      theta_w0 = fit$theta_w[fit$nmcmc])
 
 write.csv(params0, paste0("results/params0.csv"), row.names = FALSE)
 write.csv(w0, paste0("results/w0.csv"), row.names = FALSE)

### 2025-06-19 fitting/run_fit.R

@@ -19,7 +19,6 @@ library(Matrix) # bdiag
 # Read command line arguments -------------------------------------------------
 
 model <- 1 # integer 0-116 (0, 112:116 are testing)
-deep <- 1 
 
 args <- commandArgs(TRUE)
 if(length(args) > 0) 
@@ -27,7 +26,6 @@ if(length(args) > 0)
     eval(parse(text = args[[i]]))
 
 cat("model is ", model, "\n")
-if (deep) cat("model is deep \n") else cat("model is NOT deep \n")
 
 # Load data for a particular model --------------------------------------------
 
@@ -88,8 +86,8 @@ y_lra <- rowMeans(y_lo)
 # See "Weighted average from multiple computer experiments"
 # in Walsh dissertation: 3.7.1 Appendix E: Derivations
 y_avg <- (1 / Lam_z) * (Lam_pt * y_pt + Lam_lo * y_lra + Lam_hi * y_hi)
-write.csv(y_avg, file = paste0("../CosmicEmu_etc/y_avg_",model,".csv"),
-          row.names = F)
+#write.csv(y_avg, file = paste0("../CosmicEmu_etc/y_avg_",model,".csv"),
+#          row.names = F)
 
 # Scale the responses
 mean_y <- mean(y_avg)
@@ -130,20 +128,9 @@ Sigma_hat <- solve(block1 + block2 + block3)
 
 # Run MCMC --------------------------------------------------------------------
 
-if (deep) {
-  # load in initialized estimates for warping and hyperparameters for fit
-  w_0 <- read.csv("results/w0.csv")[[1]]
-  params0 <- read.csv("results/params0.csv")
-  fit <- fit_two_layer_hm(x, y_avg, nmcmc = 10000, w_0 = w_0, 
-                          theta_y_0 = params0$theta_y0,
-                          theta_w_0 = params0$theta_w0,
-                          Sigma_hat = Sigma_hat)
-} else {
-  fit <- fit_one_layer_hm(x, y_avg, nmcmc = 10000, Sigma_hat = Sigma_hat)
-}
-
+fit <- fit_two_layer_hm(x, y_avg, nmcmc = 3000, Sigma_hat = Sigma_hat)
 # plot(fit) # optionally investigate trace plots
-fit <- trim(fit, 5000, 5)
+fit <- trim(fit, 1000, 2)
 fit <- est_true(fit)
 
 # Unscale results before storing

### 2025-06-19 fitting/run_fit.R
decrease iterations

@@ -128,7 +128,7 @@ Sigma_hat <- solve(block1 + block2 + block3)
 
 # Run MCMC --------------------------------------------------------------------
 
-fit <- fit_two_layer_hm(x, y_avg, nmcmc = 3000, Sigma_hat = Sigma_hat)
+fit <- fit_two_layer_hm(x, y_avg, nmcmc = 2000, Sigma_hat = Sigma_hat)
 # plot(fit) # optionally investigate trace plots
 fit <- trim(fit, 1000, 2)
 fit <- est_true(fit)
@@ -141,7 +141,7 @@ results <- data.frame(x = log10(k),
                       lb = fit$lb * sd_y + mean_y,
                       ubb = fit$ubb * sd_y + mean_y, 
                       lbb = fit$lbb * sd_y + mean_y)
-write.csv(results, paste0("results/", ifelse(deep, "dgp", "gp"), "_", 
+write.csv(results, paste0("results/dgp_", 
                           model_name, ".csv"), row.names = FALSE)
 if(model == 1) write.table(log10(k), file = "results/logk.txt",row.names = FALSE, col.names = FALSE)
 if(model_name == "M001") save(fit, file = "results/fit_M001.rda")

### 2025-06-19 fitting/run_fit.R
chg prec to 10^8

@@ -51,7 +51,7 @@ load("../Mira-Titan-IV-Data/precision_and_indexes.Rdata")
 # Get precision info for each data type (low-res, hi-res, pert theory)
 precs_lo <- ifelse(1:n %in% index_list$lowres.ix, prec_lowres, 0) * nrun
 precs_hi <- ifelse(1:n %in% index_list$highres.ix, prec_highres, 0)
-precs_pt <- ifelse(1:n %in% index_list$pert.ix, 100000, 0)
+precs_pt <- ifelse(1:n %in% index_list$pert.ix, 10^8, 0)
 
 # Smooth the precision information (to remove steps)
 Lam_lo <- rollmean(precs_lo, k = 10, fill = "extend")

### 2025-06-19 paper/main.tex
updates to paper

@@
     \centering 
-    \includegraphics[width=\linewidth]{plot_data.png}
+    \includegraphics[width=\linewidth]{plot_data.jpeg}
     \caption{{\it Left:} The perturbation theory (solid green), low resolution runs (solid grey), 
@@
     runs are approximately unbiased.
-    {\it Right:} Same as middle, but with estimated posterior mean subtracted.
-    \textcolor{red}{(@Steve - please fix the legend so it is not going over the lines)}}
+    {\it Right:} Same as middle, but with LOESS-smooth weighted average subtracted.}
     \label{fig:plot_data}
@@
 underlying spectrum, the weighted average fluctuates too drastically.  
-This is especially apparent after centering (right panel).
-\textcolor{red}{(@Steve - how are we centering?)}
+This is especially apparent after subtracting a LOESS-smoothed weighted average (right panel).
 Although the weighted average incorporates all available observatons and 
@@
     \centering
-    \includegraphics[width=6in]{sims_logS.png}
+    \includegraphics[width=\textwidth]{sims_logS.jpeg}
     \caption{Boxplot of log scores (20 repetitions, lower is better) for each 
              simulated scenario.  Each column represents a function/covariance 
-             pair, and each row shows results for $r=5$ (top) or $r=15$ (bottom). 
-             \textcolor{red}{(This figure needs udpating.)}}
+             pair, and each row shows results for $r=5$ (top) or $r=15$ (bottom).}
     \label{fig:sims_logS}
@@
 within $\Sigma_\epsilon$ that account for the spatial dependence.
-\textcolor{red}{(@Steve - we have not provided any details on 
-the estimation of $\Sigma_\epsilon$ here.  Can you add some?)}
+Our method achieves this by first pre-scaling the functional realizations by the
+precision values (to account for nonstationarity), and then using maximum likelihood
+estimation to obtain estimates of the marginal variance and lengthscale of the 
+Mat\'ern spatial correlation function. We use these estimates and the precision 
+information to construct $\Sigma_\varepsilon$.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{CAMB_fit_model1.png}
+    \includegraphics[width=\textwidth]{CAMB_fit_model1.jpeg}
     \caption{Example of DGP.FCO model fit for the first batch of model runs from the CAMB data, 
@@
              compare the orange lines (posterior mean and dotted 95\% credible interval) to the 
-             light blue line, representing the infinite-resolution ``true" spectrum.
-             \textcolor{red}{(@Steve - can you leave the x-axis on the usual scale?
-             So -1.7 is directly visible.)}}   
+             light blue line, representing the infinite-resolution ``true" spectrum.}   
     \label{fig:fit_camb}
@@
 estimates of the CAMB power spectrum at the infinite-resolution for 
-$\log_{10}(k) < -1.7$, and then rely strictly on the low- and high-resolution 
+$\log_{10}(k) < -2.2$, and then rely strictly on the low- and high-resolution 
 runs for $k$ above this (note the transition
-in Figure \ref{fig:fit_camb} at $\log_{10}(k) = -1.7$). To compute $\bar{y}$ and $\Sigma_\epsilon$,
+in Figure \ref{fig:fit_camb} at $\log_{10}(k) = -2.2$). To compute $\bar{y}$ and $\Sigma_\epsilon$,
 we first use a log-log regression model to obtain precision estimates $\Lambda_\ell$ and $\Lambda_h$
@@
 To anchor estimates at low resolutions, we define the diagonal matrix $\Lambda_c$ with 
-$\Lambda_c^{(ii)}=10^8$ for $\log_{10}(k_i) < -1.7$ ($0$ otherwise).
+$\Lambda_c^{(ii)}=10^8$ for $\log_{10}(k_i) < -2.2$ ($0$ otherwise).
 We then compute the weighted average as: $\bar y = \Lambda^{-1}(\Lambda_c y_c + 
@@
 both the high-resolution run and the weighted average significantly deviate from the truth
-($\log_{10}(k)\approx 0.4,0.7$ - \textcolor{red}{@Steve - please adjust these numbers after 
-changing the x-axis of the figure}). Our model also provides effective UQ, appropriately shrinking
+($\log_{10}(k)\approx -1.8,-1.3$). Our model also provides effective UQ, appropriately shrinking
 credible intervals as precision increases for larger $k$.
@@
     \centering
-    \includegraphics[width=6in]{sims_MSE.png}
+    \includegraphics[width=\textwidth]{sims_MSE.jpeg}
     \caption{Boxplot of MSEs across two different functions and covariance settings 

### 2025-06-19 plotting/plot_data_fit.R
change 10^5 to 10^8 within plotting

@@ -34,7 +34,7 @@ x <- log10(k)
 # Get precision info 
 precs_lo <- ifelse(1:n %in% index_list$lowres.ix, prec_lowres, 0) * nrun
 precs_hi <- ifelse(1:n %in% index_list$highres.ix, prec_highres, 0)
-precs_pt <- ifelse(1:n %in% index_list$pert.ix, 10000, 0)
+precs_pt <- ifelse(1:n %in% index_list$pert.ix, 10^8, 0)
 Lam_lo <- rollmean(precs_lo, k = 10, fill = "extend")
 Lam_hi <- rollmean(precs_hi, k = 10, fill = "extend")
 Lam_pt <- rollmean(precs_pt, k = 10, fill = "extend")
@@ -205,3 +205,5 @@ plot.warp <- function(fit, wl = 1, wh = length(fit$x), ref.scale = 1, ...){
 load("../fitting/results/fit_M001.rda")
 par(mfrow=c(1,2))
 plot.warp(fit, xlab="X", ylab="W")
+
+plot(fit)

### 2025-06-19 plotting/plot_data_fit.R
update fit and warping for mira-titan

@@ -132,6 +132,13 @@ legend(x = "bottomright", legend = c("pert","low res","hi res", "wt avg", "UQ"),
 
 # plot_fit.png
 if(JPG) jpeg("../paper/plot_fit.jpeg", width = 12, height = 4, units = "in", res = 300)
+
+# Set layout: 2 columns, relative widths 2:1
+layout(matrix(1:2, nrow = 1), widths = c(2, 1))
+
+# Left plot (2/3 width)
+par(mar = c(4, 4, 2, 1))  # set margins: bottom, left, top, right
+
 # Plot model fit alongside data (posterior mean removed)
 matplot(log10(k), y_lo - fit$m, type="l", lty=3,
         col=cbcols[9], ylim = c(-.02,.02), xlab=expression(log[10](k)), 
@@ -145,65 +152,21 @@ lines(log10(k), fit$lb - fit$m, col=cbcols[1],lty=2,lwd=2)
 legend(x = "bottomright", legend = c("pert","low res","hi res", "wt avg", "UQ"),
        col = cbcols[c(4,9,8,2,1)], lty = c(4,3,2,1,2), 
        lwd = c(2,1,2,2,2), cex=0.9, bty="n")
-if(JPG) dev.off()
 
-# plot samples and average warping
-plot.warp <- function(fit, wl = 1, wh = length(fit$x), ref.scale = 1, ...){
-  Dx <- ncol(fit$x)
-  D <- ncol(fit$w[[1]])
-  indx <- floor(seq(from = 1, to = fit$nmcmc, length = 100))
-  if (indx[1] == 0) 
-    indx[1] <- 1
-  col <- heat.colors(100 + 10)
-  # par(mfrow = c(1, 1), mar = c(4, 4, 2, 2))
-  o <- order(fit$x)
-  plot(fit$x[o], fit$w[[indx[1]]][o] - mean(fit$w[[indx[1]]]), type = "l", 
-       xlab = "X", ylab = "W", col = col[1], #main = paste0("MCMC samples of X to W"), 
-       ylim = c(min(unlist(fit$w[indx])), max(unlist(fit$w[indx]))))
-  for (j in 2:length(indx)) {
-    lines(fit$x[o], fit$w[[indx[j]]][o] - mean(fit$w[[indx[j]]]), 
-          col = col[j])
-  }
-  
-  x <- fit$x
-  
-  for (j in 1:fit$nmcmc) {
-    y <- c(fit$w[[j]])
-    
[…]

### 2025-06-19 paper/main.tex
update sims images to say A and B, adjust rows/columns

@@
 \Lambda_p^{(ii)} = \begin{cases}
-    10^5 &\text{for } k_i < 0.04 \\
+    10^8 &\text{for } k_i < 0.04 \\
     0  &\text{otherwise}\\
@@
 for $i=1,\dots, n$.  In $\Lambda_p$, we use the sufficiently high precision 
-of $10^5$ to impose the deterministic nature of the perturbation theory output.
+of $10^8$ to impose the deterministic nature of the perturbation theory output.
 In $\Lambda_\ell$, the multiplication by 16 accounts for the sample size of the
@@
 functional realizations.  To provide insight into the nonstationarity of the power
-matter spectrum, Figure \ref{fig:plot_warp} shows the elliptical slice samples of $W$.
-These warpings depart from the identiy mapping around $x=0.4$ \textcolor{red}{(@Steve - 
-please change the x-axis to match the actual x-axis values, then adjust this number)}, 
+matter spectrum, the right panel of Figure \ref{fig:plot_fit} shows the elliptical slice samples of $W$.
+These warpings depart from the identiy mapping around $x\approx-1$, 
 stretching inputs where dynamics are shifting more rapidly for larger $k$ values.
@@
     \centering
-    \includegraphics[width=6in]{plot_fit.png}
+    \includegraphics[width=6in]{plot_fit.jpeg}
     \caption{95\% credible intervals for the power spectrum of the first Mira-Titan
     cosmology (centered by the predicted posterior mean). Perturbation theory, 
-    low-resolution, and high-resolution runs are shown for context.
-    \textcolor{red}{(@Steve - combine this plot with Figure 6.  The two panels of Figure 6
-    are redundant and unecessary (yet oddly, they don't match?).  Have this Mira-Titan 
-    plot take up 2/3 of the space, then have just the right panel from Figure 6 take 
-    up the remaining 1/3.  Update the figure caption and referencing accordingly.)}}
+    low-resolution, and high-resolution runs are shown for context.}
     \label{fig:plot_fit}
-\end{figure}
-
-\begin{figure}[ht]
-   \centering
-   \includegraphics[width=6in]{plot_warp_M001.png}
-   \caption{Left: MCMC samples of $W$ for model M001 after burnin and thinning. 
-            Right: Corresponding 95\% credible intervals for $W$, where departure 
-            from the identity line represents nonstationary behavior.
-            \textcolor{red}{(remove this figure, combine right panel with previous)}}
-   \label{fig:plot_warp}
 \end{figure}

### 2025-06-19 fitting/initialize.R
adjust prec to 10^8

@@ -34,7 +34,7 @@ load("../Mira-Titan-IV-Data/precision_and_indexes.Rdata")
 # Get precision info for each data type (low-res, hi-res, pert theory)
 precs_lo <- ifelse(1:n %in% index_list$lowres.ix, prec_lowres, 0) * nrun
 precs_hi <- ifelse(1:n %in% index_list$highres.ix, prec_highres, 0)
-precs_pt <- ifelse(1:n %in% index_list$pert.ix, 100000, 0)
+precs_pt <- ifelse(1:n %in% index_list$pert.ix, 10^8, 0)
 
 # Smooth the precision information (to remove steps)
 Lam_lo <- rollmean(precs_lo, k = 10, fill = "extend")

### 2025-06-19 plotting/plot_Sigmahat.R

@@ -36,7 +36,7 @@ load("../Mira-Titan-IV-Data/precision_and_indexes.Rdata")
 # Get precision info for each data type (low-res, hi-res, pert theory)
 precs_lo <- ifelse(1:n %in% index_list$lowres.ix, prec_lowres, 0) * nrun
 precs_hi <- ifelse(1:n %in% index_list$highres.ix, prec_highres, 0)
-precs_pt <- ifelse(1:n %in% index_list$pert.ix, 10000, 0)
+precs_pt <- ifelse(1:n %in% index_list$pert.ix, 10^8, 0)
 
 # Smooth the precision information (to remove steps)
 Lam_lo <- rollmean(precs_lo, k = 10, fill = "extend")

### 2025-06-19 paper/main.tex
update section 4

@@
     \includegraphics[width=6in]{plot_fit.jpeg}
-    \caption{95\% credible intervals for the power spectrum of the first Mira-Titan
+    \caption{{\it Left:} 95\% credible intervals for the power spectrum of the first Mira-Titan
     cosmology (centered by the predicted posterior mean). Perturbation theory, 
-    low-resolution, and high-resolution runs are shown for context.}
+    low-resolution, and high-resolution runs are shown for context. {\it Right:} Posterior draws
+    of $W$, illustrating some nonstationary behavior.}
     \label{fig:plot_fit}
@@
 
-Here, we begin with a $n \times m$ matrix composed of $m=111$ posterior means, 
-each of length $n$, denoted by $\boldsymbol\eta$. To facilitate more efficient estimation, 
-we use singular value decomposition \citep[SVD; e.g.,][]{banerjee2014linear}, 
-where $\boldsymbol\eta = UDV^T$; $U$ and $V$ are orthogonal matrices of size $n \times n$ 
-and $m\times m$, respectively. $D$ is a diagonal matrix containing the singular values. 
-For this application, $\boldsymbol\eta$ has rank $r=\text{min}\{m,n\}=111$, which 
-determines the number of non-zero elements in $D$. We find an alternative decomposition 
-where $\Sigma$ is a diagonal matrix with the non-zero singular values of $\boldsymbol\eta$, 
-and $U_1$ and $V_1$ are matrices with orthonormal columns of size $n_\eta \times r$ 
-and $m \times r$, respectively.
-
-Following \cite{higdon2008computer, higdon2010estcosmo}, we equivalently decompose 
-$\boldsymbol\eta$ into a principal component (PC) basis matrix 
-$B^* = \frac{1}{\sqrt{m}}U_1\Sigma$ along with its corresponding weights 
-$W^* = \sqrt{m}V_1$. Without loss of generality, we can perform this decomposition 
-after an overall mean trend is removed. If we use $p_\eta < r$ PCs, this will reduce 
-$B$ and $W$ to be the first $p_\eta$ columns of $B^*$ and $W^*$ respectively and 
-result in an approximation where
-
-\begin{equation}
-    \boldsymbol\eta= UDV^T = U_1\Sigma V_1^T \approx BW^T.
-\end{equation}
-
 Our goal is to predict the power spectrum for a held-out cosmology. 
 For simplicity, we will denote these eight parameters for a given cosmology (see 
-Section \ref{sec:data}) as $\psi \in \mathbb{R}^8$. 
+Section \ref{sec:data}) as $\psi \in \mathbb{R}^{p_\psi}$. With our estimated posterior means
+from DGP.FCO for each training cosmology, we now predict $S$ for unobserved cosmologies
+as a function of $\psi$.
+
+We have $m$-many training observations ($j = 1,\dots,m$) which each contain the $p_\psi$-dimensional
+input $\psi_j$ and DGP.FCO estimates $\hat{S}_j$ for the $j$th cosmology. 
+We combine all $\hat{S}_j$ estimates into a $n\times m$ matrix $\boldsymbol\eta$.
+Due to the fact that each column of $\boldsymbol\eta$ represents functional output, 
+we choose to reformulate the problem using principal components (PCs). Details of this can be found in 
+\ref{sec:apdx_basis}. Here we will use the first $p_\eta$ PCs.
+
 Figure \ref{fig:mean_PCs_oneW} illustrates this decomposition with the estimated 
-mean trend, the principal components contained in $B$, and one example of how the 
-weight for the first PC will vary dependent on the first of $p_\psi=8$ cosmological 
-parameters contained in $\psi$.
-
-In the following subsections, we consider different modeling strategies for $W$, 
-which will influence predictions and sensitivity analyses based on $\boldsymbol\eta$.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{mean_PCs_oneW.png}
-    \caption{The estimated mean trend of the 111 posterior means (left), the principal 
-    components contained in $B$ (middle), and an illustration of how the weights 
-    for the first PC will change as the first cosmological parameter varies (right) 
-    \textcolor{red}{This last plot is wonky because it was made with only the 6 test 
-    cosmologies, not a large uniform or FF design, Steve can update this}.}
-    \label{fig:mean_PCs_oneW}
-\end{figure}
+mean trend for the Mira-Titan dataset, the $p_\eta$ basis functions, and one 
+example of how the weight for the first PC will vary dependent on the first of 
+$p_\psi=8$ cosmological parameters contained in $\psi$. As a result, our goal becomes 
+estimating the weights $\Gamma$ corresponding to each of the $p_\eta$ PCs, 
+as a function of $\psi$.
 
@@
 
-Using these trained GPs from the PCs, we predict the spectra for unobserved cosmologies.
+Using these trained GPs from the PCs, we predict the spectra for each of the $N$ held-out 
+cosmologies for both the CAMB and Mira-Titan datasets.
 
@@
 
-From our 32 posterior means using DGP.FCO, these are input into SEPIA in order to predict power spectra
-for the 32 test cosmologies. With the infinite-resolution output, we have a sense of the ``best-case" 
-predictions we can have (i.e., if we trained with the 32 true power spectra, as opposed to our 
-estimated posterior means). In Figure \ref{fig:pca_preds_v_camb}, we see comparisons of the heldout 
-infinite resolution alongside two predictions: one based on DGP.FCO and the ``best-case" scenario using the
-infinite-resolution spectra of the training set. There is a high level of agreement between the pairs
-of predictions across cosmologies. Upon comparing MSEs for each of these prediction scenarios 
+From our $m=32$ posterior means using DGP.FCO, these are input into SEPIA in order to predict power spectra
+for the $N=32$ test cosmologies. In the CAMB dataset, $p_\psi=6$. With the infinite-resolution spectra, we 
+have a sense of the ``best-case" predictions we can have (i.e., if we trained with the 32 true 
+infinite-resolution spectra, as opposed to our estimated posterior means). In Figure \ref{fig:pca_preds_v_camb}, 
+we see comparisons of the heldout infinite resolution alongside two predictions: one 
+based on DGP.FCO and the ``best-case" scenario using the infinite-resolution spectra of 
+the training set. There is a high level of agreement between the pairs of predictions 
+across cosmologies. Upon comparing MSEs for each of these prediction scenarios 
 (Figure \ref{fig:mse_camb}, most are very similar across the two groups (with the two most difficult 
@@
     \includegraphics[width=4in]{preds_diff1_4_8.jpg}
-    \caption{Predictions for 16 of the 32 test cosmologies after centering and scaling. 
+    \caption{Predictions for 16 of the $N=32$ test cosmologies after centering and scaling. 
              The interpolated infinite resolution run is shown in black, the prediction 
@@
 
-Here we show prediction results for 6 held-out cosmologies. We compare our method 
-with Cosmic Emu \citep{moran2023mira}, the emulator constructed using the full suite 
-of 111 Mira-Titan model runs. Cosmic Emu uses a Bayesian approach where the smooth 
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{mean_PCs_oneW.png}
+    \caption{{\it Left:} The estimated mean trend of the $m=111$ posterior means from 
+    the Mira-Titan dataset. {\it Middle:} The $p_\eta$ principal components obtained from 
+    the posterior means. {\it Right:} An 
+    illustration of how the weights for the first PC will change as the first 
+    cosmological parameter varies \textcolor{red}{This last plot is wonky because 
+    it was made with only the $N=6$ test cosmologies, not a large uniform or FF design, 
+    Steve can update this}.}
+    \label{fig:mean_PCs_oneW}
+\end{figure}
+
+Here we show prediction results for $N=6$ held-out cosmologies for the Mira-Titan dataset. 
+We compare our method with Cosmic Emu \citep{moran2023mira}, the emulator constructed using the full suite 
+of $m=111$ Mira-Titan model runs. Cosmic Emu uses a Bayesian approach where the smooth 
 spectrum for each cosmology is modeled with a process convolution on Brownian motion. 
@@
 estimate the weights of the PCs from the estimated spectra. Below, we compare prediction 
-performance between Cosmic Emu and DGP.FCO for the 6 held-out cosmologies. 
-\textcolor{magenta}{In addition, we can also predict over a large uniform or full 
-factorial design to obtain estimated main effects for each cosmological parameter in 
-$\psi$, interactions amongst the $p_\psi=8$ parameters, and decompose the overall 
-variation by main effects, two-way interactions, and so on. Should this be included?}
+performance between Cosmic Emu and DGP.FCO for the $N=6$ held-out cosmologies. 
+% \textcolor{magenta}{In addition, we can also predict over a large uniform or full 
+% factorial design to obtain estimated main effects for each cosmological parameter in 
+% $\psi$, interactions amongst the $p_\psi=8$ parameters, and decompose the overall 
+% variation by main effects, two-way interactions, and so on. Should this be included?}
 
 Plots for the average root mean squared error (RMSE) for each prediction method over the 
-6 cosmologies and each $k$ value are shown in Figure \ref{fig:plot_rmse_k}. Comparisons 
+$N=6$ cosmologies and each $k$ value are shown in Figure \ref{fig:plot_rmse_k}. Comparisons 
 of the spectrum estimation for training cosmologies are shown as well (using RMSE against 
@@
     \includegraphics[width=6in]{rmse_by_k.png}
-    \caption{Left: Average RMSE of each prediction method across all $k$ values. 
-             Right: Average RMSE for each method when using the held-out computer 
+    \caption{{\it Left:} Average RMSE of each prediction method across all $k$ values. 
+             {\it Right:} Average RMSE for each method when using the held-out computer 
              model runs to train the model. In each case, the truth is assumed to 
@@
 
-\textcolor{magenta}{The different appendices from dissertation as necessary. 
-Can also put details of dataset in supplementary material.} 
-\textcolor{red}{Fix references as necessary, e.g. Moran et al. 2024.}
+% \textcolor{magenta}{Steve The different appendices from dissertation as necessary. 
+% Can also put details of dataset in supplementary material.} 
+% \textcolor{red}{Fix references as necessary, e.g. Moran et al. 2024.}
 
@@
 
+\section{Details on Basis Decomposition for Predictions}
+\label{sec:apdx_basis}
+
+Here, we begin with a $n \times m$ matrix composed of $m=111$ posterior means, 
+each of length $n$, denoted by $\boldsymbol\eta$. To facilitate more efficient estimation, 
+we use singular value decomposition \citep[SVD; e.g.,][]{banerjee2014linear}, 
+where $\boldsymbol\eta = UDV^T$; $U$ and $V$ are orthogonal matrices of size $n \times n$ 
+and $m\times m$, respectively. $D$ is a diagonal matrix containing the singular values. 
+For this application, $\boldsymbol\eta$ has rank $r=\text{min}\{m,n\}$ (111 for Mira-Titan, 
+32 for CAMB), which determines the number of non-zero elements in $D$. We find an alternative 
+decomposition where $\Sigma$ is a diagonal matrix with the non-zero singular values of 
+$\boldsymbol\eta$, and $U_1$ and $V_1$ are matrices with orthonormal columns of size 
+$n_\eta \times r$ and $m \times r$, respectively.
+
+Following \cite{higdon2008computer, higdon2010estcosmo}, we equivalently decompose 
+$\boldsymbol\eta$ into a principal component (PC) basis matrix 
+$B^* = \frac{1}{\sqrt{m}}U_1\Sigma$ along with its corresponding weights 
+$\Gamma^* = \sqrt{m}V_1$. Without loss of generality, we can perform this decomposition 
+after an overall mean trend is removed. If we use $p_\eta < r$ PCs, this will reduce 
+$B$ and $\Gamma$ to be the first $p_\eta$ columns of $B^*$ and $\Gamma^*$ respectively and 
+result in an approximation where
+
+\begin{equation}
+    \boldsymbol\eta= UDV^T = U_1\Sigma V_1^T \approx B\Gamma^T.
+\end{equation}
+
 \bibliographystyle{jasa}

### 2025-06-19 fitting/run.sh
all Mira-Titan models rerun

@@ -1,9 +1,9 @@
 #!/bin/bash
 
 model=1
-deep=1
 
-for (( model=12; model<=20; model++ ))
+for (( model=0; model<=116; model++ ))
 do
-  R CMD BATCH "--args model=$model deep=$deep" run_fit.R &
+  R CMD BATCH "--args model=$model" run_fit.R &
+  wait
 done

### 2025-06-19 fitting/run_fit.R

@@ -86,8 +86,8 @@ y_lra <- rowMeans(y_lo)
 # See "Weighted average from multiple computer experiments"
 # in Walsh dissertation: 3.7.1 Appendix E: Derivations
 y_avg <- (1 / Lam_z) * (Lam_pt * y_pt + Lam_lo * y_lra + Lam_hi * y_hi)
-#write.csv(y_avg, file = paste0("../CosmicEmu_etc/y_avg_",model,".csv"),
-#          row.names = F)
+write.csv(y_avg, file = paste0("../CosmicEmu_etc/y_avg_",model,".csv"),
+          row.names = F)
 
 # Scale the responses
 mean_y <- mean(y_avg)

### 2025-06-23 paper/main.tex
pass back to Steve

@@
 
-\subsection{Bayesian Inference}
+\subsection{Bayesian Inference}\label{sec:inference}
 
@@
 We then fit our DGP.FCO model on each of 111 training cosmologies, reserving 6 cosmologies
-for later hold-out testing.  To speed-up burn-in of MCMC chains, we conduct
-50,000 MCMC iterations for the first cosmology, then initialize the chains for all 
-other cosmologies with these burned-in samples, so fewer iterations are needed for the
-remaining 110 cosmologies.  Ultimately, we conduct 10,000 iterations for each of these,
-removing 5,000 for burn-in, then thinning by 5.  Figure \ref{fig:plot_fit} 
+for later hold-out testing.  Figure \ref{fig:plot_fit} 
 visualizes the DGP.FCO fit for the first training cosmology.  We center the plot
@@
 functional realizations.  To provide insight into the nonstationarity of the power
-matter spectrum, the right panel of Figure \ref{fig:plot_fit} shows the elliptical slice samples of $W$.
+matter spectrum, the right panel of Figure \ref{fig:plot_fit} shows burned-in elliptical 
+slice samples of $W$.
 These warpings depart from the identiy mapping around $x\approx-1$, 
@@
     low-resolution, and high-resolution runs are shown for context. {\it Right:} Posterior draws
-    of $W$, illustrating some nonstationary behavior.}
+    of $W$, illustrating nonstationary behavior by departing from a linear (identity) mapping.}
     \label{fig:plot_fit}
@@
 
-Our goal is to predict the power spectrum for a held-out cosmology. 
-For simplicity, we will denote these eight parameters for a given cosmology (see 
-Section \ref{sec:data}) as $\psi \in \mathbb{R}^{p_\psi}$. With our estimated posterior means
-from DGP.FCO for each training cosmology, we now predict $S$ for unobserved cosmologies
-as a function of $\psi$.
+Let $\psi \in \mathbb{R}^{p_\psi}$ denote the set of parameters that define a particular cosmology.
+Mira-Titan has 8 such parameters as described in Section \ref{sec:data}; CAMB has 6.
+Our goal is to effectively predict the power matter spectrum as a function of $\psi$, i.e. $\hat{S}(\psi)$, 
+for a held-out cosmology.  We observe $m$-many training cosmologies and fit our DGP.FCO 
+model (Section \ref{sec:inference}) to each one.  Let $\hat{S}_j$ for $j \in \{1,\ldots,m\}$ denote
+the estimated posterior mean of each training cosmology.  In total, our training data comprises
+the $p_\psi$-dimensional input $\psi_j$ and the $n$-dimensional output $\hat{S}_j$ (which is a 
+function of $x_i = \log_{10}(k_i)$) for $j \in \{1,\ldots,m\}$.
 
-We have $m$-many training observations ($j = 1,\dots,m$) which each contain the $p_\psi$-dimensional
-input $\psi_j$ and DGP.FCO estimates $\hat{S}_j$ for the $j$th cosmology. 
-We combine all $\hat{S}_j$ estimates into a $n\times m$ matrix $\boldsymbol\eta$.
-Due to the fact that each column of $\boldsymbol\eta$ represents functional output, 
-we choose to reformulate the problem using principal components (PCs). Details of this can be found in 
-\ref{sec:apdx_basis}. Here we will use the first $p_\eta$ PCs.
+To handle the functional nature of this output, we reformulate the problem using principal
+components (PCs; \textcolor{red}{@Steve - can you add a textbook reference?}).  
+We first center all $\hat{S}_j$ by subtracting $\bar{S} = \frac{1}{111}\sum_{j=1}^{111} \hat{S}_j$, then
+we combine them into an $n\times m$ matrix $\boldsymbol\eta$.  We use principal components to decompose 
+$\boldsymbol\eta = BW^\top$ where $B$ is the $n\times p_\eta$ matrix of basis functions and $W$
+is the $m\times p_\eta$ matrix of basis weights.  Details on this decomposition are 
+reserved for \ref{sec:apdx_basis}.  For the CAMB dataset, $p_\eta=6$; for Mira-Titan, $p_\eta=8$.
+As a visual, the left panel of Figure \ref{fig:mean_PCs_oneW} shows $\bar{S}$ for the Mira-Titan 
+training cosmologies.  The center panel shows the $p_\eta=8$ basis functions.
 
-Figure \ref{fig:mean_PCs_oneW} illustrates this decomposition with the estimated 
-mean trend for the Mira-Titan dataset, the $p_\eta$ basis functions, and one 
-example of how the weight for the first PC will vary dependent on the first of 
-$p_\psi=8$ cosmological parameters contained in $\psi$. As a result, our goal becomes 
-estimating the weights $\Gamma$ corresponding to each of the $p_\eta$ PCs, 
-as a function of $\psi$.
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{mean_PCs_oneW.png}
+    \caption{{\it Left:} The estimated mean trend of the $m=111$ posterior means from 
+    the Mira-Titan dataset. {\it Middle:} The $p_\eta$ principal components obtained from 
+    the posterior means. {\it Right:} An 
+    illustration of how the weights for the first PC will change as the first 
+    cosmological parameter varies \textcolor{red}{This last plot is wonky because 
+    it was made with only the $N=6$ test cosmologies, not a large uniform or FF design, 
+    Steve can update this}.
+    \textcolor{red}{This figure also needs work.  Axes need to be fixed.  Caption needs re-write.
+    Right panel needs to be redone.  It should look smooth like a GP.  Or else just remove it.}}
+    \label{fig:mean_PCs_oneW}
+\end{figure}
+
+Let $w_i^{(j)}$ denote the $ij^\textrm{th}$ element of $W$, representing the basis weight for the 
+$j^\textrm{th}$ cosmology and the $i^\textrm{th}$ PC.
+Conditional on $B$, our training data now consists of scalars $w_i^{(j)}$ as a function of $\psi_j$
+for $j \in \{1,\ldots,m\}$ and $i \in \{1,\ldots,p_\eta\}$.  If we can effectively predict these basis weights
+as a function of $\psi$ (i.e., $\hat{w}_i(\psi)$ for $i \in \{1,\ldots,p_\eta\}$), then we may obtain predictions 
+of the power spectrum as
+\[
+\hat{S}(\psi) = \sum_{i=1}^{p_\eta} \hat{w}_i(\psi) B_i,
+\]
+where $B_i$ is the $i^\textrm{th}$ basis function (column) of $B$.
 
@@
 In alignment with \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use 
-GPs to model the principal components' weights. For direct comparison alongside
+GPs to model these principal components' weights. For direct comparison alongside
 \cite{moran2023mira}, we employ a Python implementation of this approach called 
-SEPIA \citep{gattiker2020sepia}. That is, the $i$th vector of PC weights $w_i(\psi)$, 
-$i \in \{1,\ldots,p_\eta\}$, is modeled as a zero-mean Gaussian process with a 
-squared exponential (i.e., Gaussian) covariance function:
+SEPIA \citep{gattiker2020sepia}. That is, each $w_i(\psi)$ is modeled as a 
+zero-mean Gaussian process with a squared exponential (i.e., Gaussian) covariance function:
 \[
@@
 precision (set to zero if no nugget is used).
+\textcolor{red}{(@Steve - there is a lot missing here.  The inputs to the covariance kernel
+here need to be $\psi$.  Are we using a nugget?  If not, just remove it.  How are these parameters
+being estimated?  Make sure this kernel notation matches what we used earlier as neatly as possible.)}
 
 Using these trained GPs from the PCs, we predict the spectra for each of the $N$ held-out 
-cosmologies for both the CAMB and Mira-Titan datasets.
+cosmologies for both the CAMB and Mira-Titan datasets. \textcolor{red}{(@Steve - this needs work too.
+Please include the formula for the GP posterior mean, which is how we obtain $\hat{w}_i$.  We never
+had use to include it earlier, but we should use it here.  If you need, you can introduce new matrix/vector
+notation to represent the training data in this scenario.  At this point, you should refer back to the right
+panel of Figure \ref{fig:mean_PCs_oneW} and narrate how this shows the predicted GP mean for $w_1$ (for Mira-Titan)
+as a function of the first cosmological parameter $\psi_1$.)}
 
-\subsection{Predicting Spectra for CAMB Data}
+\subsection{Predicting Spectra for CAMB}
 \label{subsec:camb_pred}
 
-From our $m=32$ posterior means using DGP.FCO, these are input into SEPIA in order to predict power spectra
-for the $N=32$ test cosmologies. In the CAMB dataset, $p_\psi=6$. With the infinite-resolution spectra, we 
-have a sense of the ``best-case" predictions we can have (i.e., if we trained with the 32 true 
-infinite-resolution spectra, as opposed to our estimated posterior means). In Figure \ref{fig:pca_preds_v_camb}, 
-we see comparisons of the heldout infinite resolution alongside two predictions: one 
-based on DGP.FCO and the ``best-case" scenario using the infinite-resolution spectra of 
-the training set. There is a high level of agreement between the pairs of predictions 
-across cosmologies. Upon comparing MSEs for each of these prediction scenarios 
-(Figure \ref{fig:mse_camb}, most are very similar across the two groups (with the two most difficult 
-predictions being the exceptions), suggesting that overall DGP.FCO is estimating the true spectra 
-for each batch well.
+\textcolor{red}{(@Steve - we removed the text earlier that talked about the training/testing split in the
+wrong section.  You need to introduce it here now.)} From our $m=32$ posterior means using DGP.FCO, 
+these are input into SEPIA \textcolor{red}{(What does this mean?  Just refer to the equations from the previous
+section for how we obtain predictions.)} in order to predict power spectra
+for the $N=32$ test cosmologies.
+
+Since CAMB offers the ``true'' infinite-resolution spectra, we have the opportunity to benchmark our
+results in two ways.  First, we may compare our predictions to the true power spectrum for each held-out
+cosmology.  Second, we may leverage the infinite-resolution spectra for the training cosmologies in
+place of our estimated $\hat{S}_j$, to get a sense of the ``best-case" predictions that could originate
+from the PC decomposition.  Figure \ref{fig:pca_preds_v_camb} compares our predictions (DGP.FCO + PC, red)
+to this ``best-case'' prediction (true + PC, blue) and the actual true spectrum for \textcolor{red}{10?} 
+of the held-out test cosmologies.  There is a high level of agreement between the pairs of predictions 
+across cosmologies.  We then compare the mean squared errors of our predictions and the ``best-case'' 
+predictions across all held-out cosmologies, with results shown in Figure \ref{fig:mse_camb}.  We again
+see only minimal differences, which we take to indicate that DGP.FCO is offering effective
+estimates of the true spectra for each cosmology.
 
@@
     \includegraphics[width=4in]{preds_diff1_4_8.jpg}
-    \caption{Predictions for 16 of the $N=32$ test cosmologies after centering and scaling. 
-             The interpolated infinite resolution run is shown in black, the prediction 
-             (trained on the infinite resolution) is in blue and the prediction from 
-             the DGP.FCO posterior mean is in red.}   
+    \caption{Predictions of power matter spectra for 16 of the CAMB test cosmologies,
+             after centering and scaling. 
+             The true infinite resolution run is shown in black, the prediction 
+             (trained on the infinite resolution) is in blue, and the prediction from 
+             the DGP.FCO posterior mean is in red.
+             \textcolor{red}{(This figure does not look polished.  Make it wider.  I recommend
+             showing 2 rows of 5 plots.  Also, change line types so we are less reliant on color.
+             Missing x-axis label.  No need to number the panels.)}}   
     \label{fig:pca_preds_v_camb}
@@
     \caption{Comparison of MSE for CAMB predictions, when training with DGP.FCO posterior 
-             means (left), and training with the ``true" infinite-resolution spectra (right).}   
+             means (left), and training with the ``true" infinite-resolution spectra (right).
+             \textcolor{red}{(This figure also needs refinement.)}}   
     \label{fig:mse_camb}
@@
 
-\subsection{Predicting Spectra from Mira-Titan Data}
+\subsection{Predicting Spectra for Mira-Titan}
 \label{subsec:mira_pred}
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{mean_PCs_oneW.png}
-    \caption{{\it Left:} The estimated mean trend of the $m=111$ posterior means from 
-    the Mira-Titan dataset. {\it Middle:} The $p_\eta$ principal components obtained from 
-    the posterior means. {\it Right:} An 
-    illustration of how the weights for the first PC will change as the first 
-    cosmological parameter varies \textcolor{red}{This last plot is wonky because 
-    it was made with only the $N=6$ test cosmologies, not a large uniform or FF design, 
-    Steve can update this}.}
-    \label{fig:mean_PCs_oneW}
-\end{figure}
-
 Here we show prediction results for $N=6$ held-out cosmologies for the Mira-Titan dataset. 
-We compare our method with Cosmic Emu \citep{moran2023mira}, the emulator constructed using the full suite 
-of $m=111$ Mira-Titan model runs. Cosmic Emu uses a Bayesian approach where the smooth 
-spectrum for each cosmology is modeled with a process convolution on Brownian motion. 
+\textcolor{red}{(This sentence is misleading.... we only offer one figure that shows RMSE.)}
+We compare our method with Cosmic Emu \citep{moran2023mira}, the state-of-the-art emulator 
+constructed on the same $m=111$ Mira-Titan cosmologies. Cosmic Emu uses a Bayesian approach 
+where the ``true'' spectrum for each cosmology is modeled with a process convolution on Brownian motion. 
 Nonstationarity is permitted through modeling the bandwidth parameter with a process 
 convolution as well; this composition is referred to as a deep process convolution 
-\citep{moran2024dpc}.
+\citep{moran2024dpc}.  \textcolor{red}{(Add sentence about hwo Cosmic Emu also uses a PC
+approach to predict power matter spectra for held-out cosmologies as a function of $\psi$.)}
 
-Both Cosmic Emu and DGP.FCO are capable of modeling a true underlying functions based 
-upon multiple functional realizations, but our approach utilizes DGPs within a 
-hierarchical model in lieu of deep process convolutions. After obtaining the estimated 
-smooth spectra from the training cosmologies, each method utilizes GPs in order to 
-estimate the weights of the PCs from the estimated spectra. Below, we compare prediction 
-performance between Cosmic Emu and DGP.FCO for the $N=6$ held-out cosmologies. 
+%Both Cosmic Emu and DGP.FCO are capable of modeling a true underlying functions based 
+%upon multiple functional realizations, but our approach utilizes DGPs within a 
+%hierarchical model in lieu of deep process convolutions. After obtaining the estimated 
+%smooth spectra from the training cosmologies, each method utilizes GPs in order to 
+%estimate the weights of the PCs from the estimated spectra. Below, we compare prediction 
+%performance between Cosmic Emu and DGP.FCO for the $N=6$ held-out cosmologies. 
 % \textcolor{magenta}{In addition, we can also predict over a large uniform or full 
@@
 
-Plots for the average root mean squared error (RMSE) for each prediction method over the 
-$N=6$ cosmologies and each $k$ value are shown in Figure \ref{fig:plot_rmse_k}. Comparisons 
-of the spectrum estimation for training cosmologies are shown as well (using RMSE against 
-the weighted average of observations). For predictions, DGP.FCO achieves the greatest 
+Using our proposed approach and Cosmic Emu, we obtain two sets of predictions for $\hat{S}_t(\psi_t)$ 
+with $t\in\{1,\ldots,6\}$ indexing the six held-out cosmologies.  To summarize and compare
+performance, we evaluate the root mean squared error across $t$ for each wavenumber $x_i = \log_{10}(k_i)$.
+\textcolor{red}{(This is again missing key details.  What is the truth you are using to calculate
+RMSE???)}  The left panel of Figure \ref{fig:plot_rmse_k} shows these RMSEs for the held-out cosmologies.
+Comparisons of the spectrum estimation for training cosmologies are shown as well (using RMSE against 
+the weighted average of observations). \textcolor{red}{(What does this previous sentence mean??? We
+have discussed that the weighted average is not a good estimate of the true spectrum.  So we should not
+be comparing our estimated means to the weighted averages.  I am so confused....)}
+
+For predictions, DGP.FCO achieves the greatest 
 reduction in RMSE in the region where perturbation theory and the low-resolution runs 
@@
 lower at 56\% of $k$ values considered. The individual predictions for each method and 
-cosmology are shown in Figure \ref{fig:plot_pred_1to6}.
+cosmology are shown in Figure \ref{fig:plot_pred_1to6}.  \textcolor{red}{(Why is this figure in the appendix?
+But even more concerning, why are we subtracting the weighted average?  In every other plot we subtract our
+estimated mean.  This switch to using the weighted average in the final subsection is very concerning
+to me.  Also - what is the reader supposed to be gleaning from each of these plots?  Any time you
+have a figure, you need to include narration in the text that explains what the takeaways are.)}
 
@@
              model runs to train the model. In each case, the truth is assumed to 
-             be the weighted average of the computer model runs.}
+             be the weighted average of the computer model runs.
+             \textcolor{red}{(Shouldn't the average be across the 6 cosmologies?  Saying it 
+             is across the $k$ values is misleading.  Again, I am concerned for us to use
+             the weighted average as the truth.  This seems to go starkly against things
+             that we have said earlier in the paper.)}}
     \label{fig:plot_rmse_k}
@@
 \label{sec:disc}
+
+\textcolor{red}{(@Steve - I think you should revisit this section once the rest of the
+revisions are done.  Make sure it aligns with what the paper ends up being.  I am not going
+through this sectio myself at this time.)}
 
@@
 
-\begin{figure}
+\begin{figure}[H]
     \centering
@@
 \label{sec:apdx_basis}
+
+\textcolor{red}{(@Steve - revise the text so it works for any $m$, not just 111.  
+Also, replace $\Gamma$ with $W$ so it matches the text I revised earlier.)}
 

### 2025-06-25 plotting/plot_pca.R
updated predictions

@@ -161,9 +161,10 @@ mean(colMeans(mses) < colMeans(mses_emu))
 # Make plot of basis function decomposition
 # run predict.R first to load des_test and aps[[1]]
 # mean_PCs_oneW; hydro_plotting.R has original with GSMF
+setwd("../pca"); source("predict.R"); setwd("../plotting")
 par(mfrow=c(1,3), mar=c(4,4,1.8,1))
-plot(apply(eta,1,mean),type="l", xlab=expression(log[10](k)), ylab='script P')
-matplot(bases, type="l", xlab=expression(log[10](k)), ylab = 'deviation from average')
+plot(log10(k), apply(eta,1,mean),type="l", xlab=expression(log[10](k)), ylab='script P')
+matplot(log10(k), bases, type="l", xlab=expression(log[10](k)), ylab = 'deviation from average')
 des_test <- read.csv("../Mira-Titan-IV-Data/design_test.txt",sep="",header = F)
 x1_pred <- des_test[,1]
 plot(x1_pred[order(x1_pred)], aps[[1]]$Y_hat[order(x1_pred)], type="l", 

### 2025-06-25 fitting/run_CAMB.R
adjust nmcmc

@@ -183,17 +183,17 @@ for(model in 1:32){ # integer 1-32
     # load in initialized estimates for warping and hyperparameters for fit
     # w_0 <- read.csv("results/w0.csv")[[1]]
     # params0 <- read.csv("results/params0.csv")
-    fit <- fit_two_layer_hm(x, y_avg, nmcmc = 15000, #w_0 = w_0, 
+    fit <- fit_two_layer_hm(x, y_avg, nmcmc = 5000, #w_0 = w_0, 
                             # theta_y_0 = params0$theta_y0,
                             # theta_w_0 = params0$theta_w0,
                             # settings = list(pmx=T),
                             Sigma_hat = Sigma_hat)
   } else {
-    fit <- fit_one_layer_hm(x, y_avg, nmcmc = 10000, Sigma_hat = Sigma_hat)
+    fit <- fit_one_layer_hm(x, y_avg, nmcmc = 5000, Sigma_hat = Sigma_hat)
   }
   
   # plot(fit) # optionally investigate trace plots
-  fit <- trim(fit, 10000, 4)
+  fit <- trim(fit, 1000, 4)
   fit <- est_true(fit)
   
   coverages[model] = round(mean(fit$lb < y_cambi & fit$ub > y_cambi), 3)

### 2025-06-25 plotting/plot_pca.R
adjust pca figure

@@ -5,8 +5,8 @@
 
 library(dgp.hm)
 
-# make a PDF of plots?
-PDF <- F
+# make JPEGs of plots?
+JPG <- F
 
 # load the precision data (k, prec_highres, prec_lowres, index_list)
 load("../Mira-Titan-IV-Data/precision_and_indexes.Rdata")
@@ -144,30 +144,40 @@ legend("topright",legend = c("DPC","DGP.HM"),
 mtext(expression(log[10](k)),side=1,line=2.5,outer=T)
 mtext(paste0("Average RMSE"),side=2,line=2.5,outer=T)
 
-
-# boxplot of RMSEs by method
-par(mfrow=c(1,1))
-boxplot(c(sqrt(mses)),c(sqrt(mses_emu)),c(sqrt(mses_conv_t)),c(sqrt(mses_dgp_t)), 
-        names=c("DGP-GPPC","EMU","conv_t","dgp_t"), xlab="RMSE", horizontal = T,
-        col = c(cbcols[5],cbcols[2],cbcols[3],cbcols[6]))
-
-# compare average MSE for both prediction methods
-mean(mses_emu)
-mean(mses)
-
-# find proportion of k values where our method has lower MSE
-mean(colMeans(mses) < colMeans(mses_emu))
-
 # Make plot of basis function decomposition
-# run predict.R first to load des_test and aps[[1]]
+# run predict.R code below first to load des_test and aps[[1]] 
 # mean_PCs_oneW; hydro_plotting.R has original with GSMF
-setwd("../pca"); source("predict.R"); setwd("../plotting")
+load("trained_GP_for_pca.rda")
+
+# Read in test design
+theta_fixed <- rep(0.5, 7)  # or use colMeans(des_test)[2:8] or something else
+x1_seq <- seq(0, 1, length.out = 100)  # sweeping over theta_1
+des_test <- cbind(x1_seq, matrix(rep(theta_fixed, each = 100), ncol = 7))
+ntest <- nrow(des_test)
+
+# make the predictions for the given des_test design
+aps <- list()
+for (i in 1:n_pc) {
+  print(paste("GP pred",i))
+  aps[[i]] = predict(as[[i]],des_test)
+}
+
+# create matrix of overall mean to add back on to predictions
+mean_pred = matrix(mean_mat[,1],nrow=n_k,ncol=nrow(des_test))
+
[…]

### 2025-06-25 paper/main.tex
include GPfit description and post mean details

@@
 To handle the functional nature of this output, we reformulate the problem using principal
-components (PCs; \textcolor{red}{@Steve - can you add a textbook reference?}).  
-We first center all $\hat{S}_j$ by subtracting $\bar{S} = \frac{1}{111}\sum_{j=1}^{111} \hat{S}_j$, then
-we combine them into an $n\times m$ matrix $\boldsymbol\eta$.  We use principal components to decompose 
-$\boldsymbol\eta = BW^\top$ where $B$ is the $n\times p_\eta$ matrix of basis functions and $W$
-is the $m\times p_\eta$ matrix of basis weights.  Details on this decomposition are 
-reserved for \ref{sec:apdx_basis}.  For the CAMB dataset, $p_\eta=6$; for Mira-Titan, $p_\eta=8$.
-As a visual, the left panel of Figure \ref{fig:mean_PCs_oneW} shows $\bar{S}$ for the Mira-Titan 
-training cosmologies.  The center panel shows the $p_\eta=8$ basis functions.
+components \citep[PCs, e.g.][]{banerjee2014linear}.  
+We first center all $\hat{S}_j$ by subtracting $\bar{S} = \frac{1}{m}\sum_{j=1}^{m} \hat{S}_j$, 
+then we combine them into an $n\times m$ matrix $\boldsymbol\eta$.  We use the first $p_\eta$ 
+principal components to decompose $\boldsymbol\eta = B\Gamma^\top$ where $B$ is the 
+$n\times p_\eta$ matrix of basis functions and $\Gamma$ is the $m\times p_\eta$ matrix of basis weights.  
+Details on this decomposition are reserved for \ref{sec:apdx_basis}. For both the CAMB and Mira-Titan 
+datasets, we set $p_\eta=10$. As a visual, the left panel of Figure \ref{fig:mean_PCs_oneW} shows 
+$\bar{S}$ for the Mira-Titan training cosmologies.  The center panel shows the $p_\eta$ basis functions.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{mean_PCs_oneW.png}
+    \includegraphics[width=\textwidth]{mean_PCs_oneW.jpeg}
     \caption{{\it Left:} The estimated mean trend of the $m=111$ posterior means from 
@@
     the posterior means. {\it Right:} An 
-    illustration of how the weights for the first PC will change as the first 
-    cosmological parameter varies \textcolor{red}{This last plot is wonky because 
-    it was made with only the $N=6$ test cosmologies, not a large uniform or FF design, 
-    Steve can update this}.
-    \textcolor{red}{This figure also needs work.  Axes need to be fixed.  Caption needs re-write.
-    Right panel needs to be redone.  It should look smooth like a GP.  Or else just remove it.}}
+    illustration of how the weights for the fifth PC ($\gamma_5$) will change as the first 
+    cosmological parameter ($\psi_1$) varies.}
     \label{fig:mean_PCs_oneW}
@@
 
-Let $w_i^{(j)}$ denote the $ij^\textrm{th}$ element of $W$, representing the basis weight for the 
+Let $\gamma_i^{(j)}$ denote the $ij^\textrm{th}$ element of $\Gamma$, representing the basis weight for the 
 $j^\textrm{th}$ cosmology and the $i^\textrm{th}$ PC.
-Conditional on $B$, our training data now consists of scalars $w_i^{(j)}$ as a function of $\psi_j$
+Conditional on $B$, our training data now consists of scalars $\gamma_i^{(j)}$ as a function of $\psi_j$
 for $j \in \{1,\ldots,m\}$ and $i \in \{1,\ldots,p_\eta\}$.  If we can effectively predict these basis weights
-as a function of $\psi$ (i.e., $\hat{w}_i(\psi)$ for $i \in \{1,\ldots,p_\eta\}$), then we may obtain predictions 
+as a function of $\psi$ (i.e., $\hat{\gamma}_i(\psi)$ for $i \in \{1,\ldots,p_\eta\}$), then we may obtain predictions 
 of the power spectrum as
 \[
-\hat{S}(\psi) = \sum_{i=1}^{p_\eta} \hat{w}_i(\psi) B_i,
+\hat{S}(\psi) = \sum_{i=1}^{p_\eta} \hat{\gamma}_i(\psi) B_i,
 \]
@@
 
-% This is old: switched to SEPIA
-% We assume a powered exponential kernel for the Gaussian process and employ 
-% \texttt{GPfit} to perform a multi-start gradient based search for GP's hyperparameters 
-% \citep{macdonald2015gpfit}. 
-In alignment with \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use 
-GPs to model these principal components' weights. For direct comparison alongside
-\cite{moran2023mira}, we employ a Python implementation of this approach called 
-SEPIA \citep{gattiker2020sepia}. That is, each $w_i(\psi)$ is modeled as a 
-zero-mean Gaussian process with a squared exponential (i.e., Gaussian) covariance function:
-\[
-w_i(\psi) \sim GP\left(0,\ \frac{1}{\lambda_{z_i}} R^{(i)} + \frac{1}{\lambda_{s_i}} I \right),
-\]
-with covariance function
-\[
-R^{(i)}_{uv} = \exp\left(-\sum_{k=1}^{p_\psi} \beta_{ik} (x_{uk} - x_{vk})^2 \right),
-\]
-where $\beta_{ik}$ is the inverse squared lengthscale for the $k$th input dimension, 
-$\lambda_{z_i}$ is the marginal precision for $w_i(\psi)$, and $\lambda_{s_i}$ is the nugget 
-precision (set to zero if no nugget is used).
-\textcolor{red}{(@Steve - there is a lot missing here.  The inputs to the covariance kernel
-here need to be $\psi$.  Are we using a nugget?  If not, just remove it.  How are these parameters
-being estimated?  Make sure this kernel notation matches what we used earlier as neatly as possible.)}
-
-Using these trained GPs from the PCs, we predict the spectra for each of the $N$ held-out 
-cosmologies for both the CAMB and Mira-Titan datasets. \textcolor{red}{(@Steve - this needs work too.
-Please include the formula for the GP posterior mean, which is how we obtain $\hat{w}_i$.  We never
-had use to include it earlier, but we should use it here.  If you need, you can introduce new matrix/vector
-notation to represent the training data in this scenario.  At this point, you should refer back to the right
-panel of Figure \ref{fig:mean_PCs_oneW} and narrate how this shows the predicted GP mean for $w_1$ (for Mira-Titan)
-as a function of the first cosmological parameter $\psi_1$.)}
+% SEPIA information:
+% In alignment with \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use 
+% GPs to model these principal components' weights. For direct comparison alongside
+% \cite{moran2023mira}, we employ a Python implementation of this approach called 
+% SEPIA \citep{gattiker2020sepia}. That is, each $\gamma_i(\psi)$ is modeled as a 
+% zero-mean Gaussian process with a squared exponential (i.e., Gaussian) covariance function:
+% \[
+% \gamma_i(\psi) \sim GP\left(0,\ \frac{1}{\lambda_{z_i}} R^{(i)} + \frac{1}{\lambda_{s_i}} I \right),
+% \]
+% with covariance function
+% \[
+% R^{(i)}_{uv} = \exp\left(-\sum_{k=1}^{p_\psi} \beta_{ik} (x_{uk} - x_{vk})^2 \right),
+% \]
+% where $\beta_{ik}$ is the inverse squared lengthscale for the $k$th input dimension, 
+% $\lambda_{z_i}$ is the marginal precision for $\gamma_i(\psi)$, and $\lambda_{s_i}$ is the nugget 
+% precision (set to zero if no nugget is used).
+% \textcolor{red}{(@Steve - The inputs to the covariance kernel
+% here need to be $\psi$.  Are we using a nugget?  If not, just remove it.  How are these parameters
+% being estimated?  Make sure this kernel notation matches what we used earlier as neatly as possible.)}
+
+In alignment with \cite{higdon2010estcosmo} and \cite{moran2023mira}, we use 
+GPs to model the principal components' weights. 
+We assume a powered exponential kernel for the Gaussian process and employ 
+\texttt{GPfit} to perform a multi-start gradient based search for GP's hyperparameters 
+\citep{macdonald2015gpfit}. That is, the $i$th vector of PC weights $\gamma_i(\psi)$, 
+$i \in \{1,\ldots,p_\eta\}$, will be modeled as a zero mean GP: 
+$\gamma_i(\psi) \sim \mathrm{GP}(0, \sigma^2R)$, with
+
+\begin{equation*}
+    R_{uv} = \prod_{j=1}^{p_\psi}\exp\left\{-10^{\beta_j}|\psi_{uj}-\psi_{vj}|^\alpha\right\}.
+\end{equation*}
+
+We fix $\alpha=1.95$ and estimate $\sigma^2$ and each $\beta_j$ with $j \in \{1,...,p_\psi\}$, and repeat this
+process for all of the $p_\eta$ PCs. Using these trained GPs from the PCs, we predict the spectra for 
+each of the $N$ held-out cosmologies (such as $\psi^*$) for both the CAMB and Mira-Titan datasets.
+
+We fix the power parameter $\alpha = 1.95$ and estimate the lengthscale parameters $\beta_1, \ldots, \beta_{p_\psi}$ 
+from the data using maximum likelihood. To obtain the predicted basis weight $\hat{\gamma}_i(\psi^*)$ 
+for a new cosmology $\psi^* \in \mathbb{R}^{p_\psi}$, 
+we use the posterior mean from a zero-mean GP model fit to the training data 
+$\{(\psi_j, \gamma_i^{(j)})\}_{j=1}^m$. Under a zero-mean GP, the posterior mean is given by:
+
+\begin{equation*}
+\hat{\gamma}_i(\psi^*) = r(\psi^*)^\top R^{-1} \boldsymbol{\gamma}_i,
+\end{equation*}
+
+where $\boldsymbol{\gamma}_i = [\gamma_i^{(1)}, \gamma_i^{(2)}, \ldots, \gamma_i^{(m)}]^\top$ is the vector of 
+$m$ observed PC weights for the $i$th component, $R$ is the $m \times m$ correlation matrix between 
+all pairs of training inputs defined above, and $r(\psi^*) \in \mathbb{R}^m$ is the vector of correlations between 
+the test input $\psi^*$ and each training input $\psi_j$. To illustrate, in the right panel of Figure \ref{fig:mean_PCs_oneW}, 
+we show how the predicted GP mean for the weight of the fifth PC ($\gamma_5$) for Mira-Titan 
+varies as a function of the first cosmological parameter $\psi_1$, while the other parameters are fixed at a (scaled)
+midpoint value of 0.5.
 
@@
 
-\textcolor{red}{(@Steve - revise the text so it works for any $m$, not just 111.  
-Also, replace $\Gamma$ with $W$ so it matches the text I revised earlier.)}
-
-Here, we begin with a $n \times m$ matrix composed of $m=111$ posterior means, 
+Here, we begin with a $n \times m$ matrix composed of $m$ posterior means, 
 each of length $n$, denoted by $\boldsymbol\eta$. To facilitate more efficient estimation, 
@@
 and $m\times m$, respectively. $D$ is a diagonal matrix containing the singular values. 
-For this application, $\boldsymbol\eta$ has rank $r=\text{min}\{m,n\}$ (111 for Mira-Titan, 
-32 for CAMB), which determines the number of non-zero elements in $D$. We find an alternative 
+Here, $\boldsymbol\eta$ has rank $r=\text{min}\{m,n\}$ (for our applications, $r=111$ for Mira-Titan 
+and $r=32$ for CAMB), which determines the number of non-zero elements in $D$. We find an alternative 
 decomposition where $\Sigma$ is a diagonal matrix with the non-zero singular values of 

### 2025-06-26 paper/main.tex
update figures and manuscript

@@
 $\bar{Y}\mid S \sim \mathrm{GP}(S, \Sigma_\varepsilon)$.  While previous works
-have considered diagonal $\Sigma_\varepsilon$ \citep{moran2024dpc}, we find it
+have considered diagonal $\Sigma_\varepsilon$ \citep{moran2023mira}, we find it
 essential to account for the smoothly varying spatial dependence
@@
 We compare DGP.FCO to competing models with a simulation study where the underlying 
-true curve is known. Following \cite{moran2024dpc}, we consider two test functions 
+true curve is known. Following \cite{moran2023mira}, we consider two test functions 
 defined as
@@
 varies as a function of the first cosmological parameter $\psi_1$, while the other parameters are fixed at a (scaled)
-midpoint value of 0.5.
+midpoint value of 0.5. With these estimates $\hat{\gamma}(\psi^*)$, we predict power spectra at 
+held-out cosmologies for both the CAMB and Mira-Titan datasets.
 
@@
 
-\textcolor{red}{(@Steve - we removed the text earlier that talked about the training/testing split in the
-wrong section.  You need to introduce it here now.)} From our $m=32$ posterior means using DGP.FCO, 
-these are input into SEPIA \textcolor{red}{(What does this mean?  Just refer to the equations from the previous
-section for how we obtain predictions.)} in order to predict power spectra
-for the $N=32$ test cosmologies.
+In order to predict power spectra from the CAMB dataset, we split the 64 cosmologies such that $m=32$
+are used to train the DGP.FCO model and the remaining $N=32$ are reserved as hold-out cosmologies for
+prediction. With our $m=32$ posterior means using DGP.FCO, we obtain $N=32$ predicted power spectra from
+the functional prediction model outlined in Section \ref{subsec:pca} for the test cosmologies.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.85\textwidth]{pred_diffs_CAMB.jpeg}
+    \caption{Predictions of power matter spectra for each of the 32 CAMB test cosmologies,
+             after centering and scaling (to accentuate differences). 
+             The true infinite resolution run is shown in black, the prediction 
+             trained on the infinite resolution spectra is in dashed blue, and the prediction from 
+             the DGP.FCO posterior means is in dotted orange.}
+    \label{fig:pca_preds_v_camb}
+\end{figure}
 
@@
 place of our estimated $\hat{S}_j$, to get a sense of the ``best-case" predictions that could originate
-from the PC decomposition.  Figure \ref{fig:pca_preds_v_camb} compares our predictions (DGP.FCO + PC, red)
-to this ``best-case'' prediction (true + PC, blue) and the actual true spectrum for \textcolor{red}{10?} 
-of the held-out test cosmologies.  There is a high level of agreement between the pairs of predictions 
-across cosmologies.  We then compare the mean squared errors of our predictions and the ``best-case'' 
+from the PC decomposition.  Figure \ref{fig:pca_preds_v_camb} compares our predictions (DGP.FCO + PC, orange)
+to these ``best-case'' predictions (true + PC, blue) and the actual true spectra for each of the 32
+held-out test cosmologies.  There is a high level of agreement between the pairs of predictions across 
+cosmologies. 
+We then compare the mean squared errors of our predictions and the ``best-case'' 
 predictions across all held-out cosmologies, with results shown in Figure \ref{fig:mse_camb}.  We again
@@
     \centering
-    \includegraphics[width=4in]{preds_diff1_4_8.jpg}
-    \caption{Predictions of power matter spectra for 16 of the CAMB test cosmologies,
-             after centering and scaling. 
-             The true infinite resolution run is shown in black, the prediction 
-             (trained on the infinite resolution) is in blue, and the prediction from 
-             the DGP.FCO posterior mean is in red.
-             \textcolor{red}{(This figure does not look polished.  Make it wider.  I recommend
-             showing 2 rows of 5 plots.  Also, change line types so we are less reliant on color.
-             Missing x-axis label.  No need to number the panels.)}}   
-    \label{fig:pca_preds_v_camb}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=3in]{mse_dot.jpg}
+    \includegraphics[width=\textwidth]{mse_dot.jpeg}
     \caption{Comparison of MSE for CAMB predictions, when training with DGP.FCO posterior 
-             means (left), and training with the ``true" infinite-resolution spectra (right).
-             \textcolor{red}{(This figure also needs refinement.)}}   
+             means and training with the ``true" infinite-resolution spectra. The identity line 
+             provides equal MSE values amongst both methods.}   
     \label{fig:mse_camb}
@@
 Here we show prediction results for $N=6$ held-out cosmologies for the Mira-Titan dataset. 
-\textcolor{red}{(This sentence is misleading.... we only offer one figure that shows RMSE.)}
 We compare our method with Cosmic Emu \citep{moran2023mira}, the state-of-the-art emulator 
@@
 convolution as well; this composition is referred to as a deep process convolution 
-\citep{moran2024dpc}.  \textcolor{red}{(Add sentence about hwo Cosmic Emu also uses a PC
-approach to predict power matter spectra for held-out cosmologies as a function of $\psi$.)}
-
-%Both Cosmic Emu and DGP.FCO are capable of modeling a true underlying functions based 
-%upon multiple functional realizations, but our approach utilizes DGPs within a 
-%hierarchical model in lieu of deep process convolutions. After obtaining the estimated 
-%smooth spectra from the training cosmologies, each method utilizes GPs in order to 
-%estimate the weights of the PCs from the estimated spectra. Below, we compare prediction 
-%performance between Cosmic Emu and DGP.FCO for the $N=6$ held-out cosmologies. 
-% \textcolor{magenta}{In addition, we can also predict over a large uniform or full 
-% factorial design to obtain estimated main effects for each cosmological parameter in 
-% $\psi$, interactions amongst the $p_\psi=8$ parameters, and decompose the overall 
-% variation by main effects, two-way interactions, and so on. Should this be included?}
-
-Using our proposed approach and Cosmic Emu, we obtain two sets of predictions for $\hat{S}_t(\psi_t)$ 
-with $t\in\{1,\ldots,6\}$ indexing the six held-out cosmologies.  To summarize and compare
-performance, we evaluate the root mean squared error across $t$ for each wavenumber $x_i = \log_{10}(k_i)$.
-\textcolor{red}{(This is again missing key details.  What is the truth you are using to calculate
-RMSE???)}  The left panel of Figure \ref{fig:plot_rmse_k} shows these RMSEs for the held-out cosmologies.
-Comparisons of the spectrum estimation for training cosmologies are shown as well (using RMSE against 
-the weighted average of observations). \textcolor{red}{(What does this previous sentence mean??? We
-have discussed that the weighted average is not a good estimate of the true spectrum.  So we should not
-be comparing our estimated means to the weighted averages.  I am so confused....)}
-
-For predictions, DGP.FCO achieves the greatest 
-reduction in RMSE in the region where perturbation theory and the low-resolution runs 
-are deemed unbiased. For the region where only the high-resolution run is unbiased, 
-Cosmic Emu has a lower average RMSE. Over all $k$ values and predictions, DGP.FCO has 
-an average RMSE of 0.0089 and Cosmic Emu has an average RMSE of 0.0106. Our RMSE is 
-lower at 56\% of $k$ values considered. The individual predictions for each method and 
-cosmology are shown in Figure \ref{fig:plot_pred_1to6}.  \textcolor{red}{(Why is this figure in the appendix?
-But even more concerning, why are we subtracting the weighted average?  In every other plot we subtract our
-estimated mean.  This switch to using the weighted average in the final subsection is very concerning
-to me.  Also - what is the reader supposed to be gleaning from each of these plots?  Any time you
-have a figure, you need to include narration in the text that explains what the takeaways are.)}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=6in]{rmse_by_k.png}
-    \caption{{\it Left:} Average RMSE of each prediction method across all $k$ values. 
-             {\it Right:} Average RMSE for each method when using the held-out computer 
-             model runs to train the model. In each case, the truth is assumed to 
-             be the weighted average of the computer model runs.
-             \textcolor{red}{(Shouldn't the average be across the 6 cosmologies?  Saying it 
-             is across the $k$ values is misleading.  Again, I am concerned for us to use
-             the weighted average as the truth.  This seems to go starkly against things
-             that we have said earlier in the paper.)}}
-    \label{fig:plot_rmse_k}
-\end{figure}
-
-% \begin{figure}[ht]
-%     \centering
-%     \includegraphics[width=6in]{rmse_box.png}
-%     \caption{Boxplot of all RMSEs of each method across all six test models. 
-%.             From top to bottom: box plots for our training method, process 
-%              convolution training method, Cosmic EMU predictions, and our DGP-PC predictions.}
-%     \label{fig:plot_rmse_box}
-% \end{figure}
+\citep[DPC,][]{moran2023mira}. Cosmic Emu also uses a PC approach to predict power matter 
+spectra for held-out cosmologies as a function of $\psi$ \citep{gattiker2020sepia}.
+
+Using our proposed DGP.FCO approach and Cosmic Emu, we obtain two sets of predictions for $\hat{S}_t(\psi_t)$ 
+with $t\in\{1,\ldots,6\}$ indexing the six held-out cosmologies.  In this real-world example, we have
+no ``truth" for the Mira-Titan hold-outs against which to benchmark. In an effort to assess the 
+accuracy of these predictions, we also separately fit the DGP.FCO model on each of these hold outs
+and use the corresponding posterior means as a reference. The individual predictions for each method and 
+cosmology are shown in Figure \ref{fig:plot_pred_1to6}, where the posterior mean is subtracted so that
+predictions lines closer to zero represent a better fit.
+
+\begin{figure}[!b]
+    \centering
+    \includegraphics[width=\textwidth]{pred_1to6.jpeg}
+    \caption{Results of all six predictions for both methods on the held-out 
+    cosmologies (DGP.FCO posterior mean subtracted). CosmicEMU is dotted blue and DGP.FCO is dashed orange.
+    Closer to the dotted zero line indicates a better fit with the DGP.FCO posterior mean.}
+    \label{fig:plot_pred_1to6}
+\end{figure}
+
+From these predictions, we can calculate MSE when treating the trained DGP.FCO predictions as a benchmark. 
+Our DGP.FCO prediction achieves the greatest reduction in MSE in the region where perturbation theory
+and the low-resolution runs are deemed unbiased (see Figure \ref{fig:mse_by_k}). For the region where only 
+the high-resolution run is unbiased, Cosmic Emu generally has a slightly lower average RMSE. Our average MSE 
+is lower at 55\% of $k$ values considered. These results indicate that DGP.FCO competes favorably with the 
+state-of-the-art Cosmic Emu model.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{mse_by_k.jpeg}
+    \caption{Average MSE across each of the 6 hold out cosmologies, for each $k$ value. Cosmic Emu 
+    predictions are shown in dotted blue, and DGP.FCO predictions are shown in dashed orange.}
+    \label{fig:mse_by_k}
+\end{figure}
 
@@
 
-\textcolor{red}{(@Steve - I think you should revisit this section once the rest of the
-revisions are done.  Make sure it aligns with what the paper ends up being.  I am not going
-through this sectio myself at this time.)}
-
-In this work, we propose a novel Bayesian hierarchical model which utilizes DGPs 
-in order to estimate a smooth underlying function from correlated functional observations 
-(as cosmological computer model runs). In addition to the compositional form of the DGP, 
-we include a third component modeling the correlated errors for each spectrum. 
-With all smooth spectra estimated from the training set of 111 batches of model runs, 
-we can predict spectra for unobserved cosmologies.
-
-Multiple avenues for future research emerge: in addition to estimating dark matter 
-power spectra, batches of hydrodynamical simulations exist, and this method could 
-easily be adapted to handle such datasets. Other applications outside of cosmology 
-can be explored, wherever multiple functional outputs exist, perhaps with spatial 
-correlation (for example, observing multiple tropical cyclone forecasts). More detailed 
-sensitivity analyses for the cosmological parameters can be explored to gain insights 
-into how each of these parameters interact, and which are most significant. From a 
-modeling perspective, changes to the model could include modeling the latent layer 
-$W$ jointly across different cosmologies, as well as considering different redshift 
-values other than $z=0$. Leveraging a hybrid model between DGP.FCO and Cosmic Emu 
-may also help to obtain accurate estimates on the lower and higher ends of $k$ values, 
-respectively. Additionally, Bayesian smoothing spline ANOVA (BSS-ANOVA) or other methods 
-could be compared in addition to DGPs.
+In this work, we introduce a novel Bayesian hierarchical framework that leverages deep 
+Gaussian processes (DGPs) to model smooth latent functions from correlated functional outputs, 
+such as those arising in cosmological simulations. Our model, DGP.FCO, builds on the compositional nature 
+of DGPs and introduces a third layer to explicitly model correlated observational errors within 
+each spectrum. Through a simulation study, we demonstrate that DGP.FCO offers improved uncertainty 
+quantification compared to existing methods.
+
+We apply DGP.FCO to both the CAMB and Mira-Titan datasets, generating posterior mean estimates 
+and predicting at held-out cosmological settings. Across both datasets, DGP.FCO 
+consistently performs well against competitive baselines, showcasing strong generalization and 
+flexibility in complex, functional data regimes.
+
+Looking ahead, several avenues for future research arise. Beyond modeling dark matter power 
+spectra, DGP.FCO could be extended to hydrodynamical simulation outputs in cosmology, which 
+often share similar functional structures and error correlations. More broadly, this approach may be 
+applicable in other scientific domains involving structured functional outputs, especially those 
+with spatial or temporal dependencies.
+
+In addition, generating predictions over large cosmological grids opens the door to detailed 
+sensitivity analyses, enabling researchers to disentangle parameter interactions and identify 
+the most influential physical drivers of the spectra. From a modeling standpoint, further 
+developments could include joint modeling of latent layers across cosmologies, and extensions 
+to redshifts beyond $z=0$. Lastly, a hybrid framework combining DGP.FCO with Cosmic Emu, leveraging 
+the strengths of each at different scales of $k$, may yield even more accurate predictions.
 
@@
 
-\section{Mira-Titan Predictive Plots}
-\label{sec:apdx_mira_plots}
-
-% \textcolor{magenta}{Steve The different appendices from dissertation as necessary. 
-% Can also put details of dataset in supplementary material.} 
-% \textcolor{red}{Fix references as necessary, e.g. Moran et al. 2024.}
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=6in]{pred_1to6.png}
-    \caption{Results of all six predictions for both methods on the held-out 
-    cosmologies (weighted average subtracted). CosmicEMU is dashed, DGP.FCO is dotted, 
-    and solid gray lines show results of held-out computer model runs.}
-    \label{fig:plot_pred_1to6}
-\end{figure}
-
 \section{Simulation Study Additional Results}

### 2025-06-26 plotting/plot_pca.R

@@ -26,8 +26,6 @@ cbcols <- palette.colors(palette = "Okabe-Ito")
 #1 black, 2 orange, 3 skyblue , 4 bluishgreen, 5 yellow 
 #6 blue, 7 vermillion,  8 reddishpurple, 9 gray 
 
-if(PDF) pdf(paste0("compare_pred_",n_pc,"_",pwdExp,".pdf"))
-
 # read in predicted power matter spectra for 6 test models
 etaEmu <- read.csv(paste0("../pca/etaEmu",n_pc,"_",pwdExp,".csv"))
 colnames(etaEmu) <- c("E001", "E002", "E003", "E009", "E010", "M000")
@@ -75,6 +73,7 @@ for(i in 1:6){
   
   # select a mean term to make plots and calculate MSEs
   meanterm = y_avg
+
   
   # Make a plot showing each computer model, along with predictions
   # Average is removed to better visualize the variability and bias
@@ -98,9 +97,9 @@ for(i in 1:6){
         col=cbcols[9], lwd=3)
   # plot our prediction using DGPs and then GPs on weights of PCs (DGP-GPPC)
   lines(log10(k), (etaEmu[,model_name]) - meanterm, 
-        col=cbcols[3], lwd=2, lty=3)
+        col=cbcols[7], lwd=2, lty=3)
   # plot the cosmicEmu prediction
-  lines(log10(k),cosmscrPsz - meanterm, col=cbcols[7], lwd=2, lty=2)
+  lines(log10(k),cosmscrPsz - meanterm, col=cbcols[3], lwd=2, lty=2)
   # # plot the proc convln estimate (trained with model runs)
   # lines(log10(k), convln_pred[which(c(112:116,0)==model),] - meanterm, 
   #       col=cbcols[3],lty=2, lwd=2)
@@ -127,27 +126,88 @@ for(i in 1:6){
 mtext(expression(log[10](k)),side=1,line=2.75,outer=T)
 mtext(paste0("Average RMSE"),side=2,line=2.5,outer=T)
 
-# compare RMSEs for each k value, by method
-par(mfrow=c(1,2),oma=c(4,4,1.5,1),mar=c(0,0,0,0))
+if(JPG) jpeg("../paper/pred_1to6.jpeg", width = 12, height = 4, units = "in", res = 300)
+par(mfrow=c(2,3),oma=c(4,4,1.5,1),mar=c(0,0,0,0))
+# For each of the six test models...
+for(i in 1:6){
+  
[…]

### 2025-06-26 paper/main.tex
edits to 4.1

@@
 Mira-Titan has 8 such parameters as described in Section \ref{sec:data}; CAMB has 6.
-Our goal is to effectively predict the power matter spectrum as a function of $\psi$, i.e. $\hat{S}(\psi)$, 
+Our goal is to effectively predict the power matter spectrum as a function of $\psi$, i.e., $\hat{S}(\psi)$, 
 for a held-out cosmology.  We observe $m$-many training cosmologies and fit our DGP.FCO 
@@
 To handle the functional nature of this output, we reformulate the problem using principal
-components \citep[PCs, e.g.][]{banerjee2014linear}.  
+components \citep[PCs; e.g.,][]{banerjee2014linear}.  
 We first center all $\hat{S}_j$ by subtracting $\bar{S} = \frac{1}{m}\sum_{j=1}^{m} \hat{S}_j$, 
@@
 of the power spectrum as
-\[
+\begin{equation}\label{eq:spred}
 \hat{S}(\psi) = \sum_{i=1}^{p_\eta} \hat{\gamma}_i(\psi) B_i,
-\]
+\end{equation}
 where $B_i$ is the $i^\textrm{th}$ basis function (column) of $B$.
@@
 In alignment with \cite{higdon2010estcosmo} and \cite{moran2023mira}, we use 
-GPs to model the principal components' weights. 
-We assume a powered exponential kernel for the Gaussian process and employ 
-\texttt{GPfit} to perform a multi-start gradient based search for GP's hyperparameters 
-\citep{macdonald2015gpfit}. That is, the $i$th vector of PC weights $\gamma_i(\psi)$, 
-$i \in \{1,\ldots,p_\eta\}$, will be modeled as a zero mean GP: 
-$\gamma_i(\psi) \sim \mathrm{GP}(0, \sigma^2R)$, with
-
+GPs with a power exponential kernel to model the principal components' weights. 
+That is, each vector of PC weights $\gamma_i(\psi)$, $i \in \{1,\ldots,p_\eta\}$, 
+will be modeled as a zero mean GP: $\gamma_i(\psi) \sim \mathrm{GP}(0, \sigma^2R)$, with kernel
 \begin{equation*}
-    R_{uv} = \prod_{j=1}^{p_\psi}\exp\left\{-10^{\beta_j}|\psi_{uj}-\psi_{vj}|^\alpha\right\}.
+    R^{u,v} = \prod_{j=1}^{p_\psi}\exp\left(-10^{\beta_j}|\psi_{uj}-\psi_{vj}|^\alpha\right).
 \end{equation*}
-
-We fix $\alpha=1.95$ and estimate $\sigma^2$ and each $\beta_j$ with $j \in \{1,...,p_\psi\}$, and repeat this
-process for all of the $p_\eta$ PCs. Using these trained GPs from the PCs, we predict the spectra for 
-each of the $N$ held-out cosmologies (such as $\psi^*$) for both the CAMB and Mira-Titan datasets.
-
-We fix the power parameter $\alpha = 1.95$ and estimate the lengthscale parameters $\beta_1, \ldots, \beta_{p_\psi}$ 
-from the data using maximum likelihood. To obtain the predicted basis weight $\hat{\gamma}_i(\psi^*)$ 
-for a new cosmology $\psi^* \in \mathbb{R}^{p_\psi}$, 
-we use the posterior mean from a zero-mean GP model fit to the training data 
-$\{(\psi_j, \gamma_i^{(j)})\}_{j=1}^m$. Under a zero-mean GP, the posterior mean is given by:
+We fix $\alpha=1.95$ and estimate $\sigma^2$ and each lengthscale $\beta_1,\ldots,\beta_{p_\psi}$ 
+with the \texttt{GPfit} {\sf R}-package \citep{macdonald2015gpfit}, which employs multi-start 
+gradient-based maximum likelihood estimation.  We repeat this estimation
+process for all of the $p_\eta$ PCs.  Conditioned on kernel hyperparameters,
+posterior predictions of $\hat{\gamma}_i(\psi^*)$ for a new cosmological parameterization 
+$\psi^* \in \mathbb{R}^{p_\psi}$ follows:
 
@@
 
-where $\boldsymbol{\gamma}_i = [\gamma_i^{(1)}, \gamma_i^{(2)}, \ldots, \gamma_i^{(m)}]^\top$ is the vector of 
-$m$ observed PC weights for the $i$th component, $R$ is the $m \times m$ correlation matrix between 
-all pairs of training inputs defined above, and $r(\psi^*) \in \mathbb{R}^m$ is the vector of correlations between 
-the test input $\psi^*$ and each training input $\psi_j$. To illustrate, in the right panel of Figure \ref{fig:mean_PCs_oneW}, 
-we show how the predicted GP mean for the weight of the fifth PC ($\gamma_5$) for Mira-Titan 
-varies as a function of the first cosmological parameter $\psi_1$, while the other parameters are fixed at a (scaled)
-midpoint value of 0.5. With these estimates $\hat{\gamma}(\psi^*)$, we predict power spectra at 
-held-out cosmologies for both the CAMB and Mira-Titan datasets.
+where $\boldsymbol{\gamma}_i = [\gamma_i^{(1)}, \gamma_i^{(2)}, \ldots, \gamma_i^{(m)}]^\top$ 
+is the vector of $m$ observed PC weights for the $i^\textrm{th}$ component, 
+$R$ is the $m \times m$ correlation matrix between 
+all pairs of training inputs defined above, and $r(\psi^*) \in \mathbb{R}^m$ is 
+the vector of correlations between the test input $\psi^*$ and each training input 
+$\psi_j$. To illustrate, the right panel of Figure \ref{fig:mean_PCs_oneW} shows
+the predicted weight of the fifth PC ($\gamma_5$) for Mira-Titan as a function 
+of the first cosmological parameter ($\psi_1$), while the other parameters are fixed 
+at their midpoint.  With all estimated PC weights, $\hat{\gamma_i}(\psi^*), i=1,\ldots,p_\eta$, 
+we predict power spectra following Eq.~\ref{eq:spred}.
 

### 2025-06-26 paper/main.tex
edits to 4.2

@@
 
-In order to predict power spectra from the CAMB dataset, we split the 64 cosmologies such that $m=32$
-are used to train the DGP.FCO model and the remaining $N=32$ are reserved as hold-out cosmologies for
-prediction. With our $m=32$ posterior means using DGP.FCO, we obtain $N=32$ predicted power spectra from
-the functional prediction model outlined in Section \ref{subsec:pca} for the test cosmologies.
+For the CAMB dataset, we split the 64 available cosmologies such that $m=32$
+are used for training and $N=32$ are reserved as hold-out/test cosmologies for prediction. 
+For each training cosmology, we collect our DGP.FCO predicted spectrum, then conduct the principal
+components model just described to predict $\hat{S}(\psi)$ for each test cosmology.
 
@@
              after centering and scaling (to accentuate differences). 
-             The true infinite resolution run is shown in black, the prediction 
-             trained on the infinite resolution spectra is in dashed blue, and the prediction from 
-             the DGP.FCO posterior means is in dotted orange.}
+             The true infinite resolution run is shown in black, the PC prediction 
+             trained on the infinite resolution spectra is in dashed blue, and the PC prediction trained
+             on the DGP.FCO posterior means is in dotted orange.}
     \label{fig:pca_preds_v_camb}
@@
 cosmology.  Second, we may leverage the infinite-resolution spectra for the training cosmologies in
-place of our estimated $\hat{S}_j$, to get a sense of the ``best-case" predictions that could originate
-from the PC decomposition.  Figure \ref{fig:pca_preds_v_camb} compares our predictions (DGP.FCO + PC, orange)
-to these ``best-case'' predictions (true + PC, blue) and the actual true spectra for each of the 32
-held-out test cosmologies.  There is a high level of agreement between the pairs of predictions across 
-cosmologies. 
+place of our estimated $\hat{S}_j$ from our DGP.FCO model, to get a sense of the ``best-case" predictions 
+that could originate from the PC decomposition.  
+Figure \ref{fig:pca_preds_v_camb} compares our predictions (DGP.FCO + PC, dotted orange)
+to these ``best-case'' predictions (true + PC, dashed blue) and the actual true spectra (solid black)
+for each of the 32 held-out test cosmologies.  
+The agreement between our predictions and the ``best-case'' predictions indicates that DGP.FCO is effectively
+capturing the true power matter spectra of the training cosmologies.
+The high level of agreement between the pairs of predictions and the true spectra across cosmologies
+indicates our PC model is providing effective predictions of held-out power spectra.
 We then compare the mean squared errors of our predictions and the ``best-case'' 
 predictions across all held-out cosmologies, with results shown in Figure \ref{fig:mse_camb}.  We again
-see only minimal differences, which we take to indicate that DGP.FCO is offering effective
+see only minimal differences between the two, confirming that DGP.FCO is offering effective
 estimates of the true spectra for each cosmology.

### 2025-06-26 paper/main.tex
edits to 4.3

@@
 convolution as well; this composition is referred to as a deep process convolution 
-\citep[DPC,][]{moran2023mira}. Cosmic Emu also uses a PC approach to predict power matter 
+\citep[DPC;][]{moran2023mira}. Cosmic Emu also uses a principal components model to predict power matter 
 spectra for held-out cosmologies as a function of $\psi$ \citep{gattiker2020sepia}.
@@
 no ``truth" for the Mira-Titan hold-outs against which to benchmark. In an effort to assess the 
-accuracy of these predictions, we also separately fit the DGP.FCO model on each of these hold outs
-and use the corresponding posterior means as a reference. The individual predictions for each method and 
-cosmology are shown in Figure \ref{fig:plot_pred_1to6}, where the posterior mean is subtracted so that
-predictions lines closer to zero represent a better fit.
+accuracy of these predictions, we also separately fit our DGP.FCO model on each of these testing cosmologies
+(leveraging the perturbation theory, low resolution runs, and high resolution run to predict the underlying
+power matter spectrum).  Although these predicted ``in-sample'' spectra are not formal ``truths,'' we find them to be
+useful benchmarks for our ``out-of-sample'' predictions.  The individual predictions for each method and 
+held-out cosmology are shown in Figure \ref{fig:plot_pred_1to6}, centered by the in-sample DGP.FCO
+posterior means.  Predictions closer to zero represent a better fit.
 
@@
 
-From these predictions, we can calculate MSE when treating the trained DGP.FCO predictions as a benchmark. 
-Our DGP.FCO prediction achieves the greatest reduction in MSE in the region where perturbation theory
-and the low-resolution runs are deemed unbiased (see Figure \ref{fig:mse_by_k}). For the region where only 
-the high-resolution run is unbiased, Cosmic Emu generally has a slightly lower average RMSE. Our average MSE 
-is lower at 55\% of $k$ values considered. These results indicate that DGP.FCO competes favorably with the 
-state-of-the-art Cosmic Emu model.
+From these predictions, we can calculate MSE against the in-sample DGP.FCO predictions across the 6
+test cosmologies. 
+Our out-of-sample DGP.FCO + PC prediction achieves the greatest reduction in MSE in the region 
+where perturbation theory and the low-resolution runs are deemed unbiased (see Figure \ref{fig:mse_by_k}). 
+For the region where only the high-resolution run is unbiased, Cosmic Emu has a marginally lower
+MSE. Our MSE is lower at 55\% of $k$ values considered. These results indicate that DGP.FCO 
+competes favorably with the state-of-the-art Cosmic Emu model.
 
@@
     \includegraphics[width=\textwidth]{mse_by_k.jpeg}
-    \caption{Average MSE across each of the 6 hold out cosmologies, for each $k$ value. Cosmic Emu 
-    predictions are shown in dotted blue, and DGP.FCO predictions are shown in dashed orange.}
+    \caption{MSE across each of the 6 hold out cosmologies for each $k$ value. Cosmic Emu 
+    performance is shown in dotted blue, and DGP.FCO performance is shown in dashed orange.}
     \label{fig:mse_by_k}

### 2025-06-26 paper/main.tex
minor things in discussion

@@
 
-In this work, we introduce a novel Bayesian hierarchical framework that leverages deep 
+In this work, we introduced a novel Bayesian hierarchical framework that leverages deep 
 Gaussian processes (DGPs) to model smooth latent functions from correlated functional outputs, 
-such as those arising in cosmological simulations. Our model, DGP.FCO, builds on the compositional nature 
-of DGPs and introduces a third layer to explicitly model correlated observational errors within 
-each spectrum. Through a simulation study, we demonstrate that DGP.FCO offers improved uncertainty 
+such as those arising in cosmological simulations. Our model (DGP.FCO) builds on the 
+compositional nature of DGPs and introduces an additional outer layer to explicitly 
+model correlated observational errors within each spectrum. Through a simulation study, 
+we demonstrated that DGP.FCO offers improved uncertainty 
 quantification compared to existing methods.
 
-We apply DGP.FCO to both the CAMB and Mira-Titan datasets, generating posterior mean estimates 
+We applied DGP.FCO to both the CAMB and Mira-Titan datasets, generating posterior mean estimates 
 and predicting at held-out cosmological settings. Across both datasets, DGP.FCO 
@@
 
-Looking ahead, several avenues for future research arise. Beyond modeling dark matter power 
+Several avenues for future research arise. Beyond modeling dark matter power 
 spectra, DGP.FCO could be extended to hydrodynamical simulation outputs in cosmology, which 
@@
 to redshifts beyond $z=0$. Lastly, a hybrid framework combining DGP.FCO with Cosmic Emu, leveraging 
-the strengths of each at different scales of $k$, may yield even more accurate predictions.
+the strengths of each at different scales of $k$, may yield more accurate predictions.
 

### 2025-06-26 paper/main.tex
remove indent after equations in appendix

@@
 Here, we establish that under the model:
-
 $$
@@
 $$
-
 We can obtain the integrated likelihood $p(Y_i|W)=\int p(Y_i|S,W)p(S|W) dS$, 
@@
 $$
-
 If we focus only on terms containing $S$, we have 
@@
 $$
-
 To complete the square, we add and subtract $(\Sigma_S^{-1} \boldsymbol{\mu} 
@@
 $$
-
 With the last equality resulting from the use of the Sherman–Morrison–Woodbury formula, 
@@
 From our model specification, we have the following:
-
 \[
@@
 \]
-
 Then, to find the distribution of $S|\bar{Y}$, we apply Bayes' Theorem below, 
@@
 \]
-
 \[
@@
 \]
-
 From properties of the multivariate normal distribution \citep[e.g., Equation 7.1 of][]{hoff2009first},
@@
 and $C$ respectively, where we can solve for these values with:
-
 \[
@@
 \]
-
 Then,
@@
 \]
-
 Therefore, we have the result that $S|\bar{Y} \sim \mathcal{N}(\boldsymbol{m}, C)$.

### 2025-06-26 paper/main.tex
rm old figures, remove color on grant

@@
 
-\textcolor{red}{This work was in part supported by the U.S. Department
+This work was in part supported by the U.S. Department
 of Energy, Office of Science, Office of Advanced Scientific Computing Research, 
 Scientific Discovery through Advanced Computing (SciDAC) program through 
-Grant 420453 "Enabling Cosmic Discoveries in the Exascale Era."}
+Grant 420453 ``Enabling Cosmic Discoveries in the Exascale Era."
 

### 2025-06-26 fitting/CAMB_vs_L1300.R
rm unneeded files, init commit collect_means_CAMB.R

@@ -1,51 +1,0 @@
-library(dgp.hm)
-
-# CAMB has runs 1-64; L1300 & L2100 has runs 1-32
-for (run in 1:32) {
-
-  # growth values; need to multiply CAMB by growth^2 (only 32 tho...)
-  g = read.table("../CosmicEmu_etc/pows/growth.dat")[run,2]
-  # hubble (h) values; mult k by h, divide P(k) by h^3
-  h = read.csv("../CosmicEmu_etc/pow64/cambDesigns_32x6x2.csv")[run,2]
-  
-  # read in and plot all 15 LR runs (adjusted by h and h^3)
-  lr = read.table(paste0("../CosmicEmu_etc/pows/M0",if(run<10){"0"},run,
-                         "/L1300/PM001/output/m0",
-                         if(run<10){"0"},run,".pk.ini"))
-  if(run==1) plot(log10(h*lr[which(h*lr[,1]<=1),1]),
-                  scrP(lr[which(h*lr[,1]<=1),2]/h^3,lr[which(h*lr[,1]<=1),1]*h),
-                  col="gray", type="l", main = run,
-                  ylab = "script P (adj by h,g)",
-                  xlab = "log10(k) (adj by h)")# ylim=c(-5,-2))
-  if(run!=1) plot(log10(h*lr[which(h*lr[,1]<=1),1]),
-                  scrP(lr[which(h*lr[,1]<=1),2]/h^3,lr[which(h*lr[,1]<=1),1]*h),
-                  col="gray", main=run, type="l",
-                  ylab = "script P (adj by h,g)",
-                  xlab = "log10(k) (adj by h)")
-  lrs = matrix(NA, 15, 206)
-  lrs[1,] = scrP(lr[which(h*lr[,1]<=1),2]/h^3,lr[which(h*lr[,1]<=1),1]*h)
-  
-  for (i in 2:15) {
-    if(run ==8 & i ==9) next
-    lr = read.table(paste0("../CosmicEmu_etc/pows/M0",if(run<10){"0"},run,
-                           "/L1300/PM0",if(i<10){"0"},i,"/output/m0",
-                           if(run<10){"0"},run,".pk.ini"))
-    lrs[i,] = scrP(lr[which(h*lr[,1]<=1),2]/h^3,lr[which(h*lr[,1]<=1),1]*h)
-    lines(log10(h*lr[which(h*lr[,1]<=1),1]), lrs[i,], col="gray")
-  }
[…]

### 2025-06-26 fitting/collect_means_CAMB.R

@@ -1,0 +1,30 @@
+# collect the 32 posterior means from CAMB model fits
+post_means <- data.frame(matrix(NA, nrow = 400, ncol = 32))
+for (i in 1:32) {
+  model_name <- i
+  results <- read.csv(paste0("results/CAMB/dgp_", model_name, "postmean_int.csv"))[,2]
+  post_means[, i] <- results
+  colnames(post_means)[i] <- model_name
+}
+write.csv(post_means, "results/CAMB/post_means_int_CAMB.csv", row.names = FALSE)
+
+x_unif = read.csv(paste0("results/CAMB/dgp_", model_name, "postmean_int.csv"))[,1]
+write.csv(x_unif, "results/CAMB/xcamb_int.csv", row.names = F)
+
+
+library(dgp.hm)
+library(pracma)
+
+# for comparison, collect the 64 inf-res runs for all 64 CAMB model runs
+inf_res_ints <- data.frame(matrix(NA, nrow = 400, ncol = 64))
+for (model in 1:64) {
+  camb = read.table(paste0("../CosmicEmu_etc/pow64/RUN",model,
+                           "/oneh_matterpower.dat"))
+  h = read.csv("../CosmicEmu_etc/pow64/cambDesigns_32x6x2.csv")[model,2]
+  x_camb = log10(camb[,1]*h)
+  y_camb = scrP(camb[,2]/h^3, 10^x_camb)
+  inf_res_ints[,model] <- interp1(x_camb, y_camb, x_unif, method = "cubic")
+}
+write.table(inf_res_ints, file = "results/CAMB/inf_res_int.txt",
+            row.names = FALSE, col.names = FALSE)
+

### 2025-06-26 fitting/initialize.R

@@ -1,124 +1,0 @@
-
-###############################################################################
-# This script fits the deep Bayesian hierarchical model for M001 for many 
-# MCMC iterations and saves the resulting hyperparameters (w_0, theta_w_0, and
-# theta_y_0) to csv files.  They will be used to initialize fits for other 
-# models.
-# 
-# Outputs:
-#    writes csv file of final mcmc iteration for w and hyperparameters
-#    (theta_y, theta_w)
-#
-###############################################################################
-
-library(dgp.hm)
-library(zoo) # rollmean
-library(Matrix) # bdiag
-
-# Load data for M001 ----------------------------------------------------------
-
-model_name <- "M001"
-
-# 1st column is k, 2nd is linear pert theory, 3:18 is low-res, 19 is hi-res
-file_name <- paste0("../Mira-Titan-IV-Data/Mira-Titan-2021/STEP499/pk_", 
-                    model_name, "_test.dat")
-pk2 <- read.table(file_name)
-n <- nrow(pk2)
-nrun <- 16 # number of low res runs
-
-# Get precisions --------------------------------------------------------------
-
-# Load precision info (k, prec_highres, prec_lowres, index_list)
-load("../Mira-Titan-IV-Data/precision_and_indexes.Rdata")
-
-# Get precision info for each data type (low-res, hi-res, pert theory)
-precs_lo <- ifelse(1:n %in% index_list$lowres.ix, prec_lowres, 0) * nrun
-precs_hi <- ifelse(1:n %in% index_list$highres.ix, prec_highres, 0)
-precs_pt <- ifelse(1:n %in% index_list$pert.ix, 10^8, 0)
-
-# Smooth the precision information (to remove steps)
-Lam_lo <- rollmean(precs_lo, k = 10, fill = "extend")
-Lam_hi <- rollmean(precs_hi, k = 10, fill = "extend")
-Lam_pt <- rollmean(precs_pt, k = 10, fill = "extend")
-
-# Precision matrix for weighted average, mu_z
-Lam_z <- Lam_pt + Lam_lo + Lam_hi
-
-# Get inputs ------------------------------------------------------------------
-
-# log10 of wavenumber (k) is x
-x <- log10(k)
[…]

### 2025-06-26 fitting/run_LT.R

@@ -1,118 +1,0 @@
-# Power Spectra Simulations and linear theory (LT)
-
-library(dgp.hm)
-
-pdf("LT_sims_scrP.pdf")
-
-# All the Runs and Models are in units Mpc/h and h/Mpc.
-# To be able to plot everything in one plot, we have to convert to Mpc or 1/Mpc.
-# To convert k (column 1), we need to multiply it by hubble (to get k in units of 1/Mpc) 
-# while the power spectrum (column 2) needs to be divided by h^3 (to get it in Mpc^3).
-
-# contains model/run index, D+ values, and hubble values
-data <- read.csv("../CosmicEmu_etc/32 Models Mira Titan/models_d+_h.csv")
-Dp <- data$D.
-h <- data$Hubble
-
-# Working with runs 1-32 (aka models 11-42), but skip 17-18 (27-28)
-for(run in 1:32){
-  model <- run + 10
-  if(model %in% 27:28) next
-  # read in the linear output
-  sim <- read.table(paste0("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN",
-                           run,"/oneh_matterpower.dat"))
-  k_s <- sim$V1*h[run]
-  pk_s <- sim$V2/ (h[run])^3
-  plot(log10(k_s), scrP(pk_s, k_s), type="l", col="blue", lwd=4,
-       xlab=expression(log[10](k)), ylab="script P(k)", ylim=c(-0.2,1.4), xlim = c(-2.3,0),
-       main = paste0("LT and simulated spectra, model ", model))
-  
-  # read in first batch of model runs (model #11)
-  pk_ms <- matrix(NA, 886, 15)
-  for (i in 0:15) {
-    if(model > 10) folder = "pow.ic"
-    if(model > 19) folder = "pow_2.ic"
-    if(model > 29) folder = "pow_3.ic"
-    if(model > 39) folder = "pow_4.ic"
-    mod <- read.table(paste0("../CosmicEmu_etc/32 Models Mira Titan/",folder,"/M0", 
-                             model, "/L1300/PM0",if(i < 10){"0"}, i,
-                             "/analysis/Pow/m0", model, ".pk.ini"))
-    k_m <- mod$V1 * h[model-10]
-    pk_ms[,i] <- pk_m <- mod$V2 / (h[model-10])^3 / (Dp[model-10])^2
[…]

### 2025-06-26 pca/predict_LT.R

@@ -1,157 +1,0 @@
-#######################################################################
-# This script obtains principal components and their corresponding weights 
-# Then predict cosmologies' spectra for LINEAR THEORY test cases
-#
-# Models 11-42 (except 27-28) are considered here.
-#
-###############################################################################
-
-pdf("predict_LT_9PCs.pdf")
-
-tic <- proc.time()[3]
-
-library(GPfit)
-
-args <- commandArgs(TRUE)
-if(length(args) > 0)
-  for(i in 1:length(args))
-    eval(parse(text=args[[i]]))
-
-#########################################
-# Preprocess the data (h, interp, etc.) #
-#########################################
-
-# load the training design
-all_runs <- c(1:16,19:32)
-for(test_run in c(5,10,15,20,25,30)){ #1:16,19:32
-  print(paste0("test run is ", test_run))
-  training_runs <- setdiff(all_runs, test_run)
-  des_train_full <- read.csv("../Mira-Titan-IV-Data/design_train.txt", header=F, sep = "")
-  des_train <- des_train_full[training_runs,]
-  p = ncol(des_train)  # number of input parameters
-  
-  # contains model/run index, D+ values, and hubble values
-  # in order to appropriate scale the linear theory (and model runs)
-  data <- read.csv("../CosmicEmu_etc/32 Models Mira Titan/models_d+_h.csv")
-  Dp <- data$D.
-  h <- data$Hubble
-  
-  # Load the posterior means obtained from run_fit.R and collect_means.R
-  # Test on run #19
-  k_int <- seq(-4, 1, by=0.01)
-  eta <- matrix(NA, length(k_int), length(training_runs))
-  for (i in 1:29) {
-    run <- training_runs[i]
-    sim <- read.table(paste0("../CosmicEmu_etc/32 Models Mira Titan/linear/RUN",
-                             run,"/oneh_matterpower.dat"))
-    k_s <- sim$V1*h[run]
-    pk_s <- sim$V2/ (h[run])^3
-    eta[,i] <- approx(log10(k_s), scrP(pk_s,k_s), k_int)$y 
-  }
-  matplot(k_int, eta, type="l", lty=1, col="gray", main = paste0("test #", test_run))
-  nruns = ncol(eta)
-  
-  # Use log10 k values
-  kvals <- k_int
-  
-  #################################
[…]

### 2025-06-26 plotting/plot_Sigmahat.R
rm old plotting scripts

@@ -1,116 +1,0 @@
-
-###############################################################################
-# For a particular model, plot the Sigma_hat matrix.
-#
-# Specifications:
-#    model: model number (0-116)
-#
-###############################################################################
-
-library(dgp.hm)
-library(zoo) 
-library(Matrix) 
-library(fields)
-
-# Get model data --------------------------------------------------------------
-
-model <- 1 # integer 0-116 (0, 112:116 are testing)
-if (model <= 111) {
-  model_name <- paste0("M", if (model < 100) {"0"}, if (model < 10) {"0"}, model) 
-} else {
-  test_names <- c("E001", "E002", "E003", "E009", "E010")
-  model_name <- test_names[model - 111]
-}
-
-file_name <- paste0("../Mira-Titan-IV-Data/Mira-Titan-2021/STEP499/pk_", 
-                    model_name, "_test.dat")
-pk2 <- read.table(file_name)
-n <- nrow(pk2)
-nrun <- 16 # number of low res runs
-
-# Get precisions --------------------------------------------------------------
-
-# Load precision info (k, prec_highres, prec_lowres, index_list)
-load("../Mira-Titan-IV-Data/precision_and_indexes.Rdata")
-
-# Get precision info for each data type (low-res, hi-res, pert theory)
-precs_lo <- ifelse(1:n %in% index_list$lowres.ix, prec_lowres, 0) * nrun
-precs_hi <- ifelse(1:n %in% index_list$highres.ix, prec_highres, 0)
-precs_pt <- ifelse(1:n %in% index_list$pert.ix, 10^8, 0)
-
-# Smooth the precision information (to remove steps)
-Lam_lo <- rollmean(precs_lo, k = 10, fill = "extend")
-Lam_hi <- rollmean(precs_hi, k = 10, fill = "extend")
-Lam_pt <- rollmean(precs_pt, k = 10, fill = "extend")
-
-# Precision matrix for weighted average, mu_z
-Lam_z <- Lam_pt + Lam_lo + Lam_hi
-
-# Get inputs ------------------------------------------------------------------
-
-# log10 of wavenumber (k) is x
-x <- log10(k)
-x <- (x - min(x)) / (max(x) - min(x))
-dx <- deepgp::sq_dist(x)
-
[…]

### 2025-06-26 plotting/plot_main_effects.R

@@ -1,199 +1,0 @@
-#####################################
-# Script with plotting for PCA fits #
-# and main effects estimates, and   #
-# variance decomposition            #
-#####################################
-
-##############################
-# Load the preprocessed data #
-##############################
-
-# load the training design
-des_train <- read.csv("../Mira-Titan-IV-Data/design_train.txt", header=F, sep = "")
-p = ncol(des_train)  # number of input parameters
-
-# load the wavenumber vals
-load("../Mira-Titan-IV-Data/precision_and_indexes.Rdata")
-kvals <- log10(k)
-
-# Load the posterior means obtained from run_fit.R and collect_means.R
-eta <- read.csv("../fitting/results/post_means_train.csv")
-nruns = ncol(eta)
-
-# Read in test design
-# des_test <- read.csv("../Mira-Titan-IV-Data/design_test.txt",sep="",header = F)
-m = 100
-des_test <- read.csv(paste0("../csv/des_test_unif",m,".csv"), row.names = 1)
-ntest <- nrow(des_test)
-
-# read in predicted power matter spectra for test models
-etaEmu <- read.csv(paste0("../pca/etaEmu_unif",m,".csv"))
-# colnames(etaEmu) <- c("E001", "E002", "E003", "E009", "E010", "M000")
-ntest = ncol(etaEmu)
-
-######################
-# Obtain predictions #
-######################
-
-# n_pc: number of principal components to model
-n_pc = 5
-if(n_pc < 1 | n_pc > 111) stop("number of PCs has to be between 1 and 111")
-
-# choose the power for powered-exponential covariance function 
-pwdExp = 1.9
-
-# n_k: number of k values considered for each cosmology
-n_k <- length(kvals)
-
-# calculate overall mean of 111 posterior means
-mean_mat = matrix(apply(eta,1,mean),nrow=n_k,ncol=nruns)
-
-# convert all 111 posterior means to have zero mean
-eta0 = eta - mean_mat
-
-# use singular value decomposition
-a = svd(eta0)
-
[…]

### 2025-06-28 paper/main.tex
some quick/easy changes

@@
 \title{Bayesian Deep Gaussian Processes for Correlated Functional Data: \\
-        A Case Study in Cosmological Power Spectra}
+        A Case Study in Cosmological Matter Power Spectra}
 \author{Stephen A. Walsh\thanks{Corresponding author: Division of Natural Sciences, 
@@
         David Higdon\footnotemark[2] \and
-        Marco A.R. Ferreira\footnotemark[2] \and
         Jared Clark\footnotemark[2] \and
@@
 %inversely related to spatial scale. 
-On large scales, power spectrum behave according to linear perturbation 
+On large scales, power spectra behave according to linear perturbation 
 theory \citep{pietroni2008flowing, lesgourgues2009non}.  On smaller scales, nonlinear 
@@
 Our overarching goal is to leverage such simulation data to 
-first estimate and then predict underlying power matter spectra for various 
+first estimate and then predict underlying matter power spectra for various 
 cosmological parameterizations.
@@
 defines ``cosmologies'' according to eight cosmological parameters (more on this
-in Section \ref{sec:data}).  For a particular cosmology, the simulation of the power matter spectrum 
+in Section \ref{sec:data}).  For a particular cosmology, the simulation of the matter power spectrum 
 is rather complex.  Although we presume there exists some true matter spectrum as a function
@@
 must: handle multiple correlated functional observations, offer enough 
-flexibility to accommodate the nonstationarity present in power matter spectra 
+flexibility to accommodate the nonstationarity present in matter power spectra 
 (which vary in smoothness across wavenumbers), provide effective uncertainty
@@
 Our second objective is to leverage simulation data from a limited set of
-cosmologies in order to predict power matter spectra for unobserved cosmologies.
+cosmologies in order to predict matter power spectra for unobserved cosmologies.
 Due to the computationally expensive nature of the Mira-Titan simulation suite
@@
 \citep{santner2003design,gramacy2020surrogates} which will provide quick and effective
-predictions of power matter spectra for any cosmological parameterization.
+predictions of matter power spectra for any cosmological parameterization.
 \citet{moran2023mira} provide a state-of-the-art emulator for the Mira-Titan
-simulation suite, which was trained on 117 simulations, and is termed 
+simulation suite, which was trained on 111 simulations, and is termed 
 the ``Cosmic Emu.''\footnote{\url{https://github.com/lanl/CosmicEmu}}
@@
 
-The Mira-Titan simulation dataset consists of simulated power spectra for 
-117 unique cosmologies. Each of these cosmologies is specified by eight cosmological 
+The Mira-Titan simulation dataset consists of simulated matter power spectra for 
+117 unique cosmologies (111 for training, 6 for testing). 
+Each of these cosmologies is specified by eight cosmological 
 parameters: matter density, baryon density, amplitude of density fluctuations, 
@@
     \includegraphics[width=\linewidth]{plot_data.jpeg}
-    \caption{{\it Left:} The perturbation theory (solid green), low resolution runs (solid grey), 
+    \caption{{\it Left:} The perturbation theory (solid green), low resolution runs (solid gray), 
     and high resolution run  (solid pink) for a particular cosmology. 
@@
     {\it Middle:} Same as left, but restricted to wavenumber ($k$) values where the low resolution 
-    runs are approximately unbiased.
+    runs are approximately unbiased, i.e., the span of the gray dashed line in the left sub-figure.
     {\it Right:} Same as middle, but with LOESS-smooth weighted average subtracted.}
@@
 %cosmologies held out for testing.  Our ultimate objective is to effectively predict
-%the underlying power matter spectra for the held-out cosmologies based on their 
+%the underlying matter power spectra for the held-out cosmologies based on their 
 %8-dimensional input configuration (Section \ref{sec:pred}).  But first, we must 
@@
 
-Let $S$ represent the true (but unknown) underlying power matter spectrum for
+Let $S$ represent the true (but unknown) underlying matter power spectrum for
 a particular cosmology.  We consider $\bar{Y}$ a noisy realization of $S$.
@@
 provided in the {\tt deepgp} {\sf R}-package \citep{deepgp}.
-We then integrate Metropolis-Hastings sampling of these parameters in a Gibbs framework 
-with the ESS sampling of $W$ \citep{sauer2023active}.
+We then integrate Metropolis-Hastings sampling of these parameters in a Gibbs (i.e., one parameter 
+at a time updating) framework with the ESS sampling of $W$ \citep{sauer2023active}.
 
@@
 we may leverage Bayes' Theorem to obtain the
-posterior distribution of the true power matter spectrum for a particular cosmology as
+posterior distribution of the true matter power spectrum for a particular cosmology as
 \[
@@
 
-The CAMB \citep{lewis2011CAMB} dataset contains power spectra based
+The CAMB \citep{lewis2011CAMB} dataset contains matter power spectra based
 on six cosmological parameter settings. It includes sixty-four cosmologies,
@@
 Mira-Titan has 8 such parameters as described in Section \ref{sec:data}; CAMB has 6.
-Our goal is to effectively predict the power matter spectrum as a function of $\psi$, i.e., $\hat{S}(\psi)$, 
+Our goal is to effectively predict the matter power spectrum as a function of $\psi$, i.e., $\hat{S}(\psi)$, 
 for a held-out cosmology.  We observe $m$-many training cosmologies and fit our DGP.FCO 
@@
 
+\noindent
 where $\boldsymbol{\gamma}_i = [\gamma_i^{(1)}, \gamma_i^{(2)}, \ldots, \gamma_i^{(m)}]^\top$ 
@@
     \includegraphics[width=0.85\textwidth]{pred_diffs_CAMB.jpeg}
-    \caption{Predictions of power matter spectra for each of the 32 CAMB test cosmologies,
+    \caption{Predictions of matter power spectra for each of the 32 CAMB test cosmologies,
              after centering and scaling (to accentuate differences). 
@@
 The agreement between our predictions and the ``best-case'' predictions indicates that DGP.FCO is effectively
-capturing the true power matter spectra of the training cosmologies.
+capturing the true matter power spectra of the training cosmologies.
 The high level of agreement between the pairs of predictions and the true spectra across cosmologies
@@
 (leveraging the perturbation theory, low resolution runs, and high resolution run to predict the underlying
-power matter spectrum).  Although these predicted ``in-sample'' spectra are not formal ``truths,'' we find them to be
+matter power spectrum).  Although these predicted ``in-sample'' spectra are not formal ``truths,'' we find them to be
 useful benchmarks for our ``out-of-sample'' predictions.  The individual predictions for each method and 
@@
 
-\section{Conditional Distribution of Power Spectrum, Given Weighted Average}
+\section{Conditional Distribution of Matter Power Spectrum, Given Weighted Average}
 \label{sec:apdx_SgivenY}

### 2025-06-30 paper/main.tex
fix typos, minor adjustments

@@
 We are particularly motivated by the Mira-Titan Universe simulation
-suite which, for a specified cosmological parameterization (termed a ``cosmology''), 
+suite that, for a specified cosmological parameterization (termed a ``cosmology''), 
 provides multiple response curves of various fidelities, including correlated 
@@
 Computer simulation experiments are invaluable tools in the study of cosmology.
-Experiments which simulate the expansion of the universe are growing
-in prevalance and complexity \citep[e.g.,][]{lawrence2010coyote,derose2019aemulus,
+Experiments that simulate the expansion of the universe are growing
+in prevalence and complexity \citep[e.g.,][]{lawrence2010coyote,derose2019aemulus,
 nishimichi2019dark,angulo2021bacco,euclid2021euclid,moran2023mira}.  
@@
 is rather complex.  Although we presume there exists some true matter spectrum as a function
-of wavenumber, we are not able to directly observe it.  Rather, the simulation returns
-a convoluted patchwork of potential spectra, which vary in fidelity and are only trusted
-over particular wavenumber ranges.  Our first objective is to synthesize expert knowledge
-and all available simulation information into an effective estimate of the true underlying
+of wavenumber, we are not able to directly observe it.  Rather, the simulations for a 
+particular cosmology return a collection of multiple potential spectra, 
+which vary in fidelity and are only trusted over particular wavenumber ranges.  
+Our first objective is to synthesize expert knowledge and all available 
+simulation information into an effective estimate of the true underlying
 matter spectrum.
@@
 regarding the wavenumbers over which various outputs are trusted.  To this end, 
-we propose a Bayesian hierarchical model which treats functional simulation 
+we propose a Bayesian hierarchical model that treats functional simulation 
 outputs as realizations of a Gaussian process (GP) 
@@
 they have yet to be extended to correlated functional outputs. 
-To demonstrate the profiency of our Bayesian hierarchical DGP, 
+To demonstrate the proficiency of our Bayesian hierarchical DGP, 
 we benchmark its performance in two unique settings: with synthetic data 
-mimicing the Mira-Titan data and with real-world data from 
+mimicking the Mira-Titan data and with predictions of the linear power spectrum from 
 the Code for Anisotropies in the Microwave Background \citep[CAMB;][]{lewis2011CAMB}.
@@
 it is infeasible to evaluate the simulation for every possible eight-dimensional
-cosmological configuration which may be of interest.  Instead, we desire
+cosmological configuration that may be of interest.  Instead, we desire
 a statistical ``emulator'' or ``surrogate model'' 
-\citep{santner2003design,gramacy2020surrogates} which will provide quick and effective
+\citep{santner2003design,gramacy2020surrogates} that will provide quick and effective
 predictions of matter power spectra for any cosmological parameterization.
@@
 
-For each cosmology, Mira-Titan contains a batch of 18 simulated spectra: an inexpensive 
-spectrum estimated from linear perturbation theory ($y_p$), 16 spectra estimated from 
-low resolution simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one spectrum from 
+For each cosmology, Mira-Titan contains a batch of 18 simulated power spectra: an inexpensive 
+power spectrum estimated from perturbation theory ($y_p$), 16 power spectra estimated from 
+low resolution simulations ($y_{\ell_r}, r \in \{1,\dots,16\}$), and one power spectrum from 
 a high resolution simulation ($y_h$). 
-These spectra are represented as a function of wavenumber $k$.  Following 
+These power spectra are represented as a function of wavenumber $k$.  Following 
 \cite{moran2023mira}, we model on the emulation space 
@@
 of the output for a particular cosmology. When we focus on $k$ values where $y_{\ell_r}$ 
-is valid (center panel), the spatially correlated errors of the low-resolution and high-resolution runs 
-becomes apparent.
+is valid (center panel), the spatially correlated fluctuations of the low-resolution and 
+high-resolution runs becomes apparent.
 
 Notice, in the center/right panels of Figure \ref{fig:plot_data}, the precision of each
-low/high resolution spectra increases as $k$ increases.  The curves get progressively 
+low/high resolution spectrum increases as $k$ increases.  The curves get progressively 
 tighter from left to right.  We would like to conveniently wrap this domain-specific 
@@
 wavenumber values using a log-log regression model. A multiplier $c\approx 3.73$ for 
-the increase in precision from the low- to high- resolution output is also 
+the increase in precision from the low- to high-resolution output is also 
 estimated with this regression model. 
@@
 This is especially apparent after subtracting a LOESS-smoothed weighted average (right panel).
-Although the weighted average incorporates all available observatons and 
+Although the weighted average incorporates all available observations and 
 expert knowledge regarding the precisions, it fails to account for the spatial 
 dependence inherent in the smooth but stochastic spectra, which is pivotal to 
-effectively inferring the true spectrum.
+effectively inferring the smooth spectrum.
 
@@
 (as visualized in Figure \ref{fig:plot_data}).  We thus specify a dense 
-covariance matrix $\Sigma_\varepsilon$ that simulataneously
+covariance matrix $\Sigma_\varepsilon$ that simultaneously
 allows us to model the correlated errors and build domain-matter expertise
@@
 \label{eq:prior}
-W &\sim \mathrm{monoGP}\left(\mathbf{0}, \Sigma_W(X)\right)
+W &\sim \mathrm{monoGP}\left(\mathbf{0}, \Sigma_W(X)\right).
 \end{align}
@@
 $m_1 \sim \text{Uniform}(0.5,1.5)$, $u_1 \sim \text{Uniform}(1.5,2.5)$, and $m_2,u_2 
-\sim \text{Uniform}(0.6,1.4)$. The outputs of $f_1$ and $f_2$ (conditional on $m_1/u_1$ and 
+\sim \text{Uniform}(0.6,1.4)$. The outputs of $f_1$ and $f_2$ (conditional on $m_1,u_1$ and 
 $m_2,u_2$), determine the true function values for each simulation.
@@
 and $\Sigma_B$ for numerical stability.  Within $\Sigma_B$, $s$ introduces a degree of
-nonstationarity, more effectively mimicing the behavior of the 
-CAMB and Mira-Titan datasets.  Examples of simulation output are showin in 
+nonstationarity, more effectively mimicking the behavior of the 
+CAMB and Mira-Titan datasets.  Examples of simulation output are shown in 
 Figure \ref{fig:run_sim_f1A_f2B}.
@@
 the \texttt{hetGP} package \citep{binois2018practical, binois2021hetgp}, and
-a deep process convolution (DPC) approach, which is utilized within Cosmic 
+a deep process convolution (DPC) approach, which is utilized within the Cosmic 
 Emu \citep{moran2023mira}.  These competing methods view the training data as 
@@
 slice samples of $W$.
-These warpings depart from the identiy mapping around $x\approx-1$, 
+These warpings depart from the identity mapping around $x\approx-1$, 
 stretching inputs where dynamics are shifting more rapidly for larger $k$ values.
@@
              means and training with the ``true" infinite-resolution spectra. The identity line 
-             provides equal MSE values amongst both methods.}   
+             provides equal MSE values among both methods.}   
     \label{fig:mse_camb}
@@
     \caption{Results of all six predictions for both methods on the held-out 
-    cosmologies (DGP.FCO posterior mean subtracted). CosmicEMU is dotted blue and DGP.FCO is dashed orange.
+    cosmologies (DGP.FCO posterior mean subtracted). Cosmic Emu is dotted blue and DGP.FCO is dashed orange.
     Closer to the dotted zero line indicates a better fit with the DGP.FCO posterior mean.}
@@
 Scientific Discovery through Advanced Computing (SciDAC) program through 
-Grant 420453 ``Enabling Cosmic Discoveries in the Exascale Era."
+Grant 420453 ``Enabling Cosmic Discoveries in the Exascale Era." Work at Argonne 
+National Laboratory was supported under the U.S. Department of Energy 
+contract DE-AC02-06CH11357.
 

### 2025-06-30 paper/main.tex
minor tweaks/updates to paper

@@
 a high resolution simulation ($y_h$). 
-These power spectra are represented as a function of wavenumber $k$.  Following 
+These power spectra are represented as a function of wavenumber $k$, expressed in units of 
+$\text{Mpc}^{-1}$ (inverse megaparsecs).  Following 
 \cite{moran2023mira}, we model on the emulation space 
@@
 original spectra. Output is available for $n=351$ values of $k$, spanning
-$0.001 \leq k \leq 5$, but each data type has unique wavenumber ranges
-where it is deemed approximately unbiased. For $k<0.04$, only $y_p$ provides a reliable 
-estimate of the true spectra. $y_{\ell_r}$ are valid for $0.04 \leq k \leq 0.25$, and 
-$y_h$ is valid for $0.04 \leq k \leq 5$.  Figure \ref{fig:plot_data} shows an example 
+$0.001 \leq k \leq 5 \text{ Mpc}^{-1}$, but each data type has unique wavenumber ranges
+where it is deemed approximately unbiased. For $k<0.04 \text{ Mpc}^{-1}$, only $y_p$ provides a reliable 
+estimate of the true spectra. $y_{\ell_r}$ are valid for $0.04 \leq k \leq 0.25 \text{ Mpc}^{-1}$, and 
+$y_h$ is valid for $0.04 \leq k \leq 5 \text{ Mpc}^{-1}$.  Figure \ref{fig:plot_data} shows an example 
 of the output for a particular cosmology. When we focus on $k$ values where $y_{\ell_r}$ 

### 2025-06-30 paper/main.tex
change wording around monoGP

@@
 baryonic acoustic oscillations when $-2 \leq \log_{10}(k) \leq -1$. To account 
-for this, we incorporate a latent monotonic Gaussian process \citep[monoGP;][]{barnett2025monotonic},
-which accommodates nonstationarity by warping $X$ via 
+for this, we incorporate a latent Gaussian process that we force to be monotonic
+following \citet[][denoted ``monoGP'']{barnett2025monotonic}.  This latent layer
+accommodates nonstationarity by warping $X$ via 
 $W \sim \mathrm{monoGP}\left(\boldsymbol{\mu}_W, \Sigma_W(X)\right)$.

### 2025-06-30 pca/predict.R
specify predict for GPfit

@@ -23,7 +23,7 @@ ntest <- nrow(des_test)
 aps <- list()
 for (i in 1:n_pc) {
   print(paste("GP pred",i))
-  aps[[i]] = predict(as[[i]],des_test)
+  aps[[i]] = GPfit::predict.GP(as[[i]], des_test)
 }
 
 # create matrix of overall mean to add back on to predictions

### 2025-06-30 plotting/plot_pca.R

@@ -219,7 +219,7 @@ ntest <- nrow(des_test)
 aps <- list()
 for (i in 1:n_pc) {
   print(paste("GP pred",i))
-  aps[[i]] = predict(as[[i]],des_test)
+  aps[[i]] = GPfit::predict.GP(as[[i]], des_test)
 }
 
 # create matrix of overall mean to add back on to predictions

### 2025-06-30 paper/main.tex
remove unneeded comments and sepia details from manuscript
adjust images per coauthor suggestions

@@
 
-%To further our understanding of the structure and movement of the universe, 
-%researchers utilize numerous cosmological surveys such as the Sloan Digital Sky 
-%Survey \citep{york2000sloan} and the upcoming Nancy Grace Roman Space Telescope 
-%\citep{Dore2019WFIRST} which are continuously growing in complexity. With these 
-%tools, cosmologists aim to continue learning more about cosmic acceleration 
-%\citep{caldwell2009physics}. To this end, one important ingredient to aid in this 
-%understanding of our universe is the matter power spectrum.
-
 Computer simulation experiments are invaluable tools in the study of cosmology.
@@
 function of spatial scale. 
-%It is often represented as a function of wavenumber $k$ (units Mpc$^{-1}$), which is 
-%inversely related to spatial scale. 
 On large scales, power spectra behave according to linear perturbation 
@@
 dynamics require the use of computationally intensive simulations, yielding
-correlated functional data.
-%the Coyote Universe \citep{lawrence2010coyote} and the Mira-Titan Universe \citep{moran2023mira} 
-%are two simulation suites dedicated to this effort.  
-Our overarching goal is to leverage such simulation data to 
+correlated functional data. Our overarching goal is to leverage such simulation data to 
 first estimate and then predict underlying matter power spectra for various 
@@
 Bayesian DGP model in a public git repository.\footnote{\url{https://github.com/stevewalsh124/dgp.hm}}
-%\textcolor{cyan}{Please note: this repository is currently private until we 
-%hear back regarding permissions for sharing the Mira-Titan data alongside our code.}
-
-%Leveraging these simulations allows for emulation and prediction of the matter power 
-%spectrum under varying specifications for eight different cosmological parameters 
-%(i.e., different cosmologies). Estimating and understanding the influence of these 
-%parameters on cosmic expansion are areas of active research. In addition to 
-%presenting the Mira-Titan simulation suite, \cite{moran2023mira} builds off 
-%of previous spectrum emulation \citep{lawrence2017mira} to provide the final 
-%emulator (Cosmic Emu) based on the full suite of 117 batches of simulations.
-
-%In this work, we expand on the previous emulation approaches and propose a 
-%novel synthesis of methods to quantify uncertainty and predict matter power 
-%spectra for different cosmologies using deep Gaussian processes (DGPs). In order 
-%to handle multiple realizations of functional output (from the Mira-Titan 
-%simulation suite), we use hierarchical modeling and basis functions within our 
-%DGP framework to accurately estimate and predict power spectra for different 
-%cosmologies.
 
@@
 There is an optional ninth parameter, the redshift, which we fix at 0.
-%Different simulation output exists for different values of a ninth value, 
-%the redshift; in this work, we fix 
-%this value at 0.
 
-% after \linewidth:, trim=0 13 0 45, clip=TRUE
 \begin{figure}[ht]
@@
     \caption{{\it Left:} The perturbation theory (solid green), low resolution runs (solid gray), 
-    and high resolution run  (solid pink) for a particular cosmology. 
+    and high resolution run  (solid pink) for the first cosmology. 
     The weighted average of these is shown in dashed orange. Dotted lines at the bottom
@@
 
-%We consider a training-testing split with 111 cosmologies used for estimation and 6 
-%cosmologies held out for testing.  Our ultimate objective is to effectively predict
-%the underlying matter power spectra for the held-out cosmologies based on their 
-%8-dimensional input configuration (Section \ref{sec:pred}).  But first, we must 
-%obtain effective estimates of the underlying spectrum for each training 
-%cosmology (Section \ref{sec:hm_fit}). \textcolor{red}{(@Steve - maybe this is a
-%good place to introduce the ``hold out'' and ``within sample'' terms?)}
-
 \section{Bayesian Hierarchical Modeling for Particular Cosmologies}
@@
 random variable within the statistical model.
-%Together, $X$ and $\bar{Y}$ comprise the 
-%training data for a particular cosmology. 
+
 $\bar{Y}$ is typically pre-processed
@@
     \includegraphics[width=\textwidth]{CAMB_fit_model1.jpeg}
-    \caption{Example of DGP.FCO model fit for the first batch of model runs from the CAMB data, 
-             with a LOESS mean of the weighted average subtracted to emphasize details. We can 
+    \caption{{\it Left:} The infinite-resolution (dotted blue), 
+             low-resolution runs (solid gray), and high-resolution run (solid pink) 
+             for the first CAMB cosmology.
+             {\it Right:} Example of DGP.FCO model fit for the first CAMB cosmology, 
+             with a LOESS mean of the 
+             weighted average subtracted to emphasize details. We can 
              compare the orange lines (posterior mean and dotted 95\% credible interval) to the 
@@
 We then fit our DGP.FCO model on each of 111 training cosmologies, reserving 6 cosmologies
-for later hold-out testing.  Figure \ref{fig:plot_fit} 
-visualizes the DGP.FCO fit for the first training cosmology.  We center the plot
+for later hold-out testing.  The left panel of Figure \ref{fig:plot_fit} 
+visualizes the DGP.FCO fit for the first training cosmology (compare this with the left
+panel of Figure \ref{fig:plot_data}).  We center the plot
 by subtracting the posterior mean in order to more easily visualize the credible
@@
     \centering
-    \includegraphics[width=6in]{plot_fit.jpeg}
+    \includegraphics[width=\textwidth]{plot_fit.jpeg}
     \caption{{\it Left:} 95\% credible intervals for the power spectrum of the first Mira-Titan
@@
 from training cosmologies to predict the matter power spectra for held-out cosmologies.
-%This is achieved with modeling the weights of different basis functions as GPs.
 
@@
 where $B_i$ is the $i^\textrm{th}$ basis function (column) of $B$.
-
-% SEPIA information:
-% In alignment with \cite{heitmann2009coyote} and \cite{higdon2010estcosmo}, we use 
-% GPs to model these principal components' weights. For direct comparison alongside
-% \cite{moran2023mira}, we employ a Python implementation of this approach called 
-% SEPIA \citep{gattiker2020sepia}. That is, each $\gamma_i(\psi)$ is modeled as a 
-% zero-mean Gaussian process with a squared exponential (i.e., Gaussian) covariance function:
-% \[
-% \gamma_i(\psi) \sim GP\left(0,\ \frac{1}{\lambda_{z_i}} R^{(i)} + \frac{1}{\lambda_{s_i}} I \right),
-% \]
-% with covariance function
-% \[
-% R^{(i)}_{uv} = \exp\left(-\sum_{k=1}^{p_\psi} \beta_{ik} (x_{uk} - x_{vk})^2 \right),
-% \]
-% where $\beta_{ik}$ is the inverse squared lengthscale for the $k$th input dimension, 
-% $\lambda_{z_i}$ is the marginal precision for $\gamma_i(\psi)$, and $\lambda_{s_i}$ is the nugget 
-% precision (set to zero if no nugget is used).
-% \textcolor{red}{(@Steve - The inputs to the covariance kernel
-% here need to be $\psi$.  Are we using a nugget?  If not, just remove it.  How are these parameters
-% being estimated?  Make sure this kernel notation matches what we used earlier as neatly as possible.)}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{mse_dot.jpeg}
+    \includegraphics[width=2in]{mse_dot.jpeg}
     \caption{Comparison of MSE for CAMB predictions, when training with DGP.FCO posterior 

### 2025-06-30 plotting/plot_data_CAMB.R

@@ -1,14 +1,14 @@
 
 ###############################################################################
-# This script fits the Bayesian hierarchical model for a particular CAMB model
-# and writes the results to a csv file.
+# This script plots the data and a Bayesian hierarchical model fit 
+# for the first CAMB batch of model runs (i.e., cosmology).
 #
 # Command line arguments:
 #    model: number of particular model (1-32)
 #    deep: indicator for whether to fit a DGP or GP (1 = deep, 0 = not deep)
 # 
 # Outputs:
-#    writes csv file of predicted values to the results folder
+#    optionally outputs JPEG image files (if JPG=TRUE)
 #
 ###############################################################################
 
@@ -17,39 +17,45 @@ library(zoo) # rollmean
 library(Matrix) # bdiag
 library(pracma) #interp1
 
-JPG = FALSE
-
+# Read in the CAMB model fit for the first cosmology, in order to make plots.
 load(file = "../fitting/results/CAMB/fit_1camb_image.rda")
 
-# # Plot the first batch of model runs for CAMB
-# plot(x_lo, y_lo[,1],
-#      col="gray", type="l", main = model, xlim=c(-3,0), ylim=c(-2,2),
-#      ylab = "script P (adj by h,g)",
-#      xlab = "log10(k) (adj by h)")
-# for (i in 2:n_lo) {
-#   if(model==8 & i==9) next
-#   lr = read.table(paste0("../CosmicEmu_etc/pows/M0",if(model<10){"0"},model,
-#                          "/L1300/PM0",if(i<10){"0"},i,"/output/m0",
-#                          if(model<10){"0"},model,".pk.ini"))
-#   y_lo[,i] = scrP(lr[which(h*lr[,1]<=x_ub),2]/h^3/g^2, 10^x_lo)
-#   lines(x_lo, y_lo[,i], col="gray")
-# }
-# lines(x_hi, y_hi, col="red")#, lwd=2)
-# lines(x_camb, y_camb, col="blue1",type="l", lwd=3, lty=3)
[…]

### 2025-06-30 plotting/plot_data_fit.R

@@ -11,7 +11,8 @@
 library(dgp.hm)
 library(zoo)
 
-JPG = FALSE
+# make JPEGs of plots?
+JPG <- FALSE
 
 # Get model data
 model <- 1
@@ -134,7 +135,7 @@ legend(x = "bottomright", legend = c("pert","low res","hi res", "wt avg", "UQ"),
 if(JPG) jpeg("../paper/plot_fit.jpeg", width = 12, height = 4, units = "in", res = 300)
 
 # Set layout: 2 columns, relative widths 2:1
-layout(matrix(1:2, nrow = 1), widths = c(2, 1))
+layout(matrix(1:2, nrow = 1), widths = c(1, 1))
 
 # Left plot (2/3 width)
 par(mar = c(4, 4, 2, 1))  # set margins: bottom, left, top, right
@@ -165,7 +166,7 @@ if (indx[1] == 0) indx[1] <- 1
 o <- order(fit$x)
 w <- fit$w[o, indx]
 for (i in 1:length(indx)) w[, i] <- w[, i] - mean(w[, i])
-matplot(log10(k), w, type = "l", xlab = "X", ylab = "W", col = "grey", 
+matplot(log10(k), w, type = "l", xlab = expression(paste(log[10](k))), ylab = "W", col = "grey", 
         ylim = c(min(w), max(w)))
 if(JPG) dev.off()
 

### 2025-06-30 plotting/plot_pca.R

@@ -1,6 +1,6 @@
 #############################################################
-# compare predictions: DGP and GP-PC approach vs cosmic emu #
-# also compare estimation method from process convolution.  #
+# Compare predictions: DGP.FCO approach vs Cosmic Emu       #
+# Create plots based on the predictions for each method.    #
 #############################################################
 
 library(dgp.hm)
@@ -177,11 +177,11 @@ for(i in 1:6){
        xaxt=ifelse(i %in% c(1,2,3),"n","s"),
           ylab='script P')
   # plot the cosmic emu prediction
-  lines(log10(k), cosmscrPsz - meanterm, col=cbcols[3], lty=3)
+  lines(log10(k), cosmscrPsz - meanterm, col=cbcols[3], lty=3, lwd=2)
   # # plot the emu train
   # lines(log10(k), as.numeric(convln_pred[i,]) - meanterm, col=cbcols[6], lty=2, lwd=2)
   # plot the dgp.fco train
-  lines(log10(k), etaEmu[,model_name] - meanterm, type="l", col = cbcols[7], lty=2)
+  lines(log10(k), etaEmu[,model_name] - meanterm, type="l", col = cbcols[7], lty=2, lwd=2)
   
   mtext(model_name, line = -2, adj=0.1)
   
@@ -234,10 +234,11 @@ etaEmu <- mean_pred + Reduce('+', eta_preds)
 
 # mean_PCs_oneW.jpeg; PCA picture
 if(JPG) jpeg("../paper/mean_PCs_oneW.jpeg", width = 12, height = 4, units = "in", res = 300)
-par(mfrow=c(1,3), mar=c(4,4,1.8,1))
-plot(log10(k), apply(eta,1,mean),type="l", xlab=expression(log[10](k)), ylab='script P')
-matplot(log10(k), bases, type="l", xlab=expression(log[10](k)), ylab = 'deviation from average')
+par(mfrow=c(1,3), mar=c(4,5,1.8,1))
+plot(log10(k), apply(eta,1,mean),type="l", xlab=expression(log[10](k)), ylab=expression(paste("\U1D4AB","  (k), centered")), cex.axis=1.5, cex.lab=1.5)
+matplot(log10(k), bases, type="l", xlab=expression(log[10](k)), ylab = 'deviation from average', cex.axis=1.5, cex.lab=1.5)
 x1_pred <- des_test[,1]
[…]

### 2025-06-30 paper/main.tex
specify data can be accessed via Github for reproducibility

@@
 Section \ref{sec:disc} concludes with a discussion of our contributions and avenues for 
-future research.  We provide reproducible code and an {\sf R} package for our 
+future research.  We provide data and reproducible code and an {\sf R} package for our 
 Bayesian DGP model in a public git repository.\footnote{\url{https://github.com/stevewalsh124/dgp.hm}}
@@
 
+\section{Disclosure}
+The authors report there are no competing interests to declare.
+
 \appendix

### 2025-07-01 paper/main.tex
include Data Availability Statement
rm extra line break
add middle initial; rm extra Mpc

@@
         Jared Clark\footnotemark[2] \and
-        Kelly Moran\thanks{Los Alamos National Laboratory} \and
+        Kelly R. Moran\thanks{Los Alamos National Laboratory} \and
         Katrin Heitmann\thanks{Argonne National Laboratory}}
@@
 $0.001 \leq k \leq 5 \text{ Mpc}^{-1}$, but each data type has unique wavenumber ranges
-where it is deemed approximately unbiased. For $k<0.04 \text{ Mpc}^{-1}$, only $y_p$ provides a reliable 
-estimate of the true spectra. $y_{\ell_r}$ are valid for $0.04 \leq k \leq 0.25 \text{ Mpc}^{-1}$, and 
-$y_h$ is valid for $0.04 \leq k \leq 5 \text{ Mpc}^{-1}$.  Figure \ref{fig:plot_data} shows an example 
+where it is deemed approximately unbiased. For $k<0.04$, only $y_p$ provides a reliable 
+estimate of the true spectra. $y_{\ell_r}$ are valid for $0.04 \leq k \leq 0.25$, and 
+$y_h$ is valid for $0.04 \leq k \leq 5$.  Figure \ref{fig:plot_data} shows an example 
 of the output for a particular cosmology. When we focus on $k$ values where $y_{\ell_r}$ 
@@
 random variable within the statistical model.  
-
 $\bar{Y}$ is typically pre-processed
@@
 
+\section{Data Availability Statement}
+We provide data and reproducible code and an {\sf R} package for our 
+Bayesian DGP.FCO model in a public git repository: \url{https://github.com/stevewalsh124/dgp.hm}
+
 \appendix
